### 2021-11-18 Paper/main.tex
commit draft 1 before Philippe/Clem comment

%%%%%%%%%%%%%%%%%%%%%%% Boundary-Layer Meteorology 2019 Template %%%%%%%%%%%%%%%%%%%%%%%%%

% \begin{filecontents*}{example.eps}
% gsave
% newpath
%   20 20 moveto
%   20 220 lineto
%   220 220 lineto
%   220 20 lineto
% closepath
% 2 setlinewidth
% gsave
%   .4 setgray fill
% grestore
% stroke
% grestore
% \end{filecontents*}

\RequirePackage{fix-cm}
\documentclass[smallextended]{svjour3}
\smartqed
\usepackage{appendix}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{lineno}
\usepackage{array}
\usepackage{longtable}
\usepackage{natbib}
\setcitestyle{aysep={}}
\linenumbers

\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
\expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
\expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
\renewenvironment{#1}%
{\linenomath\csname old#1\endcsname}%
{\csname oldend#1\endcsname\endlinenomath}}%
\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
\patchAmsMathEnvironmentForLineno{#1}%
\patchAmsMathEnvironmentForLineno{#1*}}%
\AtBeginDocument{%
\patchBothAmsMathEnvironmentsForLineno{equation}%
\patchBothAmsMathEnvironmentsForLineno{align}%
\patchBothAmsMathEnvironmentsForLineno{flalign}%
\patchBothAmsMathEnvironmentsForLineno{alignat}%
\patchBothAmsMathEnvironmentsForLineno{gather}%
\patchBothAmsMathEnvironmentsForLineno{multline}%
}

\usepackage[dvipsnames]{xcolor}
\newcommand{\cg}{\textcolor{WildStrawberry}}

\begin{document}

\title{Significant wind disturbances induced by giant dunes.}

\author{First Author         \and
        Second Author \and Third Author \and \{...\}
}

\institute{F. Author \at
              first address \\
              Tel.: +123-45-678910\\
              Fax: +123-45-678910\\
              \email{fauthor@example.com}
           \and
           S. Author \at
              second address
            \and
            T. Author \at
            third address
}

\date{Received: DD Month YEAR / Accepted: DD Month YEAR}

\maketitle

\begin{abstract}
  abstract
\keywords{Boundary layer \and Turbulent flow \and Sand dunes \and Fluide-structures interactions}
\end{abstract}

\newpage

\section{Introduction}
\section{Wind regimes across the Namib Sand Sea}

  \begin{figure}
    \centering
    \includegraphics[scale=1]{Figures/Figure1.pdf}
    \caption{Wind data used in this study \textbf{a}: Location of the studied sites. \textbf{b--e}: Satellite images of the studied sites (Google-Earth, Maxar Technologies, CNES/Airbus). In each subplot, the left and right wind roses represent the data from the ERA5Land climate reanalysis and the in situ stations, respectively. Note that the bars show the direction towards which the wind blows. The red dots show the location of the in situ stations.}
    \label{Fig1}
  \end{figure}

  \subsection{Datasets}

  Two wind datasets are used in this study. First, local winds are provided by stations situated in different environments (see Fig.~\ref{Fig1}). The wind strength and direction are measured every 10 minutes by cup anemometers and wind vanes, at heights between $n$ and $n$ meters depending on the station. The available period of measurements ranges from 1 to 5 discontinuous years distributed between 2012 and 2020 (see Fig.~\ref{Fig1_supp}). We checked that at least one complete seasonal cycle is available at each station.
  %
  Then, regional winds are extracted at the same locations and periods from the ERA5-Land dataset, which is a replay at a smaller spatial resolution of ERA5, the latest climate reanalysis from the ECMWWF \citep{Hersbach2020, munoz2021}. It provides hourly estimates of the 10-m wind velocity and direction at a spatial resolution of $\sim 9~\textup{km}$ ($0.1^\circ\times0.1^\circ$).

  For comparison, the local measurements are averaged into 1-hr bins centered on the temporal scale of the ERA5-Land estimates (see Fig.~\ref{Fig2_supp}). As the wind velocities of both datasets are provided a different heights, we convert them into shear velocities (see SI section~\ref{}), characteristic of the whole turbulent wind profile within the atmospheric boundary layer, which are then used together with the wind direction for further analysis. The resulting wind data are shown on the wind roses of Fig.~\ref{Fig1}(b--e).

  Finally, the dune properties are computed using an autocorrelation method from the 30-m Digital Elevation Models (DEMs) of the shuttle radar topography mission \citep{Farr2007}. For the South Namib and Deep Sea stations, we obtain respectively orientations of $85^\circ$ and $125^\circ$, wavelengths of $2.6~\textup{km}$ and $2.3~\textup{km}$ and amplitudes of $45~\textup{m}$ and $20~\textup{m}$ (see SI for additional details).

  \subsection{Agreement between local and regional winds}

  \begin{figure}
    \centering
    \includegraphics[scale=1]{Figures/Figure2.pdf}
    \caption{Temporal comparison between the wind data coming from the Era5Land climate reanalysis (orange lines) and from the in situ measurements (blue lines). \textbf{a--b}: Huab station. \textbf{c--d}: Deep Sea station in winter. \textbf{e--f}: Deep Sea station in summer.}
    \label{Fig2}
  \end{figure}

  The regional wind patterns in the Namib are strongly controlled by the see breeze, resulting in strong northward components present in all regional wind roses of figure~\ref{Fig1}, also deviated by the large scale topography \citep{lancaster1985}. These daily winds are the strongest during the second-half of the year (Septembre-January). In winter, an additional easterly component can be recorded during the night, induced by the combination of katabatic winds forming on the mountains, and infrequent `berg' winds, which are responsible of the high wind velocities observed \citep{lancaster1984}. The frequency of these easterly components decreases from the inland to the coast, resulting in three bidirectional wind regimes within the Namib Sand Sea and at the Adamax salt pan (Fig.~\ref{Fig1}(c, d, e)) and one unidirectional wind regime on the coast at the outlet of the Huab River.

  In the case of the Adamax and Huab stations, these regional wind roses qualitatively matches their respective in situ local measurements. However, for the Deep Sea and South Namib stations, the local wind roses exhibit additional components aligned with the giant dune orientation visible on the satellite images.

  Indeed, the analysis of the wind speed and direction time series shows that the matching between the local and regional winds is verified for the stations with flat environment at all times (Fig.~\ref{Fig2}(a--b)), as well as for the stations within the giant dune field, but only during Septembre-January (Fig.~\ref{Fig2}(c--d). This agreement is statistically verified, as shown by the Fig.~\ref{Fig5_supp}.

  \subsection{Influence of the giant dunes on local wind regimes}

  In winter, the local and regional winds match only during the morning for the stations within the giant dune field, when the southerly/southwesterly sea breeze dominates (see Fig.~\ref{Fig2}(e--f), Fig.~\ref{Fig3} and Fig.~\ref{Fig6_supp}). In the late afternoon and during the night, when the northwesterly `berg' and katabatic winds blow, the two datasets differs. More specifically, this mismatch mainly occurs during at low wind velocities. In this case, the angular wind distribution of the local measurements exhibits two additional modes separated of $\simeq 180^\circ$, each corresponding to the giant dune alignement (purple frame in Fig.~\ref{Fig3} and Fig.~\ref{Fig6_supp}). This deviation at low wind velocities is statistically verified, and is also associated with a global attenuation of the wind strength (Fig.~\ref{Fig6_supp} and Fig.~\ref{Fig7_supp}).

  \begin{figure}
    \centering
    \includegraphics[scale=1]{Figures/Figure3.pdf}
    \caption{Distributions of wind direction at the Deep Sea Station for the Era5Land climate reanalysis (orange) and the in situ measurements (blue). In each subplot, both distributions are plotted from the same time steps, selected with constraints on the wind direction (columns) and/or wind velocity (rows) in the Era5Land dataset. The vertical gray dashed lines indicate the dune orientation, and the top right numbers the percentage of the total number of time steps selected in each subplot. The purple frame highlights the regime (small wind velocities, nocturnal summer wind) during which the wind data from both datasets differs. A similar figure can be obtained for the Deep Sea station (see Fig.~\ref{Fig6_supp}).}
    \label{Fig3}
  \end{figure}

  \section{Influence of the circadian cycle of the atmospheric boundary layer}

  Previous studies have linked meteorological flow around and over obstacles to the vertical structure of the atmosphere \citep{Stull1988}. At the dune scale, its lower part, the turbulent atmospheric boundary layer (ABL), is directly influenced by the planetary surface \citep{Stull1988}. When the latter is warmer than the ambient air (typically during the day), a strong convection takes place, resulting in a logarithmic wind profile and a vertically constant potential temperature $\theta$ (which corresponds to the air density corrected from the vertical pressure variation). During the night, the surface cools faster than the ambient air, which can result in a stable layer close to the surface (typically the first $200$ meters). Above, the free atmosphere (FA) is a stably stratified zone in which turbulence is negligible, and where the flow is usually considered as incompressible and inviscid. In the middle, a transitional layer, also known as entrainment zone, is characterized by a sharp increase of the potential temperature, which traps turbulence resulting fro the surface friction below it. As it is much shallow than the other two layers, we consider it as capping singularity inducing a discontinuity in the potential temperature profile.

  \begin{figure}
    \centering
    \includegraphics[scale=1]{Figures/Figure4.pdf}
    \caption{\textbf{a}: Vertical profiles of the virtual potential temperature at 2 different time steps (blue - 29/11/2012 - 1100 UTC, red - 21/03/2017 - 1200 UTC) at the Deep Sea station. Dots: data from the ERA5 reanalysis. Transparent dashed lines: boundary layer height given by the ERA5 reanalysis, calculated from the bulk Richardson number \citep{seidel2012}. Plain lines: vertical (boundary layer) and linear (free atmosphere) fits to estimate the stratification properties. \textbf{b--d}: Sketches representing the interaction between the giant dunes and the atmospheric flow for different meteorological conditions. \textbf{e}: Streamlines qualitatively representing the effect of flow confinement, in the case of the Deep Sea station. The red and straight blue lines are calculated from the unconfined case, representing the situations \textbf{b--c}. The sinuous blue line represents the confined case of \textbf{d}. For details on their derivation, see Appendix.}
    %
    %
    % calculated from equation~\eqref{Tau}. Note that topography and wind geometry represent the case of the Deep Sea station. For the red and straight blue lines, $(\mathcal{A}_{0}, \mathcal{B}_{0}) = (3.8, 1.3)$, calculated from the unbounded case of \citep{Fourriere2010}, and representing the situations \textbf{b--c}. For the sinuous blue line, $(\mathcal{A}_{0}, \mathcal{B}_{0}) = (8, 2)$, representing the confined case of \textbf{d}.}
    \label{Fig4}
  \end{figure}

  \subsection{Relevant non-dimensional parameters and physical modeling}
  \label{theoretical_framework}

  As a result, the giant dunes evolve in a turbulent flow capped by a stratified atmosphere, with a transitional layer acting as a free surface in between. When the dune are small enough, typically in the case of dune emergence or early coarsening (Fig.~\ref{Fig4}b), they do no interact with the top of the ABL \citep{Fourriere2010}. However, the giant dunes present in the Deep Sea and South Namib stations have wavelengths comparable or larger than typical ABL depths, inducing their interaction with the capping layer and the free atmosphere above \citep{andreotti2009}. However, as sketched in Fig.~\ref{Fig4}c--d, the nature of this interaction depends on two non-dimensional number, both called Froude numbers in the literature~\citep{Vosper2004, Sheridan2006, Hunt2006, Jiang2014}.

  The ability of the capping layer to deform under the presence of an underlying topography is controlled by the \emph{surface Froude number}, typically used in the context of bedforms in rivers \citep{Fourriere2010, Andreotti2012}:
  \begin{equation}
    \mathcal{F}r_{\textup{S}} = \displaystyle\frac{U}{\sqrt{\displaystyle\frac{\Delta\rho}{\rho}gH}},
  \end{equation}
  where $U$ is the wind velocity at the top of the ABL, $H$ its depth, $\rho$ its average density and $\Delta\rho$ the density jump between the ABL and the FA.
  %
  Likewise, the FA stratification ability to deform is controlled by the \emph{internal Froude number}, typically used in the context of flow disturbances induced by mountains \citep{Stull1988}:
  \begin{equation}
    \mathcal{F}r_{\textup{I}} = \displaystyle\frac{U}{H N},
  \end{equation}
  where $N$ is the Brunt-Väisälä frequency, characteristic of a stratified fluid.
  %
  When the flow inertia dominates over its weight ($\mathcal{F}r_{\textup{S}} \gg 1$) and the weight of the stratification ($\mathcal{F}r_{\textup{I}} \gg 1$), both the capping layer and the stratification deform and adapt to the underlying topography (Fig.~\ref{Fig4}d), resulting in an unconfined flow \citep{refs}. However, an opposite force balance induces a rigid capping layer and/or stratification (Fig.~\ref{Fig4}c), resulting in confined flows. In this case, the confined flow has trouble going smoothly over the obstacle, and are then typically associated with large deviations and modulation of the wind velocity, as qualitatively illustrated in the case of the Deep Sea station by Fig.~\ref{Fig4}e \citep{refs}.

  \subsection{Flow regime diagrams}

  In the spirit of \citet{Sheridan2006}, we aim to compute flow regime diagrams in the space defined by the three relevant non-dimensional numbers presented above, $\left(k H,\, \mathcal{F}r_{\textup{S}}, \, \mathcal{F}r_{\textup{I}}\right)$. They are calculated from the time series of the geopotential, temperature and specific humidity vertical profiles available in the ERA5 climate reanalysis (see SI~\ref{?}). The relative velocity modulation is computed as
  %
  \begin{equation}
    \delta_{\textup{u}} = \frac{u_{*}^{\textup{ERA}} -  u_{*}^{\textup{station}}}{u_{*}^{\textup{ERA}}},
  \end{equation}
  %
  and the flow deviation as the minimal angle between the wind orientation in the two datasets:
  %
  \begin{equation}
    \delta_{\theta} = \left\vert \textup{min}\left(\left[\theta_{\textup{ERA}} - \theta_{\textup{station}}\right] \, \textup{mod.} \, 360, \left[\theta_{\textup{station}} - \theta_{\textup{ERA}}\right] \, \textup{mod.} \, 360\right) \right\vert,
  \end{equation}
  %

  When these two variables are represented in the marginal spaces $(k H,\, \mathcal{F}r_{\textup{S}})$ and $(k H,\, \mathcal{F}r_{\textup{I}})$, different regime emerges (Fig.~\ref{Fig5}). The small wind disturbances ($\delta_{\theta} \to 0$, $\delta_{\textup{u}} \to 0$) are located in the top-right part of the diagrams, corresponding to a regime mixing low-interactions ($k H$ large enough, Fig.~\ref{Fig4}b) and low-confinement ($\mathcal{F}r_{\textup{S}}, \, \mathcal{F}r_{\textup{I}}$ large enough, Fig.~\ref{Fig4}c).

  Lower values of $k H$ (stronger interaction) or Froude numbers (stronger confinement) then both leads to an increase in wind disturbances, both in terms of orientation and velocity. Interestingly, the limit of no-interactions between the topography and the boundary layer structure ($k H \gg 1$), in which the properties of the capping layer and the stratification become irrelevant (Fig.~\ref{Fig4}b--\ref{Fig4}c), is never reached here, in the case of giant dunes.
  %
  Below a threshold value of $k H \simeq 0.3$, wind disturbance occurs independently of the Froude numbers value. However, the latter seem to control a transition between from damped to amplified wind velocities within the interdune (Fig.~\ref{Fig5}c--\ref{Fig5}d), for which we do not have an explanation.

  \begin{figure}
    \centering
    \includegraphics[scale=1]{Figures/Figure5.pdf}
    \caption{Regime diagrams of the wind deviation $\delta_{\theta}$ and relative attenuation/amplification $\delta_{u}$, containing the data from both the Deep Sea and South Namib stations. Blue dashed lines empirically delimit the different regimes. The point density in each bin of the diagrams is shown in Fig.~\ref{Fig11_supp}.}
    \label{Fig5}
  \end{figure}

\section{Discussion}

 The comparison of local and regional wind data gives a direct evidence of the giant dunes feedback on the flow. In flat areas, the matching between both datasets highlights the ability of the latest generation of climate reanalysis to predict the wind flow up to scales $\sim 10~\textrm{km}$, i.e the grid model. When smaller scale topographies are present (giant dunes in our case), locally measured wind regimes may significantly differ from the regional ones. Furthermore, we link these disturbances induced by the dunes to their interaction with the lower part of the atmospheric vertical structure, and more specifically to its circadian variability. During the day, the top of the ABL is high enough to limit the interaction of the capping layer and the FA stratification with the giant dunes, resulting in a low flow confinement, and thus small wind disturbances. During the night, the small ABL height induces a stronger flow confinement, associated with large wind deviation and acceleration or deceleration. Interestingly, we also found that this effect could be counterbalanced by the presence of large wind velocities, capable of deforming the capping layer and/or the FA stratification and thus decreasing the confinement effect.

 Simple linear model such as the one of \citet{andreotti2009} also suggests that larger wind disturbances occur under strong flow confinement such as described above. However, they are unable to reproduce the magnitude of the observed deviations, probably due to the presence of hydrodynamical non-linear effects, negligible in low confinement situations, but not otherwise (see Fig.~\ref{Fig12_supp} and Appendix 1). Another limit in the comparison between theoretical predictions and measured is induces by the the single-point measurements. To have reliable representations of the flow structures related to wind disturbances, additional measurements in different places on and near the same topographical obstacle are needed.

 This study highlights the interaction between giant dunes and the atmospheric boundary layer, thus supporting for example the way the capping layer acts as a bounding surface limiting dune growth \citep{andreotti2009, gunn2021}. This interaction also have implications at smaller scales, where bedforms then develop from the disturbed wind instead of the regional one. Differences between larger and smaller scale (thus older and more recent) dune patterns are observed ubiquitously, and have sometimes in the literature been attributed to climatic changes in wind regimes \citep{refs}. Here, we suggest using this feedback mechanism that current winds can explain dune patterns at all scales, such as the linear dunes ($\sim 50$ m -wide) elongating within the interdune between two giant linear dunes ($\sim 2$ km -wide) in the Namib Sand Sea (see Fig.~\ref{Fig6}).

\begin{figure}
  \centering
  \includegraphics[scale=1]{Figures/Figure6.pdf}
  \caption{Implications for smaller scale patterns in (a) the South Namib and (b) Deep Sea. The ellipses indicates the different types of elongating dunes, at large (plain) and small scale (dashed). The dune orientations are calculated using the model of \citet{Cour14} from the sand flux angular distributions, shown here for typical sand quartz grains of $180~\mu$m. The double blue and single red arrows correspond to the two possible dune growth mechanisms, bed instability and elongation, respectively. Likewise, plain arrows are calculated from the ERA5-Land datasets, and dashed arrows from the in situ measurements. Wedges show the uncertainty on the orientation calculation, and the arrows correspond to typical parameters found in the literature, i.e a grain diameter of $180~\mu$m and a flux-up ratio of 1.6. The green dots indicate the position of the measurement stations. See Appendix 2 for details.}
  \label{Fig6}
\end{figure}

\begin{acknowledgements}
These should follow the concluding section of the paper and precede the References and any appendices, if they are present. The acknowledgements section does not require a section number.
\end{acknowledgements}

\section*{Appendix 1: ABL turbulent wind model}

Following the work of \citet{Fourriere2010} and \citet{Andreotti2012}, we briefly expose in this section the linear response of a turbulent flow to a small aspect ratio perturbation of the topography $\xi$. As this topography can be decomposed into several sinusoidal modes, we focus on the response to a sinusoidal topography as:
\begin{equation}
  \xi = \xi_{0}\cos\left[k\left(\cos(\alpha)x + \sin(\alpha)y\right)\right],
\end{equation}
which is also a good approximation to the giant dunes observed in the Deep Sea and South Namib Station (see Fig~\ref{Fig1} and Fig~\ref{Fig4_supp}). Here, $x$ and $y$ are the streamwise and spanwise coordinates, $k=2\pi/\lambda$ the wavenumber of the sinusoidal perturbation, and $\alpha$ its crest orientation, calculated with respect to the $y$--direction.

In terms of basal shear stress $\tau = \rho u_{*}^{2}$, the flow response can then generally be written in as:
\begin{align}
  \tau_{x} & = \tau_{0}\left(1 + k\xi_{0}\sqrt{\mathcal{A}_{x}^{2} + \mathcal{B}_{x}^{2}}\cos\left[k\left(\cos(\alpha)x + \sin(\alpha)y\right) + \phi_{x}\right]\right), \\
  \tau_{y} & = \tau_{0}k\xi_{0}\sqrt{\mathcal{A}_{y}^{2} + \mathcal{B}_{y}^{2}}\cos\left[k\left(\cos(\alpha)x + \sin(\alpha)y\right) + \phi_{y}\right],
\end{align}
where $\tau_{0}$ is the basal shear stress on a flat bed, and $\phi_{x, y} = \tan^{-1}\left(\mathcal{B}_{x, y}/\mathcal{A}_{x, y}\right)$. The in--phase and in--quadrature hydrodynamical coefficients $\mathcal{A}_{x, y}$ and $\mathcal{B}_{x, y}$ are functions of the flow conditions, i.e the bottom roughness, the free surface or the incident flow direction \citep{Fourriere2010, andreotti2009, Andreotti2012, Charru2013}.

\citet{Andreotti2012} have shown that the impact of the incident wind direction can be well approximated by the following expressions:
\begin{align}
  \mathcal{A}_{x} & = \mathcal{A}_{0}\cos^{2}\alpha, \\
  \mathcal{B}_{x} & = \mathcal{B}_{0}\cos^{2}\alpha, \\
  \mathcal{A}_{y} & = \displaystyle\frac{1}{2}\mathcal{A}_{0}\cos\alpha \sin\alpha, \\
  \mathcal{B}_{y} & = \displaystyle\frac{1}{2}\mathcal{B}_{0}\cos\alpha \sin\alpha,
\end{align}
where $\mathcal{A}_{0}$ and $\mathcal{B}_{0}$ are now two coefficients independent of the dune orientation $\alpha$. In the case of a fully turbulent boundary layer capped by a free atmosphere capping, they now only depend on $k H$, $k z_{0}$, $\mathcal{F}r_{\textup{I}}$ and $\mathcal{F}r_{\textup{S}}$, as detailed by \citet{andreotti2009}. More specifically, their variation in the marginal spaces $(k H,\, \mathcal{F}r_{\textup{S}})$ and $(k H,\, \mathcal{F}r_{\textup{I}})$ are shown in Fig.~\ref{Fig12_supp}.

Typical values for the unconfined case are therefore $\mathcal{A}_{0} = 3.4$ and $\mathcal{B}_{0} = 1$. In our case of giant dunes with $k\xi_{0} \sim 0.1$, significant wind disturbances are then expected when $\sqrt{\mathcal{A}_{0}^{2} + \mathcal{B}_{0}^{2}} \sim 10$. However, this is also the limit of the linear regime where this theoretical model is applicable, as hydrodynamical non-linearities become significant when $k\xi_{0}\sqrt{\mathcal{A}_{0}^{2} + \mathcal{B}_{0}^{2}} \sim 1$.

\section*{Appendix 2: Sediment transport and dune morphodynamics}

Here, we briefly detail the sediment transport and dune morphodynamics theoretical framework leading to the prediction of sand fluxes and dune orientations from wind data.

The sediment fluxes can been directly linked to the wind basal shear stress at each time steps $t$ from transport laws, whose exact forms depends on the sediment transport mechanisms taken into account. In this work, we following the recent work of \citet{Pahtz2020}, where the sediment flux $q_{\textup{sat}}$ on a flat bed made of loose sand can be expressed as:
\begin{equation}
  \label{transport_law}
  \frac{q_{\textup{sat, t}}}{Q} = \frac{2\sqrt{\Theta_{\textup{th}}}}{\kappa\mu}\left(\Theta_{t} - \Theta_{\textup{th}}\right)\left(1 +\frac{C_{\textup{M}}}{\mu}\left[\Theta_{t} - \Theta_{\textup{th}}\right]\right),
\end{equation}
where $\kappa = 0.4$ is the von Kármán constant, $C_{\rm M}=1.7$ a constant, $Q = d\sqrt{(\rho_{\textup{s}} - \rho)gd/\rho}$ is a characteristic flux, with $\rho_{\textup{s}} = 2.6~\textup{g}~\textup{cm}^{-3}$ and $d=180~\mu\textup{m}$ the grain density and diameter, and $g$ the gravitational acceleration. The friction coefficient $\mu$ is taken to be the avalanche slope of the granular material, i.e $\sim 0.6$. Finally, the Shields number is defined as $\Theta = \rho u_{*, t}^{2}/(\rho_{\textup{s}} - \rho)gd$, and its threshold value for incipient sediment transport as been calibrated using laboratory experiments to $\Theta_{\textup{th}} = 0.0035$.

The dune orientations are then predicted from the dimensional model of \citet{Courrech2014}. The orientation $\alpha$ corresponding the bed instability is then the one that maximizes the following growth rate:
\begin{equation}
  \sigma \propto \frac{1}{H W T}\int_{t}  q_{\textup{crest}, t}\vert \sin\left(\theta_{t} - \alpha\right) \vert,
\end{equation}
where $H$ and $W$ are dimensional constants representing the dune height and width, respectively. The flux at the crest is expressed as:
\begin{equation}
  q_{\textup{crest}, t} = q_{\textup{sat}, t}\left[1 + \gamma\vert\sin\left(\theta_{t} - \alpha\right)\vert\right],
\end{equation}
where the flux-up ratio $\gamma$ has been calibrated to 1.6 using field studies, underwater laboratory experiments and numerical simulations. Similarly, the dune orientation corresponding to the elongation mechanism is the one that verifies:
\begin{equation}
  \tan(\alpha) = \frac{\langle q_{\textup{crest}, t}(\alpha) \boldsymbol{e}_{\theta_{t}}\rangle \cdot \boldsymbol{e}_{WE} }{ \langle q_{\textup{crest}, t}(\alpha) \boldsymbol{e}_{\theta_{t}} \rangle \cdot \boldsymbol{e}_{SN}},
\end{equation}
where $\langle.\rangle$ denotes a vectorial time average. The unitary vectors $\boldsymbol{e}_{WE}$, $\boldsymbol{e}_{SN}$ and $\boldsymbol{e}_{\theta_{t}}$ are in the West--East, South--North and wind direction, respectively.

The computed dune orientations, blue and red arrows in figure~\ref{Fig6}, are however depending on a large number of parameters, for which we took typical values for eolian desert on Earth. We therefore run a sensibility test by calculating the dune orientations for grain diameters ranging from $100~\mu\textup{m}$ to $400~\mu\textup{m}$ and the speed-up ratio from $0.1$ to $10$ (wedges on figure~\ref{Fig6}). We also checked the sensibility the transport low by repeating the process with the quadratic transport also used for comparison in \citet{Pahtz2020}, which led to no more than $n\%$ of variation with respect to \eqref{transport_law}.

\clearpage

\bibliographystyle{spbasic_updated}
\bibliography{biblio}

\newpage
\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{figure}{0}

\section*{Supplementary Material for \textit{Boundary-Layer Meteorology} Sample Paper: Instructions for Authors}

{\textbf{First Author* $\cdot$ Second Author $\cdot$ Third Author \\}}
\\
\text{*}Affiliation and email address for the corresponding author only (note that the corresponding author does not need to be the first author).

\begin{figure}
  \centering
  \includegraphics[scale=1]{Figures/Figure1_supp.pdf}
  \caption{Gant chart representing the usable time steps for the two data sets, for all stations.}
  \label{Fig1_supp}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[scale=1]{Figures/Figure2_supp.pdf}
  \caption{Comparison between raw in situ wind measurements, and centered averaged data over one hour for the South Namib station. \textbf{a}: wind direction. \textbf{b}: wind velocity at the measurement height, 2.6 m.}
  \label{Fig2_supp}
\end{figure}

\section*{1. Calibration of the hydrodynamical roughness}

For each station, the hydrodynamic roughness is calibrated by finding the one that minimizes the relative difference $\delta$ between the wind vectors of both datasets:
\begin{equation}
  \label{metric_roughness}
      \delta = \frac{\sqrt{\langle\| \boldsymbol{u}_{*, \textrm{era}} - \boldsymbol{u}_{*, \textrm{station}} \|^{2}\rangle_{t}}}{\sqrt{ \langle \| \boldsymbol{u}_{*, \textrm{era}} \| \rangle_{t}\langle \| \boldsymbol{u}_{*, \textrm{station}} \| \rangle_{t}}}
\end{equation}

This $\delta$--parameter is computed for hydrodynamic roughness values ranging from $10^{-5}$~m to $10^{-2}$~m for the different stations. A shown by figure~\ref{Fig3_supp}, the minimum of $\delta$ in the space ($z_{0,\, \textup{Era}}$, $z_{0,\, \textup{in situ}}$) forms a line. We thus take the roughness of the Era5Land dataset as the typical value when sediment transport occurs, $10^{-3}$~m, corresponding to the thickness of the transport layer \citep{Sherman2007}. It leads for the Adamax, Deep Sea, Huab and South Namib stations values of $2.7~\textup{mm}$, $0.76~\textup{mm}$, $0.12~\textup{mm}$ and $0.48~\textup{mm}$, respectively.

The choice of the hydrodynamic roughness values only impacts the calculated shear velocities, but note the wind directions. As such, most of our conclusions are then independent of such a choice, and only the magnitude of the wind velocity attenuation in confined situation might be affected.

\begin{figure}
  \centering
  \includegraphics[scale=1]{Figures/Figure3_supp.pdf}
  \caption{Calibration of the hydrodynamic roughnesses. The metric $\delta$ defined in \eqref{metric_roughness} is represented in colorscale as a function of the hydrodynamic roughnesses chosen for the Era5-Land and in situ datasets, for the Adamax (\textbf{a}), Deep Sea (\textbf{b}), Huab (\textbf{c}) and South Namib (\textbf{d}) Stations. The red dashed and plain lines shows the minima of $\delta$ and the identity line. The orange lines and dots highlights the chosen the hydrodynamic roughnesses for the in situ datasets by imposing $z_{0, \textup{\textup{ERA}}} = 1~\textup{mm}$, leading for each station to $2.7~\textup{mm}$, $0.76~\textup{mm}$, $0.12~\textup{mm}$ and $0.48~\textup{mm}$, respectively.}
  \label{Fig3_supp}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[scale=1]{Figures/Figure4_supp.pdf}
  \caption{Analysis of the DEMs of the Deep Sea (left column -- \textbf{a}, \textbf{b}, \textbf{c}) and South Namib (right column -- \textbf{d}, \textbf{e}, \textbf{f}) stations. \textbf{a--c}: Detrended topography (a second order polynomial is first fitted and then removed) . \textbf{b--d}: autocorrelation matrix shown in colorscale. The black line shows the detected orientation, and the red line the profile along which the wavelength is calculated, shown in \textbf{e--f}. The blue lines and dots show the first peak of the autocorrelation profile, whose horizontal position gives the characteristic wavelength of the dune pattern.}
  \label{Fig4_supp}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[scale=1]{Figures/Figure5_supp.pdf}
  \caption{Statistical agreement of the wind orientation (\textbf{a}) and velocity (\textbf{b}) between the Era5Land dataset and the in situ measurements for the Huab and Adamax stations. Note how the points are clustered around identity lines, black and dashed.}
  \label{Fig5_supp}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[scale=1]{Figures/Figure6_supp.pdf}
  \caption{Distributions of wind direction at the South Namib Station for the Era5Land climate reanalysis (orange) and the in situ measurements (blue). In each subplot, both distributions are plotted from the same time steps, selected with constraints on the wind direction (columns) and/or wind velocity (rows) in the Era5Land dataset. The grey dashed vertical lines indicate the dune orientation. The purple frame highlights the regime (small wind velocities, nocturnal summer wind) during which the wind data from both datasets differs. A similar figure can be obtained for the South Namib station (see Fig.~\ref{Fig3}).}
  \label{Fig6_supp}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[scale=1]{Figures/Figure7_supp.pdf}
  \caption{Statistical comparison of the wind orientation between the Era5Land dataset and in situ measurements for the South Namib and Deep Sea stations, for different velocity ranges. \textbf{a}: $u_{*, \textup{\textup{ERA}}} < 0.1~\textup{m}~\textup{s}^{-1}$. \textbf{b}: $0.1 < u_{*, \textup{\textup{ERA}}} \leq 0.25~\textup{m}~\textup{s}^{-1}$. \textbf{c}: $u_{*, \textup{\textup{ERA}}} \geq 0.25~\textup{m}~\textup{s}^{-1}$. Note that the dune orientations measured are substracted to the wind orientation, which allows to plot both stations on the same graph. Black dashed lines indicates in situ orientations aligned with the dune crests (here $0^\circ$, $180^\circ$ and $360^\circ$ -- \textbf{a}, \textbf{b}), as well as the identity lines (\textbf{b}, \textbf{c}).}
  \label{Fig7_supp}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[scale=1]{Figures/Figure8_supp.pdf}
  \caption{Statistical comparison of the wind velocity between the Era5Land dataset and in situ measurements for the South Namib and Deep Sea stations. \textbf{a}: Nocturnal summer easterly wind. \textbf{b}: Diurnal southerly wind. Black dashed lines are identity lines. The angle ranges corresponding to diurnal and nocturnal summer winds are those taken in Fig.~\ref{Fig3} and Fig.~\ref{Fig6_supp}.}
  \label{Fig8_supp}
\end{figure}

\section*{2. Extraction of the ABL properties}

In order to estimate the relevant non-dimensional numbers, one need to estimate in addition to the wind and dune properties some parameters of the ABL. The Era5 dataset provides a direct bulk estimate of the ABL depth $H$ from a bulk Richardson number calculation, as well as vertical profiles of the geopotential $\phi$, temperature $T$ and specific humidity $e_{\textup{w}}$ at given pressure levels $P$. From these quantities, the virtual potential temperature, which takes into account the vertical pressure and humidity changes, can be calculated as:
\begin{equation}
  T_{\textup{vp}} = T\left(1 + \left[R_{\textup{M}} - 1\right]e_{\textup{w}}\right)\left(\frac{P_{0}}{P}\right)^{P_{\textup{c}}(1 - 0.24e_{\textup{w}})},
\end{equation}
where $P_{0} = 10^{5}~\textup{Pa}$ is the standard pressure, $P_{\textup{c}} = 0.2854$ the Poisson coefficient for dry air and $R_{\textup{M}}=1.61$ is the ratio between the melocular masses of dry air and water. The vertical coordinates are calculated as:
\begin{equation}
  z = \frac{\phi R_{\textup{t}}}{g R_{\textup{t}} - \phi},
\end{equation}
where $R_{\textup{t}} = 6356766~\textup{m}$ is the average Earth radius, and $g = 9.81~\textup{m}~\textup{s}^{-2}$ the gravitational acceleration.

Example of obtained vertical profiles of the virtual potential temperature are shown in Fig.~\ref{Fig9_supp}. On each of them, an average is computed below the ABL depth given by the Era5 dataset, and a linear function is fitted above.

Under the Boussinesq approximation, the temperature variations are assumed to induce most of those of the density, leading to $\Delta\rho/\rho \simeq \Delta T_{\textup{vp}}/T_{\textup{vp}}$. Here, $T_{\textup{vp}}/T_{\textup{vp}}$ is the relative virtual potential temperature jump at the capping, directly measured on the vertical profiles.

Following \citet{Tritton2012}, the relative density jump at the capping layer

\begin{figure}
  \centering
  \includegraphics[scale=1]{Figures/Figure9_supp.pdf}
  \caption{\textbf{a}: Vertical profiles of the virtual potential temperature at 3 different time steps (blue - 29/11/2012 - 1100 UTC, orange - 21/03/2017 - 1200 UTC, green - 21/03/2017 - 2000 UTC) at the South Namib station. Dots: data from the ERA5 reanalysis. Transparent dashed lines: boundary layer height given by the ERA5 reanalysis, calculated from the bulk Richardson number \citep{seidel2012}. Plain lines: vertical (boundary layer) and linear (free atmosphere) fits to estimate the quantities in Fig.~\ref{Fig10_supp}. \textbf{b}: Examples of ill-processed vertical profiles at 3 different time steps (blue - 2/12/2013 - 2300 UTC, orange - 20/03/2017 - 0000 UTC, green - 14/07/2017 - 1400 UTC) at the South Namib station. \textbf{c}: Hourly distribution of ill-processed vertical profiles. \textbf{d}: Monthly distribution of ill-processed vertical profiles. \cg{These profiles are ill-processed because the temperature found at the boundary layer from the linear fit in the free-atm is smaller than the average one inside the boundary layer. This is an unstable situation, which does not allow to calculate the surface Froude number.}}
  \label{Fig9_supp}
\end{figure}

\begin{figure}
  \centering
  \includegraphics{Figures/Figure10_supp.pdf}
  \caption{Distributions of the meteorological parameters resulting from the processing of the Era5-Land data for the South Namib (blue) and the Deep Sea (orange) stations.}
  \label{Fig10_supp}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[scale=1]{Figures/Figure11_supp.pdf}
  \caption{Non-dimensional parameters distributions. For the marginal distributions, the orange correspond to the South Namib station, and the blue to the Deep Sea station.}
  \label{Fig11_supp}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[scale=1]{Figures/Figure12_supp.pdf}
  \caption{Physical interpretation of the flow disturbance. (a) and (b) Magnitude of the disturbance induced by a sinusoidal topography calculated from the time series of the non-dimensional numbers presented in Figures~\ref{Fig4} and \ref{Fig5} using the linear model of \citet{andreotti2009}. (c) Shear velocity streamlines represented in the case of the Deep Sea station, for increasing values of $\sqrt{\mathcal{A}_{0}^{2} + \mathcal{B}_{0}^{2}}$. From the upper to the lower streamline, values of $\left(kH,\, Fr_{\textup{surface}},\, Fr_{\textup{internal}},\, \mathcal{A}_{0},\, \mathcal{B}_{0},\, \sqrt{\mathcal{A}_{0}^{2} + \mathcal{B}_{0}^{2}}\right)$ are (1.9, 0.6, 1.5, 3.4, 1.0, 3.5), (1.5, 0.3, 0.4, 4.8, 1.4, 5.0), (0.1, 3.5, 1.0, 8.6, 0.1, 8.6), (0.5, 0.05, 0.04, 9.6, 2.5, 9.9).}
  \label{Fig12_supp}
\end{figure}

\end{document}

### 2021-11-18 Paper/template/BLM_LaTeX_template_bibdesk.tex
commit draft 1 before Philippe/Clem comment

%%%%%%%%%%%%%%%%%%%%%%% Boundary-Layer Meteorology 2019 Template %%%%%%%%%%%%%%%%%%%%%%%%%

\begin{filecontents*}{example.eps}
gsave
newpath
  20 20 moveto
  20 220 lineto
  220 220 lineto
  220 20 lineto
closepath
2 setlinewidth
gsave
  .4 setgray fill
grestore
stroke
grestore
\end{filecontents*}

\RequirePackage{fix-cm}
\documentclass[smallextended]{svjour3}
\smartqed
\usepackage{appendix}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{lineno}
\usepackage{array}
\usepackage{longtable}
\usepackage{natbib}
\setcitestyle{aysep={}}
\linenumbers

\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
\expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
\expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
\renewenvironment{#1}%
{\linenomath\csname old#1\endcsname}%
{\csname oldend#1\endcsname\endlinenomath}}%
\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
\patchAmsMathEnvironmentForLineno{#1}%
\patchAmsMathEnvironmentForLineno{#1*}}%
\AtBeginDocument{%
\patchBothAmsMathEnvironmentsForLineno{equation}%
\patchBothAmsMathEnvironmentsForLineno{align}%
\patchBothAmsMathEnvironmentsForLineno{flalign}%
\patchBothAmsMathEnvironmentsForLineno{alignat}%
\patchBothAmsMathEnvironmentsForLineno{gather}%
\patchBothAmsMathEnvironmentsForLineno{multline}%
}

\begin{document}

\title{Sample Paper for \textit{Boundary-Layer Meteorology}: Instructions for Authors
}

\author{First Author         \and
        Second Author \and Third Author \and \{...\}
}

\institute{F. Author \at
              first address \\
              Tel.: +123-45-678910\\
              Fax: +123-45-678910\\
              \email{fauthor@example.com}
           \and
           S. Author \at
              second address
            \and
            T. Author \at
            third address
}

\date{Received: DD Month YEAR / Accepted: DD Month YEAR}

\maketitle

\begin{abstract}Limit the Abstract to 250 words. The Abstract should not be overly descriptive, should focus on main results and conclusions, and should not contain any undefined abbreviations. Acronyms, if needed, must be defined at first use. Avoid citing literature, but if absolutely necessary, the reference should be given as, e.g., ``based on Gheynani and Taylor (Boundary-Layer Meteorology, 2010, Vol. 137, 223--236)''. The use of mathematical symbols in the Abstract should be avoided.
\keywords{Alphabetical order \and Boundary-layer meteorology \and Five \and \LaTeX \and Manuscript preparation
\newline
 \{Keywords should be in alphabetical order with the first letter of each keyword in upper case. No more than five keywords should be used and the terms themselves should be no more than three words in length, as a rule.\}}
\end{abstract}

\section{Introduction}
\label{intro}
Start writing the Introduction here. Carry on to the next page, ensuring that author information remains at the bottom of the first page. Lines and pages should be numbered. The font used should be clearly legible, and symbols in the font (including subscripts and superscripts) should be checked for legibility.

\section{Section Title}
\label{sec:1}
The remaining body of the text should be placed here, divided appropriately into sections. Individual words in all section, subsection, and secondary subsection titles should start with upper case letters. Avoid hanging titles by keeping the title and section text together on the same page. Do not use acronyms in section titles. Sections should be referred to in the text as Sect. 1, unless starting a new sentence, in which case Section 1 should be used. Multiple sections should be referred as Sects. 1 and 2, or Sects. 3–5.

\section{Next Section Title}
Text can be further divided into subsections as demonstrated below.

\subsection{Acronyms}
All acronyms should be defined at first use, both within the Abstract and in the main text. If an acronym is defined in the Abstract, it should be defined again at first use in the main body of text. Acronyms should not be used in manuscript titles, and excessive usage of acronyms should be avoided, particularly of those which are not commonly employed in meteorological literature. Two-letter acronyms should be used in exceptional cases and only for well-established word combinations. When making acronyms plural, make sure to use an ``s" (e.g., the term ``low-level jets" becomes ``LLJs"). Acronyms that are used as variables should be written in Italic (e.g., $TKE$, $LAI$), however, one-letter notation for variables, such as $e$ for turbulence kinetic energy (TKE), is preferable.

\subsection{Spelling and Grammar}
Generally, British (UK) spelling should be used. Spelling examples for frequently used terms include: airflow, anticyclonic, autocorrelation, behaviour, centre, colour, cospectrum, covariance, cross-section, cross-spectrum, dataset, daytime, freestream, grey, lidar, metre, nonlinear, night-time, point-of-view, set-up, subgrid, subrange, time scale, timestep, turbulence intensity, turbulence kinetic energy.

It should be noted that though British spelling is generally used, when it comes to words ending in yse/yze and ise/ize, the American z-form is used, such as in analyze, characterize, idealize, normalize, and parametrize.

Personal names should be spelled and formatted correctly, in particular: Boussinesq, Kolmogorov, Obukhov, Prandtl, V{\"a}is{\"a}l{\"a}, and von K{\'a}rm{\'a}n.

Above ground level, above sea level, left-hand side, random mean square, and right-hand side are abbreviated, respectively, as a.g.l., a.s.l., l.h.s., r.m.s., and r.h.s.

Geographical directions should be written as south, north-west, south-east, north-north-west, etc. A clause involving two words should be hyphenated (-) when used as an adjective, but not when otherwise used (e.g., boundary layer, boundary-layer depth, wind tunnel, wind-tunnel observations, turbulence of small scale, small-scale turbulence, 10-m wind speed). The en dash (--) should be used in Monin--Obukhov theory, Brunt--V{\"a}is{\"a}l{\"a} frequency, London--Paris railway, linear--log plot, atmosphere--ocean interaction, north--south, in the period 1970--2000, and in other cases when it could be replaced with the words ``and", ``to", ``through", etc. Use of the Oxford comma is required.

\subsection{Units}
The International System of Units (SI) and derived SI units should be used (e.g., m, km, s). The units should be typed in Roman font, not in Italic. Units requiring an exponent should be typed with a space between the portions of the unit, and using superscripts for the power, e.g., m~s$^{-1}$, kg~m$^{-3}$, J~kg$^{-1}$~K$^{-1}$ (do not write these as m/s, kg/m$^3$, J/kg/K).

\subsection{Variables and Symbols}
All variables should be typed in an appropriate font (see Sect. 3.1), and written consistently throughout the main text, the figure captions, figure axis legends, and in tables. Generally, variables should be written in Italic (e.g., $p$, $T$, $\rho$, $\beta$, $\gamma$, $\theta$, $e$, $LAI$), except for Greek capital letters, which should not be italicized and vectors, which are Bold (e.g., \textbf{v}, \textbf{F}). Mathematical signs used in the text should have a space on either side of the sign (e.g., write $x$ = 0.1 m, $\beta$ $<$ 3, $z/L$ $\geq$ 5, etc.). Functions (e.g., log, ln, exp, sin, cos, tan, arcsin, arccos, arctan, min, max, etc.), e (base of natural logarithm), i (imaginary unit), and $\pi$ (the number Pi) should be in Roman font. Dimensionless parameters/numbers, like Reynolds number ($Re$), Richardson number ($Ri$), Rossby number ($Ro$), etc., should be written in Italic.

When writing numbers in scientific notation, use the multiplication symbol rather than the letter x (e.g., write 4 $\times$ 10$^{-3}$ rather than 4 x 10$^{-3}$). To indicate approximate equality, use the symbol $\approx$ rather than the symbol $\sim$, which should be used to indicate ``on the order of". The symbol $\propto$ is reserved to indicate proportionality. Avoid beginning sentences with mathematical symbols or expressions.

In \textit{Boundary-Layer Meteorology}, ``Obukhov length'' is used rather than ``Monin--Obukhov length''. The surface-layer and boundary-layer `star' variables (scales) should be written in the format $T_*$, $u_*$, $q_*$, and $w_*$, i.e., with a subscript asterisk.

\subsection{Equations}
Line equations should be centred in the line. Equations to which reference is made elsewhere in the text must be numbered sequentially, starting with (1), and equations that are not referenced in the text need not be numbered. The numbering should continue through the text and into the appendices, if present. Symbols used in the equations should appear in the same format as in the text. Where an equation (e.g., number 10) has several parts, these parts should be indicated as (10a), (10b), (10c), etc., with each part on a separate line. Equations should be included within sentence structures, if possible, with surrounding punctuation used as appropriate. All variables that appear in the equations for the first time should be explained.

A numbered line equation example:
\begin{equation}
\overline{(\delta{T})^2}(\textbf{r},t)=\overline{[T(\textbf{x},t)-T(\textbf{x}+\textbf{r},t)]^2},
\end{equation}
where $T$ is temperature, $\overline{(\delta{T})^2}$ is the temperature structure function, \textbf{x} is a position vector, \textbf{r} is a separation vector, the overbar denotes spatial averaging, and $t$ is time. A non-numbered equation example:
\begin{equation*}
\overline{(\delta{T})^2}(\textbf{r},t)=\overline{[T(\textbf{x},t)-T(\textbf{x}+\textbf{r},t)]^2}.
\end{equation*}
The paragraph following a line equation should not be indented. Equations should be referred to in the text as Eq. 1, unless starting a new sentence, in which case Equation 1 should be used. Referring to equations by their number, e.g., ``as indicated by (1)", or ``the right-hand side of (25)" is also acceptable.

Equations presented in the appendices should continue the sequential numbering from the main text, i.e., if the last equation in the main text is (20), then the first equation in the appendices should be (21). If multiple appendices with equations are included then the sequential numbering continues throughout the appendices in the order the equations are presented.

\subsection{Times and Dates}
Times should be written in the format 0000 UTC, 1523 UTC, etc., using the 24-hour clock and no colons. If times correspond to local time, then at first introduction LT should be defined and the difference from UTC should be stated, e.g., ``at 1945 LT (local time = UTC $-$ 6 h)". The acronym UTC does not need to be defined.

Dates in the main text of a manuscript should be written in a date-month-year format, e.g., 23 April 2011, 7 January 2016. Dates that are being included in a table or figure can be shortened to the format 23/04/2011, 07/01/2016, with day before month.

\subsection{Instruments}
The make and model of instruments used in experimental campaigns reported in the manuscript should be listed. For example, ``An eddy-covariance gas analyzer (LI-7500DS, LI-COR, Lincoln, Nebraska, USA) was used to measure the water vapour density" or ``wind profiles were recorded using a PCS.2000 Doppler sodar (Metek, Elmshorn, Germany)".

\subsection{Citations}
Citations should be presented in an appropriate format, for example: ``as found by \cite{Mason+1987}'', ``\cite{Garratt1994} demonstrated that...'', ``as found in previous studies \citep{Mason+1987,Garratt1994,Wyngaard2004,Marusic++2011}", and should be given in order of year. Note that commas are not used to separate author name and citation year. This is accomplished by adding \verb\setcitestyle{aysep={}}\ after \verb\usepackage{natbib}\ at the beginning of the document.

If two papers within a group of citations are from the same authors, these should be listed together and ordered by the oldest cited article, for example, ``\citep{Mason+1987,Beljaars+1990,Beljaars+1991,Garratt1994}". Please note that articles currently in preparation or in review should not be cited, but those available via early online release may be cited using the DOI.

\subsubsection{Further Subsections}
If secondary subsections are required, the headings should be italicized.

\section{Figures}

Figures should be included as demonstrated by Fig. \ref{fig1} and numbered sequentially, starting at number 1. Figures with multiple panels should have panels labelled as a, b, c, etc. When referring to a figure in the text, use ``Fig." unless starting a new sentence, when ``Figure" is appropriate. For example, ``... as illustrated by the blue dashed line in Fig. \ref{fig1}." or ``Figure \ref{fig1} shows that ...". Multiple-panel figures can be referred to using, for example, ``... as illustrated by the blue dashed line in Fig. 2b, d." or ``Figure 2c shows that ...".

All figures should include a caption. The figures should be placed within the appropriate section in the main text. The number of figures should not exceed 15. All figures should be checked for legibility and consistency of the figure contents, axes labels, and any legends. It is preferable to use mathematical notation for variables in the labels rather than their names. Variables should be defined not in the figure legend but in the text describing the figure in the paper or in the caption.

Units and variables used within figures should be in the same format and font as in the main text, i.e., variables should be written in Italic font and units in Roman font. Figures included in any appendices should continue the sequential numbering from the main text, e.g., if there are 11 figures in the main text then the first figure in the first appendix should be labelled as Fig. 12 and not Fig. A1. File types that can be used are: .pdf, .jpeg, .jpg, .png, .eps, .epsf, .epsi, .pgf, .tikz, .ps.

\begin{figure}
\centering
  \includegraphics{Fig1.png}
\caption{Write an appropriate figure caption here. Discussions of the implications of the results shown in the figure should be left for the main text. Captions do not end with periods}
\label{fig1}
\end{figure}

\section{Tables}
Tables should be clearly presented, easy to read, and should be numbered sequentially. Variables presented in the tables should be formatted in the same manner as in the text (e.g., Italic variables and Bold vectors). Each table should include a preceding caption. Tables should be cited as ``Table 1", e.g., ``Table 1 shows ..." or ``the sensible heat flux values are presented in Table 1". If tables are included in the appendices, then these tables should be numbered sequentially continuing from the last table in the main text. Limit the number of tables to five.

\begin{table}
\caption{Write an appropriate caption for the table here. Captions do not end with periods}
\label{tab:1}
\begin{tabular}{lll}
\hline\noalign{\smallskip}
Variable & Number & Date  \\
\noalign{\smallskip}\hline\noalign{\smallskip}
\textit{x} & 1 & 15/05/2019 \\
\textit{y} & 2 & 16/05/2019 \\
\textit{z} & 3 & 17/05/2019 \\
\noalign{\smallskip}\hline
\end{tabular}
\end{table}

\section{Reference Formatting}
References should be presented in alphabetical order (not in the order of their appearance in the text). They should not be numbered. The total number of pages is not required for book references. List all authors (editors) of referred publications. References can be entered manually using the bibitem structure, or through BibTex (recommended). BibTex users should use the spbasic bibliography style and the natbib package. Sample references of several types are shown below: a journal article \citep{Mason+1987}, a book \citep{Garratt1994}, a book chapter \citep{Wyngaard2004}, a dissertation \citep{Fedorovich1986,Salesky2014}, a technical report \citep{Newsom++2015}, and a paper in conference proceedings \citep{Kaimal1979,Batchvarova+2003, Marusic++2011}.

\begin{acknowledgements}
These should follow the concluding section of the paper and precede the References and any appendices, if they are present. The acknowledgements section does not require a section number.
\end{acknowledgements}

\section*{Appendix 1: Title of Appendix}
\{Appendices are optional, as are appendix titles\}\\
Appendices should precede the references and should be numbered (if there is more than one appendix). Equations, tables, and figures contained within the appendices should be numbered sequentially following those in the main text.

\bibliographystyle{spbasic_updated}
\bibliography{sample_library}

\newpage

\section*{Supplementary Material for \textit{Boundary-Layer Meteorology} Sample Paper: Instructions for Authors}

{\textbf{First Author* $\cdot$ Second Author $\cdot$ Third Author \\}}
\\
\text{*}Affiliation and email address for the corresponding author only (note that the corresponding author does not need to be the first author).

\section*{1 Supplementary Electronic Materials}
Supplementary multimedia files and other supplementary materials are also accepted for online publication in \textit{Boundary-Layer Meteorology }alongside an article. The supplementary files should be provided in standard file formats.

To accommodate user downloads, please keep in mind that larger-sized files may require very long download times and that some users may experience other problems during downloading.

\subsection*{1.1 Audio, Video, and Animations}
Video and animation files should be provided at an aspect ratio of 16:9 or 4:3. The maximum file size that can be accommodated is 25 GB. The minimum allowable video length is 1 s. The supported file formats include avi, wmv, mp4, mov, m2p, mp2, mpg, mpeg, flv, mxf, mts, m4v, and 3gp. Video files should not contain more than three flashes per second.

\subsection*{1.2 Presentations, Text Files, and Spreadsheets}
Supplemental text files and presentations should be submitted as pdf files. Files in doc or ppt format cannot be accepted. Spreadsheets should also be converted to pdf format if they are intended for viewing only. If readers are encouraged to download and use the spreadsheet, then it can be provided in xls format.

\subsection*{1.3 Specialized Formats}
Other specialized file formats can also be supplied (e.g., tex, pdb, wrl, nb). It is also possible to provide multiple files within a zip or gz file.

\section*{2 General Information}
All supplementary materials should be specifically cited within the main text of the manuscript, in a manner similar to citing tables and figures. The supplementary materials should be cited as ``Online Resource", e.g., ``... as shown in the animation (Online Resource 3)", ``... additional data are provided in Online Resource 4". If more than one supplementary file is provided, these files should be numbered sequentially following the order they are cited in the main text, e.g., ``ESM\_1.mpg", ``ESM\_2.avi". Each supplementary file also requires a concise caption that describes the contents of the file. These captions should be listed at the end of the manuscript at initial submission.

Authors should note that supplementary materials will be published without any conversion, editing, or reformatting.
\clearpage

\section*{Journal Abbreviations used in \textit{Boundary-Layer Meteorology}}

\begin{longtable}{| p{8 cm} | p{6 cm} |}
\hline
Journal Name & Abbreviation used in BLM \\
\hline
ACM Transactions of Mathematical Software & ACM Trans Math Soft \\
Acoustics Australia & Acoust Aust \\
Acta Geophysica & Acta Geophys \\
Acta Mechanica Synica & Acta Mech Sinica \\
Acta Mechanica Supplement & Acta Mech Suppl \\
Advances in Atmospheric Science & Adv Atmos Sci \\
Advances in Ecological Research & Adv Ecol Res \\
Advances in Meteorology & Adv Meteorol \\
Advances in Science and Research & Adv Sci Res \\
Advances in Water Resources & Adv Water Resour \\
Aeolian Research & Aeolian Res \\
Aerospace Science and Technology & Aerosp Sci Technol \\
Agricultural Meteorology & Agric Meteorol \\
Agricultural and Forest Meteorology & Agric For Meteorol \\
Agricultural Water Management & Agric Water Manag \\
American Institute of Aeronautics and Astronautics & Am Inst Aeronaut Astronaut \\
Annals of Glaciology & Ann Glaciol \\
Annalen der Meteorologie & Ann Meteorol \\
Annals of Statistics & Ann Stat \\
Antarctic Science & Antarct Sci \\
Annual Review of Fluid Mechanics & Annu Rev Fluid Mech \\
Applied Energy & Appl Energy \\
Applied Mechanics Review & Appl Mech Rev \\
Applied Numerical Mathematics & Appl Numer Math \\
Applied Physics B & Appl Phys B \\
Applied Optics & Appl Opt \\
Aquatic Botany & Aquat Bot \\
Archiv f\"{u}r Meteorologie Geophysik und Bioklimatologie Serie A-Meteorologie und Geophysik & Arch Meteorol Geophys Bioklim Ser A \\
Archiv fur Hydrobiologie & Arch Hydrobiol \\
Artificial Intelligence & Artif Intell \\
Astronomy \& Astrophysics & Astron Astrophys \\
Atmospheric Measurement Techniques & Atmos Meas Tech \\
Atmosphere-Ocean & Atmos-Ocean \\
Atmospheric Research & Atmos Res \\
Atmospheric Science Letters & Atmos Sci Lett \\
Australian Journal of Physics & Aust J Phys \\
Australian Journal of Botany & Aust J Bot \\
\hline
Beitraege zur Physik der Atmosphaere & Beitr Phys Atmos \\
Biogeosciences & Biogeosciences \\
Biometrika & Biometrika \\
Biosystems Engineering & Biosyst Eng \\

\hline
\end{longtable}
\newpage

\begin{longtable}{| p{8 cm} | p{6 cm} |}
\hline
Journal Name & Abbreviation used in BLM \\
\hline
Boreal Environment Research & Boreal Environ Res \\
Boundary-Layer Meteorology & Boundary-Layer Meteorol \\
Building and Environment & Build Environ \\
Bulletin of the American Meteorological Society & Bull Am Meteorol Soc \\
\hline
Climate Research & Clim Res \\
Cold Regions Science and Technology & Cold Reg Sci Technol \\
Communications in Agricultural and Applied Biological Sciences & Commun Agric Appl Biol Sci \\
Communications in Mathematical Physics & Commun Math Phys \\
Communications on Pure and Applied Mathematics & Commun Pure Appl Math \\
Comptes Rendus Physique & C R Phys \\
Computers and Electronics in Agriculture & Comput Electron Agric \\
Computing and Informatics & Comput Inf \\
Computer Methods in Applied Mechanical Engineering & Comput Methods Appl Mech Eng \\
Computational Statistics and Data Analysis & Comput Stat Data Anal \\
Contributions to Atmospheric Physics & Contr Atmos Phys \\
Crop Protection & Crop Prot \\
\hline
Deep Sea Research Part II & Deep Sea Res II \\
Dynamics of Atmpsheres and Oceans & Dyn Atmos Oceans \\
\hline
Earth System Science Data Discussions & Earth Syst Sci Data Discuss \\
Earth Surface Processes and Landforms & Earth Surf Process Landf \\
Ecological Applications & Ecol Appl \\
Ecological Indicators & Ecol Indic \\
Ecological Modelling & Ecol Model \\
Ecology & Ecology \\
Electronic Journal of Operational Meteorology & Electron J Oper Meteorol \\
Energies & Energies \\
Energy and Buildings & Energy Buil \\
Energy Conversion and Management & Energy Convers Manag \\
Environmental Fluid Mechanics & Environ Fluid Mech \\
Environmental Modelling and Software & Environ Modell Softw \\
Environmental Pollution & Environ Pollut \\
Environmental Research Letters & Environ Res Lett \\
Environmental Science and Technology & Environ Sci Technol \\
Environmental Software & Environ Softw \\

Eos, Transactions, American Geophysical Union & Eos Trans AGU \\
European Journal of Forest Research & Eur J For Res \\
Experiments in Fluids & Exp Fluids \\
\hline
Fisheries Research & Fish Res \\
Flow Turbulence and Combustion & Flow Turbul Combust \\
Forestry & Forestry \\
Freshwater Biology & Freshwater Biol \\
\hline
\end{longtable}

\pagebreak[4]

\begin{longtable}{| p{8 cm} | p{6 cm} |}
\hline
Journal Name & Abbreviation used in BLM \\
\hline
Functional Ecology & Funct Ecol \\
\hline
Acta Geodaetica et Geophysica Hungarica & Geod Geophys \\
Geografiska Annaler Series A & Geogr Ann Ser A \\
Geography Compass & Geogr Compass \\
Geomorphology & Geomorphology \\
Geophysical Research Letters & Geophys Res Lett \\
Geoscientific Instrumentation, Methods and Data Systems & Geosci Instrum Method Data Syst \\
Geoscientific Model Development & Geosci Model Dev \\
Global Biogeochemical Sciences & Glob Biogeochem Cycles \\
Global Change Biology & Glob Change Biol \\
\hline
Hydrology and Earth System Sciences & Hydrol Earth Syst Sci \\
Hydrological Processes & Hydrol Proc \\
\hline
IEEE Journal of Ocean Engineering & IEEE J Ocean Eng \\
IEEE Transactions on Geoscience and Remote Sensing & IEEE Trans Geosci Remote \\
International Journal of Climatology & Int J Climatol \\
International Journal of Wildland Fire & Int J Wildland Fire \\
International Journal of Heat and Fluid Flow & Int J Heat Fluid Flow \\
International Journal of Numerical Methods for Fluids & Int J Numer Methods Fluids \\
International Journal of Remote Sensing & Int J Remote Sens \\
Izvestiya, Atmospheric and Oceanic Physics & Izv Atmos Ocean Phys \\
\hline
Journal of Advances in Modeling Earth Systems & J Adv Model Earth Syst \\
Journal of Aerosol Science & J Aerosol Sci \\
Journal of Agricultural Engineering Research & J Agric Eng Res \\
Journal of the Air Pollution Control Association & J Air Pollut Control Assoc \\
Journal of Aircraft & J Aircr \\
Journal of Applied Meteorology and Climatology & J Appl Meteorol Clim \\
Journal of Applied Meteorology & J Appl Meteorol \\
Journal of Aquatic Plant Management & J Aquat Plant Manag \\
Journal of Arid Environments & J Arid Environ \\
Journal of Atmospheric and Oceanic Technology & J Atmos Ocean Technol \\
Journal of Atmospheric Science & J Atmos Sci \\
Journal of Climate & J Clim \\
Journal of Computational Physics & J Comput Phys \\
Journal of Earth Simulation & J Earth Simul \\
Journal of Earth System Science & J Earth Syst Sci \\
Journal of Environmental Engineering & J Environ Eng \\
Journal of Experimental Botany & J Exp Bot \\
Journal of the Faculty of Science Hokkaido University & J Fac Sci Hokkaido Univ \\
\hline
\end{longtable}

\pagebreak[4]

\begin{longtable}{| p{8 cm} | p{6 cm} |}
\hline
Journal Name & Abbreviation used in BLM \\
\hline

Journal of Field Robotics & J Field Robot \\
Journal of Fluid Mechanics & J Fluid Mech \\
Journal of Geophysical Research & J Geophys Res \\
Journal of Geophysical Research-Atmospheres & J Geophys Res Atmos \\
Journal of Glaciology & J Glaciol \\
Journal of Great Lakes Research & J Great Lakes Res \\
Journal of Hazardous Materials & J Hazard Mater A \\
Journal of Heat Transfer & J Heat Transf \\
Journal of Hydraulic Engineering & J Hydraul Eng \\
Journal of Hydrology & J Hydrol \\
Journal of Hydrometeorology & J Hydrometeorol \\
Journal of Marine Research & J Mar Res \\
Journal of Marine Systems & J Mar Syst \\
Journal de Mathematiques Pures et Appliquees & J Math Pures Appl \\
Journal of Meteorology & J Meteorol \\
Journal of the Meteorological Society of Japan & J Meteorol Soc Jpn \\
Journal of Oceanography & J Oceanogr \\
Journal of Operational Oceanography & J Oper Oceanogr \\
Journal of the operational Research Society & J Oper Res Soc \\
Journal of the Optical Society of America & J Opt Soc Am \\
Journal of Plankton Research & J Plankton Res \\
Journal of Solar Energy Engineering & J Sol Energy Eng \\
Journal of Quantitative Spectroscopy and Radiative Transfer & J Quant Spectrosc Radiat Transf \\
Journal of Renewable and Sustainable Energy & J Renew Sust Energy \\
Journal of Scientific Statistical Computing & J Sci Stat Comput \\
Journal of Statistical Physics & J Stat Phys \\
Journal of Thermophysics and Heat Transfer & J Thermophys Heat Transf \\
Journal of Tropical Ecology & J Trop Ecol \\
Journal of Turbulence & J Turbul \\
Journal of Wind Engineering and Industrial Aerodynamics & J Wind Eng Ind Aerodyn \\
\hline
Landscape and Urban Planning & Landsc Urban Plan \\
Limnology and Oceanography & Limnol Oceanogr \\
Low Temperature Science & Low Temp Sci \\
\hline
Machine Learning & Mach Learn \\
Marine Chemistry & Mar Chem \\
Mathematische Annalen & Math Ann \\
Meteorological Applications & Meteorol Appl \\
Meteorology and Atmospheric Physics & Meteorol Atmos Phys \\
Meteorologische Zeitschrift & Meteorol Z \\
Monthly Weather Review & Mon Weather Rev \\
\hline
Natural Hazards and Earth System Sciences & Nat Hazards Earth Syst Sci \\
\hline
\end{longtable}

\pagebreak[4]

\begin{longtable}{| p{8 cm} | p{6 cm} |}
\hline
Journal Name & Abbreviation used in BLM \\
\hline
Nature Climate Change & Nat Clim Change \\
Nature Letters & Nat Lett \\
Nature Geoscience & Nat Geosci \\
Neural Computation & Neural Comput \\
Nonlinear Processes in Geophysics & Nonlin Process Geophys \\
New Zealand Journal of Science & N Z J Sci \\
\hline
Oceanography & Oceanography \\
Ocean Dynamics & Ocean Dyn \\
Ocean Engineering Science & Ocean Eng Sci \\
Ocean Modeling & Ocean Model \\
\hline
Papers in Physical Oceanography and Meteorology & Pap Phys Oceanogr Meteorol \\
Particle \& Particle Systems Characterization & Part Part Syst Charact \\
Particuology & Particuology \\
Philosophical Transactions of the Royal Society of London & Philos Trans R Soc \\
Photogrammetric Engineering and Remote Sensing & Photogramm Eng Remote Sens \\
Physical Review Letters & Phys Rev Lett \\
Physical Review E & Phys Rev E \\
Physics and Chemistry of the Earth & Phys Chem Earth \\
Physics of Fluids & Phys Fluids \\
Physics A - Statistical Mechanics and its Applications & Physica A Stat Mech Appl \\
Physica D & Physica D \\
Plant Biosystems & Plant Biosyst \\
PLOS One & PLOS One \\
Powder technology & Powder Technol \\
Proceedings of the Royal Society & Proc Roy Soc \\
Progress in Aerospace Science & Prog Aerosp Sci \\
Progress in Heat and Mass Transfer & Prog Heat Mass Transf \\
Progress in Physical Geography & Prog Phys Geogr \\
Pure and Applied Geophysics & Pure Appl Geophys \\
\hline
Quarterly Journal of the Royal Meteorological Society & Q J R Meteorol Soc \\
\hline
Remote Sensing & Remote Sens \\
Remote Sensing of Environment & Remote Sens Environ \\
Renewable Energy & Renew Energy \\
Reviews of Geophysics & Rev Geophys \\
Reviews of Geophysics and Space Physics & Rev Geophys Space Phys \\
Review of Scientific Instruments & Rev Sci Inst \\
\hline
Science & Science \\
Science of the Total Environment & Sci Tot Environ \\
Sedimentology & Sedimentol \\
Siam Journal on Applied Mathematics & SIAM J Appl Math \\
\hline
\end{longtable}

\pagebreak[4]

\begin{longtable}{| p{8 cm} | p{6 cm} |}
\hline
Journal Name & Abbreviation used in BLM \\
\hline
Tellus & Tellus \\
Tellus Series B - Chemical and Physical Meteorology & Tellus Ser B Chem Phys Meteorol \\
Theoretical and Applied Climatology & Theor Appl Climatol \\
Theoretical and Computational Fluid Dynamics & Theor Comput Fluid Dyn \\
Theoretical Computational Fluid Dynamics & Theor Comput Fluid Mech \\
Thermal Science Engineering & Therm Sci Eng \\
Transactions of the American Society of Agricultural Engineers & Trans ASAE \\
Tree Physiology & Tree Physiol \\
Trudy Geofizicheskogo Instituta, Akademiya Nauk SSSR & Trudy Geofiz Inst AN SSSR \\
\hline
Urban Climate & Urban Clim \\
\hline
Water Air and Soil Pollution & Water Air Soil Pollut \\
Water Resources Research & Water Resour Res \\
Waterway Port Coastal and Ocean Engineering & Waterw Port Coast Ocean Eng \\
Weather & Weather \\
Weather and Forecasting & Weather Forecast \\
Wind Energy & Wind Energy \\
Wind Engineering & Wind Eng \\
\hline
Zeitschrift f\"{u}r Angewandte Mathematik und Mechanik & Z Agnew Math Mech \\
\hline

\hline
\end{longtable}

\end{document}

### 2021-11-18 Analysis/data_analysis/DEM_analysis_plot.py

@@ -34,6 +34,7 @@ def DEM_analysis(file):
 
 
 Stations = ['Deep_Sea_Station', 'South_Namib_Station']
+Data_DEM = {}
 #
 # Paths
 path_savefig = '../../static/output_data/figures/'
@@ -43,6 +44,11 @@ path_inputdata = '../../static/input_data/'
 for i, station in enumerate(Stations):
     file = os.path.join(path_inputdata, station, 'DEM_' + station + '.npy')
     orientation, wavelength, amplitude, p0, p1, transect, C, topo, lon, lat, km_step = DEM_analysis(file)
+    Data_DEM[station] = {'orientation': orientation, 'wavelength': wavelength,
+                         'amplitude': amplitude, 'p0': p0, 'p1': p1,
+                         'transect': transect, 'C': C, 'topo': topo, 'lat': lat,
+                         'lon': lon, 'km_step': km_step}
+
     fig_width = theme.fig_width
     fig_height = 0.6*fig_width
     fig = plt.figure(figsize=(fig_width, fig_height), tight_layout=True)
@@ -82,3 +88,6 @@ for i, station in enumerate(Stations):
         plt.show()
     else:
         plt.close()
+
+
+np.save(os.path.join(path_outputdata, 'Data_DEM.npy'), Data_DEM)

### 2021-11-18 Analysis/data_analysis/meteo_data_analysis_plot.py

@@ -109,13 +109,12 @@ np.save(os.path.join(path_outputdata, 'Data_final.npy'), Data)
 # Plotting a few vertical profiles for the Deep Sea station
 # ------------------------
 
-def plot_fitprofile(t):
-    grad, theta, blh = gradient_free_atm[t], theta_free_atm[t], BLH[t]
-    mask_H = (height_sort[:, t] >= blh) & (height_sort[:, t] <= Hmax_fit)
+def plot_fitprofile(height, Virtual_potential_temperature, grad_free_atm, theta_free_atm, blh, theta_ground, Hmax_fit):
+    mask_H = (height >= blh) & (height <= Hmax_fit)
     #
-    plt.plot(Virtual_potential_temperature_sort[:, t], height_sort[:, t]/1e3, '+', label='data', zorder=-1)
-    plt.vlines(theta_ground[t], 0, blh/1e3, linewidth=2, label='boundary layer', zorder=0, color='tab:orange')
-    plt.plot(np.poly1d([grad, theta])(height_sort[:, t][mask_H]), height_sort[:, t][mask_H]/1e3, linewidth=2, label='free atm.', color='tab:green')
+    plt.plot(Virtual_potential_temperature, height/1e3, '+', label='data', zorder=-1)
+    plt.vlines(theta_ground, 0, blh/1e3, linewidth=2, label='boundary layer', zorder=0, color='tab:orange')
+    plt.plot(np.poly1d([grad_free_atm, theta_free_atm])(height[mask_H]), height[mask_H]/1e3, linewidth=2, label='free atm.', color='tab:green')
     plt.xlabel('Virtual potential temp. [K]')
     plt.ylabel('Height [km]')
     plt.gca().set_ylim(0, top=0.7*Hmax_fit/1e3)
@@ -124,12 +123,12 @@ def plot_fitprofile(t):
 
 
 times = [2012, 30254, 2024, 30266]
-plt.figure(figsize=(theme.fig_width, 0.9*theme.fig_width))
+plt.figure(figsize=(theme.fig_width, 0.9*theme.fig_width), constrained_layout=True)
 for i, t in enumerate(times):
     ax = plt.subplot(2, 2, i+1)
-    plot_fitprofile(t)
+    plot_fitprofile(height_sort[:, t], Virtual_potential_temperature_sort[:, t], gradient_free_atm[t],
+                    theta_free_atm[t], BLH[t], theta_ground[t], Hmax_fit)
[…]

### 2021-11-18 Analysis/data_preprocessing/data_preprocessing_plot.py

@@ -4,7 +4,7 @@ Preprocessing of the data
 =========================
 
 For each station, we follow these preprocessing steps:
-    - correction for constant shifts in some data (time, orientation).
+    - put the wind orientation angles in the trigonometric referential.
     - averaging of the in situ data in 1-hr bins centered on the time stamps of the Era5Land dataset.
     - filtering unusued data (NaNs, 0 velocity)
 

### 2021-11-18 Analysis/data_presentation/regime_diagram_plot.py

@@ -1,7 +1,7 @@
 """
-=======================================================
-Plotting the different diagram regime for both stations
-=======================================================
+========================
+Plotting regime diagrams
+========================
 
 """
 
@@ -10,13 +10,12 @@ import matplotlib.pyplot as plt
 import matplotlib.gridspec as gridspec
 import matplotlib.ticker as mticker
 import matplotlib.colors as mpcolors
-from scipy.stats import binned_statistic_2d
 import os
 import sys
 sys.path.append('../../')
 import python_codes.theme as theme
 from python_codes.general import smallestSignedAngleBetween, find_mode_distribution
-from python_codes.plot_functions import log_tick_formatter, rgba_to_rgb
+from python_codes.plot_functions import log_tick_formatter, plot_regime_diagram
 theme.load_style()
 
 
@@ -43,13 +42,12 @@ mode_delta = np.array([find_mode_distribution(Delta, i) for i in np.arange(150, 
 delta_angle = np.abs(Delta)
 delta_u = (U_era - U_station)/U_era
 #
-ax_labels = {'Froude': r'Froude number, $ U/\sqrt{(\Delta\rho/\rho) g H}$', 'kH': '$k H$', 'kLB': r'$k U/N$'}
+ax_labels = {'Froude': r'$Fr_{\textup{surface}} =  U/\sqrt{(\Delta\rho/\rho) g H}$', 'kH': '$k H$', 'kLB': r'$Fr_{\textup{internal}} = k U/N$'}
 lims = {'Froude': (5.8e-3, 450), 'kLB': (0.009, 7.5), 'kH': (2.2e-2, 10.8)}
-mask = ~np.isnan(numbers['Froude'])
-plot_idx = np.random.permutation(np.arange(delta_angle[mask].size))  # to plot the points of the scatter plot in random order
+plot_idx = np.random.permutation(np.arange(delta_angle.size))  # to plot the points of the scatter plot in random order
 couples = [('Froude', 'kH'), ('kLB', 'kH')]
-cmaps = ['plasma', theme.diverging_cmap]
-norms = [mpcolors.Normalize(vmin=0, vmax=70),
+cmaps = [theme.cmap_delta_theta, theme.cmap_delta_u]
[…]

### 2021-11-18 Analysis/data_presentation/wind_regime_comparison_plot.py

@@ -237,7 +237,10 @@ bin_hr = np.arange(-0.5, 24.5, 1)
 #
 label_x = [r'Hour of the day', r'$u_{*, \textup{Era5Land}}$']
 label_y = [r'$\delta_{\theta}$ [deg.]', r'$\delta_{u}$']
-
+#
+vlines_x = [[14, 19], [0.23]]
+vlines_y = [(55, 90), (0.5, 1)]
+#
 fig, axs = plt.subplots(2, 2, figsize=(theme.fig_width, 0.8*theme.fig_width), constrained_layout=True)
 #
 for i, (bin_quantity, quantity) in enumerate(zip([bin_delta_angle, bin_delta_u], [delta_angle, delta_u])):
@@ -248,6 +251,10 @@ for i, (bin_quantity, quantity) in enumerate(zip([bin_delta_angle, bin_delta_u],
         X, Y = np.meshgrid(x_center, y_center)
         vmax = 550 if i == 0 else 450
         a = axs[i, j].pcolormesh(x_edge, y_edge, counts.T, snap=True, vmax=vmax)
+        a.set_edgecolor('face')
+        #
+        for x_line in vlines_x[j]:
+            axs[i, j].vlines(x_line, vlines_y[i][0], vlines_y[i][1], linestyle='--', color='tab:orange')
         if i > 0:
             axs[i, j].set_xlabel(label_x[j])
         else:
@@ -260,4 +267,50 @@ for i, (bin_quantity, quantity) in enumerate(zip([bin_delta_angle, bin_delta_u],
             cbar = fig.colorbar(a, ax=axs[i, j], label=r'$N_{\textup{points}}$', location='right')
 
 plt.savefig(os.path.join(path_savefig, 'deltaU__hr_velocity_diagrams.pdf'))
+plt.show()
+
+# %%
+# Relative difference between the wind vectors
+# --------------------------------------------
+Stations = ['Adamax_Station', 'Huab_Station']
+
+Orientation_era = np.concatenate([Data[station]['Orientation_era'] for station in Stations])
+Orientation_station = np.concatenate([Data[station]['Orientation_station'] for station in Stations])
+U_era = np.concatenate([Data[station]['U_star_era'] for station in Stations])
+U_station = np.concatenate([Data[station]['U_star_station'] for station in Stations])
+time = np.concatenate([Data[station]['time'] for station in Stations])
+hrs = np.array([i.hour for i in time])
+#
[…]

### 2021-11-18 Analysis/linear_theory/_parameter_space_exploration.py

@@ -1,0 +1,119 @@
+r"""
+=============================================================
+Parameter space exploration for the hydrodyanmic coefficients
+=============================================================
+
+Note that the parameter space exploration takes a certain amount of times to run, and as such is separated from the plotting of te results.
+
+"""
+
+import numpy as np
+import os
+import sys
+sys.path.append('../../')
+from python_codes.linear_theory import calculate_solution
+
+# Paths
+path_outputdata = '../../static/output_data/data/'
+
+# Importing non-dimensional numbers calculated
+Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
+Stations = ['South_Namib_Station', 'Deep_Sea_Station']
+numbers = {key: np.concatenate([Data[station][key] for station in Stations]) for key in ('Froude', 'kH', 'kLB')}
+
+# ##### Loading pattern characteristics
+Data_pattern = {}
+for station in ['Deep_Sea_Station', 'South_Namib_Station']:
+    Data_pattern[station] = np.load(os.path.join(path_outputdata, 'Analysis_DEM_' + station + '.npy'), allow_pickle=True).item()
+
+# Parameters
+Kappa = 0.4  # Von Kàrmàn constant
+k = 2*np.pi/np.mean([Data_pattern[station]['wavelength']*1e3 for station in Stations])  # mean wavelength [m]
+z0 = 1e-3  # hydrodynamic roughness, [m]
+eta_0 = k*1e-3
+eta = 0  # non dimensional position where to calculate the solution (bottom)
+
+# Parameter space
+Npoints = 100
+Dic = {}
+Dic['kH_vals'] = np.logspace(-2, 1.1, Npoints)
+Dic['Froude_vals'] = np.logspace(-2.3, 2.5, Npoints)
+Dic['kLB_vals'] = np.logspace(-2.1, 1, Npoints)
+
+# # %%
+# # Plane (Froude - kH)
+# # ------------------------
+# print(('Froude', 'kH'))
+# Dic[('Froude', 'kH')] = {}
[…]

### 2021-11-18 Analysis/linear_theory/hydro_coeff_time_series.py

@@ -12,14 +12,12 @@ import sys
 sys.path.append('../../')
 from python_codes.linear_theory import calculate_solution
 
-
 # Paths
 path_outputdata = '../../static/output_data/data/'
 
 # Importing non-dimensional numbers calculated
 Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
 Stations = ['South_Namib_Station', 'Deep_Sea_Station']
-numbers = {key: np.concatenate([Data[station][key] for station in Stations]) for key in ('Froude', 'kH', 'kLB')}
 
 # ##### Loading pattern characteristics
 Data_pattern = {}
@@ -33,16 +31,19 @@ z0 = 1e-3  # hydrodynamic roughness, [m]
 eta_0_vals = k*1e-3
 eta = 0  # non dimensional position where to calculate the solution (bottom)
 
-hydro_Coeffs = np.zeros((2, k.size))
-
-for i, (eta_0, eta_H, Froude, eta_B) in enumerate(zip(eta_0_vals, numbers['kH'], numbers['Froude'], numbers['kLB'])):
-    max_z = 0.9999*eta_H
-    if not np.isnan([eta_H, Froude, eta_B]).any():
-        Sol = calculate_solution(eta, eta_H, eta_0, eta_B, Froude, max_z, Kappa=0.4)
-        #
-        Ax, Bx = np.real(Sol[2]), np.imag(Sol[2])
-    else:
-        Ax, Bx = np.nan, np.nan
-    hydro_Coeffs[:, i] = [Ax, Bx]
+hydro_Coeffs = {}
+for station in Stations:
+    hydro_Coeffs[station] = np.zeros((2, Data[station]['kH'].size))
+    k = 2*np.pi/(Data_pattern[station]['wavelength']*1e3)
+    #
+    for i, (eta_0, eta_H, Froude, eta_B) in enumerate(zip(eta_0_vals, Data[station]['kH'], Data[station]['Froude'], Data[station]['kLB'])):
+        max_z = 0.9999*eta_H
+        if not np.isnan([eta_H, Froude, eta_B]).any():
+            Sol = calculate_solution(eta, eta_H, eta_0, eta_B, Froude, max_z, Kappa=0.4)
+            #
+            Ax, Bx = np.real(Sol[2]), np.imag(Sol[2])
+        else:
+            Ax, Bx = np.nan, np.nan
[…]

### 2021-11-18 Analysis/linear_theory/regime_diagram_hydro_coeff_time_series_plot.py

@@ -1,0 +1,324 @@
+"""
+=====================================================
+Plot theoretical regime diagrams from the time series
+=====================================================
+
+"""
+
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+import matplotlib.ticker as mticker
+import matplotlib.colors as mpcolors
+import os
+import sys
+sys.path.append('../../')
+from python_codes.linear_theory import Cisaillement_basal
+from python_codes.plot_functions import log_tick_formatter, plot_regime_diagram
+import python_codes.theme as theme
+theme.load_style()
+
+
+# Paths
+path_savefig = '../../static/output_data/figures/'
+path_outputdata = '../../static/output_data/data/'
+
+# ##### Loading meteo data
+Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
+
+# ##### Loading time series of hydrodyamic coefficients
+Data_coeff = np.load(os.path.join(path_outputdata, 'time_series_hydro_coeffs.npy'), allow_pickle=True).item()
+Data_coeff = np.concatenate([Data_coeff[station] for station in Data_coeff.keys()], axis=1)
+
+# ##### Loading pattern characteristics
+Data_pattern = {}
+for station in ['Deep_Sea_Station', 'South_Namib_Station']:
+    Data_pattern[station] = np.load(os.path.join(path_outputdata, 'Analysis_DEM_' + station + '.npy'), allow_pickle=True).item()
+
+
+Stations = ['South_Namib_Station', 'Deep_Sea_Station']
+# Stations = ['South_Namib_Station']
+
+# %%
+# Plotting the regime diagrams for the modulus -- scatter plots
+# -------------------------------------------------------------
+numbers = {key: np.concatenate([Data[station][key] for station in Stations]) for key in ('Froude', 'kH', 'kLB')}
+modulus = np.linalg.norm(Data_coeff, axis=0)
+# phase = np.arctan2(Data_coeff[1], Data_coeff[0])
+#
[…]

### 2021-11-18 Analysis/linear_theory/regime_diagrams_theoretical_plot.py

@@ -1,7 +1,7 @@
 r"""
-======================================================================
-Plot the parameter space exploration of the hydrodyanamic coefficients
-======================================================================
+=================================================================
+Plot theoretical regime diagrams from parameter space exploration
+=================================================================
 
 """
 
@@ -19,14 +19,21 @@ theme.load_style()
 
 # Paths
 path_outputdata = '../../static/output_data/data/'
+path_savefig = '../../static/output_data/figures/'
 
 # Loading parameter space exploration
 Dic = np.load(os.path.join(path_outputdata, 'parameter_exploration_hydro_coeff_3D.npy'), allow_pickle=True).item()
+#
+axes_av = [3, 2]  # axes over which selecting/averaging is perform for plane representation
+index = [64, 45]  # when selection, selecting these indexes in the relevant `axes_av`
+#
+hydro_coeffs_3D = Dic[('Froude', 'kH', 'kLB')]['hydro_coeffs'].swapaxes(1, 2)
+# axes are now (A/B, kH, Froude, kLB)
 
 # %%
 # Plane (Froude - kH)
 # ------------------------
-hydro_coeffs = Dic[('Froude', 'kH', 'kLB')]['hydro_coeffs'].mean(axis=-1).swapaxes(1, 2)
+hydro_coeffs = hydro_coeffs_3D.take(index[0], axis=axes_av[0])
 
 # #### figure
 quantities = [hydro_coeffs[0, :, :], hydro_coeffs[1, :, :],
@@ -62,7 +69,7 @@ plt.show()
 # %%
 # Plane (kLB - kH)
 # ------------------------
-hydro_coeffs = Dic[('Froude', 'kH', 'kLB')]['hydro_coeffs'].mean(axis=1)
+hydro_coeffs = hydro_coeffs_3D.take(index[1], axis=axes_av[1])
 
 # #### figure
 quantities = [hydro_coeffs[0, :, :], hydro_coeffs[1, :, :],
@@ -98,19 +105,16 @@ plt.show()
 # %%
 # Computing theoretical regime diagrams
 # -------------------------------------
[…]

### 2021-11-18 Analysis/linear_theory/test_streamlines.py

@@ -1,0 +1,66 @@
+import numpy as np
+import os
+import sys
+import matplotlib.pyplot as plt
+sys.path.append('../../')
+from python_codes.linear_theory_old import calculate_solution, _mu_prime
+from python_codes.meteo_analysis import mu
+
+
+def focus_point(eta_0, Uho_Ustar, Kappa=0.4):
+    return eta_0*np.exp(Kappa*Uho_Ustar)
+
+
+eta_H = 0.5
+eta_0 = (2*np.pi*1e-3)/2.5e3
+eta_B = 0.1
+Fr = 10
+
+max_z = 0.99*eta_H
+k_xi = 0.1
+k_x = np.linspace(0, 2*2*np.pi, 300)
+
+eta = np.logspace(np.log10(eta_0), np.log10(max_z), 100)
+
+# list = calculate_solution(eta, eta_H, eta_lim, eta_0, eta_B, Fr, max_z, Kappa=0.4, output='full')
+list = calculate_solution(eta, eta_H, eta_0, eta_B, Fr, max_z, Kappa=0.4, output='full')
+
+Ux = np.real(mu(eta, eta_0)[None, :] + k_xi*np.exp(1j*k_x[:, None]) * list[0][0, :][None, :])
+Uz = np.real(k_xi*np.exp(1j*k_x[:, None]) * list[0][1, :][None, :])
+
+U, W, St, Sn = list[0]
+Ax, Bx = np.real(St), np.imag(St)
+Cx, Cy = np.real(Sn), np.imag(Sn)
+
+Results = list[1]
+coeffs = list[-1]
+
+
+# plt.semilogy(np.real(U) + _mu_prime(eta, eta_0), eta, '.')
+# plt.semilogy(np.imag(U), eta, '.')
+plt.semilogy(np.real(W), eta, '.')
+plt.semilogy(np.imag(W), eta, '.')
+# plt.semilogy(Ax, eta, '.')
+# plt.semilogy(Bx, eta, '.')
+plt.show()
+
+# #### Free atmosphere
+
+
+def q(eta_B):
+    np.piecwise(eta_B, [eta_B >= 1, eta_B < 1],
+                [lambda eta_B: -np.sqrt(1 - 1/eta_B**2),
+                 lambda eta_B: 1j*np.sqrt(1/eta_B**2 - 1)])
+
+
[…]

### 2021-11-18 Paper_figure/Figure3.py

@@ -1,0 +1,76 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import sys
+import os
+sys.path.append('../')
+import python_codes.theme as theme
+from python_codes.plot_functions import make_nice_histogram
+
+# Loading figure theme
+theme.load_style()
+
+# path
+path_imgs = '../static/images/'
+path_savefig = '../Paper/Figures'
+path_outputdata = '../static/output_data/data/'
+
+# Loading wind data
+Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
+Stations = sorted(Data.keys())
+
+# Figure properties
+# station = 'South_Namib_Station'
+# #
+# theta_bins = [[0, 130], [150, 250]]
+# velocity_bins = [[0.05, 0.2], [0.3, 10]]
+# Data_pattern = np.load(os.path.join(path_ouputdata, 'Analysis_DEM_' + station + '.npy'), allow_pickle=True).item()
+#
+station = 'Deep_Sea_Station'
+#
+theta_bins = [[0, 90], [150, 230]]
+velocity_bins = [[0.05, 0.25], [0.3, 10]]
+Data_pattern = np.load(os.path.join(path_outputdata, 'Analysis_DEM_' + station + '.npy'), allow_pickle=True).item()
+
+
+color_ax = 'purple'
+
+# ################ Figure
+fig, axarr = plt.subplots(3, 3, figsize=(theme.fig_width, 0.9*theme.fig_width), constrained_layout=True, sharex=True)
+for i in range(3):  # Loop over velocites
+    if i < 2:
+        mask_U = (Data[station]['U_star_era'] >= velocity_bins[i][0]) & (Data[station]['U_star_era'] <= velocity_bins[i][1])
+        label_u = r'$u_{*, \, \textup{ERA}} < ' + str(velocity_bins[i][1]) + '$' if i == 0 else r'$u_{*, \, \textup{ERA}} > ' + str(velocity_bins[i][0]) + '$'
+    else:
+        mask_U = (Data[station]['U_star_era'] < 10)  # take all velocities
+        label_u = 'all velocities'
+    axarr[i, -1].set_ylabel(label_u)
+    axarr[i, -1].yaxis.set_label_position("right")
+    for j in range(3):  # loop over angles
+        if j < 2:
[…]

### 2021-11-18 Paper_figure/Figure4.py

@@ -1,0 +1,196 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import sys
+import os
+sys.path.append('../')
+import python_codes.theme as theme
+from python_codes.general import cosd, sind
+from python_codes.linear_theory import Cisaillement_basal_rotated_wind, coeffA0, coeffB0
+
+
+def perturb(x, amp, lamb, shift):
+    return np.cos(2*np.pi*x/lamb)*amp + shift
+
+
+def plot_vertical_profile(ax, height, Virtual_potential_temperature, grad_free_atm, theta_free_atm, blh, theta_ground, Hmax_fit, color='tab:blue', label=None):
+    Hfit = np.linspace(blh, Hmax_fit, 100)
+    #
+    line = ax.vlines(theta_ground, 0, blh/1e3, color=color, label=label, zorder=-3)
+    ax.axhline(blh/1e3, color=color, ls='--')
+    ax.plot(np.poly1d([grad_free_atm, theta_free_atm])(Hfit), Hfit/1e3, color=line.get_color(), zorder=-2)
+    ax.plot(Virtual_potential_temperature, height/1e3, '.', color=line.get_color(), zorder=-1)
+    # ax.scatter(theta_ground, blh/1e3, s=30, facecolors=line.get_color(), edgecolors='k', linewidth=2, zorder=0)
+
+
+def plot_streamlines(ax, z_pos, x, amp, lamb, **kwargs):
+    for z in z_pos:
+        ax.plot(x, perturb(x, amp, lamb, z), **kwargs)
+
+
+def topo(x, y, alpha, k, xi):
+    return xi*np.cos(k*(cosd(alpha)*x + sind(alpha)*y))
+
+
+# Loading figure theme
+theme.load_style()
+
+# path
+path_imgs = '../static/images/'
+path_savefig = '../Paper/Figures'
+path_outputdata = '../static/output_data/data/'
+
+# ##### Loading meteo data
+Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
+Stations = ['South_Namib_Station', 'Deep_Sea_Station']
+
+# #### figure parameters
+# ## sketch parameters
+flow_color = 'powderblue'
[…]

### 2021-11-18 Paper_figure/Supplementary_Figures/Figure10_supp.py

@@ -1,0 +1,64 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import sys
+import os
+sys.path.append('../../')
+import python_codes.theme as theme
+from python_codes.meteo_analysis import mu
+from python_codes.plot_functions import make_nice_histogram
+
+
+# Loading figure theme
+theme.load_style()
+
+# path
+path_imgs = '../../static/images/'
+path_savefig = '../../Paper/Figures'
+path_outputdata = '../../static/output_data/data/'
+
+# ##### Loading meteo data
+Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
+Stations = ['South_Namib_Station', 'Deep_Sea_Station']
+
+# ## histograms parameters
+Stations = ['South_Namib_Station', 'Deep_Sea_Station']
+g = 9.81  # [m/s2] gravitational constant
+z0_era = 1e-3  # [m] hydrodynamic roughness
+#
+labels = [r'\textbf{a}', r'\textbf{b}', r'\textbf{c}', r'\textbf{d}']
+nbins = 80
+
+# #### Figure
+fig, axarr = plt.subplots(2, 2, figsize=(theme.fig_width, 0.9*theme.fig_width),
+                          constrained_layout=True, sharey=True)
+
+for station in Stations:
+    make_nice_histogram(Data[station]['Boundary layer height'], nbins, axarr[0, 0], alpha=0.4, label=' '.join(station.split('_')[:-1]), density=False, scale_bins='log')
+    #
+    N = np.sqrt(g*Data[station]['gradient_free_atm']/Data[station]['theta_ground'])
+    make_nice_histogram(N, nbins, axarr[0, 1], alpha=0.4, label=' '.join(station.split('_')[:-1]), density=False)
+    #
+    U_H = Data[station]['U_star_era']*mu(Data[station]['Boundary layer height'], z0_era)
+    make_nice_histogram(U_H, nbins, axarr[1, 0], alpha=0.4, label=' '.join(station.split('_')[:-1]), density=False)
+    #
+    make_nice_histogram(Data[station]['delta_theta']/Data[station]['theta_ground'], nbins, axarr[1, 1], alpha=0.4, label=' '.join(station.split('_')[:-1]), density=False)
+
+
[…]

### 2021-11-18 Paper_figure/Supplementary_Figures/Figure1_supp.py

@@ -1,0 +1,65 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.patches as mpatches
+import sys
+import os
+sys.path.append('../../')
+import python_codes.theme as theme
+from datetime import timedelta
+#
+theme.load_style()
+
+
+def make_range_broken_barh(time, dt):
+    diff = np.diff(time)
+    t_diff = np.concatenate((time[1:][diff > dt_threshold], time[:-1][diff > dt_threshold]))
+    t_diff = np.insert(t_diff, [0, t_diff.size], [time[0], time[-1]])
+    t_diff = np.array(sorted(t_diff))
+    return [(tstart, tspan) for tstart, tspan in zip(t_diff[::2], np.diff(t_diff)[::2])]
+
+
+# Paths
+path_savefig = '../../Paper/Figures'
+path_inputdata = '../../static/input_data/'
+
+
+Stations = ['Adamax_Station', 'Huab_Station', 'Deep_Sea_Station', 'South_Namib_Station']
+dat_types = ['Era5Land_wind_data_', 'in_situ_wind_data_']
+colors = ['tab:blue', 'tab:orange', 'tab:red']
+
+dt_threshold = timedelta(minutes=60)
+height_rect = 0.75
+height_delta = 1
+height_plot = 0
+centers = []
+
+fig_width = theme.fig_width
+fig_height = 0.45*fig_width
+fig = plt.figure(figsize=(fig_width, fig_height), constrained_layout=True)
+for station in Stations:
+    for i, data_type in enumerate(dat_types):
+        name = os.path.join(path_inputdata, station, data_type + station + '.npy')
+        if os.path.exists(name):
+            data = np.load(name, allow_pickle=True).item()
+            time = data['time']
+            if data_type == 'in_situ_wind_data_':
+                orientation, velocities = data['direction'], data['velocity']
+                mask = (~(np.isnan(velocities) | np.isnan(orientation))) & (velocities > 0)
+            else:
+                mask = np.ones(time.size).astype(bool)
+            xranges = make_range_broken_barh(time[mask], dt_threshold)
+            plt.broken_barh(xranges, (height_plot, height_rect), facecolor=colors[i])
+            height_plot += height_rect
+    centers.append(height_plot - height_rect)
+    height_plot += height_delta
+
+plt.xlabel('time [years]')
[…]

### 2021-11-18 Paper_figure/Supplementary_Figures/Figure3_supp.py

@@ -1,0 +1,53 @@
+import numpy as np
+import os
+import matplotlib.pyplot as plt
+from matplotlib.colors import Normalize
+import sys
+sys.path.append('../../')
+import python_codes.theme as theme
+
+theme.load_style()
+
+# paths
+path_savefig = '../../Paper/Figures'
+path_outputdata = '../../static/output_data/data/'
+path_inputdata = '../../static/input_data'
+
+Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
+Data_roughness = np.load(os.path.join(path_outputdata, 'Data_calib_roughness.npy'),
+                         allow_pickle=True).item()
+
+labels = [r'\textbf{a}', r'\textbf{b}', r'\textbf{c}', r'\textbf{d}']
+norm = Normalize(vmin=0.3, vmax=1, clip=True)
+
+# #### Figure
+fig, axrr = plt.subplots(2, 2, figsize=(theme.fig_width, 1.1*theme.fig_width),
+                         sharex=True, sharey=True, constrained_layout=True)
+#
+for i, (ax, p, metric, station, label) in enumerate(zip(axrr.flatten(), Data_roughness['Pvals'],
+                                                        Data_roughness['Metrics'], Data_roughness['Stations'], labels)):
+    cs = ax.contourf(Data_roughness['z0_era_vals'], Data_roughness['z0_station_vals'], metric, levels=50, norm=norm)
+    for c in cs.collections:
+        c.set_edgecolor("face")
+    ax.plot(Data_roughness['z0_era_vals'], Data_roughness['z0_station_vals'], color='tab:red', label='$y=x$')
+    ax.plot(Data_roughness['z0_era_vals'], np.exp(p[1])*Data_roughness['z0_era_vals']**p[0], 'r--', label='minimum line')
+    a, = ax.plot([Data_roughness['z0_era_vals'].min(), 1e-3], [Data[station]['z0_station'], Data[station]['z0_station']], color='tab:orange')
+    ax.plot([1e-3, 1e-3], [Data_roughness['z0_station_vals'].min(), Data[station]['z0_station']], color=a.get_color())
+    ax.plot(1e-3, Data[station]['z0_station'], '.', color=a.get_color())
+    #
[…]

### 2021-11-18 Paper_figure/Supplementary_Figures/Figure4_supp.py

@@ -1,0 +1,72 @@
+import numpy as np
+import os
+import matplotlib.pyplot as plt
+import sys
+from types import SimpleNamespace
+sys.path.append('../../')
+import python_codes.theme as theme
+from python_codes.general import cosd, sind
+
+theme.load_style()
+
+# paths
+path_savefig = '../../Paper/Figures'
+path_outputdata = '../../static/output_data/data/'
+path_inputdata = '../../static/input_data'
+
+Data_DEM = np.load(os.path.join(path_outputdata, 'Data_DEM.npy'),
+                   allow_pickle=True).item()
+
+labels = [(r'\textbf{a}', r'\textbf{d}'), (r'\textbf{b}', r'\textbf{e}'), (r'\textbf{c}', r'\textbf{f}')]
+
+fig, axrr = plt.subplots(3, 2, figsize=(theme.fig_width, 0.75*theme.fig_height_max),
+                         constrained_layout=True, gridspec_kw={'width_ratios': (0.9, 1)})
+for i, station in enumerate(Data_DEM.keys()):
+    # loading into namespace from data dictionnary to shorten call
+    n = SimpleNamespace(**Data_DEM[station])
+    # ax0: Topo
+    cs = axrr[0, i].contourf(n.lon, n.lat, n.topo, levels=50)
+    for c in cs.collections:
+        c.set_edgecolor("face")
+        c.set_rasterized(True)
+    cb = fig.colorbar(cs, ax=axrr[0, i], label='$h$~[m]', location='top')
+    cb.ax.locator_params(nbins=8)
+    axrr[0, i].set_xlabel(r'longitude [$^{\circ}$]')
+    axrr[0, i].set_ylabel(r'latitude [$^{\circ}$]')
+    axrr[0, i].set_aspect('equal')
+    #
+    # ax1: Autocorrelation map
+    x = list(-(n.lon - n.lon[0])[:: -1]) + list((n.lon - n.lon[0])[1:])
+    y = list(-(n.lat - n.lat[0])[:: -1]) + list((n.lat - n.lat[0])[1:])
+    cs = axrr[1, i].contourf(x, y, n.C, levels=50)
+    for c in cs.collections:
+        c.set_edgecolor("face")
+        c.set_rasterized(True)
+    #
[…]

### 2021-11-18 Paper_figure/Supplementary_Figures/Figure6_supp.py

@@ -1,0 +1,69 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import sys
+import os
+sys.path.append('../../')
+import python_codes.theme as theme
+from python_codes.plot_functions import make_nice_histogram
+
+# Loading figure theme
+theme.load_style()
+
+# path
+path_imgs = '../../static/images/'
+path_savefig = '../../Paper/Figures'
+path_outputdata = '../../static/output_data/data/'
+
+# Loading wind data
+Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
+Stations = sorted(Data.keys())
+
+# Figure properties
+station = 'South_Namib_Station'
+#
+theta_bins = [[0, 140], [150, 260]]
+velocity_bins = [[0.05, 0.2], [0.2, 10]]
+Data_pattern = np.load(os.path.join(path_outputdata, 'Analysis_DEM_' + station + '.npy'), allow_pickle=True).item()
+
+
+color_ax = 'purple'
+
+# ################ Figure
+fig, axarr = plt.subplots(3, 3, figsize=(theme.fig_width, 0.9*theme.fig_width), constrained_layout=True, sharex=True)
+for i in range(3):  # Loop over velocites
+    if i < 2:
+        mask_U = (Data[station]['U_star_era'] >= velocity_bins[i][0]) & (Data[station]['U_star_era'] <= velocity_bins[i][1])
+        label_u = r'$u_{*} < ' + str(velocity_bins[i][1]) + '$' if i == 0 else r'$u_{*} > ' + str(velocity_bins[i][0]) + '$'
+    else:
+        mask_U = (Data[station]['U_star_era'] < 10)  # take all velocities
+        label_u = 'all velocities'
+    axarr[i, -1].set_ylabel(label_u)
+    axarr[i, -1].yaxis.set_label_position("right")
+    for j in range(3):  # loop over angles
+        if j < 2:
+            mask_theta = (Data[station]['Orientation_era'] >= theta_bins[j][0]) & (Data[station]['Orientation_era'] <= theta_bins[j][1])
+            label_theta = r'$' + str(theta_bins[j][0]) + r'< \theta < ' + str(theta_bins[j][-1]) + '$'
+        else:
+            mask_theta = Data[station]['Orientation_era'] < 400  # take all orientations
+            label_theta = 'all angles'
[…]

### 2021-11-18 Paper_figure/Supplementary_Figures/Figure7_supp.py

@@ -1,0 +1,74 @@
+import numpy as np
+import os
+import matplotlib.pyplot as plt
+import sys
+sys.path.append('../../')
+import python_codes.theme as theme
+from python_codes.plot_functions import plot_scatter_surrounded
+
+
+theme.load_style()
+
+# paths
+path_savefig = '../../Paper/Figures'
+path_outputdata = '../../static/output_data/data/'
+path_inputdata = '../../static/input_data'
+
+# Loading data
+Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
+
+labels = [r'\textbf{a}', r'\textbf{b}']
+
+# preparing data
+Stations = ['South_Namib_Station', 'Deep_Sea_Station']
+
+Dune_orientations = [np.load(os.path.join(path_outputdata, 'Analysis_DEM_' + station + '.npy'), allow_pickle=True).item()['orientation']
+                     for station in Stations]
+
+velocity_thresholds = [0.1, 0.25]
+
+# variables
+x1 = np.concatenate([Data[station]['Orientation_era'][Data[station]['U_star_era'] < velocity_thresholds[0]] - Dune_orientations[Stations.index(station)]
+                     for station in Stations])
+y1 = np.concatenate([Data[station]['Orientation_station'][Data[station]['U_star_era'] < velocity_thresholds[0]] - Dune_orientations[Stations.index(station)]
+                     for station in Stations])
+
+x2 = np.concatenate([Data[station]['Orientation_era'][(Data[station]['U_star_era'] >= velocity_thresholds[0]) & (Data[station]['U_star_era'] < velocity_thresholds[1])] - Dune_orientations[Stations.index(station)]
+                     for station in Stations])
+y2 = np.concatenate([Data[station]['Orientation_station'][(Data[station]['U_star_era'] >= velocity_thresholds[0]) & (Data[station]['U_star_era'] < velocity_thresholds[1])] - Dune_orientations[Stations.index(station)]
+                     for station in Stations])
+
+x3 = np.concatenate([Data[station]['Orientation_era'][(Data[station]['U_star_era'] >= velocity_thresholds[1])] - Dune_orientations[Stations.index(station)]
+                     for station in Stations])
[…]

### 2021-11-18 Paper_figure/Supplementary_Figures/Figure8_supp.py

@@ -1,0 +1,61 @@
+import numpy as np
+import os
+import matplotlib.pyplot as plt
+import sys
+sys.path.append('../../')
+import python_codes.theme as theme
+from python_codes.plot_functions import plot_scatter_surrounded
+
+theme.load_style()
+
+# paths
+path_savefig = '../../Paper/Figures'
+path_outputdata = '../../static/output_data/data/'
+path_inputdata = '../../static/input_data'
+
+# Loading data
+Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
+
+labels = [r'\textbf{a}', r'\textbf{b}']
+
+# preparing data
+Stations = ['South_Namib_Station', 'Deep_Sea_Station']
+
+Nocturnal_wind = {'South_Namib_Station': [150, 260], 'Deep_Sea_Station': [150, 230]}
+
+# variables
+x1 = np.concatenate([Data[station]['U_star_era'][(Data[station]['Orientation_era'] > Nocturnal_wind[station][0]) & (Data[station]['Orientation_era'] < Nocturnal_wind[station][1])]
+                     for station in Stations])
+y1 = np.concatenate([Data[station]['U_star_station'][(Data[station]['Orientation_era'] > Nocturnal_wind[station][0]) & (Data[station]['Orientation_era'] < Nocturnal_wind[station][1])]
+                     for station in Stations])
+#
+x2 = np.concatenate([Data[station]['U_star_era'][~((Data[station]['Orientation_era'] > Nocturnal_wind[station][0]) & (Data[station]['Orientation_era'] < Nocturnal_wind[station][1]))]
+                     for station in Stations])
+y2 = np.concatenate([Data[station]['U_star_station'][~((Data[station]['Orientation_era'] > Nocturnal_wind[station][0]) & (Data[station]['Orientation_era'] < Nocturnal_wind[station][1]))]
+                     for station in Stations])
+
+X = [x1, x2]
+Y = [y1, y2]
+
+# #### Figure
+pad_angle = 2
+labels = [r'\textbf{a}', r'\textbf{b}']
+alphas = [0.075, 0.045]
+
+fig, axarr = plt.subplots(1, 2, figsize=(theme.fig_width, 0.53*theme.fig_width),
[…]

### 2021-11-18 Paper_figure/Supplementary_Figures/Figure9_supp.py

@@ -1,0 +1,97 @@
+import numpy as np
+import os
+import matplotlib.pyplot as plt
+import sys
+sys.path.append('../../')
+import python_codes.theme as theme
+from python_codes.plot_functions import make_nice_histogram
+
+
+def plot_vertical_profile(ax, height, Virtual_potential_temperature, grad_free_atm, theta_free_atm, blh, theta_ground, Hmax_fit, color='tab:blue', label=None):
+    Hfit = np.linspace(blh, Hmax_fit, 100)
+    #
+    line = ax.vlines(theta_ground, 0, blh/1e3, color=color, label=label, zorder=-3)
+    ax.axhline(blh/1e3, alpha=0.5, color=color, ls='--')
+    ax.plot(np.poly1d([grad_free_atm, theta_free_atm])(Hfit), Hfit/1e3, color=line.get_color(), zorder=-2)
+    ax.plot(Virtual_potential_temperature, height/1e3, '.', color=line.get_color(), zorder=-1)
+    # ax.scatter(theta_ground, blh/1e3, s=30, facecolors=line.get_color(), edgecolors='k', linewidth=2, zorder=0)
+
+
+theme.load_style()
+
+# paths
+path_savefig = '../../Paper/Figures'
+path_outputdata = '../../static/output_data/data/'
+path_inputdata = '../../static/input_data'
+
+# Loading data
+Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
+
+labels = [r'\textbf{a}', r'\textbf{b}', r'\textbf{c}', r'\textbf{d}']
+
+# ## vertical profiles parameters
+station = 'Deep_Sea_Station'
+time_steps_bad = [10856, 30266, 33463]
+time_steps_good = [2012, 30302, 30310]
+colors = ['tab:blue', 'tab:orange', 'tab:green']
+Hmax_fit = 10000  # [m]
+
+# ## Distribution parameters
+Stations = ['South_Namib_Station', 'Deep_Sea_Station']
+
+fig, axarr = plt.subplots(2, 2, figsize=(theme.fig_width, 1*theme.fig_width),
+                          constrained_layout=True, gridspec_kw={'height_ratios': [2, 1]})
+
+
+# ill-processed vertical profiles
+for i, t in enumerate(time_steps_bad):
[…]

### 2021-11-18 doc/auto_examples/data_presentation/regime_diagram_plot.py

@@ -1,7 +1,7 @@
 """
-=======================================================
-Plotting the different diagram regime for both stations
-=======================================================
+========================
+Plotting regime diagrams
+========================
 
 """
 
@@ -10,13 +10,12 @@ import matplotlib.pyplot as plt
 import matplotlib.gridspec as gridspec
 import matplotlib.ticker as mticker
 import matplotlib.colors as mpcolors
-from scipy.stats import binned_statistic_2d
 import os
 import sys
 sys.path.append('../../')
 import python_codes.theme as theme
 from python_codes.general import smallestSignedAngleBetween, find_mode_distribution
-from python_codes.plot_functions import log_tick_formatter, rgba_to_rgb
+from python_codes.plot_functions import log_tick_formatter, plot_regime_diagram
 theme.load_style()
 
 
@@ -43,14 +42,13 @@ mode_delta = np.array([find_mode_distribution(Delta, i) for i in np.arange(150, 
 delta_angle = np.abs(Delta)
 delta_u = (U_era - U_station)/U_era
 #
-ax_labels = {'Froude': r'Froude number, $ U/\sqrt{(\Delta\rho/\rho) g H}$', 'kH': '$k H$', 'kLB': r'$k L_{\textup{B}}$'}
+ax_labels = {'Froude': r'$Fr_{\textup{surface}} =  U/\sqrt{(\Delta\rho/\rho) g H}$', 'kH': '$k H$', 'kLB': r'$Fr_{\textup{internal}} = k U/N$'}
 lims = {'Froude': (5.8e-3, 450), 'kLB': (0.009, 7.5), 'kH': (2.2e-2, 10.8)}
-mask = ~np.isnan(numbers['Froude'])
-plot_idx = np.random.permutation(np.arange(delta_angle[mask].size))  # to plot the points of the scatter plot in random order
+plot_idx = np.random.permutation(np.arange(delta_angle.size))  # to plot the points of the scatter plot in random order
 couples = [('Froude', 'kH'), ('kLB', 'kH')]
-cmaps = ['plasma', 'seismic']
-norms = [mpcolors.Normalize(vmin=0, vmax=70),
-         mpcolors.TwoSlopeNorm(vmin=-1, vcenter=0, vmax=1)]
[…]

### 2021-11-18 doc/auto_examples/data_presentation/wind_regime_comparison_plot.py

@@ -228,16 +228,19 @@ hrs = np.array([i.hour for i in time])
 Delta = smallestSignedAngleBetween(Orientation_era, Orientation_station)
 mode_delta = np.array([find_mode_distribution(Delta, i) for i in np.arange(150, 350)]).mean()
 delta_angle = np.abs(Delta)
-delta_u = np.abs(U_era - U_station)/U_era
+delta_u = (U_era - U_station)/U_era
 #
-bin_delta_u = np.linspace(0, 1, 20)
+bin_delta_u = np.linspace(-1, 1, 20)
 bin_delta_angle = np.linspace(0, 90, 17)
 bin_U_era = np.linspace(0, 0.475, 30)
 bin_hr = np.arange(-0.5, 24.5, 1)
 #
 label_x = [r'Hour of the day', r'$u_{*, \textup{Era5Land}}$']
 label_y = [r'$\delta_{\theta}$ [deg.]', r'$\delta_{u}$']
-
+#
+vlines_x = [[14, 19], [0.23]]
+vlines_y = [(55, 90), (0.5, 1)]
+#
 fig, axs = plt.subplots(2, 2, figsize=(theme.fig_width, 0.8*theme.fig_width), constrained_layout=True)
 #
 for i, (bin_quantity, quantity) in enumerate(zip([bin_delta_angle, bin_delta_u], [delta_angle, delta_u])):
@@ -248,6 +251,10 @@ for i, (bin_quantity, quantity) in enumerate(zip([bin_delta_angle, bin_delta_u],
         X, Y = np.meshgrid(x_center, y_center)
         vmax = 550 if i == 0 else 450
         a = axs[i, j].pcolormesh(x_edge, y_edge, counts.T, snap=True, vmax=vmax)
+        a.set_edgecolor('face')
+        #
+        for x_line in vlines_x[j]:
+            axs[i, j].vlines(x_line, vlines_y[i][0], vlines_y[i][1], linestyle='--', color='tab:orange')
         if i > 0:
             axs[i, j].set_xlabel(label_x[j])
         else:

### 2021-11-18 doc/auto_examples/linear_theory/regime_diagrams_plot.py

@@ -1,0 +1,130 @@
+r"""
+======================================================================
+Plot the parameter space exploration of the hydrodyanamic coefficients
+======================================================================
+
+"""
+
+import numpy as np
+import matplotlib.pyplot as plt
+from matplotlib.colors import TwoSlopeNorm, SymLogNorm
+from scipy.ndimage import gaussian_filter
+import os
+import sys
+sys.path.append('../../')
+import python_codes.theme as theme
+theme.load_style()
+
+# Paths
+path_outputdata = '../../static/output_data/data/'
+
+# Loading parameter space exploration
+Dic = np.load(os.path.join(path_outputdata, 'parameter_exploration_hydro_coeff.npy'), allow_pickle=True).item()
+
+# %%
+# Plane (Froude - kH)
+# ------------------------
+hydro_coeffs = Dic[('Froude', 'kH')]['hydro_coeffs']
+
+# #### figure
+quantities = [hydro_coeffs[0, :, :], hydro_coeffs[1, :, :],
+              np.linalg.norm(hydro_coeffs, axis=0), np.arctan2(hydro_coeffs[1, :, :], hydro_coeffs[0, :, :])]
+labels = [r'$\mathcal{A}_{0}$', r'$\mathcal{B}_{0}$', r'$\sqrt{\mathcal{A}_{0}^{2} + \mathcal{B}_{0}^{2}}$',
+          r'$\textup{arctan}\left(\mathcal{B}_{0}/\mathcal{A}_{0}\right)$']
+#
+fig, axs = plt.subplots(2, 2, figsize=(theme.fig_width, 0.8*theme.fig_width), constrained_layout=True)
+for i, (ax, label, quantity) in enumerate(zip(axs.flatten(), labels, quantities)):
+    smoothed = gaussian_filter(quantity, 5)
+    if i in [0, 2]:
+        cmap = 'viridis'
+        norm = None
+    else:
+        cmap = 'seismic'
+        norm = TwoSlopeNorm(vcenter=0, vmin=smoothed.min(), vmax=smoothed.max())
+    cnt = ax.contourf(Dic['Froude_vals'], Dic['kH_vals'], smoothed, levels=100, norm=norm, cmap=cmap)
+    for c in cnt.collections:
[…]

### 2021-11-18 doc/auto_examples/linear_theory/regime_diagrams_theoretical_plot.py

@@ -1,7 +1,7 @@
 r"""
-======================================================================
-Plot the parameter space exploration of the hydrodyanamic coefficients
-======================================================================
+=================================================================
+Plot theoretical regime diagrams from parameter space exploration
+=================================================================
 
 """
 
@@ -19,14 +19,21 @@ theme.load_style()
 
 # Paths
 path_outputdata = '../../static/output_data/data/'
+path_savefig = '../../static/output_data/figures/'
 
 # Loading parameter space exploration
 Dic = np.load(os.path.join(path_outputdata, 'parameter_exploration_hydro_coeff_3D.npy'), allow_pickle=True).item()
+#
+axes_av = [3, 2]  # axes over which selecting/averaging is perform for plane representation
+index = [60, 48]  # when selection, selecting these indexes in the relevant `axes_av`
+#
+hydro_coeffs_3D = Dic[('Froude', 'kH', 'kLB')]['hydro_coeffs'].swapaxes(1, 2)
+# axes are now (A/B, kH, Froude, kLB)
 
 # %%
 # Plane (Froude - kH)
 # ------------------------
-hydro_coeffs = Dic[('Froude', 'kH', 'kLB')]['hydro_coeffs'].mean(axis=-1).swapaxes(1, 2)
+hydro_coeffs = hydro_coeffs_3D.take(index[0], axis=axes_av[0])
 
 # #### figure
 quantities = [hydro_coeffs[0, :, :], hydro_coeffs[1, :, :],
@@ -62,7 +69,7 @@ plt.show()
 # %%
 # Plane (kLB - kH)
 # ------------------------
-hydro_coeffs = Dic[('Froude', 'kH', 'kLB')]['hydro_coeffs'].mean(axis=1)
+hydro_coeffs = hydro_coeffs_3D.take(index[1], axis=axes_av[1])
 
 # #### figure
 quantities = [hydro_coeffs[0, :, :], hydro_coeffs[1, :, :],
@@ -98,19 +105,16 @@ plt.show()
 # %%
 # Computing theoretical regime diagrams
 # -------------------------------------
[…]

### 2021-11-18 python_codes/CourrechDuPont2014.py

@@ -1,0 +1,260 @@
+r"""
+Dune growth mechanism theory developped by Courrech du Pont et al. 2014.
+
+References
+----------
+[1] Courrech du Pont, S., Narteau, C., & Gao, X. (2014). Two modes for dune orientation. Geology, 42(9), 743-746.
+"""
+
+# Note that it is here derived is the fixed referential math:`(x,y)`
+# used by Gadal et al. 2019. The dune orientation math:`\alpha` is then calculated with respect to the `y`-axis.
+# [2] Gadal, C., Narteau, C., Du Pont, S. C., Rozier, O., & Claudin, P. (2019). Incipient bedforms in a bidirectional wind regime. Journal of Fluid Mechanics, 862, 490-516.
+
+import numpy as np
+from python_codes.general import Vector_average, cosd, sind
+
+
+############################################################################
+########################### Fluxes at the crest ############################
+############################################################################
+
+
+def Flux_at_crest(alpha, theta, Q0, gamma=1.6):
+    r"""Calculation of the sand flux at the dune crest:
+
+    .. math::
+
+        Q_{\rm crest} = Q_{0}\left[1 + \gamma\sin\vert\theta-\alpha\vert\right]
+
+    Parameters
+    ----------
+    alpha : scalar, numpy array
+        Dune orientation :math:`\alpha`.
+    theta : scalar, numpy array
+        Flux orientation :math:`\theta` in degrees.
+    Q0 : scalar, numpy array
+        Flux at the bottom of the dune :math:`Q_{0}`.
+    gamma : scalar, numpy array
+        Flux-up ratio :math:`\gamma` (the default is 1.6).
+
+    Returns
+    -------
+    scalar, numpy array
+        Flux at the dune crest :math:`Q_{\rm crest}`
+
+    Examples
+    --------
+    >>> import numpy as np
+    >>> alpha = 10
+    >>> theta = np.random.random((1000,))*360
+    >>> Q0 = np.random.random((1000,))*50
+    >>> Qcrest = Flux_at_crest(alpha, theta, Q0)
+
+    """
+    return theta, Q0*(1 + gamma*np.abs(sind(theta - alpha)))
+
+
+def Resultant_flux_at_crest(alpha, theta, Q0, gamma=1.6, **kwargs):
+    r"""Resultant flux (i.e vectorial average) of the sand flux at the crest.
+
+    Parameters
[…]

### 2021-11-18 python_codes/general.py

@@ -1,4 +1,5 @@
 import numpy as np
+from xhistogram.core import histogram
 
 
 def cosd(theta):
@@ -11,6 +12,11 @@ def sind(theta):
 
 def tand(theta):
     return np.tan(theta*np.pi/180)
+
+
+def Vector_average(Direction, Norm, axis=-1):
+    average = np.nanmean(Norm*np.exp(1j*Direction*np.pi/180), axis=axis)
+    return np.angle(average)*180/np.pi, np.absolute(average)
 
 
 def smallestSignedAngleBetween(x, y):
@@ -53,3 +59,9 @@ def find_mode_distribution(data, bin_number):
     counts, bins_edges = np.histogram(data, bins=bin_number)
     bin_centers = bins_edges[:-1] + (bins_edges[1] - bins_edges[0])
     return bin_centers[counts.argmax()]
+
+
+def Make_angular_PDF(angles, weight, bin_edges=np.linspace(0, 360, 361), axis=-1):
+    hist = histogram(angles, bins=bin_edges, density=1, weights=weight, axis=axis)
+    bin_centers = bin_edges[1:] - (bin_edges[1] - bin_edges[0])/2
+    return hist, bin_centers

### 2021-11-18 python_codes/linear_theory_asymptotic.py

@@ -1,0 +1,388 @@
+"""
+Linear theory of a turbulent flow over a sinusoidal bottom.
+"""
+
+import numpy as np
+from scipy.integrate import solve_ivp
+from python_codes.general import cosd, sind
+from python_codes.meteo_analysis import mu
+
+# %%
+# Geometrical model
+# -----------------
+# Geometrical model for the scaling of the coefficient as the function of the orientation of the bottom perturbation
+
+
+def Ax(alpha, A0):
+    r"""Calculate the hydrodynamic coefficient :math:`\mathcal{A}_{x}` using the geometrical model:
+
+    .. math::
+
+        \mathcal{A}_{x} = \mathcal{A}_{0}\cos^{2}\alpha.
+
+    Parameters
+    ----------
+    alpha : array, scalar
+        Dune orientation with respect to the perpendicular to the flow direction (in degree).
+    A0 : array, scalar
+        value of the hydrodynamic coefficient for :math:`\alpha = 0`, i.e. for a dune orientation perpendicular to the flow direction.
+
+    Returns
+    -------
+    array, scalar
+         the hydrodynamic coefficient.
+    """
+    return A0*cosd(alpha)**2
+
+
+def Ay(alpha, A0):
+    r"""Calculate the hydrodynamic coefficient :math:`\mathcal{A}_{y}` using the geometrical model:
+
+    .. math::
+
+        \mathcal{A}_{y} = 0.5\mathcal{A}_{0}\cos\alpha\sin\alpha.
+
+    Parameters
+    ----------
+    alpha : array, scalar
+        Dune orientation with respect to the perpendicular to the flow direction (in degree).
+    A0 : array, scalar
+        value of the hydrodynamic coefficient for :math:`\alpha = 0`, i.e. for a dune orientation perpendicular to the flow direction.
+
+    Returns
+    -------
+    array, scalar
+         the hydrodynamic coefficient.
+    """
+    return A0*cosd(alpha)*sind(alpha)/2
+
+
+def Bx(alpha, B0):
+    r"""Calculate the hydrodynamic coefficient :math:`\mathcal{B}_{x}` using the geometrical model:
+
+    .. math::
+
+        \mathcal{B}_{x} = \mathcal{B}_{0}\cos^{2}\alpha.
+
+    Parameters
[…]

### 2021-11-18 python_codes/linear_theory_focalpoint.py

@@ -1,0 +1,388 @@
+"""
+Linear theory of a turbulent flow over a sinusoidal bottom.
+"""
+
+import numpy as np
+from scipy.integrate import solve_ivp
+from python_codes.general import cosd, sind
+from python_codes.meteo_analysis import mu
+
+# %%
+# Geometrical model
+# -----------------
+# Geometrical model for the scaling of the coefficient as the function of the orientation of the bottom perturbation
+
+
+def Ax(alpha, A0):
+    r"""Calculate the hydrodynamic coefficient :math:`\mathcal{A}_{x}` using the geometrical model:
+
+    .. math::
+
+        \mathcal{A}_{x} = \mathcal{A}_{0}\cos^{2}\alpha.
+
+    Parameters
+    ----------
+    alpha : array, scalar
+        Dune orientation with respect to the perpendicular to the flow direction (in degree).
+    A0 : array, scalar
+        value of the hydrodynamic coefficient for :math:`\alpha = 0`, i.e. for a dune orientation perpendicular to the flow direction.
+
+    Returns
+    -------
+    array, scalar
+         the hydrodynamic coefficient.
+    """
+    return A0*cosd(alpha)**2
+
+
+def Ay(alpha, A0):
+    r"""Calculate the hydrodynamic coefficient :math:`\mathcal{A}_{y}` using the geometrical model:
+
+    .. math::
+
+        \mathcal{A}_{y} = 0.5\mathcal{A}_{0}\cos\alpha\sin\alpha.
+
+    Parameters
+    ----------
+    alpha : array, scalar
+        Dune orientation with respect to the perpendicular to the flow direction (in degree).
+    A0 : array, scalar
+        value of the hydrodynamic coefficient for :math:`\alpha = 0`, i.e. for a dune orientation perpendicular to the flow direction.
+
+    Returns
+    -------
+    array, scalar
+         the hydrodynamic coefficient.
+    """
+    return A0*cosd(alpha)*sind(alpha)/2
+
+
+def Bx(alpha, B0):
+    r"""Calculate the hydrodynamic coefficient :math:`\mathcal{B}_{x}` using the geometrical model:
+
+    .. math::
+
+        \mathcal{B}_{x} = \mathcal{B}_{0}\cos^{2}\alpha.
+
+    Parameters
+    ----------
[…]

### 2021-11-19 Paper/main.tex
Adding a new figure for other regime diagrams in SI

@@
     \includegraphics[scale=1]{Figures/Figure5.pdf}
-    \caption{Regime diagrams of the wind deviation $\delta_{\theta}$ and relative attenuation/amplification $\delta_{u}$, containing the data from both the Deep Sea and South Namib stations.
+    \caption{Regime diagrams of the wind deviation $\delta_{\theta}$ and relative attenuation/amplification $\delta_{u}$ in the space $(\mathcal{F}r_{\textup{S}}, \, kH)$, containing the data from both the Deep Sea and South Namib stations.
 Blue dashed lines empirically delimit the different regimes.
-The point density in each bin of the diagrams is shown in Fig.~\ref{Fig11_supp}.}
+The point density in each bin of the diagrams is shown in Fig.~\ref{Fig11_supp}.
+The regime diagrams in the spaces $(\mathcal{F}r_{\textup{I}}, \, kH)$ and $(\mathcal{F}r_{\textup{I}}, \, \mathcal{F}r_{\textup{S}})$ are shown in Fig.~\ref{Fig12_supp}.}
     \label{Fig5}
@@
   \includegraphics[scale=1]{Figures/Figure12_supp.pdf}
+  \caption{Regime diagrams of the wind deviation $\delta_{\theta}$ and relative attenuation/amplification $\delta_{u}$ in the spaces $(\mathcal{F}r_{\textup{I}}, \, kH)$ and $(\mathcal{F}r_{\textup{I}}, \, \mathcal{F}r_{\textup{S}})$, containing the data from both the Deep Sea and South Namib stations.
+Blue dashed lines empirically delimit the different regimes.
+The point density in each bin of the diagrams is shown in Fig.~\ref{Fig11_supp}.
+The regime diagrams in the space $(\mathcal{F}r_{\textup{S}}, \, kH)$ are shown in Fig.~\ref{Fig5_supp}.}
+  \label{Fig12_supp}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[scale=1]{Figures/Figure13_supp.pdf}
   \caption{Physical interpretation of the flow disturbance.
@@
 From the upper to the lower streamline, values of $\left(kH,\, Fr_{\textup{surface}},\, Fr_{\textup{internal}},\, \mathcal{A}_{0},\, \mathcal{B}_{0},\, \sqrt{\mathcal{A}_{0}^{2} + \mathcal{B}_{0}^{2}}\right)$ are (1.9, 0.6, 1.5, 3.4, 1.0, 3.5), (1.5, 0.3, 0.4, 4.8, 1.4, 5.0), (0.1, 3.5, 1.0, 8.6, 0.1, 8.6), (0.5, 0.05, 0.04, 9.6, 2.5, 9.9).}
-  \label{Fig12_supp}
+  \label{Fig13_supp}
 \end{figure}

### 2021-11-19 Paper_figure/Figure5.py

@@ -49,11 +49,11 @@ regime_line_color = 'tab:blue'
 
 
 var1, var2 = 'Froude', 'kH'
-xlabel = r'$Fr_{\textup{surface}} =  U/\sqrt{(\Delta\rho/\rho) g H}$'
+xlabel = r'$\mathcal{F}r_{\textup{S}} =  U/\sqrt{(\Delta\rho/\rho) g H}$'
 
 # #### Figure
 fig, axarr = plt.subplots(1, 2, figsize=(theme.fig_width, 0.375*theme.fig_height_max),
-                          constrained_layout=True, sharey=True)
+                          constrained_layout=True)
 
 for i, (ax, quantity, cmap, norm) in enumerate(zip(axarr.flatten(), quantities, cmaps, norms)):
     ylabel = '$k H$' if i == 0 else None
@@ -74,5 +74,8 @@ for i, (ax, quantity, cmap, norm) in enumerate(zip(axarr.flatten(), quantities, 
 
     # colorbar
     sm = plt.cm.ScalarMappable(cmap=cmap, norm=norm)
-    cb = fig.colorbar(sm, ax=ax, location='top')
+    cb = plt.colorbar(sm, ax=ax, location='top')
     cb.set_label(cbar_labels[i])
+
+plt.savefig(os.path.join(path_savefig, 'Figure5.pdf'))
+plt.show()

### 2021-11-19 Paper/main.tex
Re-organisation of doc

@@
 
-\author{First Author         \and
-        Second Author \and Third Author \and \{...\}
-}
+\author{Cyril Gadal \and Pauline Delorme \and Clément Narteau \and
+        Giles Wiggs \and Matthew Baddock \and Joanna M. Nield \and Philippe Claudin }
 
-\institute{F. Author \at
-              first address \\
-              Tel.: +123-45-678910\\
-              Fax: +123-45-678910\\
-              \email{fauthor@example.com}
+\institute{C. Gadal \at
+              Institut de Mécanique des Fluides de Toulouse (IMFT), Université de Toulouse, CNRS, INPT, UPS, Toulouse, France \\
+              \email{cyril.gadal@imft.fr}
            \and

### 2021-11-19 Analysis/data_preprocessing/data_preprocessing_plot.py

@@ -1,142 +1,0 @@
-"""
-=========================
-Preprocessing of the data
-=========================
-
-For each station, we follow these preprocessing steps:
-    - put the wind orientation angles in the trigonometric referential.
-    - averaging of the in situ data in 1-hr bins centered on the time stamps of the Era5Land dataset.
-    - filtering unusued data (NaNs, 0 velocity)
-
-
-"""
-
-
-import numpy as np
-import os
-from scipy.stats import binned_statistic
-from datetime import datetime, timedelta
-import matplotlib.dates as mdates
-import matplotlib.pyplot as plt
-import sys
-sys.path.append('../../')
-from python_codes.general import cosd, sind
-import python_codes.theme as theme
-#
-import warnings
-warnings.filterwarnings("ignore", category=RuntimeWarning)
-
-theme.load_style()
-
-# paths
-path_savefig = '../../static/output_data/figures/'
-path_ouputdata = '../../static/output_data/data/'
-path_inputdata = '../../static/input_data'
-
-Stations = ['Adamax_Station', 'Deep_Sea_Station', 'Huab_Station', 'South_Namib_Station']
-
-Data = {}
-
-for station in Stations:
-    Data[station] = {}
-    ############################################################################
-    # Loading data
-    ############################################################################
-    #
-    # ###### Era5Land wind data
-    path_Era5_land = os.path.join(path_inputdata, station, 'Era5Land_wind_data_' + station + '.npy')
-    Data_era = np.load(path_Era5_land, allow_pickle=True).item()
-    #
-    t_era = Data_era['time']
-    U_era = np.sqrt(Data_era['Uwind']**2 + Data_era['Vwind']**2).squeeze()
-    Orientation_era = (np.arctan2(Data_era['Vwind'], Data_era['Uwind'])*180/np.pi).squeeze() % 360
-    # ###### in situ wind data
-    path_insitu = os.path.join(path_inputdata, station, 'in_situ_wind_data_' + station + '.npy')
-    Data_insitu = np.load(path_insitu, allow_pickle=True).item()
-    #
-    t_station = Data_insitu['time']
-    U_station = Data_insitu['velocity']
-    # putting angles in trigo. ref.
-    Orientation_station = (270 - Data_insitu['direction']) % 360
-    #
-    ############################################################################
-    # Averaging in situ data over 1hr
[…]

### 2021-11-19 Analysis/data_preprocessing/roughness_calibration_plot.py

@@ -1,126 +1,0 @@
-r"""
-===========
-Calibration of the hydrodynamic roughness
-===========
-
-For each station, the hydrodynamic roughness is calibrated by finding the one that minimizes the difference between the wind vectors of both datasets.
-
-    - we compute the difference between wind vectors using hydrodynamic roughnesses ranging from :math:`10^{-5}` m to :math:`10^{-2}` m.
-    - we find the minimum in this space, which is a line.
-    - we impose an hydrodynamic roughness of :math:`10^{-3}` m for the Era5Land dataset, and compute the corresponding roughness for the in situ dataset.
-
-The chosen metric for comparison is then:
-
-.. math::
-
-    \delta = \frac{\sqrt{\langle\| \boldsymbol{u}_{*, \textrm{era}} - \boldsymbol{u}_{*, \textrm{station}} \|^{2}\rangle_{t}}}{\sqrt{ \langle \| \boldsymbol{u}_{*, \textrm{era}} \| \rangle_{t}\langle \| \boldsymbol{u}_{*, \textrm{station}} \| \rangle_{t}}}
-"""
-
-
-import numpy as np
-import matplotlib.pyplot as plt
-import os
-import sys
-sys.path.append('../../')
-import python_codes.theme as theme
-from python_codes.general import smallestSignedAngleBetween, cosd, sind, find_mode_distribution
-from python_codes.meteo_analysis import mu
-theme.load_style()
-#
-import warnings
-warnings.filterwarnings("ignore", category=RuntimeWarning)
-
-
-# paths
-path_savefig = '../../static/output_data/figures/'
-path_ouputdata = '../../static/output_data/data/'
-
-# Parameters
-z_era = 10      # height of ERA5Land wind data [m]
-z0_era = 1e-3  # hydrodynamic roughness chosen for the Era5Land dataset [m]
-g = 9.81  # gravitational acceleration [m2/s]
-z_station = {'Adamax_Station': 3.18, 'Deep_Sea_Station': 2.62, 'Huab_Station': 2, 'South_Namib_Station': 2.62}  # Height of wind measurements [m]
-angle_tolerance = 15  # tolerance in selecting the wind orientation matching between both datasets
-
-# parameter space exploration
-z0_station_vals = np.logspace(-5, -2, 50)
[…]

### 2021-11-19 Analysis/data_presentation/regime_diagram_plot.py

@@ -1,221 +1,0 @@
-"""
-========================
-Plotting regime diagrams
-========================
-
-"""
-
-import numpy as np
-import matplotlib.pyplot as plt
-import matplotlib.gridspec as gridspec
-import matplotlib.ticker as mticker
-import matplotlib.colors as mpcolors
-import os
-import sys
-sys.path.append('../../')
-import python_codes.theme as theme
-from python_codes.general import smallestSignedAngleBetween, find_mode_distribution
-from python_codes.plot_functions import log_tick_formatter, plot_regime_diagram
-theme.load_style()
-
-
-# Paths
-path_savefig = '../../static/output_data/figures/'
-path_outputdata = '../../static/output_data/data/'
-
-# ##### Loading meteo data
-Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
-Stations = ['South_Namib_Station', 'Deep_Sea_Station']
-
-# %%
-# Plotting the regime diagrams -- scatter plots
-# ---------------------------------------------
-
-Orientation_era = np.concatenate([Data[station]['Orientation_era'] for station in Stations])
-Orientation_station = np.concatenate([Data[station]['Orientation_station'] for station in Stations])
-U_era = np.concatenate([Data[station]['U_star_era'] for station in Stations])
-U_station = np.concatenate([Data[station]['U_star_station'] for station in Stations])
-numbers = {key: np.concatenate([Data[station][key] for station in Stations]) for key in ('Froude', 'kH', 'kLB')}
-#
-Delta = smallestSignedAngleBetween(Orientation_era, Orientation_station)
-mode_delta = np.array([find_mode_distribution(Delta, i) for i in np.arange(150, 350)]).mean()
-delta_angle = np.abs(Delta)
-delta_u = (U_era - U_station)/U_era
-#
-ax_labels = {'Froude': r'$Fr_{\textup{surface}} =  U/\sqrt{(\Delta\rho/\rho) g H}$', 'kH': '$k H$', 'kLB': r'$Fr_{\textup{internal}} = k U/N$'}
[…]

### 2021-11-19 Analysis/data_presentation/temporal_coverage_plot.py

@@ -1,74 +1,0 @@
-"""
-===========
-Temporal coverage of the data
-===========
-
-Plot the temporal coverage of the two different sets of data for each station using a Gantt chart.
-"""
-
-
-import numpy as np
-import matplotlib.pyplot as plt
-import matplotlib.patches as mpatches
-import sys
-import os
-sys.path.append('../../')
-import python_codes.theme as theme
-from datetime import timedelta
-#
-theme.load_style()
-
-
-def make_range_broken_barh(time, dt):
-    diff = np.diff(time)
-    t_diff = np.concatenate((time[1:][diff > dt_threshold], time[:-1][diff > dt_threshold]))
-    t_diff = np.insert(t_diff, [0, t_diff.size], [time[0], time[-1]])
-    t_diff = np.array(sorted(t_diff))
-    return [(tstart, tspan) for tstart, tspan in zip(t_diff[::2], np.diff(t_diff)[::2])]
-
-
-# Paths
-path_savefig = '../../static/output_data/figures/'
-path_inputdata = '../../static/input_data/'
-
-
-Stations = ['Adamax_Station', 'Huab_Station', 'Deep_Sea_Station', 'South_Namib_Station']
-dat_types = ['Era5Land_wind_data_', 'in_situ_wind_data_']
-colors = ['tab:blue', 'tab:orange', 'tab:red']
-
-dt_threshold = timedelta(minutes=60)
-height_rect = 0.75
-height_delta = 1
-height_plot = 0
-centers = []
-
-fig_width = theme.fig_width
-fig_height = 0.45*fig_width
-fig = plt.figure(figsize=(fig_width, fig_height), constrained_layout=True)
-for station in Stations:
-    for i, data_type in enumerate(dat_types):
-        name = os.path.join(path_inputdata, station, data_type + station + '.npy')
-        if os.path.exists(name):
-            data = np.load(name, allow_pickle=True).item()
-            time = data['time']
-            if data_type == 'in_situ_wind_data_':
-                orientation, velocities = data['direction'], data['velocity']
-                mask = (~(np.isnan(velocities) | np.isnan(orientation))) & (velocities > 0)
-            else:
-                mask = np.ones(time.size).astype(bool)
-            xranges = make_range_broken_barh(time[mask], dt_threshold)
-            plt.broken_barh(xranges, (height_plot, height_rect), facecolor=colors[i])
[…]

### 2021-11-19 Analysis/data_presentation/wind_regime_comparison_plot.py

@@ -1,316 +1,0 @@
-"""
-===========
-Comparison between the Era5Land and the in situ datasets
-===========
-"""
-
-import numpy as np
-import matplotlib.pyplot as plt
-import matplotlib.colors as Mplcolors
-import matplotlib.gridspec as gridspec
-import os
-import sys
-sys.path.append('../../')
-import python_codes.theme as theme
-from python_codes.plot_functions import plot_scatter_surrounded, plot_wind_rose
-from python_codes.meteo_analysis import compute_circadian_annual_cycle
-from python_codes.general import smallestSignedAngleBetween, find_mode_distribution
-from scipy.stats import binned_statistic_2d
-theme.load_style()
-
-# paths
-path_savefig = '../../static/output_data/figures/'
-path_ouputdata = '../../static/output_data/data/'
-
-# Loading wind data
-Data = np.load(os.path.join(path_ouputdata, 'Data_final.npy'), allow_pickle=True).item()
-Stations = sorted(Data.keys())
-
-# Loading pattern characteristics
-Data_pattern = {}
-for station in ['Deep_Sea_Station', 'South_Namib_Station']:
-    Data_pattern[station] = np.load(os.path.join(path_ouputdata, 'Analysis_DEM_' + station + '.npy'), allow_pickle=True).item()
-
-
-# %%
-# Wind roses
-# ------------------------
-
-h = 115
-bins = [0.03, 0.15, 0.2, 0.25, 0.3, 0.35, 0.4]
-
-fig_width = theme.fig_width
-fig_height = 0.75*fig_width
-
-fig = plt.figure(figsize=(fig_width, fig_height))
-gs = gridspec.GridSpec(3, 2, height_ratios=[0.25, 1, 1], width_ratios=[1, 1])
-gs.update(left=0.025, right=0.97, bottom=0.03, top=0.97, wspace=0.15, hspace=0.25)
-for i, station in enumerate(Stations):
-    gs_sub = gs[i + 2].subgridspec(2, 2, height_ratios=[0.01, 1])
-    #
-    ax = plt.subplot(gs_sub[0, :])
-    plt.axis('off')
-    ax.text(0.5, 0.5, Stations[i][:-8].replace('_', ' '), ha='center', va='center', transform=ax.transAxes)
[…]

### 2021-11-19 Analysis/linear_theory/_parameter_space_exploration.py

@@ -1,119 +1,0 @@
-r"""
-=============================================================
-Parameter space exploration for the hydrodyanmic coefficients
-=============================================================
-
-Note that the parameter space exploration takes a certain amount of times to run, and as such is separated from the plotting of te results.
-
-"""
-
-import numpy as np
-import os
-import sys
-sys.path.append('../../')
-from python_codes.linear_theory import calculate_solution
-
-# Paths
-path_outputdata = '../../static/output_data/data/'
-
-# Importing non-dimensional numbers calculated
-Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
-Stations = ['South_Namib_Station', 'Deep_Sea_Station']
-numbers = {key: np.concatenate([Data[station][key] for station in Stations]) for key in ('Froude', 'kH', 'kLB')}
-
-# ##### Loading pattern characteristics
-Data_pattern = {}
-for station in ['Deep_Sea_Station', 'South_Namib_Station']:
-    Data_pattern[station] = np.load(os.path.join(path_outputdata, 'Analysis_DEM_' + station + '.npy'), allow_pickle=True).item()
-
-# Parameters
-Kappa = 0.4  # Von Kàrmàn constant
-k = 2*np.pi/np.mean([Data_pattern[station]['wavelength']*1e3 for station in Stations])  # mean wavelength [m]
-z0 = 1e-3  # hydrodynamic roughness, [m]
-eta_0 = k*1e-3
-eta = 0  # non dimensional position where to calculate the solution (bottom)
-
-# Parameter space
-Npoints = 100
-Dic = {}
-Dic['kH_vals'] = np.logspace(-2, 1.1, Npoints)
-Dic['Froude_vals'] = np.logspace(-2.3, 2.5, Npoints)
-Dic['kLB_vals'] = np.logspace(-2.1, 1, Npoints)
-
-# # %%
-# # Plane (Froude - kH)
-# # ------------------------
-# print(('Froude', 'kH'))
-# Dic[('Froude', 'kH')] = {}
[…]

### 2021-11-19 Analysis/linear_theory/_regime_diagrams_plot.py

@@ -1,130 +1,0 @@
-r"""
-======================================================================
-Plot the parameter space exploration of the hydrodyanamic coefficients
-======================================================================
-
-"""
-
-import numpy as np
-import matplotlib.pyplot as plt
-from matplotlib.colors import TwoSlopeNorm
-from scipy.ndimage import gaussian_filter
-import os
-import sys
-sys.path.append('../../')
-import python_codes.theme as theme
-theme.load_style()
-
-# Paths
-path_outputdata = '../../static/output_data/data/'
-
-# Loading parameter space exploration
-Dic = np.load(os.path.join(path_outputdata, 'parameter_exploration_hydro_coeff.npy'), allow_pickle=True).item()
-
-# %%
-# Plane (Froude - kH)
-# ------------------------
-hydro_coeffs = Dic[('Froude', 'kH')]['hydro_coeffs']
-
-# #### figure
-quantities = [hydro_coeffs[0, :, :], hydro_coeffs[1, :, :],
-              np.linalg.norm(hydro_coeffs, axis=0), np.arctan2(hydro_coeffs[1, :, :], hydro_coeffs[0, :, :])]
-labels = [r'$\mathcal{A}_{0}$', r'$\mathcal{B}_{0}$', r'$\sqrt{\mathcal{A}_{0}^{2} + \mathcal{B}_{0}^{2}}$',
-          r'$\textup{arctan}\left(\mathcal{B}_{0}/\mathcal{A}_{0}\right)$']
-#
-fig, axs = plt.subplots(2, 2, figsize=(theme.fig_width, 0.8*theme.fig_width), constrained_layout=True)
-for i, (ax, label, quantity) in enumerate(zip(axs.flatten(), labels, quantities)):
-    smoothed = gaussian_filter(quantity, 5)
-    if i in [0, 2]:
-        cmap = 'viridis'
-        norm = None
-    else:
-        cmap = 'seismic'
-        norm = TwoSlopeNorm(vcenter=0, vmin=smoothed.min(), vmax=smoothed.max())
-    cnt = ax.contourf(Dic['Froude_vals'], Dic['kH_vals'], smoothed, levels=100, norm=norm, cmap=cmap)
-    for c in cnt.collections:
[…]

### 2021-11-19 doc/auto_examples/data_analysis/meteo_data_analysis_plot.py

@@ -1,185 +1,0 @@
-"""
-==================================
-Processing the meteorological data
-==================================
-
-
-"""
-
-import numpy as np
-import matplotlib.pyplot as plt
-import os
-import sys
-sys.path.append('../../')
-import python_codes.theme as theme
-from python_codes.meteo_analysis import mu
-theme.load_style()
-
-# Paths
-path_savefig = '../../static/output_data/figures/'
-path_outputdata = '../../static/output_data/data/'
-
-# ##### Loading meteo data
-Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
-Stations = ['South_Namib_Station', 'Deep_Sea_Station']
-
-# ##### Loading pattern characteristics
-Data_pattern = {}
-for station in ['Deep_Sea_Station', 'South_Namib_Station']:
-    Data_pattern[station] = np.load(os.path.join(path_outputdata, 'Analysis_DEM_' + station + '.npy'), allow_pickle=True).item()
-
-
-# %%
-# Calculating the relevant meteorological quantities
-# ------------------------
-#
-# We calculate the relevant meteorological quantities from the Era5 pressure levels data.
-
-# Parameters
-g = 9.81  # gravitational acceleration [m2/s]
-Rt = 6356766  # Average Earth radius [m]
-Kelvin = 273.15  # Kelvin shift
-P0 = 1000        # Standard pressure [hPa]
-Md = 0.029       # Molecular mass of dry air [kg/mol]
-Mw = 0.018       # Molecular mass of water [kg/mol]
-R_M = Md/Mw
-R = 8.314  # Gaz constant
-Pc = 0.2854      # Poisson coefficient for dry air R/Cp
-z0_era = 1e-3  # hydrodynamic roughness chosen for the Era5Land dataset [m]
-
-# #### Calculating relevant meteorological quantities
-for station in Stations:
-    Data[station]['height'] = Data[station]['Geopotential']*Rt/(g*Rt - Data[station]['Geopotential'])
-    Data[station]['Potential_temperature'] = Data[station]['Temperature']*(P0/Data[station]['Pressure levels'][:, None])**(Pc*(1 - 0.24*Data[station]['Specific humidity']))
[…]

### 2021-11-19 doc/auto_examples/data_presentation/wind_regime_comparison_plot.py

@@ -1,270 +1,0 @@
-"""
-===========
-Comparison between the Era5Land and the in situ datasets
-===========
-"""
-
-import numpy as np
-import matplotlib.pyplot as plt
-import matplotlib.colors as Mplcolors
-import matplotlib.gridspec as gridspec
-import os
-import sys
-sys.path.append('../../')
-import python_codes.theme as theme
-from python_codes.plot_functions import plot_scatter_surrounded, plot_wind_rose
-from python_codes.meteo_analysis import compute_circadian_annual_cycle
-from python_codes.general import smallestSignedAngleBetween, find_mode_distribution
-from scipy.stats import binned_statistic_2d
-theme.load_style()
-
-# paths
-path_savefig = '../../static/output_data/figures/'
-path_ouputdata = '../../static/output_data/data/'
-
-# Loading wind data
-Data = np.load(os.path.join(path_ouputdata, 'Data_final.npy'), allow_pickle=True).item()
-Stations = sorted(Data.keys())
-
-# Loading pattern characteristics
-Data_pattern = {}
-for station in ['Deep_Sea_Station', 'South_Namib_Station']:
-    Data_pattern[station] = np.load(os.path.join(path_ouputdata, 'Analysis_DEM_' + station + '.npy'), allow_pickle=True).item()
-
-
-# %%
-# Wind roses
-# ------------------------
-
-h = 115
-bins = [0.03, 0.15, 0.2, 0.25, 0.3, 0.35, 0.4]
-
-fig_width = theme.fig_width
-fig_height = 0.75*fig_width
-
-fig = plt.figure(figsize=(fig_width, fig_height))
-gs = gridspec.GridSpec(3, 2, height_ratios=[0.25, 1, 1], width_ratios=[1, 1])
-gs.update(left=0.025, right=0.97, bottom=0.03, top=0.97, wspace=0.15, hspace=0.25)
-for i, station in enumerate(Stations):
-    gs_sub = gs[i + 2].subgridspec(2, 2, height_ratios=[0.01, 1])
-    #
-    ax = plt.subplot(gs_sub[0, :])
-    plt.axis('off')
-    ax.text(0.5, 0.5, Stations[i][:-8].replace('_', ' '), ha='center', va='center', transform=ax.transAxes)
[…]

### 2021-11-19 doc/auto_examples/linear_theory/hydro_coeff_time_series.py

@@ -1,48 +1,0 @@
-r"""
-============================================
-Time series of the hydrodyanmic coefficients
-============================================
-
-Here, we calculate the time series of the hydrodynamic coefficients from the time series of the non dimensional parameters extracted from the meteorological and elevation data.
-"""
-
-import numpy as np
-import os
-import sys
-sys.path.append('../../')
-from python_codes.linear_theory import calculate_solution
-
-
-# Paths
-path_outputdata = '../../static/output_data/data/'
-
-# Importing non-dimensional numbers calculated
-Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
-Stations = ['South_Namib_Station', 'Deep_Sea_Station']
-numbers = {key: np.concatenate([Data[station][key] for station in Stations]) for key in ('Froude', 'kH', 'kLB')}
-
-# ##### Loading pattern characteristics
-Data_pattern = {}
-for station in ['Deep_Sea_Station', 'South_Namib_Station']:
-    Data_pattern[station] = np.load(os.path.join(path_outputdata, 'Analysis_DEM_' + station + '.npy'), allow_pickle=True).item()
-
-# Parameters
-Kappa = 0.4  # Von Kàrmàn constant
-k = np.concatenate([np.zeros(Data[station]['Froude'].shape) + 2*np.pi/(Data_pattern[station]['wavelength']*1e3) for station in Stations])  # vector of wavelength [m]
-z0 = 1e-3  # hydrodynamic roughness, [m]
-eta_0_vals = k*1e-3
-eta = 0  # non dimensional position where to calculate the solution (bottom)
-
-hydro_Coeffs = np.zeros((2, k.size))
-
-for i, (eta_0, eta_H, Froude, eta_B) in enumerate(zip(eta_0_vals, numbers['kH'], numbers['Froude'], numbers['kLB'])):
-    max_z = 0.9999*eta_H
-    if not np.isnan([eta_H, Froude, eta_B]).any():
-        Sol = calculate_solution(eta, eta_H, eta_0, eta_B, Froude, max_z, Kappa=0.4)
-        #
-        Ax, Bx = np.real(Sol[2]), np.imag(Sol[2])
-    else:
-        Ax, Bx = np.nan, np.nan
-    hydro_Coeffs[:, i] = [Ax, Bx]
-#
-np.save(os.path.join(path_outputdata, 'time_series_hydro_coeffs.npy'), hydro_Coeffs)

### 2021-11-19 doc/auto_examples/linear_theory/regime_diagrams_theoretical_plot.py

@@ -1,171 +1,0 @@
-r"""
-=================================================================
-Plot theoretical regime diagrams from parameter space exploration
-=================================================================
-
-"""
-
-import numpy as np
-import matplotlib.pyplot as plt
-from matplotlib.colors import TwoSlopeNorm, Normalize
-import matplotlib.gridspec as gridspec
-from scipy.ndimage import gaussian_filter
-import os
-import sys
-sys.path.append('../../')
-from python_codes.linear_theory import Cisaillement_basal
-import python_codes.theme as theme
-theme.load_style()
-
-# Paths
-path_outputdata = '../../static/output_data/data/'
-path_savefig = '../../static/output_data/figures/'
-
-# Loading parameter space exploration
-Dic = np.load(os.path.join(path_outputdata, 'parameter_exploration_hydro_coeff_3D.npy'), allow_pickle=True).item()
-#
-axes_av = [3, 2]  # axes over which selecting/averaging is perform for plane representation
-index = [60, 48]  # when selection, selecting these indexes in the relevant `axes_av`
-#
-hydro_coeffs_3D = Dic[('Froude', 'kH', 'kLB')]['hydro_coeffs'].swapaxes(1, 2)
-# axes are now (A/B, kH, Froude, kLB)
-
-# %%
-# Plane (Froude - kH)
-# ------------------------
-hydro_coeffs = hydro_coeffs_3D.take(index[0], axis=axes_av[0])
-
-# #### figure
-quantities = [hydro_coeffs[0, :, :], hydro_coeffs[1, :, :],
-              np.linalg.norm(hydro_coeffs, axis=0), np.arctan2(hydro_coeffs[1, :, :], hydro_coeffs[0, :, :])]
-labels = [r'$\mathcal{A}_{0}$', r'$\mathcal{B}_{0}$', r'$\sqrt{\mathcal{A}_{0}^{2} + \mathcal{B}_{0}^{2}}$',
-          r'$\textup{arctan}\left(\mathcal{B}_{0}/\mathcal{A}_{0}\right)$']
-#
-fig, axs = plt.subplots(2, 2, figsize=(theme.fig_width, 0.8*theme.fig_width), constrained_layout=True)
[…]

### 2021-11-26 Paper/main.tex
temp commit
First draft Introduction

@@
 \section{Introduction}
+
+Whenever a flow encounters an obstacle, various types of interactions can arise depending the different time and lengths scales involved.
+In the case of atmospheric flows, this depends mainly on the part of the vertical structure of the atmosphere, schematically composed of a turbulent boundary layer topped by a turbulence-free part, with which the obstacle interacts.
+%
+At the largest scale, the feedback of mountains on the stratified flow of the free atmosphere results in wave generation as well as significant wind disturbances, such as foehn winds in the lee side \citep{refs}.
+Inside the boundary layer, the interaction between a turbulent flow and hilly surfaces is for example key to the understanding ocean surface wind-driven waves, or eolian beforms in desert \citep{Belcher1998, Sullivan2010, Courrech2015}.
+
+Indeed, eolian sand dunes typically emerge from the interaction by the turbulent flow and the topography \citep{Charru2013}.
+Later on, when the dune reaches an intermediate size, it may induce large deflections that can significantly impact sediment pathways of coastal systems \citep{Hesp2015}.
+Likewise, the collective behavior of dune populations may also be affected by long-range interactions due to flow disturbances induced by each individual \citep{Smith2017, Bacik2020}.
+As the dunes increase in size by collisions and coarsening, they sometimes reach a giant size, comparable to the boundary layer depth, thus inducing interactions not only with the turbulent flow of the ABL, but also with the free atmosphere~\citep{andreotti2009}.
+However, the wind disturbances induced by these giant dunes have never been quantified.
+
+The study of the impact of obstacles on the atmospheric flows allow its incorporation within meteorological numerical model.
+Therefore, they mainly become limited by the precision of the included topographical data, as well as the spatial grid of the model.
+For example, the latest climate reanalysis, ERA5-Land, is limited by its $9~\textup{km}$ spatial resolution, while including the data $30$-m Digital Elevation Models (DEMs) of the shuttle radar topography mission \citep{Farr2007, munoz2021}.
+As such, it can not reproduce the flow disturbances induced by giant dunes, which have a typical length scale $\sim 1~\textup{km}$.
+
+Here, we compare the wind predictions from the ERA5-Land dataset to local measurements in four different places across the Namib desert.
+In places with no significant topographies smaller than the model grid, we show that both wind datasets agree with each other.
+On the contrary, in places with giant dunes, we show that they may differ for some specific meteorological conditions, that we link to the circadian cycle of the ABL.
+We thus highlight the importance of the mid-scale topographies for local wind regimes, and its implications in the case of sand seas for smaller-scale eolian bedforms.
+
 \section{Wind regimes across the Namib Sand Sea}

### 2021-11-26 Paper/main.tex
Clean start for the diffrenet branches

@@
 Whenever a flow encounters an obstacle, various types of interactions can arise depending the different time and lengths scales involved.
-In the case of atmospheric flows, this depends mainly on the part of the vertical structure of the atmosphere, schematically composed of a turbulent boundary layer topped by a turbulence-free part, with which the obstacle interacts.
+In the case of atmospheric flows, this depends mainly on the part of the vertical structure of the atmosphere, schematically composed of a turbulent boundary layer topped by a turbulence-free part, with which the obstacle interacts \citep{Stull1988}.
 %

### 2021-11-26 Paper/main.tex
Smoothing begining of section 2 and 2.2

@@
 
+In this study, we focus on four places across and nearby the Namib desert, highlighting different environments (see Fig.~\ref{Fig1}).
+%
+The Adamax station is located near the Adamax salt pan, in a highly vegetated area.
+The Huab station, located on the coast at the outlet of the Huab river is an arid environment exhibiting $60$-m scale barchan dunes.
+While these two stations are in environments with no mid-scale topography, this is not the case for the Deep Sea and South Namib stations.
+Both are located in the interdune between giant linear dunes with kilometric wavelengths and more than $40$-m high.
+In this section, we describe and compare the wind regimes resulting from the available datasets in each station.
+
   \begin{figure}
@@
   Two wind datasets are used in this study.
-First, local winds are provided by stations situated in different environments (see Fig.~\ref{Fig1}).
-The wind strength and direction are measured every 10 minutes by cup anemometers and wind vanes, at heights between $n$ and $n$ meters depending on the station.
+First, local winds are provided by stations situated in the four different places (see Fig.~\ref{Fig1}).
+The wind strength and direction are measured every 10 minutes by cup anemometers and wind vanes, at heights between $2$~m and $3$~m depending on the station.
 The available period of measurements ranges from 1 to 5 discontinuous years distributed between 2012 and 2020 (see Fig.~\ref{Fig1_supp}).
@@
   For comparison, the local measurements are averaged into 1-hr bins centered on the temporal scale of the ERA5-Land estimates (see Fig.~\ref{Fig2_supp}).
-As the wind velocities of both datasets are provided a different heights, we convert them into shear velocities (see SI section~\ref{}), characteristic of the whole turbulent wind profile within the atmospheric boundary layer, which are then used together with the wind direction for further analysis.
+As the wind velocities of both datasets are provided at different heights, we convert them into shear velocities (see SI section 1), characteristic of the whole turbulent wind profile within the atmospheric boundary layer, which are then used together with the wind direction for further analysis.
 The resulting wind data are shown on the wind roses of Fig.~\ref{Fig1}(b--e).
@@
   Finally, the dune properties are computed using an autocorrelation method from the 30-m Digital Elevation Models (DEMs) of the shuttle radar topography mission \citep{Farr2007}.
-For the South Namib and Deep Sea stations, we obtain respectively orientations of $85^\circ$ and $125^\circ$, wavelengths of $2.6~\textup{km}$ and $2.3~\textup{km}$ and amplitudes of $45~\textup{m}$ and $20~\textup{m}$ (see SI for additional details).
+For the South Namib and Deep Sea stations, we obtain respectively orientations of $85^\circ$ and $125^\circ$, wavelengths of $2.6~\textup{km}$ and $2.3~\textup{km}$ and amplitudes of $45~\textup{m}$ and $20~\textup{m}$ (see SI section 2 for additional details).
 
@@
 
-\section*{1. Calibration of the hydrodynamical roughness}
+\section*{1. Shear velocity and calibration of the hydrodynamical roughness}
+\label{calib_z0}
 

### 2021-12-02 Paper/main.tex
Re-ecriture jusqu'à section 3.1

@@
 
-Indeed, eolian sand dunes typically emerge from the interaction by the turbulent flow and the topography \citep{Charru2013}.
-Later on, when the dune reaches an intermediate size, it may induce large deflections that can significantly impact sediment pathways of coastal systems \citep{Hesp2015}.
-Likewise, the collective behavior of dune populations may also be affected by long-range interactions due to flow disturbances induced by each individual \citep{Smith2017, Bacik2020}.
+Indeed, eolian sand dunes typically emerge from the feedback of the topography on the turbulent flow, which speeds up close to the dune crest \citep{Rubin1987, Charru2013, Courrech2014}.
+Later on, when the dune reaches an intermediate size, it may also induce significant wind deflections.
+This can impact the sediment pathways of coastal systems \citep{Hesp2015}, or affect the collective behavior of dune populations with long-range interactions due to flow disturbances induced by each individual \citep{Smith2017, Bacik2020}.
 As the dunes increase in size by collisions and coarsening, they sometimes reach a giant size, comparable to the boundary layer depth, thus inducing interactions not only with the turbulent flow of the ABL, but also with the free atmosphere~\citep{andreotti2009}.
@@
 While these two stations are in environments with no mid-scale topography, this is not the case for the Deep Sea and South Namib stations.
-Both are located in the interdune between giant linear dunes with kilometric wavelengths and more than $40$-m high.
+Both are located in the interdune between giant linear dunes with kilometric wavelengths and superimposed patterns.
 In this section, we describe and compare the wind regimes resulting from the available datasets in each station.
@@
 
-  Finally, the dune properties are computed using an autocorrelation method from the 30-m Digital Elevation Models (DEMs) of the shuttle radar topography mission \citep{Farr2007}.
-For the South Namib and Deep Sea stations, we obtain respectively orientations of $85^\circ$ and $125^\circ$, wavelengths of $2.6~\textup{km}$ and $2.3~\textup{km}$ and amplitudes of $45~\textup{m}$ and $20~\textup{m}$ (see SI section 2 for additional details).
+  Finally, the dune properties are computed using autocorrelation on the 30-m Digital Elevation Models (DEMs) of the shuttle radar topography mission \citep{Farr2007}.
+For the South Namib and Deep Sea stations, we obtain respectively orientations of $85^\circ$ and $125^\circ$, wavelengths of $2.6~\textup{km}$ and $2.3~\textup{km}$ and amplitudes of $45~\textup{m}$ and $20~\textup{m}$ (see Fig.~\ref{Fig4_supp} for more details).
 
@@
 
-  The regional wind patterns in the Namib are strongly controlled by the see breeze, resulting in strong northward components present in all regional wind roses of figure~\ref{Fig1}, also deviated by the large scale topography \citep{lancaster1985}.
-These daily winds are the strongest during the second-half of the year (Septembre-January).
+  The obtained wind regimes are shown in figure~\ref{Fig1}.
+In the Namib, the regional wind patterns are essentially controlled by the see breeze, resulting in strong northward components (sometimes slightly deviated by the large scale topography) present in all regional wind roses \citep{lancaster1985}.
+These daily winds are dominant during the second-half of the year (Septembre-January).
 In winter, an additional easterly component can be recorded during the night, induced by the combination of katabatic winds forming on the mountains, and infrequent `berg' winds, which are responsible of the high wind velocities observed \citep{lancaster1984}.
-The frequency of these easterly components decreases from the inland to the coast, resulting in three bidirectional wind regimes within the Namib Sand Sea and at the Adamax salt pan (Fig.~\ref{Fig1}(c, d, e)) and one unidirectional wind regime on the coast at the outlet of the Huab River.
+The frequency of these easterly components decreases from the inland to the coast, resulting in bidirectional wind regimes within the Namib Sand Sea and at the Adamax salt pan (Fig.~\ref{Fig1}b, \ref{Fig1}d and \ref{Fig1}e) and a unidirectional wind regime on the coast at the outlet of the Huab River (Fig.~\ref{Fig1}c).
 
-  In the case of the Adamax and Huab stations, these regional wind roses qualitatively matches their respective in situ local measurements.
-However, for the Deep Sea and South Namib stations, the local wind roses exhibit additional components aligned with the giant dune orientation visible on the satellite images.
-
-  Indeed, the analysis of the wind speed and direction time series shows that the matching between the local and regional winds is verified for the stations with flat environment at all times (Fig.~\ref{Fig2}(a--b)), as well as for the stations within the giant dune field, but only during Septembre-January (Fig.~\ref{Fig2}(c--d).
-This agreement is statistically verified, as shown by the Fig.~\ref{Fig5_supp}.
+  In the case of the Adamax and Huab stations, the regional wind roses qualitatively match those corresponding to the local in situ measurements.
+However, for the Deep Sea and South Namib stations, the local wind roses exhibit additional components aligned with the giant dune orientation visible on the satellite images (Fig.~\ref{Fig1}c--d).
+  %
+  Indeed, the analysis of the wind speed and direction time series shows that the agreement between the local and regional datasets is always verified when no mid-scale topography are present (Fig.~\ref{Fig2}a--b) and Fig.~\ref{Fig5_supp}).
+In contrast, for the stations within the giant dune field, we observe that this agreement is limited to the Septembre--January time periods (Fig.~\ref{Fig2}c--d).
 
@@
 
-  In winter, the local and regional winds match only during the morning for the stations within the giant dune field, when the southerly/southwesterly sea breeze dominates (see Fig.~\ref{Fig2}(e--f), Fig.~\ref{Fig3} and Fig.~\ref{Fig6_supp}).
-In the late afternoon and during the night, when the northwesterly `berg' and katabatic winds blow, the two datasets differs.
-More specifically, this mismatch mainly occurs during at low wind velocities.
-In this case, the angular wind distribution of the local measurements exhibits two additional modes separated of $\simeq 180^\circ$, each corresponding to the giant dune alignement (purple frame in Fig.~\ref{Fig3} and Fig.~\ref{Fig6_supp}).
-This deviation at low wind velocities is statistically verified, and is also associated with a global attenuation of the wind strength (Fig.~\ref{Fig6_supp} and Fig.~\ref{Fig7_supp}).
+  When giant dunes are present, in the February--August period, the local and regional winds match only during the morning, i.e when the southerly/southwesterly sea breeze dominates (see Fig.~\ref{Fig2}(e--f), Fig.~\ref{Fig3} and Fig.~\ref{Fig6_supp}).
+In the late afternoon and during the night, when the northwesterly `berg' and katabatic winds blow, the two datasets differ.
+In this case, the angular wind distribution of the local measurements exhibits two additional modes separated of $\simeq 180^\circ$, each corresponding to the giant dune alignement (purple frame in Fig.~\ref{Fig3} and Fig.~\ref{Fig6_supp}, as well as Fig.~\ref{Fig7_supp}).
+This deviation is also associated with a global attenuation of the wind strength (Fig.~\ref{Fig8_supp}).
+Remarkably, all these figures show that this process occurs for low wind velocities, typically for $u_{*} < 0.1~\textrm{m}~\textrm{s}^{-1}$.
+For shear velocities larger than $0.25~\textrm{m}~\textrm{s}^{-1}$, this wind reorientation does not occur.
+Finally, for intermediate shear velocities, both reorientation along the dune crest and no reorientation are observed (Fig.~\ref{Fig7_supp}).
 
@@
 
-  Previous studies have linked meteorological flow around and over obstacles to the vertical structure of the atmosphere \citep{Stull1988}.
-At the dune scale, its lower part, the turbulent atmospheric boundary layer (ABL), is directly influenced by the planetary surface \citep{Stull1988}.
-When the latter is warmer than the ambient air (typically during the day), a strong convection takes place, resulting in a logarithmic wind profile and a vertically constant potential temperature $\theta$ (which corresponds to the air density corrected from the vertical pressure variation).
-During the night, the surface cools faster than the ambient air, which can result in a stable layer close to the surface (typically the first $200$ meters).
+  In the case of linear ridges, dune-induced flow disturbances have mainly been related to the incident wind direction~\citep{Walker2009, Hesp2015}.
+In our case, it is unlikely to be the dominant parameter, as the most deflected wind for both stations is the most perpendicular, where it should be winds with a $30^{\circ}--70^{\circ}$ incident direction~\citep{Hesp2015}.
+  An important observation is the difference in behavior between low and high wind velocities, which suggests a change in the hydrodynamical regime.
+
+  Previous studies have linked meteorological flow around and over topographical obstacles to the vertical structure of the atmosphere \citep{Stull1988}.
+More particularly, dunes evolves in its lower part, the turbulent atmospheric boundary layer (ABL), typically characterized by a logarithmic wind profile and a vertically constant potential temperature $\theta$.
 Above, the free atmosphere (FA) is a stably stratified zone in which turbulence is negligible, and where the flow is usually considered as incompressible and inviscid.
-In the middle, a transitional layer, also known as entrainment zone, is characterized by a sharp increase of the potential temperature, which traps turbulence resulting fro the surface friction below it.
-As it is much shallow than the other two layers, we consider it as capping singularity inducing a discontinuity in the potential temperature profile.
+In the middle, a transitional layer, also known as entrainment zone, is characterized by a sharp increase of the potential temperature, which traps turbulence resulting from the surface friction below it.
+
+  In the following, we sum-up the dominant numbers leading to different hydrodynamical interactions with the obstacle, and interpret the data with respect to the corresponding physical mechanisms.
 

### 2021-12-03 Paper/main.tex
Rewriting paper up to the end of 3.1

@@
 \usepackage{appendix}
-\usepackage{amsmath}
+\usepackage{amsmath,amssymb}
 \usepackage{graphicx}
@@
 At the largest scale, the feedback of mountains on the stratified flow of the free atmosphere results in wave generation as well as significant wind disturbances, such as foehn winds in the lee side \citep{refs}.
-Inside the boundary layer, the interaction between a turbulent flow and hilly surfaces is for example key to the understanding ocean surface wind-driven waves, or eolian beforms in desert \citep{Belcher1998, Sullivan2010, Courrech2015}.
+Inside the boundary layer, the interaction between a turbulent flow and hilly surfaces is for example key to the understanding ocean surface wind-driven waves, or eolian bedforms in desert \citep{Belcher1998, Sullivan2010, Courrech2015}.
 
@@
   \subsection{Influence of the giant dunes on local wind regimes}
+  \label{section_data_feedback}
 
@@
   In the case of linear ridges, dune-induced flow disturbances have mainly been related to the incident wind direction~\citep{Walker2009, Hesp2015}.
-In our case, it is unlikely to be the dominant parameter, as the most deflected wind for both stations is the most perpendicular, where it should be winds with a $30^{\circ}--70^{\circ}$ incident direction~\citep{Hesp2015}.
+In our case, it is unlikely to be the dominant parameter, as the most deflected wind for both stations is the most perpendicular, where it should be winds with incident directions between $30^{\circ}$ and $70^{\circ}$~\citep{Hesp2015}.
 An important observation is the difference in behavior between low and high wind velocities, which suggests a change in the hydrodynamical regime.
 
-  Previous studies have linked meteorological flow around and over topographical obstacles to the vertical structure of the atmosphere \citep{Stull1988}.
-More particularly, dunes evolves in its lower part, the turbulent atmospheric boundary layer (ABL), typically characterized by a logarithmic wind profile and a vertically constant potential temperature $\theta$.
+  Previous studies have linked atmospheric flow around and over topographical obstacles to the vertical structure of the atmosphere \citep{Stull1988}.
+More particularly, dunes evolves in its lower part, the turbulent atmospheric boundary layer (ABL), typically characterized by a logarithmic wind profile and a vertically constant potential temperature.
 Above, the free atmosphere (FA) is a stably stratified zone in which turbulence is negligible, and where the flow is usually considered as incompressible and inviscid.
-In the middle, a transitional layer, also known as entrainment zone, is characterized by a sharp increase of the potential temperature, which traps turbulence resulting from the surface friction below it.
+In the middle, a transitional layer, also known as entrainment zone, is characterized by a sharp increase of the potential temperature, which traps the turbulence resulting from the surface friction below it.
 
-  In the following, we sum-up the dominant numbers leading to different hydrodynamical interactions with the obstacle, and interpret the data with respect to the corresponding physical mechanisms.
+  In the following, we sum-up the dominant numbers leading to different hydrodynamical interactions with topographical obstacles, and interpret the data with respect to the corresponding physical mechanisms.
 
@@
 
-  As a result, the giant dunes evolve in a turbulent flow capped by a stratified atmosphere, with a transitional layer acting as a free surface in between.
-When the dune are small enough, typically in the case of dune emergence or early coarsening (Fig.~\ref{Fig4}b), they do no interact with the top of the ABL \citep{Fourriere2010}.
-However, the giant dunes present in the Deep Sea and South Namib stations have wavelengths comparable or larger than typical ABL depths, inducing their interaction with the capping layer and the free atmosphere above \citep{andreotti2009}.
-However, as sketched in Fig.~\ref{Fig4}c--d, the nature of this interaction depends on two non-dimensional number, both called Froude numbers in the literature~\citep{Vosper2004, Sheridan2006, Hunt2006, Jiang2014}.
+  Flow deflection over ridges can be simplistically understood from a balance between inertia and pressure gradients~\citep{Hesp2015}.
+As the flow approaches the ridge crest, the compression of the streamlines results in larger flow velocities, and thus lower pressures~\citep{Rubin1987}.
+An incident flow oblique to the ridge is then deflected towards lower pressure zones, i.e towards the crest.
+Turbulent dissipation at the bottom and non-linearities tends to increase this effect downstream, resulting in along the crest wind deflection in the lee side~\citep{Hesp2015, Gadal2019}.
 
-  The ability of the capping layer to deform under the presence of an underlying topography is controlled by the \emph{surface Froude number}, typically used in the context of bedforms in rivers \citep{Fourriere2010, Andreotti2012}:
+  Another way to increase the flow deflection is its confinement below a capping surface, that result in further streamline compression.
+This happens when the flow disturbance induced by the obstacle reaches the surface.
+As obstacles typically disturb flow over a characteristic height similar to their width, this is well captured by the parameter $k H$, where $k = 2\pi/\lambda$ is the wavenumber and $H$ the ABL depth.
+Here, the giant dunes have kilometric wavelengths, such that $0.02 \lesssim k H \lesssim 5$, and they interact most of the time with the capping layer and the stratified free atmosphere above \citep{andreotti2009}.
+
+  Note that the ability of the capping layer and stratification to accommodate a perturbation induced by the topography directly impacts the strength of this confinement effect (Fig.~\ref{Fig4}).
+This is typically quantified using surface and internal Froude numbers
+  \citep{Vosper2004, Stull2006, Sheridan2006, Hunt2006, Jiang2014}:
   \begin{equation}
-    \mathcal{F}r_{\textup{S}} = \displaystyle\frac{U}{\sqrt{\displaystyle\frac{\Delta\rho}{\rho}gH}},
+        \mathcal{F}r_{\textup{S}} = \displaystyle\frac{U}{\sqrt{\displaystyle\frac{\Delta\rho}{\rho}gH}}, \, \mathcal{F}r_{\textup{I}} = \displaystyle\frac{k U}{N},
   \end{equation}
-  where $U$ is the wind velocity at the top of the ABL, $H$ its depth, $\rho$ its average density and $\Delta\rho$ the density jump between the ABL and the FA.
-  %
-  Likewise, the FA stratification ability to deform is controlled by the \emph{internal Froude number}, typically used in the context of flow disturbances induced by mountains \citep{Stull1988}:
-  \begin{equation}
-    \mathcal{F}r_{\textup{I}} = \displaystyle\frac{U}{H N},
-  \end{equation}
-  where $N$ is the Brunt-Väisälä frequency, characteristic of a stratified fluid.
-  %
-  When the flow inertia dominates over its weight ($\mathcal{F}r_{\textup{S}} \gg 1$) and the weight of the stratification ($\mathcal{F}r_{\textup{I}} \gg 1$), both the capping layer and the stratification deform and adapt to the underlying topography (Fig.~\ref{Fig4}d), resulting in an unconfined flow \citep{refs}.
-However, an opposite force balance induces a rigid capping layer and/or stratification (Fig.~\ref{Fig4}c), resulting in confined flows.
-In this case, the confined flow has trouble going smoothly over the obstacle, and are then typically associated with large deviations and modulation of the wind velocity, as qualitatively illustrated in the case of the Deep Sea station by Fig.~\ref{Fig4}e \citep{refs}.
+  where $U$ is the wind velocity at the top of the ABL, $\rho$ its average density, $\Delta\rho$ the density jump between the ABL and the FA and $N$ is the Brunt-Väisälä frequency, characteristic of the stratification.
+
+  The smallest wind disturbances are expected during the day, when the ABL depth is comparable to the dune wavelength ($k H \gtrsim 1$) and for large wind velocities, which correspond to a weak confinement situation (Fig.~\ref{Fig4b}).
+On the contrary, large wind disturbances are expected to occur during the night, when the confinement is mainly induced by shallow ABL (Fig.~\ref{Fig4d--f}).
+Note that this strong confinement can be somewhat reduced in the case of strong winds (corresponding to large Froude numbers, see Fig.~\ref{Fig4f}), explaining the threshold in velocity observed in the data (see section~\ref{section_data_feedback}).
 

### 2021-12-10 Paper/main.tex
Rewriting section 3.2

@@
 
-  Another way to increase the flow deflection is its confinement below a capping surface, that result in further streamline compression.
+  Another way to increase the flow deflection is its confinement below a capping surface, that results in further streamline compression.
 This happens when the flow disturbance induced by the obstacle reaches the surface.
-As obstacles typically disturb flow over a characteristic height similar to their width, this is well captured by the parameter $k H$, where $k = 2\pi/\lambda$ is the wavenumber and $H$ the ABL depth.
+As obstacles typically disturb flow over a characteristic height similar to their width, the potential of interaction between the dunes and the overlying atmospheric structure is well captured by the parameter $k H$, where $k = 2\pi/\lambda$ is the wavenumber and $H$ the ABL depth.
 Here, the giant dunes have kilometric wavelengths, such that $0.02 \lesssim k H \lesssim 5$, and they interact most of the time with the capping layer and the stratified free atmosphere above \citep{andreotti2009}.
@@
 
-  The smallest wind disturbances are expected during the day, when the ABL depth is comparable to the dune wavelength ($k H \gtrsim 1$) and for large wind velocities, which correspond to a weak confinement situation (Fig.~\ref{Fig4b}).
-On the contrary, large wind disturbances are expected to occur during the night, when the confinement is mainly induced by shallow ABL (Fig.~\ref{Fig4d--f}).
-Note that this strong confinement can be somewhat reduced in the case of strong winds (corresponding to large Froude numbers, see Fig.~\ref{Fig4f}), explaining the threshold in velocity observed in the data (see section~\ref{section_data_feedback}).
+  The smallest wind disturbances are expected during the day, when the ABL depth is comparable to the dune wavelength ($k H \gtrsim 1$) and for large wind velocities, which correspond to a weak confinement situation (Fig.~\ref{Fig4}b).
+On the contrary, large wind disturbances are expected to occur during the night, when the confinement is mainly induced by shallow ABL (Fig.~\ref{Fig4}d--f).
+Note that this strong confinement can be somewhat reduced in the case of strong winds (corresponding to large Froude numbers, see Fig.~\ref{Fig4}f), explaining the threshold in velocity observed in the data (see section~\ref{section_data_feedback}).
 
@@
 
-  In the spirit of \citet{Sheridan2006}, we aim to compute flow regime diagrams in the space defined by the three relevant non-dimensional numbers presented above, $\left(k H,\, \mathcal{F}r_{\textup{S}}, \, \mathcal{F}r_{\textup{I}}\right)$.
-They are calculated from the time series of the geopotential, temperature and specific humidity vertical profiles available in the ERA5 climate reanalysis (see SI~\ref{?}).
+  To highlight these different regimes from our data, we compute wind disturbance diagrams in the space defined by the three relevant non-dimensional numbers presented above, $\left(k H,\, \mathcal{F}r_{\textup{S}}, \, \mathcal{F}r_{\textup{I}}\right)$.
+Those are calculated from the time series of the geopotential, temperature and specific humidity vertical profiles available in the ERA5 climate reanalysis (see SI section 2).
+  %
+  Flow deviation is computed as the minimal angle between the wind orientations from the two datasets:
+  \begin{equation}
+    \delta_{\theta} = \left\vert \textup{min}\left(\left[\theta_{\textup{ERA}} - \theta_{\textup{station}}\right] \, \textup{mod.} \, 360, \left[\theta_{\textup{station}} - \theta_{\textup{ERA}}\right] \, \textup{mod.} \, 360\right) \right\vert.
+  \end{equation}
+  %
   The relative velocity modulation is computed as
@@
   \begin{equation}
-    \delta_{\textup{u}} = \frac{u_{*}^{\textup{ERA}} -  u_{*}^{\textup{station}}}{u_{*}^{\textup{ERA}}},
+    \delta_{\textup{u}} = \frac{u_{*}^{\textup{ERA}} -  u_{*}^{\textup{station}}}{u_{*}^{\textup{ERA}}}.
   \end{equation}
+
+  As both Froude numbers have qualitatively the same impact on the flow confinement, we first leave aside the internal Froude number, and focus on the space $(k H,\, \mathcal{F}r_{\textup{S}})$.
+When representing the two variables $\delta_{\theta}$ and $\delta_{\textup{u}}$ in this space, different regime emerges (Fig.~\ref{Fig5}).
+Small wind disturbances ($\delta_{\theta} \to 0$, $\delta_{\textup{u}} \to 0$) are located in the top-right part of the diagrams, corresponding to a regime mixing low-interaction and low-confinement ($k H$ and $\mathcal{F}r_{\textup{S}}$ large enough, Fig.~\ref{Fig4}c).
   %
-  and the flow deviation as the minimal angle between the wind orientation in the two datasets:
-  %
-  \begin{equation}
-    \delta_{\theta} = \left\vert \textup{min}\left(\left[\theta_{\textup{ERA}} - \theta_{\textup{station}}\right] \, \textup{mod.} \, 360, \left[\theta_{\textup{station}} - \theta_{\textup{ERA}}\right] \, \textup{mod.} \, 360\right) \right\vert,
-  \end{equation}
-  %
+  Lower values of $k H$ (stronger interaction) or Froude numbers (stronger confinement) then both leads to an increase in wind disturbances, both in terms of orientation and velocity.
+Below a threshold value of $k H \simeq 0.3$, wind disturbance occurs independently of the Froude numbers value.
+Furthermore, this also seems to control a transition between from damped to amplified wind velocities within the interdune (Fig.~\ref{Fig5}b), for which we do not have an explanation.
+Note that the same interpretation can be done with the diagrams including the internal Froude number $\mathcal{F}r_{\textup{I}}$, as shown by Fig.~\ref{Fig12_supp}.
 
-  When these two variables are represented in the marginal spaces $(k H,\, \mathcal{F}r_{\textup{S}})$ and $(k H,\, \mathcal{F}r_{\textup{I}})$, different regime emerges (Fig.~\ref{Fig5}).
-The small wind disturbances ($\delta_{\theta} \to 0$, $\delta_{\textup{u}} \to 0$) are located in the top-right part of the diagrams, corresponding to a regime mixing low-interactions ($k H$ large enough, Fig.~\ref{Fig4}b) and low-confinement ($\mathcal{F}r_{\textup{S}}, \, \mathcal{F}r_{\textup{I}}$ large enough, Fig.~\ref{Fig4}c).
-
-  Lower values of $k H$ (stronger interaction) or Froude numbers (stronger confinement) then both leads to an increase in wind disturbances, both in terms of orientation and velocity.
   Interestingly, the limit of no-interactions between the topography and the boundary layer structure ($k H \gg 1$), in which the properties of the capping layer and the stratification become irrelevant (Fig.~\ref{Fig4}b--\ref{Fig4}c), is never reached here, in the case of giant dunes.
-  %
-  Below a threshold value of $k H \simeq 0.3$, wind disturbance occurs independently of the Froude numbers value.
-However, the latter seem to control a transition between from damped to amplified wind velocities within the interdune (Fig.~\ref{Fig5}c--\ref{Fig5}d), for which we do not have an explanation.
 

### 2021-12-13 Paper/main.tex
Reviewing figure captions

@@
 \textbf{b--e}: Satellite images of the studied sites (Google-Earth, Maxar Technologies, CNES/Airbus).
-In each subplot, the left and right wind roses represent the data from the ERA5Land climate reanalysis and the in situ stations, respectively.
+In each subplot, the left and right wind roses represent the data from the ERA5Land climate reanalysis and the local wind stations, respectively.
 Note that the bars show the direction towards which the wind blows.
-The red dots show the location of the in situ stations.}
+The red dots show the location of local wind stations.}
     \label{Fig1}
@@
     \includegraphics[scale=1]{Figures/Figure2.pdf}
-    \caption{Temporal comparison between the wind data coming from the Era5Land climate reanalysis (orange lines) and from the in situ measurements (blue lines).
+    \caption{Temporal comparison between the wind data coming from the Era5Land climate reanalysis (orange lines) and from the local measurements (blue lines).
 \textbf{a--b}: Huab station.
@@
     \includegraphics[scale=1]{Figures/Figure3.pdf}
-    \caption{Distributions of wind direction at the Deep Sea Station for the Era5Land climate reanalysis (orange) and the in situ measurements (blue).
-In each subplot, both distributions are plotted from the same time steps, selected with constraints on the wind direction (columns) and/or wind velocity (rows) in the Era5Land dataset.
-The vertical gray dashed lines indicate the dune orientation, and the top right numbers the percentage of the total number of time steps selected in each subplot.
-The purple frame highlights the regime (small wind velocities, nocturnal summer wind) during which the wind data from both datasets differs.
+    \caption{Distributions of wind direction at the Deep Sea Station for the Era5Land climate reanalysis (orange) and the local measurements (blue).
+In each subplot, both distributions are plotted from the same time steps, selected with constraints on the wind direction (columns) and/or wind velocity (rows) of the Era5Land dataset.
+The gray vertical dashed lines indicate the dune orientation, and the top right numbers the percentage of the total number of time steps selected in each subplot.
+The purple frame highlights the regime (small wind velocities, nocturnal summer wind) in which the data from both datasets differs.
 A similar figure can be obtained for the Deep Sea station (see Fig.~\ref{Fig6_supp}).}
@@
     \includegraphics[scale=1]{Figures/Figure4.pdf}
-    \caption{\textbf{a}: Vertical profiles of the virtual potential temperature at 2 different time steps (blue - 29/11/2012 - 1100 UTC, red - 21/03/2017 - 1200 UTC) at the Deep Sea station.
+    \caption{\textbf{a--b}: Vertical profiles of the virtual potential temperature at 2 different time steps (blue - 29/11/2012 - 1100 UTC, red - 21/03/2017 - 1200 UTC) at the Deep Sea station.
 Dots: data from the ERA5 reanalysis.
@@
 Plain lines: vertical (boundary layer) and linear (free atmosphere) fits to estimate the stratification properties.
-\textbf{b--d}: Sketches representing the interaction between the giant dunes and the atmospheric flow for different meteorological conditions.
-\textbf{e}: Streamlines qualitatively representing the effect of flow confinement, in the case of the Deep Sea station.
-The red and straight blue lines are calculated from the unconfined case, representing the situations \textbf{b--c}.
-The sinuous blue line represents the confined case of \textbf{d}.
-For details on their derivation, see Appendix.}
+\textbf{c--f}: Sketches representing the interaction between the giant dunes and the atmospheric flow for different meteorological conditions.
+\textbf{g}: Streamlines qualitatively representing the effect of flow confinement, in the case of the Deep Sea station.
+The red lines are calculated from the unconfined case similar to \textbf{d}.
+The sinuous blue line represents a confined case similar to \textbf{f}.
+For details on the streamline derivation, see Appendix~\ref{turbulent_wind_model}.}
     %
@@
 
-  The smallest wind disturbances are expected during the day, when the ABL depth is comparable to the dune wavelength ($k H \gtrsim 1$) and for large wind velocities, which correspond to a weak confinement situation (Fig.~\ref{Fig4}b).
-On the contrary, large wind disturbances are expected to occur during the night, when the confinement is mainly induced by shallow ABL (Fig.~\ref{Fig4}d--f).
+  The smallest wind disturbances are expected during the day, when the ABL depth is comparable to the dune wavelength ($k H \gtrsim 1$) and for large wind velocities, which correspond to a weak confinement situation (Fig.~\ref{Fig4}d).
+On the contrary, large wind disturbances are expected to occur during the night, when the confinement is mainly induced by shallow ABL (Fig.~\ref{Fig4}e--f).
 Note that this strong confinement can be somewhat reduced in the case of strong winds (corresponding to large Froude numbers, see Fig.~\ref{Fig4}f), explaining the threshold in velocity observed in the data (see section~\ref{section_data_feedback}).
@@
 When representing the two variables $\delta_{\theta}$ and $\delta_{\textup{u}}$ in this space, different regime emerges (Fig.~\ref{Fig5}).
-Small wind disturbances ($\delta_{\theta} \to 0$, $\delta_{\textup{u}} \to 0$) are located in the top-right part of the diagrams, corresponding to a regime mixing low-interaction and low-confinement ($k H$ and $\mathcal{F}r_{\textup{S}}$ large enough, Fig.~\ref{Fig4}c).
+Small wind disturbances ($\delta_{\theta} \to 0$, $\delta_{\textup{u}} \to 0$) are located in the top-right part of the diagrams, corresponding to a regime mixing low-interaction and low-confinement ($k H$ and $\mathcal{F}r_{\textup{S}}$ large enough, Fig.~\ref{Fig4}d).
   %
@@
 
-  Interestingly, the limit of no-interactions between the topography and the boundary layer structure ($k H \gg 1$), in which the properties of the capping layer and the stratification become irrelevant (Fig.~\ref{Fig4}b--\ref{Fig4}c), is never reached here, in the case of giant dunes.
+  Interestingly, the limit of no-interactions between the topography and the boundary layer structure ($k H \gg 1$), in which the properties of the capping layer and the stratification become irrelevant, is never reached here, in the case of giant dunes.
 
@@
 The ellipses indicates the different types of elongating dunes, at large (plain) and small scale (dashed).
-The dune orientations are calculated using the model of \citet{Cour14} from the sand flux angular distributions, shown here for typical sand quartz grains of $180~\mu$m.
+The dune orientations are calculated using the model of \citet{Courrech2014} from the sand flux angular distributions, shown here for typical sand quartz grains of $180~\mu$m.
 The double blue and single red arrows correspond to the two possible dune growth mechanisms, bed instability and elongation, respectively.
@@
 The green dots indicate the position of the measurement stations.
-See Appendix 2 for details.}
+See Appendix 2 for additional details.}
   \label{Fig6}
@@
 \section*{Appendix 1: ABL turbulent wind model}
+\label{turbulent_wind_model}
 
@@
   \includegraphics[scale=1]{Figures/Figure1_supp.pdf}
-  \caption{Gant chart representing the usable time steps for the two data sets, for all stations.}
+  \caption{Gant chart representing the valid time steps for the two data sets, for all stations.}
   \label{Fig1_supp}
@@
   \includegraphics[scale=1]{Figures/Figure2_supp.pdf}
-  \caption{Comparison between raw in situ wind measurements, and centered averaged data over one hour for the South Namib station.
+  \caption{Comparison between raw local wind measurements, and centered averaged data over one hour for the South Namib station.
 \textbf{a}: wind direction.
@@
 The red dashed and plain lines shows the minima of $\delta$ and the identity line.
-The orange lines and dots highlights the chosen the hydrodynamic roughnesses for the in situ datasets by imposing $z_{0, \textup{\textup{ERA}}} = 1~\textup{mm}$, leading for each station to $2.7~\textup{mm}$, $0.76~\textup{mm}$, $0.12~\textup{mm}$ and $0.48~\textup{mm}$, respectively.}
+The orange lines and dots highlight the chosen hydrodynamic roughnesses for the local datasets, by imposing $z_{0, \textup{\textup{ERA}}} = 1~\textup{mm}$, leading for each station to $2.7~\textup{mm}$, $0.76~\textup{mm}$, $0.12~\textup{mm}$ and $0.48~\textup{mm}$, respectively.}
   \label{Fig3_supp}
@@
   \caption{Analysis of the DEMs of the Deep Sea (left column -- \textbf{a}, \textbf{b}, \textbf{c}) and South Namib (right column -- \textbf{d}, \textbf{e}, \textbf{f}) stations.
-\textbf{a--c}: Detrended topography (a second order polynomial is first fitted and then removed) .
-\textbf{b--d}: autocorrelation matrix shown in colorscale.
-The black line shows the detected orientation, and the red line the profile along which the wavelength is calculated, shown in \textbf{e--f}.
-The blue lines and dots show the first peak of the autocorrelation profile, whose horizontal position gives the characteristic wavelength of the dune pattern.}
+\textbf{a--d}: Detrended topography (a second order polynomial is first fitted and then removed) .
+\textbf{b--e}: autocorrelation matrix shown in colorscale.
+The black line shows the detected orientation, and the red line the profile along which the wavelength is calculated, shown in \textbf{c--f}.
+The blue lines and dots show the first peak of the autocorrelation profile, whose abscissa gives the characteristic wavelength of the dune pattern.}
   \label{Fig4_supp}
@@
   \includegraphics[scale=1]{Figures/Figure5_supp.pdf}
-  \caption{Statistical agreement of the wind orientation (\textbf{a}) and velocity (\textbf{b}) between the Era5Land dataset and the in situ measurements for the Huab and Adamax stations.
-Note how the points are clustered around identity lines, black and dashed.}
+  \caption{Statistical agreement of the wind orientation (\textbf{a}) and velocity (\textbf{b}) between the Era5Land dataset and the local measurements for the Huab and Adamax stations.
+Note how the points are clustered around identity lines (dashed and black).}
   \label{Fig5_supp}
@@
   \includegraphics[scale=1]{Figures/Figure7_supp.pdf}
-  \caption{Statistical comparison of the wind orientation between the Era5Land dataset and in situ measurements for the South Namib and Deep Sea stations, for different velocity ranges.
+  \caption{Statistical comparison of the wind orientation between the Era5Land dataset and the local measurements for the South Namib and Deep Sea stations, for different velocity ranges.
 \textbf{a}: $u_{*, \textup{\textup{ERA}}} < 0.1~\textup{m}~\textup{s}^{-1}$.
@@
 \textbf{c}: $u_{*, \textup{\textup{ERA}}} \geq 0.25~\textup{m}~\textup{s}^{-1}$.
-Note that the dune orientations measured are substracted to the wind orientation, which allows to plot both stations on the same graph.
-Black dashed lines indicates in situ orientations aligned with the dune crests (here $0^\circ$, $180^\circ$ and $360^\circ$ -- \textbf{a}, \textbf{b}), as well as the identity lines (\textbf{b}, \textbf{c}).}
+Note that the measured dune orientations are substracted to the wind orientation, which allows to plot both stations on the same graph.
+Black dashed lines indicate locally measured orientations aligned with the dune crests (here $0^\circ$, $180^\circ$ and $360^\circ$ -- \textbf{a}, \textbf{b}), as well as the identity lines (\textbf{b}, \textbf{c}).}
   \label{Fig7_supp}
@@
   \includegraphics[scale=1]{Figures/Figure8_supp.pdf}
-  \caption{Statistical comparison of the wind velocity between the Era5Land dataset and in situ measurements for the South Namib and Deep Sea stations.
+  \caption{Statistical comparison of the wind velocity between the Era5Land dataset and the local measurements for the South Namib and Deep Sea stations.
 \textbf{a}: Nocturnal summer easterly wind.
@@
 Black dashed lines are identity lines.
-The angle ranges corresponding to diurnal and nocturnal summer winds are those taken in Fig.~\ref{Fig3} and Fig.~\ref{Fig6_supp}.}
+The angle ranges used to select diurnal and nocturnal summer winds are those taken in Fig.~\ref{Fig3} and Fig.~\ref{Fig6_supp}.}
   \label{Fig8_supp}
@@
 The point density in each bin of the diagrams is shown in Fig.~\ref{Fig11_supp}.
-The regime diagrams in the space $(\mathcal{F}r_{\textup{S}}, \, kH)$ are shown in Fig.~\ref{Fig5_supp}.}
+The regime diagrams in the space $(\mathcal{F}r_{\textup{S}}, \, kH)$ are shown in Fig.~\ref{Fig5} of the main article.}
   \label{Fig12_supp}

### 2021-12-15 Paper/main.tex
Rewriting conclusion

@@
 
- The comparison of local and regional wind data gives a direct evidence of the giant dunes feedback on the flow.
-In flat areas, the matching between both datasets highlights the ability of the latest generation of climate reanalysis to predict the wind flow up to scales $\sim 10~\textrm{km}$, i.e the grid model.
+ The comparison of local (direct measurements) and regional (climate reanalysis) wind data reveals the giant dunes feedback on the flow.
+In flat areas, the matching between both datasets confirms the ability of the ERA5Land climate reanalysis to predict the wind flow down to scales $\sim 10~\textrm{km}$, i.e the grid model.
 When smaller scale topographies are present (giant dunes in our case), locally measured wind regimes may significantly differ from the regional ones.
 Furthermore, we link these disturbances induced by the dunes to their interaction with the lower part of the atmospheric vertical structure, and more specifically to its circadian variability.
-During the day, the top of the ABL is high enough to limit the interaction of the capping layer and the FA stratification with the giant dunes, resulting in a low flow confinement, and thus small wind disturbances.
-During the night, the small ABL height induces a stronger flow confinement, associated with large wind deviation and acceleration or deceleration.
-Interestingly, we also found that this effect could be counterbalanced by the presence of large wind velocities, capable of deforming the capping layer and/or the FA stratification and thus decreasing the confinement effect.
+During the night, the presence of a shallow atmospheric boundary layer (ABL) induces a strong flow confinement, associated with large wind deviation and acceleration or deceleration.
+During the day, the capping layer is high enough to prevent its interaction with the giant dunes, resulting in a low flow confinement, and thus smaller wind disturbances.
+Interestingly, we also found that this effect could be counterbalanced by the presence of large wind velocities, capable of deforming the capping layer and/or the FA stratification, thus decreasing the confinement effect.
 
- Simple linear model such as the one of \citet{andreotti2009} also suggests that larger wind disturbances occur under strong flow confinement such as described above.
+ Simple linear models such as the one of \citet{andreotti2009} also suggests that larger wind disturbances occur under strong flow confinement such as described above.
 However, they are unable to reproduce the magnitude of the observed deviations, probably due to the presence of hydrodynamical non-linear effects, negligible in low confinement situations, but not otherwise (see Fig.~\ref{Fig12_supp} and Appendix 1).
-Another limit in the comparison between theoretical predictions and measured is induces by the the single-point measurements.
-To have reliable representations of the flow structures related to wind disturbances, additional measurements in different places on and near the same topographical obstacle are needed.
+Additionally, note that the spatial flow structures associated with the dune feedback on the wind can not be studied with our single-point measurements.
+Measurements in different places on and near the same topographical obstacle are then needed for further comparisons with models.
 
- This study highlights the interaction between giant dunes and the atmospheric boundary layer, thus supporting for example the way the capping layer acts as a bounding surface limiting dune growth \citep{andreotti2009, gunn2021}.
-This interaction also have implications at smaller scales, where bedforms then develop from the disturbed wind instead of the regional one.
-Differences between larger and smaller scale (thus older and more recent) dune patterns are observed ubiquitously, and have sometimes in the literature been attributed to climatic changes in wind regimes \citep{refs}.
-Here, we suggest using this feedback mechanism that current winds can explain dune patterns at all scales, such as the linear dunes ($\sim 50$ m -wide) elongating within the interdune between two giant linear dunes ($\sim 2$ km -wide) in the Namib Sand Sea (see Fig.~\ref{Fig6}).
+ This study highlights the interaction between giant dunes and the atmospheric boundary layer, supporting as well the way the capping layer acts as a bounding surface limiting dune growth \citep{andreotti2009, gunn2021}.
+This interaction also have implications at smaller scales, where bedforms then develop from the wind disturbed by the giant dunes, instead of the regional wind regime.
+Differences between larger and smaller scale dune patterns are observed ubiquitously.
+As larger dunes are often older than smaller ones, sometimes this discrepancy have been previously attributed to climatic changes in wind regimes \citep{refs}.
+Here, using this feedback mechanism, we suggest that current winds are also able to explain dune patterns at all scales.
+For example, this seems to be the case for the linear dunes ($\sim 50$ m -wide) elongating within the interdune between two giant linear dunes ($\sim 2$ km -wide) in the Namib Sand Sea (see Fig.~\ref{Fig6}).
 

### 2021-12-15 Paper/main.tex
Rewriting Introduction

@@
 
-Whenever a flow encounters an obstacle, various types of interactions can arise depending the different time and lengths scales involved.
-In the case of atmospheric flows, this depends mainly on the part of the vertical structure of the atmosphere, schematically composed of a turbulent boundary layer topped by a turbulence-free part, with which the obstacle interacts \citep{Stull1988}.
+Whenever a flow encounters an obstacle, different interactions can arise depending on the different time and length scales involved.
+In the case of atmospheric flows, this mainly depends on the part of the atmosphere, schematically composed of a turbulent boundary layer topped by a part free of turbulence, with which the obstacle interacts \citep{Stull1988}.
 %
-At the largest scale, the feedback of mountains on the stratified flow of the free atmosphere results in wave generation as well as significant wind disturbances, such as foehn winds in the lee side \citep{refs}.
+At the largest scale, the feedback of mountains on the stratified flow of the free atmosphere results in wave generation as well as significant wind disturbances, such as downslope winds in the lee side \citep{Duran1990}.
 Inside the boundary layer, the interaction between a turbulent flow and hilly surfaces is for example key to the understanding ocean surface wind-driven waves, or eolian bedforms in desert \citep{Belcher1998, Sullivan2010, Courrech2015}.
@@
 Later on, when the dune reaches an intermediate size, it may also induce significant wind deflections.
-This can impact the sediment pathways of coastal systems \citep{Hesp2015}, or affect the collective behavior of dune populations with long-range interactions due to flow disturbances induced by each individual \citep{Smith2017, Bacik2020}.
-As the dunes increase in size by collisions and coarsening, they sometimes reach a giant size, comparable to the boundary layer depth, thus inducing interactions not only with the turbulent flow of the ABL, but also with the free atmosphere~\citep{andreotti2009}.
+For exemple, this can affect the sediment pathways of coastal systems \citep{Hesp2015}, or the collective behavior of dune populations, through long-range interactions due to flow disturbances induced by each individual \citep{Smith2017, Bacik2020}.
+As the dunes increase in size by collisions and coarsening, they sometimes reach a giant size comparable to the boundary layer depth~\citep{andreotti2009}.
 However, the wind disturbances induced by these giant dunes have never been quantified.
 
-The study of the impact of obstacles on the atmospheric flows allow its incorporation within meteorological numerical model.
-Therefore, they mainly become limited by the precision of the included topographical data, as well as the spatial grid of the model.
-For example, the latest climate reanalysis, ERA5-Land, is limited by its $9~\textup{km}$ spatial resolution, while including the data $30$-m Digital Elevation Models (DEMs) of the shuttle radar topography mission \citep{Farr2007, munoz2021}.
+The study of the feedback of obstacles on atmospheric flows allows its incorporation into numerical meteorological models.
+The latter then become mainly limited by the accuracy of the included topographical data, as well as by the spatial grid of the model.
+For example, the ERA5-Land climate reanalysis is limited by its $9~\textup{km}$ spatial resolution, while including the data $30$-m Digital Elevation Models (DEMs) of the shuttle radar topography mission \citep{Farr2007, munoz2021}.
 As such, it can not reproduce the flow disturbances induced by giant dunes, which have a typical length scale $\sim 1~\textup{km}$.
 
-Here, we compare the wind predictions from the ERA5-Land dataset to local measurements in four different places across the Namib desert.
-In places with no significant topographies smaller than the model grid, we show that both wind datasets agree with each other.
-On the contrary, in places with giant dunes, we show that they may differ for some specific meteorological conditions, that we link to the circadian cycle of the ABL.
-We thus highlight the importance of the mid-scale topographies for local wind regimes, and its implications in the case of sand seas for smaller-scale eolian bedforms.
+Here, we compare the wind speed and direction of the ERA5-Land dataset to local measurements in four different places across the Namib desert.
+Where no significant topographies smaller than the model grid are present, we show that the two wind datasets agree with each other.
+On the contrary, in places with giant dunes, we show that they may differ for some specific meteorological conditions, that we link to the circadian cycle of the atmospheric boundary layer.
+In doing so, we highlight the importance of medium-scale topographies for local wind regimes and, in the case of sand seas, draw implications for smaller-scale eolian bedforms.
 

### 2021-12-21 Paper/main.tex
Modifying all figures -- introducing day and night + color scheme across figures

@@
     \caption{Temporal comparison between the wind data coming from the Era5Land climate reanalysis (orange lines) and from the local measurements (blue lines).
-\textbf{a--b}: Huab station.
+Color swathes indicate day (between 1000 UTC and 2200 UTC) and night (before 1000 UTC or after 2200 UTC) \textbf{a--b}: Huab station.
 \textbf{c--d}: Deep Sea station in winter.
@@
 In each subplot, both distributions are plotted from the same time steps, selected with constraints on the wind direction (columns) and/or wind velocity (rows) of the Era5Land dataset.
-The gray vertical dashed lines indicate the dune orientation, and the top right numbers the percentage of the total number of time steps selected in each subplot.
+The gray vertical dashed lines indicate the dune orientation.
+The numbers at the top right give the percentage of time steps selected in each sub-range, as well as the percentage of them corresponding to the day (between 1000 UTC and 2200 UTC).
 The purple frame highlights the regime (small wind velocities, nocturnal summer wind) in which the data from both datasets differs.
@@
 \textbf{c--f}: Sketches representing the interaction between the giant dunes and the atmospheric flow for different meteorological conditions.
-\textbf{g}: Streamlines qualitatively representing the effect of flow confinement, in the case of the Deep Sea station.
-The red lines are calculated from the unconfined case similar to \textbf{d}.
-The sinuous blue line represents a confined case similar to \textbf{f}.
+\textbf{g}: Streamlines qualitatively representing the effect of low, medium and high flow confinement.
 For details on the streamline derivation, see Appendix~\ref{turbulent_wind_model}.}
-    %
-    %
-    % calculated from equation~\eqref{Tau}.
-Note that topography and wind geometry represent the case of the Deep Sea station.
-For the red and straight blue lines, $(\mathcal{A}_{0}, \mathcal{B}_{0}) = (3.8, 1.3)$, calculated from the unbounded case of \citep{Fourriere2010}, and representing the situations \textbf{b--c}.
-For the sinuous blue line, $(\mathcal{A}_{0}, \mathcal{B}_{0}) = (8, 2)$, representing the confined case of \textbf{d}.}
     \label{Fig4}
@@
 In each subplot, both distributions are plotted from the same time steps, selected with constraints on the wind direction (columns) and/or wind velocity (rows) in the Era5Land dataset.
-The grey dashed vertical lines indicate the dune orientation.
+The gray vertical dashed lines indicate the dune orientation.
+The numbers at the top center give the percentage of time steps selected in each sub-range, as well as the percentage of them corresponding to the day (between 1000 UTC and 2200 UTC).
 The purple frame highlights the regime (small wind velocities, nocturnal summer wind) during which the wind data from both datasets differs.

### 2021-12-21 Paper_figure/Figure02.py

@@ -58,18 +58,18 @@ for i, (subfig, yr, mth, dy, station) in enumerate(zip(subfigs, years, months, d
     tmin = datetime(yr, mth, dy[0])
     tmax = datetime(yr, mth, dy[1])
     for j, (ax, var, label) in enumerate(zip(axarr, variables, labels[i])):
-        l1, = ax.plot(Data[station]['time'], Data[station][var + '_station'], label='measurements')
-        l2, = ax.plot(Data[station]['time'], Data[station][var + '_era'], label='Era5Land')
+        l1, = ax.plot(Data[station]['time'], Data[station][var + '_station'], label='measurements', color=theme.color_insitu)
+        l2, = ax.plot(Data[station]['time'], Data[station][var + '_era'], label='Era5Land', color=theme.color_Era5Land)
         ax.set_xlim(tmin, tmax)
         tick_formatter(ax)
         #
         # #### plot nights
         tstart = tmin - timedelta(days=1)
-        tstart = tstart.replace(hour=6)
+        tstart = tstart.replace(hour=10)
         x_night = [tstart + timedelta(days=i) for i in range((tmax-tmin).days + 2)]
-        for dawn in x_night:
-            a1 = ax.axvspan(dawn, dawn + timedelta(hours=12), facecolor='sandybrown', alpha=0.1, edgecolor=None, label=r'\faSun')
-            a2 = ax.axvspan(dawn - timedelta(hours=12), dawn, facecolor='lightgrey', alpha=0.25, edgecolor=None, label=r'\faMoon')
+        for daylight in x_night:
+            a1 = ax.axvspan(daylight, daylight + timedelta(hours=12), facecolor=theme.color_day, alpha=0.1, edgecolor=None, label=theme.Icon_day)
+            a2 = ax.axvspan(daylight - timedelta(hours=12), daylight, facecolor=theme.color_night, alpha=0.1, edgecolor=None, label=theme.Icon_night)
         #
         ax.set_ylabel(label_var[var])
         ax.set_xlabel('Days in {} {:d}'.format(month_calendar[tmin.month], tmin.year))
@@ -80,6 +80,8 @@ for i, (subfig, yr, mth, dy, station) in enumerate(zip(subfigs, years, months, d
         else:
             ax.set_ylim((0, 360))
             ax.set_yticks((0, 90, 180, 270, 360))
+#
+# a1.set_edgecolor((0, 0, 0, 1))
[…]

### 2021-12-21 python_codes/theme.py

@@ -51,3 +51,11 @@ cmap_topo = cmocean.cm.turbid_r
 # #### colors
 color_dune_orientation = 'grey'
 flux_color = matplotlib.colors.ListedColormap('navajowhite')
+color_day = '#c3b632'
+color_night = '#3d134f'
+color_Era5Land = 'tab:orange'
+color_insitu = 'tab:blue'
+
+# #### special Icons
+Icon_day = r'\faSun'
+Icon_night = r'\faMoon'

### 2021-12-31 Paper/main.tex
Rewieving all paper + appendices

@@
 
-\title{Significant wind disturbances induced by giant dunes.}
+\title{Local wind regime induced by the presence of giant linear dunes.}
 
@@
 \begin{abstract}
-  abstract
+
   \keywords{Boundary layer \and Turbulent flow \and Sand dunes \and Fluide-structures interactions}
@@
 
-Whenever a flow encounters an obstacle, different interactions can arise depending on the different time and length scales involved.
-In the case of atmospheric flows, this mainly depends on the part of the atmosphere, schematically composed of a turbulent boundary layer topped by a part free of turbulence, with which the obstacle interacts \citep{Stull1988}.
+As a flow encounters an obstacle, different interactions can arise depending on the different time and length scales involved.
+In the case of atmospheric flows, this mainly depends on the part of the atmosphere, schematically composed of a turbulent boundary layer topped by a stably stratified one, with which the obstacle interacts \citep{Stull1988}.
 %
-At the largest scale, the feedback of mountains on the stratified flow of the free atmosphere results in wave generation as well as significant wind disturbances, such as downslope winds in the lee side \citep{Duran1990}.
-Inside the boundary layer, the interaction between a turbulent flow and hilly surfaces is for example key to the understanding ocean surface wind-driven waves, or eolian bedforms in desert \citep{Belcher1998, Sullivan2010, Courrech2015}.
+At the largest scale, the feedback of mountains on the stratified part results in wave generation as well as significant wind disturbances, such as downslope winds in the lee side \citep{Duran1990}.
+Inside the boundary layer, the interaction between a turbulent flow and hilly surfaces is for example key to the understanding ocean surface wind-driven waves~\citep{Sullivan2010}, or eolian bedforms in desert \citep{Courrech2015}.
 
-Indeed, eolian sand dunes typically emerge from the feedback of the topography on the turbulent flow, which speeds up close to the dune crest \citep{Rubin1987, Charru2013, Courrech2014}.
-Later on, when the dune reaches an intermediate size, it may also induce significant wind deflections.
-For exemple, this can affect the sediment pathways of coastal systems \citep{Hesp2015}, or the collective behavior of dune populations, through long-range interactions due to flow disturbances induced by each individual \citep{Smith2017, Bacik2020}.
-As the dunes increase in size by collisions and coarsening, they sometimes reach a giant size comparable to the boundary layer depth~\citep{andreotti2009}.
-However, the wind disturbances induced by these giant dunes have never been quantified.
+Looking at the flow close to the surface, two different components of the topography feedback can be isolated.
+First, the flow accelerates on the upwind slope, and slows down on the downwind one, with a maximum velocity slightly upwind of the crest~\citep{Jackson1975, Sykes1980, Hunt1988}.
+Several measurements are available, performed in subaqueous flumes~\citep{Zilker1977, Zilker1979, Frederick1988} and in eolian field conditions~\citep{Claudin2013, Lu2021}.
+Second, deflection occurs when the incident flow direction is not perpendicular to the ridge crest.
+While predicted to be small (less than $10^{\circ}$) in the linear regime~\citep{Gadal2019}, significant flow steering has been reported in the field on the downwind side of larger aspect ratios obstacles, such as mountain ranges~\citep{Kim2000, Lewis2008, Fernando2019} or sand dunes~\citep{Walker2009, Walker2013, Hesp2015, Smith2017}.
 
-The study of the feedback of obstacles on atmospheric flows allows its incorporation into numerical meteorological models.
-The latter then become mainly limited by the accuracy of the included topographical data, as well as by the spatial grid of the model.
-For example, the ERA5-Land climate reanalysis is limited by its $9~\textup{km}$ spatial resolution, while including the data $30$-m Digital Elevation Models (DEMs) of the shuttle radar topography mission \citep{Farr2007, munoz2021}.
-As such, it can not reproduce the flow disturbances induced by giant dunes, which have a typical length scale $\sim 1~\textup{km}$.
+In the case of aeolian bedforms, the measurements mentioned above were made on dunes a few meters high only.
+The resulting wind disturbances have then been shown to significantly impact the sediment pathways of the system \citep{Hesp2015}, but also to affect the collective behaviour of dune populations trough long-range interactions \citep{Bacik2020}.
+%
+However, eolian dunes coexists in sand seas at different scales, each representative of their different evolution stages, from emergence at twenty meters wavelengths to giant size, corresponding to kilometric wavelengths and heights of several dozen meters \citep{McKee1979}.
+Quantifying these two effects for giant dunes is then all the more important for several reasons.
+%
+First, larger wind disturbances are expected from larger obstacles.
+%
+Second, the flow--structure coupling is not straightforward due to vertical structure of the atmosphere with which the giant dunes interact due to their sizes \citep{andreotti2009}.
+%
+Finally, the morphodynamics of aeolian bedforms are strongly dependent on the wind regime~\citep{Livingstone1996}.
+Thus, small-scale bedforms are sensitive to wind disturbances induced by larger-scale bedforms.
+This has been demonstrated for impact ripples on dunes \citep{Howard1977,Hood2021}, but has never been studied for medium-scale dunes near giant dunes.
 
-Here, we compare the wind speed and direction of the ERA5-Land dataset to local measurements in four different places across the Namib desert.
-Where no significant topographies smaller than the model grid are present, we show that the two wind datasets agree with each other.
-On the contrary, in places with giant dunes, we show that they may differ for some specific meteorological conditions, that we link to the circadian cycle of the atmospheric boundary layer.
-In doing so, we highlight the importance of medium-scale topographies for local wind regimes and, in the case of sand seas, draw implications for smaller-scale eolian bedforms.
+In the literature, arid areas have been studied at the desert scale from climate reanalyses based on global atmospheric models \citep{Blumberg1996, Livingstone2010, Ashkenazy2012, Jolivet2021, Hu2021}, such as ERA-40, ERA-Interim or ERA-5 \citep{Uppala2005, Dee2011, Hersbach2020}.
+However, the spatial resolution (tens of kilometers) of these reanalyses implies average quantities that do not resolve the smaller scales, ranging from the study of individual dunes to the border of mountains \citep{Livingstone2010}.
+Lately, the release of ERA5-Land allow to push back this limitation by providing up to 70 years of hourly wind predictions at a 9~km spatial resolution \citep{munoz2021}.
+However, due to its recent nature, the applicability limitations of this dataset remain to be studied.
+
+Here, we compare local wind speeds and directions measured in four different places across the Namib desert to the regional predictions of the ERA5-Land climate reanalysis.
+When the measurement stations are surrounded by a relatively flat environment, we show that local measurements and regional predictions agree with each other.
+On the contrary, in the sand sea, we show that they differ due to the presence of giant dunes.
+Furthermore, we link the magnitude of these discrepancies to the circadian cycle of the atmospheric boundary layer.
+Finally, we draw implications for smaller-scale eolian bedforms.
 
@@
 
-In this study, we focus on four places across and nearby the Namib desert, highlighting different environments (see Fig.~\ref{Fig1}).
+  We focus on four places across and nearby the Namib desert, highlighting different environments (see Fig.~\ref{Fig1}).
   %
   The Adamax station is located near the Adamax salt pan, in a highly vegetated area.
-The Huab station, located on the coast at the outlet of the Huab river is an arid environment exhibiting $60$-m scale barchan dunes.
-While these two stations are in environments with no mid-scale topography, this is not the case for the Deep Sea and South Namib stations.
-Both are located in the interdune between giant linear dunes with kilometric wavelengths and superimposed patterns.
-In this section, we describe and compare the wind regimes resulting from the available datasets in each station.
+The Huab station, on the coast at the outlet of the Huab river, is in an arid environment exhibiting $60$-m scale barchan dunes.
+While the surroundings of these two stations are relatively flat, this is not the case for the Deep Sea and South Namib stations.
+Both are in the interdune between tens of meters high giant linear dunes with kilometric wavelengths and superimposed patterns.
+In this section, we describe and compare winds from local measurements and climate reanalysis predictions.
 
@@
 Note that the bars show the direction towards which the wind blows.
-The red dots show the location of local wind stations.}
+The black dots show the location of local wind stations.}
     \label{Fig1}
@@
 
-  Two wind datasets are used in this study.
-First, local winds are provided by stations situated in the four different places (see Fig.~\ref{Fig1}).
-The wind strength and direction are measured every 10 minutes by cup anemometers and wind vanes, at heights between $2$~m and $3$~m depending on the station.
+  Local winds are provided by measurement stations located in the four different places (see black dots in Fig.~\ref{Fig1}).
+The wind strength and direction are sampled every 10 minutes by cup anemometers and wind vanes, at heights between $2$~m and $3$~m depending on the station.
 The available period of measurements ranges from 1 to 5 discontinuous years distributed between 2012 and 2020 (see Fig.~\ref{Fig1_supp}).
@@
   %
-  Then, regional winds are extracted at the same locations and periods from the ERA5-Land dataset, which is a replay at a smaller spatial resolution of ERA5, the latest climate reanalysis from the ECMWWF \citep{Hersbach2020, munoz2021}.
-It provides hourly estimates of the 10-m wind velocity and direction at a spatial resolution of $\sim 9~\textup{km}$ ($0.1^\circ\times0.1^\circ$).
+  Regional winds are extracted at the same locations and periods from the ERA5-Land dataset, which is a replay at a smaller spatial resolution of ERA5, the latest climate reanalysis from the ECMWWF \citep{Hersbach2020, munoz2021}.
+It provides hourly predictions of the 10-m wind velocity and direction at a spatial resolution of $\sim 9~\textup{km}$ ($0.1^\circ\times0.1^\circ$).
 
   For comparison, the local measurements are averaged into 1-hr bins centered on the temporal scale of the ERA5-Land estimates (see Fig.~\ref{Fig2_supp}).
-As the wind velocities of both datasets are provided at different heights, we convert them into shear velocities (see SI section 1), characteristic of the whole turbulent wind profile within the atmospheric boundary layer, which are then used together with the wind direction for further analysis.
+As the wind velocities of both datasets are provided at different heights, we convert them into shear velocities (see SI section 1), characteristic of the turbulent wind profile, which are then used together with the wind direction for further analysis.
 The resulting wind data are shown on the wind roses of Fig.~\ref{Fig1}(b--e).
@@
 In the Namib, the regional wind patterns are essentially controlled by the see breeze, resulting in strong northward components (sometimes slightly deviated by the large scale topography) present in all regional wind roses \citep{lancaster1985}.
-These daily winds are dominant during the second-half of the year (Septembre-January).
+These daytime winds are dominant during the second-half of the year (Septembre-January).
 In winter, an additional easterly component can be recorded during the night, induced by the combination of katabatic winds forming on the mountains, and infrequent `berg' winds, which are responsible of the high wind velocities observed \citep{lancaster1984}.
@@
 
-  In the case of the Adamax and Huab stations, the regional wind roses qualitatively match those corresponding to the local in situ measurements.
-However, for the Deep Sea and South Namib stations, the local wind roses exhibit additional components aligned with the giant dune orientation visible on the satellite images (Fig.~\ref{Fig1}c--d).
+  In the case of the Adamax and Huab stations, the wind roses from the regional predictions qualitatively match those corresponding to the local measurements.
+However, for the Deep Sea and South Namib stations, the measured wind roses exhibit additional components aligned with the giant dune orientation visible on the satellite images (Fig.~\ref{Fig1}c--d).
   %
-  Indeed, the analysis of the wind speed and direction time series shows that the agreement between the local and regional datasets is always verified when no mid-scale topography are present (Fig.~\ref{Fig2}a--b) and Fig.~\ref{Fig5_supp}).
-In contrast, for the stations within the giant dune field, we observe that this agreement is limited to the Septembre--January time periods (Fig.~\ref{Fig2}c--d).
+  The time series of wind speed and direction show that this agreement in the case of Adamax and Huab stations is always verified (Fig.~\ref{Fig2}a--b) and Fig.~\ref{Fig5_supp}).
+In contrast, for the stations within the giant dune field, we observe that this agreement is limited to Septembre--January time periods (Fig.~\ref{Fig2}c--d).
 
@@
 
-  When giant dunes are present, in the February--August period, the local and regional winds match only during the morning, i.e when the southerly/southwesterly sea breeze dominates (see Fig.~\ref{Fig2}(e--f), Fig.~\ref{Fig3} and Fig.~\ref{Fig6_supp}).
-In the late afternoon and during the night, when the northwesterly `berg' and katabatic winds blow, the two datasets differ.
-In this case, the angular wind distribution of the local measurements exhibits two additional modes separated of $\simeq 180^\circ$, each corresponding to the giant dune alignement (purple frame in Fig.~\ref{Fig3} and Fig.~\ref{Fig6_supp}, as well as Fig.~\ref{Fig7_supp}).
+  In the February--August period, when giant dunes are present, the local and regional winds match during daytime only, i.e when the southerly/southwesterly sea breeze dominates (see Fig.~\ref{Fig2}(e--f), Fig.~\ref{Fig3} and Fig.~\ref{Fig6_supp}).
+In the late afternoon and during the night, when the northwesterly `berg' and katabatic winds blow, measurements and predictions differ.
+In this case, the angular wind distribution of the local measurements exhibits two additional modes separated of $\simeq 180^\circ$, each corresponding to the giant dune alignement (see the purple frame in Fig.~\ref{Fig3} and Fig.~\ref{Fig6_supp}, as well as Fig.~\ref{Fig7_supp}).
 This deviation is also associated with a global attenuation of the wind strength (Fig.~\ref{Fig8_supp}).
@@
 
-  In the case of linear ridges, dune-induced flow disturbances have mainly been related to the incident wind direction~\citep{Walker2009, Hesp2015}.
-In our case, it is unlikely to be the dominant parameter, as the most deflected wind for both stations is the most perpendicular, where it should be winds with incident directions between $30^{\circ}$ and $70^{\circ}$~\citep{Hesp2015}.
-An important observation is the difference in behavior between low and high wind velocities, which suggests a change in the hydrodynamical regime.
+  For linear ridges, dune-induced flow disturbances have mainly been related to the angle between wind direction and crest orientation, with a maximum for angles between $30^{\circ}$ and $70^{\circ}$~\citep{Walker2009, Hesp2015}.
+In our case, the most deflected wind for both stations is the most perpendicular, such that the incident wind direction does not seem to be the dominant parameter controlling the wind deflection.
+In contrast, a different behavior is observed between  low and high wind velocities, suggesting a change in hydrodynamical regime.
 
-  Previous studies have linked atmospheric flow around and over topographical obstacles to the vertical structure of the atmosphere \citep{Stull1988}.
-More particularly, dunes evolves in its lower part, the turbulent atmospheric boundary layer (ABL), typically characterized by a logarithmic wind profile and a vertically constant potential temperature.
-Above, the free atmosphere (FA) is a stably stratified zone in which turbulence is negligible, and where the flow is usually considered as incompressible and inviscid.
-In the middle, a transitional layer, also known as entrainment zone, is characterized by a sharp increase of the potential temperature, which traps the turbulence resulting from the surface friction below it.
-
-  In the following, we sum-up the dominant numbers leading to different hydrodynamical interactions with topographical obstacles, and interpret the data with respect to the corresponding physical mechanisms.
+  In the following, we discuss the relevant parameters leading to different hydrodynamical interactions with topographical obstacles, and interpret the data with respect to the corresponding physical mechanisms.
 
@@
     \includegraphics[scale=1]{Figures/Figure4.pdf}
-    \caption{\textbf{a--b}: Vertical profiles of the virtual potential temperature at 2 different time steps (blue - 29/11/2012 - 1100 UTC, red - 21/03/2017 - 1200 UTC) at the Deep Sea station.
+    \caption{\textbf{a--b}: Vertical profiles of the virtual potential temperature at 2 different time steps (day - 31/03/2017 - 1200 UTC, night - 21/03/2017 - 2200 UTC) at the Deep Sea station.
 Dots: data from the ERA5 reanalysis.
@@
 
-  \subsection{Relevant non-dimensional parameters and physical modeling}
+  \subsection{Relevant non-dimensional parameters and physical considerations}
   \label{theoretical_framework}
 
-  Flow deflection over ridges can be simplistically understood from a balance between inertia and pressure gradients~\citep{Hesp2015}.
+  Flow deflection over ridges can be understood from the Bernoulli principle~\citep{Hesp2015}.
 As the flow approaches the ridge crest, the compression of the streamlines results in larger flow velocities, and thus lower pressures~\citep{Rubin1987}.
@@
   Another way to increase the flow deflection is its confinement below a capping surface, that results in further streamline compression.
-This happens when the flow disturbance induced by the obstacle reaches the surface.
-As obstacles typically disturb flow over a characteristic height similar to their width, the potential of interaction between the dunes and the overlying atmospheric structure is well captured by the parameter $k H$, where $k = 2\pi/\lambda$ is the wavenumber and $H$ the ABL depth.
-Here, the giant dunes have kilometric wavelengths, such that $0.02 \lesssim k H \lesssim 5$, and they interact most of the time with the capping layer and the stratified free atmosphere above \citep{andreotti2009}.
+This is the case for bedforms forming in open channel flows such as rivers~\citep{Fourriere2010, Unsworth2018}, but also for eolian dunes.
+These dunes evolve in the turbulent atmospheric boundary layer (ABL), which is capped by a transitional layer separating it from the stratified atmosphere above (see Fig.~\ref{Fig4}).
+Two different mechanisms control the possibility of this additional streamline compression.
 
-  Note that the ability of the capping layer and stratification to accommodate a perturbation induced by the topography directly impacts the strength of this confinement effect (Fig.~\ref{Fig4}).
+  On one hand, it depends if the flow disturbance induced by the underlying topography reach the surface.
+As obstacles typically disturb flow over a characteristic height similar to their length, the potential of interaction between the dunes and the overlying surface is well captured by the parameter $k H$, where $k = 2\pi/\lambda$ is the wavenumber and $H$ the ABL depth.
+Note that $H$ is directly related to the radiative fluxes at the Earth surface, and thus varies with the circadian and seasonal cycles.
+Here, the giant dunes have kilometric wavelengths, such that $0.02 \lesssim k H \lesssim 5$, and they interact most of the time with the capping layer and the stratified free atmosphere (FA) above \citep{andreotti2009}.
+Interestingly, the limit of no-interactions between the topography and the boundary layer structure ($k H \gg 1$), in which the properties of the overlying atmospheric structure are irrelevant, is never reached here, in the case of giant dunes.
+
+  On the other hand, it depends on rigidity of the capping surface, as its deformation releases the confinement effect inducing streamline compression.
+This is typically quantified using the Froude number~\citep{Vosper2004, Stull2006, Sheridan2006, Hunt2006, Jiang2014}:
+  \begin{equation}
+        \mathcal{F} = \displaystyle\frac{U}{\sqrt{\displaystyle\frac{\Delta\rho}{\rho}gH}},
+  \end{equation}
+  where $U$ is the wind velocity at the top of the ABL, $\rho$ its average density, $\Delta\rho$ the density jump between the ABL and the FA.
+
+  % Note that the ability of the capping layer and stratification to accommodate a perturbation induced by the topography directly impacts the strength of this confinement effect (Fig.~\ref{Fig4}).
 This is typically quantified using surface and internal Froude numbers
-  \citep{Vosper2004, Stull2006, Sheridan2006, Hunt2006, Jiang2014}:
-  \begin{equation}
-        \mathcal{F}r_{\textup{S}} = \displaystyle\frac{U}{\sqrt{\displaystyle\frac{\Delta\rho}{\rho}gH}}, \, \mathcal{F}r_{\textup{I}} = \displaystyle\frac{k U}{N},
-  \end{equation}
-  where $U$ is the wind velocity at the top of the ABL, $\rho$ its average density, $\Delta\rho$ the density jump between the ABL and the FA and $N$ is the Brunt-Väisälä frequency, characteristic of the stratification.
+  % \citep{Vosper2004, Stull2006, Sheridan2006, Hunt2006, Jiang2014}:
+  % \begin{equation}
+  %       \mathcal{F} = \displaystyle\frac{U}{\sqrt{\displaystyle\frac{\Delta\rho}{\rho}gH}}, \, \mathcal{F}_{\textup{I}} = \displaystyle\frac{k U}{N},
+  % \end{equation}
+  % where $U$ is the wind velocity at the top of the ABL, $\rho$ its average density, $\Delta\rho$ the density jump between the ABL and the FA and $N$ is the Brunt-Väisälä frequency, characteristic of the stratification.
 
@@
 On the contrary, large wind disturbances are expected to occur during the night, when the confinement is mainly induced by shallow ABL (Fig.~\ref{Fig4}e--f).
-Note that this strong confinement can be somewhat reduced in the case of strong winds (corresponding to large Froude numbers, see Fig.~\ref{Fig4}f), explaining the threshold in velocity observed in the data (see section~\ref{section_data_feedback}).
+Note that this strong confinement can be somewhat reduced in the case of strong winds (corresponding to large Froude numbers, see Fig.~\ref{Fig4}f), explaining the transition from deflected to non-deflected winds related to low and high velocities observed in the data (see section~\ref{section_data_feedback}).
 
@@
 
-  To highlight these different regimes from our data, we compute wind disturbance diagrams in the space defined by the three relevant non-dimensional numbers presented above, $\left(k H,\, \mathcal{F}r_{\textup{S}}, \, \mathcal{F}r_{\textup{I}}\right)$.
+  To highlight these different regimes from our data, we compute wind disturbance diagrams in the space defined by the two relevant non-dimensional numbers presented above, $\left(k H,\, \mathcal{F}\right)$.
 Those are calculated from the time series of the geopotential, temperature and specific humidity vertical profiles available in the ERA5 climate reanalysis (see SI section 2).
   %
-  Flow deviation is computed as the minimal angle between the wind orientations from the two datasets:
-  \begin{equation}
-    \delta_{\theta} = \left\vert \textup{min}\left(\left[\theta_{\textup{ERA}} - \theta_{\textup{station}}\right] \, \textup{mod.} \, 360, \left[\theta_{\textup{station}} - \theta_{\textup{ERA}}\right] \, \textup{mod.} \, 360\right) \right\vert.
-  \end{equation}
-  %
-  The relative velocity modulation is computed as
+  Flow deviation is computed as the minimal angle between the wind orientations from the local measurements, and the regional predictions. The relative velocity modulation is computed as
   %
@@
 
-  As both Froude numbers have qualitatively the same impact on the flow confinement, we first leave aside the internal Froude number, and focus on the space $(k H,\, \mathcal{F}r_{\textup{S}})$.
   When representing the two variables $\delta_{\theta}$ and $\delta_{\textup{u}}$ in this space, different regime emerges (Fig.~\ref{Fig5}).
-Small wind disturbances ($\delta_{\theta} \to 0$, $\delta_{\textup{u}} \to 0$) are located in the top-right part of the diagrams, corresponding to a regime mixing low-interaction and low-confinement ($k H$ and $\mathcal{F}r_{\textup{S}}$ large enough, Fig.~\ref{Fig4}d).
+Small wind disturbances ($\delta_{\theta} \to 0$, $\delta_{\textup{u}} \to 0$) are located in the top-right part of the diagrams, corresponding to a regime mixing low-interaction and low-confinement ($k H$ and $\mathcal{F}$ large enough, Fig.~\ref{Fig4}d).
   %
-  Lower values of $k H$ (stronger interaction) or Froude numbers (stronger confinement) then both leads to an increase in wind disturbances, both in terms of orientation and velocity.
-Below a threshold value of $k H \simeq 0.3$, wind disturbance occurs independently of the Froude numbers value.
-Furthermore, this also seems to control a transition between from damped to amplified wind velocities within the interdune (Fig.~\ref{Fig5}b), for which we do not have an explanation.
-Note that the same interpretation can be done with the diagrams including the internal Froude number $\mathcal{F}r_{\textup{I}}$, as shown by Fig.~\ref{Fig12_supp}.
+  Lower values of $k H$ (stronger interaction) or Froude number (stronger confinement) then both lead to an increase in wind disturbances, both in terms of orientation and velocity.
+Below a threshold value of $k H \simeq 0.3$, wind disturbance occurs independently of the Froude number value, probably due to enhanced non-linear effects linked to strong flow modulation by the obstacle in this part of the diagram.
+The Froude number also controls a transition from damped to amplified wind velocities in the interdune, with a transition at $\mathcal{F} \simeq 0.4$ (Fig.~\ref{Fig5}b).
+This may be linked to a transition in the flow regime in the lee side of the obstacle (lee waves, hydraulic jumps, rotors) but further measurements are needed in order to assess this \citep{baines1995, Vosper2004}.
 
-  Interestingly, the limit of no-interactions between the topography and the boundary layer structure ($k H \gg 1$), in which the properties of the capping layer and the stratification become irrelevant, is never reached here, in the case of giant dunes.
+  %
+  % Furthermore, this also seems to control a transition between from damped to amplified wind velocities within the interdune (Fig.~\ref{Fig5}b), for which we do not have an explanation.
+Note that the same interpretation can be done with the diagrams including the internal Froude number $\mathcal{F}_{\textup{I}}$, as shown by Fig.~\ref{Fig12_supp}.
 
@@
     \includegraphics[scale=1]{Figures/Figure5.pdf}
-    \caption{Regime diagrams of the wind deviation $\delta_{\theta}$ and relative attenuation/amplification $\delta_{u}$ in the space $(\mathcal{F}r_{\textup{S}}, \, kH)$, containing the data from both the Deep Sea and South Namib stations.
+    \caption{Regime diagrams of the wind deviation $\delta_{\theta}$ and relative attenuation/amplification $\delta_{u}$ in the space $(\mathcal{F}, \, kH)$, containing the data from both the Deep Sea and South Namib stations.
 Blue dashed lines empirically delimit the different regimes.
 The point density in each bin of the diagrams is shown in Fig.~\ref{Fig11_supp}.
-The regime diagrams in the spaces $(\mathcal{F}r_{\textup{I}}, \, kH)$ and $(\mathcal{F}r_{\textup{I}}, \, \mathcal{F}r_{\textup{S}})$ are shown in Fig.~\ref{Fig12_supp}.}
+The regime diagrams in the spaces $(\mathcal{F}_{\textup{I}}, \, kH)$ and $(\mathcal{F}_{\textup{I}}, \, \mathcal{F})$ are shown in Fig.~\ref{Fig12_supp}.}
     \label{Fig5}
@@
 
+  \subsection{On the influence of the stratification of the free atmosphere}
+
+  The presence of a stratification in the free atmosphere can also impact the flow confinement, depending on its ability to deform under the presence of an underlying obstacle.
+This can be quantified using the internal Froude number~\citep{Vosper2004, Stull2006, Sheridan2006, Hunt2006, Jiang2014}:
+  \begin{equation}
+    \mathcal{F}_{\textup{I}} = \displaystyle\frac{k U}{N},
+  \end{equation}
+  where $N = \sqrt{-(g/\rho)(\partial \rho/\partial z)}$ is characteristic of the stratification.
+Both Froude numbers have the same qualitative effect on flow confinement, as they quantify the rigidity of the overlying layers.
+This is confirmed by figure~\ref{Fig12_supp}, where we can also find the different regimes related to wind disturbances described previously for the Froude number $\mathcal{F}$.
+
 \section{Discussion}
-
- The comparison of local (direct measurements) and regional (climate reanalysis) wind data reveals the giant dunes feedback on the flow.
-In flat areas, the matching between both datasets confirms the ability of the ERA5Land climate reanalysis to predict the wind flow down to scales $\sim 10~\textrm{km}$, i.e the grid model.
-When smaller scale topographies are present (giant dunes in our case), locally measured wind regimes may significantly differ from the regional ones.
-Furthermore, we link these disturbances induced by the dunes to their interaction with the lower part of the atmospheric vertical structure, and more specifically to its circadian variability.
-During the night, the presence of a shallow atmospheric boundary layer (ABL) induces a strong flow confinement, associated with large wind deviation and acceleration or deceleration.
-During the day, the capping layer is high enough to prevent its interaction with the giant dunes, resulting in a low flow confinement, and thus smaller wind disturbances.
-Interestingly, we also found that this effect could be counterbalanced by the presence of large wind velocities, capable of deforming the capping layer and/or the FA stratification, thus decreasing the confinement effect.
-
- Simple linear models such as the one of \citet{andreotti2009} also suggests that larger wind disturbances occur under strong flow confinement such as described above.
-However, they are unable to reproduce the magnitude of the observed deviations, probably due to the presence of hydrodynamical non-linear effects, negligible in low confinement situations, but not otherwise (see Fig.~\ref{Fig12_supp} and Appendix 1).
-Additionally, note that the spatial flow structures associated with the dune feedback on the wind can not be studied with our single-point measurements.
-Measurements in different places on and near the same topographical obstacle are then needed for further comparisons with models.
-
- This study highlights the interaction between giant dunes and the atmospheric boundary layer, supporting as well the way the capping layer acts as a bounding surface limiting dune growth \citep{andreotti2009, gunn2021}.
-This interaction also have implications at smaller scales, where bedforms then develop from the wind disturbed by the giant dunes, instead of the regional wind regime.
-Differences between larger and smaller scale dune patterns are observed ubiquitously.
-As larger dunes are often older than smaller ones, sometimes this discrepancy have been previously attributed to climatic changes in wind regimes \citep{refs}.
-Here, using this feedback mechanism, we suggest that current winds are also able to explain dune patterns at all scales.
-For example, this seems to be the case for the linear dunes ($\sim 50$ m -wide) elongating within the interdune between two giant linear dunes ($\sim 2$ km -wide) in the Namib Sand Sea (see Fig.~\ref{Fig6}).
 
@@
     \caption{Implications for smaller scale patterns in (a) the South Namib and (b) Deep Sea.
-The ellipses indicates the different types of elongating dunes, at large (plain) and small scale (dashed).
+The ellipses indicates the diffe  rent types of elongating dunes, at large (plain) and small scale (dashed).
 The dune orientations are calculated using the model of \citet{Courrech2014} from the sand flux angular distributions, shown here for typical sand quartz grains of $180~\mu$m.
 The double blue and single red arrows correspond to the two possible dune growth mechanisms, bed instability and elongation, respectively.
-Likewise, plain arrows are calculated from the ERA5-Land datasets, and dashed arrows from the in situ measurements.
+Likewise, plain arrows are calculated from the ERA5-Land data, and dashed arrows from the local measurements.
 Wedges show the uncertainty on the orientation calculation, and the arrows correspond to typical parameters found in the literature, i.e a grain diameter of $180~\mu$m and a flux-up ratio of 1.6.
-The green dots indicate the position of the measurement stations.
+The black dots indicate the position of the measurement stations.
 See Appendix 2 for additional details.}
@@
 
+ The comparison of local (direct measurements) and regional (climate reanalysis) wind data reveals the giant dunes feedback on the flow.
+In flat areas, the matching between measurements and prediction confirms the ability of the ERA5Land climate reanalysis to predict the wind flow down to scales $\sim 10~\textrm{km}$, i.e the grid model.
+When smaller scale topographies are present (giant dunes in our case), locally measured wind regimes can significantly differ from the regionally predicted ones.
+Furthermore, we link these disturbances induced by the dunes to their interaction with the lower part of the atmospheric vertical structure, and more specifically to its circadian variability.
+During the night, the presence of a shallow atmospheric boundary layer (ABL) induces a strong confinement of the flow, associated with large wind deviation and acceleration or deceleration.
+During the day, the capping layer is high enough to prevent its interaction with the giant dunes, resulting in a low confinement of the flow, and thus smaller wind disturbances.
+Interestingly, we also found that this effect could be counterbalanced by the presence of large wind velocities, capable of deforming the capping layer and/or the FA stratification, thus decreasing the confinement effect.
+
+ Simple linear models also suggest that larger wind disturbances occur under strong flow confinement such as described above \citet{andreotti2009, Andreotti2012}.
+However, they are unable to reproduce the magnitude of the observed deviations, probably due to the presence of hydrodynamical non-linear effects, all the more present in high confinement situations linked to strong flow modulations (see Fig.~\ref{Fig12_supp} and Appendix 1).
+They also predict different spatial flow structures such as lee waves and rotors~\citep{baines1995, Vosper2004}, which are likely to be complicated by these non-linearities, and which cannot be observed by our single point measurements.
+Measurements in different places along and across the ridge are then needed in order to properly map these flow structures, and allow further comparisons with models.
+
+ This study highlights the interaction between giant dunes and the atmospheric boundary layer.
+It then supports the debated idea that the capping layer acts as a bounding surface limiting dune growth~\citep{andreotti2009}, as opposed to an unconstrained growth ever-slower with size \citep{Eastwood2011, gunn2021}.
+Once validated, this mechanism would then allow inference of the ABL depth from the giant bedforms spacing where measurements are not feasible or available, as performed by~\citet{Lorenz2010} on Titan.
+
+ This interaction also have strong implications for smaller scales bedforms, as illustrated in Fig.~\ref{Fig6}.
+In the Namib Sand Sea, small linear dunes ($\sim 50$ m -wide) are present in the interdune between giant linear dunes ($\sim 2$ km -wide).
+While differences between larger and smaller scale dune patterns are observed ubiquitously, they are now largely attributed to the presence of two different dune growth mechanisms, leading to two different dune patterns (orientations and/or morphologies) for the same wind regime~\citep{Courrech2014, Runyon2017, lu2017, Song2019, Hu2021}.
+Here, coupling sediment transport and dune growth models, we show that the orientations of the small and giant linear dunes can be predicted from the same dune growth mechanism, using the locally measured and regionally predicted winds, respectively (red arrows in Fig.~\ref{Fig6}).
+The giant dune feedback on the flow described in this study then provides a mechanism for the existence of these small linear dunes elongating across the interdune, as yet unresolved.
+While further studies are needed, these dune type could provide additional strong constraints for the inference of local winds from bedforms, as currently performed on Mars using ripple orientations \citep{Liu2015, Hood2021}.
+
 \begin{acknowledgements}
-These should follow the concluding section of the paper and precede the References and any appendices, if they are present. The acknowledgements section does not require a section number.
+[citing all grants ...]\\
+Finally, we would like to acknowledge the contributors of the following open-source python librairies, Matplotlib \citep{Hunter2007}, Numpy \citep{Harris2020} and Scipy \citep{Virtanen2020}, which provide an incredibly efficient ecosystem allowing scientific research in Python.
 \end{acknowledgements}
@@
 
-Following the work of \citet{Fourriere2010} and \citet{Andreotti2012}, we briefly expose in this section the linear response of a turbulent flow to a small aspect ratio perturbation of the topography $\xi$.
-As this topography can be decomposed into several sinusoidal modes, we focus on the response to a sinusoidal topography as:
+Following the work of \citet{Fourriere2010}, \citet{Andreotti2012} anc \citet{andreotti2009}, we briefly expose in this section the linear response of a turbulent flow to a small aspect ratio perturbation of the underlying topography.
+As this topography can be decomposed into several sinusoidal modes, we focus on a sinusoidal topography:
 \begin{equation}
@@
 \end{equation}
-which is also a good approximation to the giant dunes observed in the Deep Sea and South Namib Station (see Fig~\ref{Fig1} and Fig~\ref{Fig4_supp}).
+which is also a good approximation for the giant dunes observed in the Deep Sea and South Namib Station (see Fig~\ref{Fig1} and Fig~\ref{Fig4_supp}).
 Here, $x$ and $y$ are the streamwise and spanwise coordinates, $k=2\pi/\lambda$ the wavenumber of the sinusoidal perturbation, and $\alpha$ its crest orientation, calculated with respect to the $y$--direction.
 
-In terms of basal shear stress $\tau = \rho u_{*}^{2}$, the flow response can then generally be written in as:
+The two components of the basal shear stress $\boldsymbol{\tau} = \rho u_{*}\boldsymbol{u}_{*}$, constant in a flat bottom situation, can then be written without loss of generality as:
 \begin{align}
@@
 where $\tau_{0}$ is the basal shear stress on a flat bed, and $\phi_{x, y} = \tan^{-1}\left(\mathcal{B}_{x, y}/\mathcal{A}_{x, y}\right)$.
-The in--phase and in--quadrature hydrodynamical coefficients $\mathcal{A}_{x, y}$ and $\mathcal{B}_{x, y}$ are functions of the flow conditions, i.e the bottom roughness, the free surface or the incident flow direction \citep{Fourriere2010, andreotti2009, Andreotti2012, Charru2013}.
+The in--phase and in--quadrature hydrodynamical coefficients $\mathcal{A}_{x, y}$ and $\mathcal{B}_{x, y}$ are functions of the flow conditions, i.e the bottom roughness, the vertical flow structure or the incident flow direction \citep{Fourriere2010, andreotti2009, Andreotti2012}.
 
-\citet{Andreotti2012} have shown that the impact of the incident wind direction can be well approximated by the following expressions:
+Following \citet{Andreotti2012}, the impact of the incident wind direction can be approximated by the following expressions:
 \begin{align}
@@
 where $\mathcal{A}_{0}$ and $\mathcal{B}_{0}$ are now two coefficients independent of the dune orientation $\alpha$.
-In the case of a fully turbulent boundary layer capped by a free atmosphere capping, they now only depend on $k H$, $k z_{0}$, $\mathcal{F}r_{\textup{I}}$ and $\mathcal{F}r_{\textup{S}}$, as detailed by \citet{andreotti2009}.
-More specifically, their variation in the marginal spaces $(k H,\, \mathcal{F}r_{\textup{S}})$ and $(k H,\, \mathcal{F}r_{\textup{I}})$ are shown in Fig.~\ref{Fig12_supp}.
+In the case of a fully turbulent boundary layer capped by a stratified atmosphere, they now only depend on $k H$, $k z_{0}$, $\mathcal{F}$ and $\mathcal{F}_{\textup{I}}$ \citet{andreotti2009}.
+In this study, we assume a constant hydrodynamic roughness $z_{0} \sim 1~\textup{mm}$, leading to a constant value of $k z_{0} \sim 10^{-6}$.
+Measured values of $z_{0}$ in the field indeed reports a variation of $z_{0}$ between 0.1 mm and 10 mm \citep{Sherman2008, Field2018}, but $\mathcal{A}_{0}$ and $\mathcal{B}_{0}$ does not vary much in the corresponding range of $k z_{0}$ \citep{Fourriere2010}.
+%
+Note that the linearity assumption of this theoretical framework requires $\left(\vert \tau \vert - \tau_{0}\right)/\tau_{0} \ll 1$, which is satisfied by $k\xi\sqrt{\mathcal{A}_{0}^{2} + \mathcal{B}_{0}^{2}} \ll 1$.
+In our case, the giant dune morphology gives $k\xi \simeq 0.1$, setting the upper bound of the coefficient modulus $\sqrt{\mathcal{A}_{0}^{2} + \mathcal{B}_{0}^{2}}$ to 10.
 
-Typical values for the unconfined case are therefore $\mathcal{A}_{0} = 3.4$ and $\mathcal{B}_{0} = 1$.
-In our case of giant dunes with $k\xi_{0} \sim 0.1$, significant wind disturbances are then expected when $\sqrt{\mathcal{A}_{0}^{2} + \mathcal{B}_{0}^{2}} \sim 10$.
-However, this is also the limit of the linear regime where this theoretical model is applicable, as hydrodynamical non-linearities become significant when $k\xi_{0}\sqrt{\mathcal{A}_{0}^{2} + \mathcal{B}_{0}^{2}} \sim 1$.
+Additionlly, we also calculate the time series of the hydrodyamical coefficients from the time series of the non-dimensional numbers used in this study.
+The results, shown Fig.~\ref{Fig12_supp} under the similar form of the regime diagrams presented in Fig.~\ref{Fig5} and Fig.~\ref{Fig12_supp}, exhibit a qualitative matching with the observations presented in this study.
+%
+Small values of these coefficients ($\mathcal{A}_{0} \simeq 3.4$ and $\mathcal{B}_{0} \simeq 1$) are found in low confinement cases, i.e for $k H \gg 1$ (no interaction between the dunes and the capping layer) or for $k H \gg 1$ but large enough Froude numbers (reduced flow confinement due to the deformation of the olverlying capping layer and stratification).
+In contrast, larger values are obtained in high confinement cases (small $k H$ and Froude numbers).
+However, not that most of this part of the diagrams are outside of the linear limit discussed above ($\sqrt{\mathcal{A}_{0}^{2} + \mathcal{B}_{0}^{2}} \gtrsim 10$), which does not allow further quantitative comparison with the data.
 
@@
 
-Here, we briefly detail the sediment transport and dune morphodynamics theoretical framework leading to the prediction of sand fluxes and dune orientations from wind data.
+Here, we briefly describe the sediment transport and dune morphodynamics theoretical framework leading to the prediction of sand fluxes and dune orientations from wind data.
 
-The sediment fluxes can been directly linked to the wind basal shear stress at each time steps $t$ from transport laws, whose exact forms depends on the sediment transport mechanisms taken into account.
-In this work, we following the recent work of \citet{Pahtz2020}, where the sediment flux $q_{\textup{sat}}$ on a flat bed made of loose sand can be expressed as:
+\paragraph{Sediment transport}
+The prediction of sand fluxes from wind data has been a long standing issue in geomorphological studies~\citep{Fryberger79, Pearce2005, Sherman2012, Shen2019}.
+Based on laboratory studies in wind tunnels \citep{Rasmussen96, Iversen99, Creyssels2009, Ho2011}, as well as physical considerations \citep{Ungar1987, Andreotti2004bis, Duran2011, Pahtz2020}, it has been shown that the steady saturated sand flux over a flat sand bed depends linearly on the shear stress:
 \begin{equation}
-  \label{transport_law}
+  \label{transport_law_quadratic}
+  \frac{q_{\textup{sat, t}}}{Q} = \Omega\sqrt{\Theta_{\textup{th}}}\left(\Theta_{t} - \Theta_{\textup{th}}\right),
+\end{equation}
+where $\Omega$ is a proportionality constant, $Q = d\sqrt{(\rho_{\textup{s}} - \rho)gd/\rho}$ is a characteristic flux, $\Theta = \rho u_{*, t}^{2}/(\rho_{\textup{s}} - \rho)gd$ the Shields number, and $\Theta_{\textup{th}}$ its threshold value for incipient sediment transport.
+Here, $\rho_{\textup{s}} = 2.6~\textup{g}~\textup{cm}^{-3}$ and $d=180~\mu\textup{m}$ are the grain density and diameter, and $g$ is the gravitational acceleration.
+
+Recently, \citet{Pahtz2020} suggested a quadratic dependency on the shear stress by taking into account grain--grain interactions within the transport layer, performing better at reproducing laboratory data at high wind velocities:
+\begin{equation}
+  \label{transport_law_quartic}
   \frac{q_{\textup{sat, t}}}{Q} = \frac{2\sqrt{\Theta_{\textup{th}}}}{\kappa\mu}\left(\Theta_{t} - \Theta_{\textup{th}}\right)\left(1 +\frac{C_{\textup{M}}}{\mu}\left[\Theta_{t} - \Theta_{\textup{th}}\right]\right),
 \end{equation}
-where $\kappa = 0.4$ is the von Kármán constant, $C_{\rm M}=1.7$ a constant, $Q = d\sqrt{(\rho_{\textup{s}} - \rho)gd/\rho}$ is a characteristic flux, with $\rho_{\textup{s}} = 2.6~\textup{g}~\textup{cm}^{-3}$ and $d=180~\mu\textup{m}$ the grain density and diameter, and $g$ the gravitational acceleration.
-The friction coefficient $\mu$ is taken to be the avalanche slope of the granular material, i.e $\sim 0.6$.
-Finally, the Shields number is defined as $\Theta = \rho u_{*, t}^{2}/(\rho_{\textup{s}} - \rho)gd$, and its threshold value for incipient sediment transport as been calibrated using laboratory experiments to $\Theta_{\textup{th}} = 0.0035$.
+where $\kappa = 0.4$ is the von Kármán constant, $C_{\rm M}=1.7$ a constant and $\mu$ a friction coefficient, taken to be the avalanche slope of the granular material, i.e $\sim 0.6$.
+The fit of this law to the experimental data of \citet{Creyssels2009} and \citet{Ho2011} gives $\Theta_{\textup{th}} = 0.0035$.
+The sand flux angular distributions and the dune orientations in Fig.~\ref{Fig6} are calculated using this quartic law \eqref{transport_law_quartic}.
+However, we verified that using the quadratic law \eqref{transport_law_quadratic} instead did not change the predicted dune orientations by more than a few degrees.
 
-The dune orientations are then predicted from the dimensional model of \citet{Courrech2014}.
+\paragraph{Dune orientations}
+
+The dune orientations are predicted from the computed sand flux time series, using the dimensional model of \citet{Courrech2014}.
+Two orientations are possible depending on the mechanism dominating the dune growth: elongation or bed instability (the latter is also known as the rule of maximum gross bedform-normal transport from \citet{Rubin1987}).
+
 The orientation $\alpha$ corresponding the bed instability is then the one that maximizes the following growth rate:
 \begin{equation}
-  \sigma \propto \frac{1}{H W T}\int_{t}  q_{\textup{crest}, t}\vert \sin\left(\theta_{t} - \alpha\right) \vert,
+  \sigma \propto \frac{1}{H_{d} W_{d} T}\int_{t}  q_{\textup{crest}, t}\vert \sin\left(\theta_{t} - \alpha\right) \vert,
 \end{equation}
-where $H$ and $W$ are dimensional constants representing the dune height and width, respectively. The flux at the crest is expressed as:
+where $H_{d}$ and $W_{d}$ are dimensional constants representing the dune height and width, respectively. The flux at the crest is expressed as:
 \begin{equation}
@@
 
-The computed dune orientations, blue and red arrows in figure~\ref{Fig6}, are however depending on a large number of parameters, for which we took typical values for eolian desert on Earth.
+The resulting computed dune orientations, blue and red arrows in figure~\ref{Fig6}, are then depending on a certain number of parameters (grain properties, flux-up ratio), for which we took typical values for eolian desert on Earth.
+Due to the lack of measurements in the studied places, significant uncertainties can however be expected.
 We therefore run a sensibility test by calculating the dune orientations for grain diameters ranging from $100~\mu\textup{m}$ to $400~\mu\textup{m}$ and the speed-up ratio from $0.1$ to $10$ (wedges on figure~\ref{Fig6}).
-We also checked the sensibility the transport low by repeating the process with the quadratic transport also used for comparison in \citet{Pahtz2020}, which led to no more than $n\%$ of variation with respect to \eqref{transport_law}.
 
@@
 A shown by figure~\ref{Fig3_supp}, the minimum of $\delta$ in the space ($z_{0,\, \textup{Era}}$, $z_{0,\, \textup{in situ}}$) forms a line.
-We thus take the roughness of the Era5Land dataset as the typical value when sediment transport occurs, $10^{-3}$~m, corresponding to the thickness of the transport layer \citep{Sherman2007}.
+We thus take the roughness of the Era5Land dataset as the typical value when sediment transport occurs, $10^{-3}$~m, corresponding to the thickness of the transport layer \citep{Duran2011}.
 It leads for the Adamax, Deep Sea, Huab and South Namib stations values of $2.7~\textup{mm}$, $0.76~\textup{mm}$, $0.12~\textup{mm}$ and $0.48~\textup{mm}$, respectively.
@@
   \includegraphics[scale=1]{Figures/Figure12_supp.pdf}
-  \caption{Regime diagrams of the wind deviation $\delta_{\theta}$ and relative attenuation/amplification $\delta_{u}$ in the spaces $(\mathcal{F}r_{\textup{I}}, \, kH)$ and $(\mathcal{F}r_{\textup{I}}, \, \mathcal{F}r_{\textup{S}})$, containing the data from both the Deep Sea and South Namib stations.
+  \caption{Regime diagrams of the wind deviation $\delta_{\theta}$ and relative attenuation/amplification $\delta_{u}$ in the spaces $(\mathcal{F}_{\textup{I}}, \, kH)$ and $(\mathcal{F}_{\textup{I}}, \, \mathcal{F})$, containing the data from both the Deep Sea and South Namib stations.
 Blue dashed lines empirically delimit the different regimes.
 The point density in each bin of the diagrams is shown in Fig.~\ref{Fig11_supp}.
-The regime diagrams in the space $(\mathcal{F}r_{\textup{S}}, \, kH)$ are shown in Fig.~\ref{Fig5} of the main article.}
+The regime diagrams in the space $(\mathcal{F}, \, kH)$ are shown in Fig.~\ref{Fig5} of the main article.}
   \label{Fig12_supp}

### 2021-12-31 Paper_figure/Figure05.py

@@ -56,7 +56,7 @@ regime_line_color = 'tab:blue'
 
 
 var1, var2 = 'Froude', 'kH'
-xlabel = r'$\mathcal{F}r_{\textup{S}} =  U/\sqrt{(\Delta\rho/\rho) g H}$'
+xlabel = r'$\mathcal{F} =  U/\sqrt{(\Delta\rho/\rho) g H}$'
 
 # #### Figure
 fig, axarr = plt.subplots(1, 2, figsize=(theme.fig_width, 0.375*theme.fig_height_max),

### 2021-12-31 Paper_figure/Supplementary_Figures/Figure11_supp.py

@@ -43,7 +43,7 @@ fig, axarr = plt.subplots(4, 3, figsize=(theme.fig_width, 0.95*theme.fig_width),
                           # constrained_layout=True,
                           gridspec_kw={'height_ratios': [0.2, 1.3, 2, 2],
                                        'width_ratios': [2, 2, 1]})
-# Plotting diagrams
+# Plotting density diagrams
 ax_list = [axarr[2, 0], axarr[3, 0], axarr[3, 1]]
 for j, (ax, (var1, var2)) in enumerate(zip(ax_list, couples)):
     ax.set_xscale('log')
@@ -71,6 +71,7 @@ for j, (ax, (var1, var2)) in enumerate(zip(ax_list, couples)):
     else:
         ax.set_yticklabels([])
 
+# #### Plotting marginal distributions
 for i, (ax, var) in enumerate(zip([axarr[1, 0], axarr[2, 1], axarr[3, 2]], ['Froude', 'kLB', 'kH'])):
     orientation = 'vertical' if i < 2 else 'horizontal'
     make_nice_histogram(Data['South_Namib_Station'][var], 150, ax, alpha=0.4, density=False, scale_bins='log', orientation=orientation)

### 2021-12-31 python_codes/plot_functions.py

@@ -5,7 +5,8 @@ from scipy.stats import binned_statistic_2d
 
 
 def plot_wind_rose(theta, U, bins, ax, fig, label=None,
-                   props=dict(boxstyle='round', facecolor=(1, 1, 1, 0.9), edgecolor=(1, 1, 1, 1), pad=0)):
+                   props=dict(boxstyle='round', facecolor=(1, 1, 1, 0.9), edgecolor=(1, 1, 1, 1), pad=0),
+                   **kwargs):
     """Plot a wind rose from one dimensional time series.
 
     Parameters
@@ -22,6 +23,8 @@ def plot_wind_rose(theta, U, bins, ax, fig, label=None,
         figure on which the wind rose is plotted
     label : str or None
         if not None, label plotted below the wind rose (default is None).
+    **kwargs :
+        Optional parameters passed to :func:`windrose.WindroseAxes.bar <windrose.WindroseAxes>`.
 
     Returns
     -------
@@ -33,7 +36,7 @@ def plot_wind_rose(theta, U, bins, ax, fig, label=None,
     ax_rose.set_position(ax.get_position(), which='both')
     Angle = (90 - theta) % 360
     ax_rose.bar(Angle, U, bins=bins, normed=True, zorder=20, opening=1, edgecolor=None,
-                linewidth=0.5, nsector=60, cmap=plt.cm.viridis)
+                linewidth=0.5, nsector=60, **kwargs)
     ax_rose.grid(True, linewidth=0.4, color='k', linestyle='--')
     ax_rose.patch.set_alpha(0.6)
     ax_rose.set_axisbelow(True)
@@ -67,8 +70,8 @@ def plot_flux_rose(angles, distribution, ax, fig, nbins=20, withaxe=0, label=Non
         Define if the polar axes are plotted or not (the default is 0).
     label : str
         If not None, sets a label at the bottom of the flux rose (the default is None).
-    **kwargs : type
-        Description of parameter `**kwargs`.
+    **kwargs :
+        Optional parameters passed to :func:`windrose.WindroseAxes.bar <windrose.WindroseAxes>`.
 
     Returns
     -------
@@ -187,17 +190,6 @@ def plot_regime_diagram(ax, quantity, vars, lims, xlabel, ylabel, type='scatter'
         ax.set_yticklabels([])
     return a
 
-
-# def make_nice_histogram(data, nbins, ax, vmin=None, vmax=None, scale_bins='lin', density=True, **kwargs):
[…]

### 2022-01-03 Paper/main.tex
Reviewing all text + adding abstract and acknowledgments

@@
 \author{Cyril Gadal \and Pauline Delorme \and Clément Narteau \and
-        Giles Wiggs \and Matthew Baddock \and Joanna M. Nield \and Philippe Claudin }
+        Giles F.S. Wiggs \and Matthew Baddock \and Joanna M. Nield \and Philippe Claudin}
 
@@
 \begin{abstract}
+  In large-scale studies of arid areas, sediment fluxes and resulting geomorphological quantities (sediment pathways, erosion rates, dune morphodynamics) are often derived from wind data.
+  %
+  However, the feedback of nearby topography on the flow, such as kilometric giant dunes, is rarely taken into account because barely studied.
+  %
+  Here, we compare locally measured wind data to the predictions of the Era5Land climate reanalysis in four different places across and around the Namib sand sea.
+In places with flat environments, we show that the two wind datasets agree with each other.
+In contrast, within the sand sea, we show that they significantly differ due to the presence of giant linear dunes.
+We quantify this discrepancy, and link it to dynamics of the atmospheric boundary layer.
+  %
+  Aside setting limits to the applicability of the Era5land dataset, this study highlights the giant dunes feedback on the wind, and discuss its implications for smaller scale bedforms.
+  %
+  All codes used, from the raw data to the figures of the paper, are publicly available with full documentation at https://github.com/Cgadal/GiantDunes (will be made public upon acceptance of this manuscript for publication).
 
@@
   \begin{equation}
-        \mathcal{F} = \displaystyle\frac{U}{\sqrt{\displaystyle\frac{\Delta\rho}{\rho}gH}},
+        \mathcal{F} = \displaystyle\frac{U}{\sqrt{\displaystyle\frac{\Delta\rho}{\rho_{0}}gH}},
   \end{equation}
-  where $U$ is the wind velocity at the top of the ABL, $\rho$ its average density, $\Delta\rho$ the density jump between the ABL and the FA.
+  where $U$ is the wind velocity at the top of the ABL, $\rho_{0}$ its average density, $\Delta\rho$ the density jump between the ABL and the FA.
 
@@
   \end{equation}
-  where $N = \sqrt{-(g/\rho)(\partial \rho/\partial z)}$ is characteristic of the stratification.
+  where $N = \sqrt{-(g/\rho_{0})(\partial \rho/\partial z)}$ is characteristic of the stratification.
 Both Froude numbers have the same qualitative effect on flow confinement, as they quantify the rigidity of the overlying layers.
@@
 \begin{acknowledgements}
+We would like to acknowledge the contributors of the following open-source python librairies, Matplotlib \citep{Hunter2007}, Numpy \citep{Harris2020} and Scipy \citep{Virtanen2020}, which provide an incredibly efficient ecosystem allowing scientific research in Python.
+
+Era5 and Era5Land datasets are publicly available at the Copernicus Climate Change Service (C3S) Climate Data Store.
+The locally measured wind data can be found at \cg{upload on public data repository}.
+The digital elevation models from the Shuttle Radar Topography Mission are publicly available from Nasa servers, and can be downloaded at https://dwtkns.com/srtm30m/.
+Fully documented codes used to analyze this study are available at https://github.com/Cgadal/GiantDunes (will be made public upon acceptance of this manuscript for publication).
+
 [citing all grants ...]\\
-Finally, we would like to acknowledge the contributors of the following open-source python librairies, Matplotlib \citep{Hunter2007}, Numpy \citep{Harris2020} and Scipy \citep{Virtanen2020}, which provide an incredibly efficient ecosystem allowing scientific research in Python.
 \end{acknowledgements}
@@
 
-The two components of the basal shear stress $\boldsymbol{\tau} = \rho u_{*}\boldsymbol{u}_{*}$, constant in a flat bottom situation, can then be written without loss of generality as:
+The two components of the basal shear stress $\boldsymbol{\tau} = \rho_{0} u_{*}\boldsymbol{u}_{*}$, constant in a flat bottom situation, can then be written without loss of generality as:
 \begin{align}
@@
 \end{equation}
-where $\Omega$ is a proportionality constant, $Q = d\sqrt{(\rho_{\textup{s}} - \rho)gd/\rho}$ is a characteristic flux, $\Theta = \rho u_{*, t}^{2}/(\rho_{\textup{s}} - \rho)gd$ the Shields number, and $\Theta_{\textup{th}}$ its threshold value for incipient sediment transport.
+where $\Omega$ is a proportionality constant, $Q = d\sqrt{(\rho_{\textup{s}} - \rho_{0})gd/\rho_{0}}$ is a characteristic flux, $\Theta = \rho_{0} u_{*, t}^{2}/(\rho_{\textup{s}} - \rho_{0})gd$ the Shields number, and $\Theta_{\textup{th}}$ its threshold value for incipient sediment transport.
 Here, $\rho_{\textup{s}} = 2.6~\textup{g}~\textup{cm}^{-3}$ and $d=180~\mu\textup{m}$ are the grain density and diameter, and $g$ is the gravitational acceleration.
@@
 
+As the regionally predicted and locally measured velocities are available at different heights, we can not compare them directly.
+We then convert all velocities into shear velocities $u_{*}$, characteristic of the turbulent velocity profile \citep{Spalding1961, Stull1988}:
+\begin{equation}
+  \frac{u(z)}{u_{*}} = \frac{1}{\kappa}\ln\left(\frac{z}{z_{0}}\right),
+\end{equation}
+where $z$ is the vertical coordinate, $\kappa \approx 0.4$ the von Kármán constant and $z_{0}$ the hydrodynamic roughness.
+Several field measurements of hydrodynamic roughnesses are available.
+In the absence of sediment transport, it scales with the geometric roughness of the topography \citep{Pelletier2016}.
+When transport occurs, it then scales with the thickness of the transport layer ($\sim 1$~mm), depending on the wind velocity and grain properties \citep{Sherman2008, Zhang2016, Field2018}.
+
+The different environments of our stations (vegetated, arid, sandy) makes it difficult to develop a precise model for the computation of the hydrodyamic roughness.
+Furthermore, we do not have precise enough topographical measurements that would allow to compute the geometric roughness.
+We choose to leave aside this complexity, and use a different approach.
+The selected hydrodynamic roughness is then the one that leads to the best possible matching between the regionally predicted and locally measured winds.
+
 For each station, the hydrodynamic roughness is calibrated by finding the one that minimizes the relative difference $\delta$ between the wind vectors of both datasets:
@@
 \end{equation}
+This $\delta$--parameter is computed for hydrodynamic roughness values ranging from $10^{-5}$~m to $10^{-2}$~m for the different stations.
+Note that for the Deep Sea and South Namib stations, where the giant dunes feedback presumably affect the wind, we take into account the non-deflected winds only in the calculation of $\delta$ (with a $15\circ$ tolerance).
 
-This $\delta$--parameter is computed for hydrodynamic roughness values ranging from $10^{-5}$~m to $10^{-2}$~m for the different stations.
-A shown by figure~\ref{Fig3_supp}, the minimum of $\delta$ in the space ($z_{0,\, \textup{Era}}$, $z_{0,\, \textup{in situ}}$) forms a line.
+A shown by figure~\ref{Fig3_supp}, the minimum of $\delta$ in the space ($z_{0}^{\textup{Era5Land}}$, $z_{0}^{\textup{local}}$) forms a line.
 We thus take the roughness of the Era5Land dataset as the typical value when sediment transport occurs, $10^{-3}$~m, corresponding to the thickness of the transport layer \citep{Duran2011}.
@@
 
-The choice of the hydrodynamic roughness values only impacts the calculated shear velocities, but note the wind directions.
-As such, most of our conclusions are then independent of such a choice, and only the magnitude of the wind velocity attenuation in confined situation might be affected.
+The choice of the hydrodynamic roughness values impacts the calculated shear velocities only, but not the wind directions.
+As such, most of our conclusions are independent of such a choice.
+However, it may affect the magnitude of the wind velocity attenuation/amplification in flow confinement situations.
 
@@
   \caption{Calibration of the hydrodynamic roughnesses.
-The metric $\delta$ defined in \eqref{metric_roughness} is represented in colorscale as a function of the hydrodynamic roughnesses chosen for the Era5-Land and in situ datasets, for the Adamax (\textbf{a}), Deep Sea (\textbf{b}), Huab (\textbf{c}) and South Namib (\textbf{d}) Stations.
-The red dashed and plain lines shows the minima of $\delta$ and the identity line.
-The orange lines and dots highlight the chosen hydrodynamic roughnesses for the local datasets, by imposing $z_{0, \textup{\textup{ERA}}} = 1~\textup{mm}$, leading for each station to $2.7~\textup{mm}$, $0.76~\textup{mm}$, $0.12~\textup{mm}$ and $0.48~\textup{mm}$, respectively.}
+The metric $\delta$ defined in \eqref{metric_roughness} is represented in colorscale as a function of the hydrodynamic roughnesses chosen for the Era5-Land and local winds, for the (\textbf{a}) Adamax, (\textbf{b}) Deep Sea, (\textbf{c}) Huab and (\textbf{d}) South Namib stations.
+The red dashed and plain lines shows the minima of $\delta$ and the identity line, respectively.
+The orange lines and dots highlight the chosen hydrodynamic roughnesses for the local winds, by imposing $z_{0}^{\textup{Era5Land}} = 1~\textup{mm}$.
+It leads for for each station to $2.7~\textup{mm}$, $0.76~\textup{mm}$, $0.12~\textup{mm}$ and $0.48~\textup{mm}$, respectively.}
   \label{Fig3_supp}
@@
   \includegraphics[scale=1]{Figures/Figure6_supp.pdf}
-  \caption{Distributions of wind direction at the South Namib Station for the Era5Land climate reanalysis (orange) and the in situ measurements (blue).
+  \caption{Distributions of wind direction at the South Namib Station for the Era5Land climate reanalysis (orange) and the local measurements (blue).
 In each subplot, both distributions are plotted from the same time steps, selected with constraints on the wind direction (columns) and/or wind velocity (rows) in the Era5Land dataset.
@@
 
-In order to estimate the relevant non-dimensional numbers, one need to estimate in addition to the wind and dune properties some parameters of the ABL.
-The Era5 dataset provides a direct bulk estimate of the ABL depth $H$ from a bulk Richardson number calculation, as well as vertical profiles of the geopotential $\phi$, temperature $T$ and specific humidity $e_{\textup{w}}$ at given pressure levels $P$.
-From these quantities, the virtual potential temperature, which takes into account the vertical pressure and humidity changes, can be calculated as:
+The estimation of the non-dimensional numbers requires the computation of meteorological quantities representative of the current atmospheric boundary layer.
+In arid areas, the vertical structure of the atmosphere can be approximated by well mixed convective boundary layer of height $H$, topped by the stratified free atmosphere \citep{Stull1988, Shao2008}.
+%
+In this type of structure, the virtual potential temperature is constant inside the boundary layer, and then increases linearly in the stratified free atmosphere (Fig.~\ref{Fig9_supp}a):
+\begin{equation}
+  T_{\textup{vp}}(z) = \begin{cases}
+                          T_{0} &\text{for $z < H$},\\
+                          T_{0}\left(1 + \displaystyle\frac{\Delta T_{\textup{vp}}}{T_{0}} + \frac{N^{2}}{g}(z - H)\right) &\text{for $ z > H$},
+                        \end{cases}
+\end{equation}
+where, $T_{0}$ is the surface virtual potential temperature, $\Delta T_{\textup{vp}}$ the discontinuity at the capping layer and $N$ the Brunt-Väisälä frequency, characteristic of the stratification.
+
+The Era5 dataset provides a direct estimate of the ABL depth $H$ from a bulk Richardson number calculation, as well as vertical profiles of the geopotential $\phi$, temperature $T$ and specific humidity $e_{\textup{w}}$ at given pressure levels $P$.
+From these quantities, the virtual potential temperature can be calculated as:
 \begin{equation}
@@
 \end{equation}
-where $P_{0} = 10^{5}~\textup{Pa}$ is the standard pressure, $P_{\textup{c}} = 0.2854$ the Poisson coefficient for dry air and $R_{\textup{M}}=1.61$ is the ratio between the melocular masses of dry air and water.
+where $P_{0} = 10^{5}~\textup{Pa}$ is the standard pressure, $P_{\textup{c}} = 0.2854$ the Poisson coefficient for dry air and $R_{\textup{M}}=1.61$ is the ratio between the molecular masses of dry air and water.
 The vertical coordinates are calculated as:
@@
 
-Example of obtained vertical profiles of the virtual potential temperature are shown in Fig.~\ref{Fig9_supp}.
-On each of them, an average is computed below the ABL depth given by the Era5 dataset, and a linear function is fitted above.
-
-Under the Boussinesq approximation, the temperature variations are assumed to induce most of those of the density, leading to $\Delta\rho/\rho \simeq \Delta T_{\textup{vp}}/T_{\textup{vp}}$.
-Here, $T_{\textup{vp}}/T_{\textup{vp}}$ is the relative virtual potential temperature jump at the capping, directly measured on the vertical profiles.
-
-Following \citet{Tritton2012}, the relative density jump at the capping layer
+Example of obtained vertical profiles of the virtual potential temperature are shown in Fig.~\ref{Fig9_supp}a.
+On each of them, an average is computed below the ABL depth directly given by the Era5 dataset, and a linear function is fitted above.
+Note that under the Boussinesq approximation, the temperature variations are assumed to induce most of those of the density, leading to $\Delta\rho/\rho_{0} \simeq \Delta T_{\textup{vp}}/T_{0}$ (see supplementary material of \citep{andreotti2009}).
+Note that we removed some profiles that displayed a vertical structure that could not be approximated by the simple model used here, leading for example to negative values of $\Delta T_{\textup{vp}}$ (see Fig.~\ref{Fig9_supp}b).
+While these profiles dominantly occurs in winter, when the two winds blows, they are evenly spread across the hours of the day, and represent 12~\% of our data only (see Fig.~\ref{Fig9_supp}c--d).
+For these reasons, we are confident this does not affect pour conclusions.
 
@@
 Transparent dashed lines: boundary layer height given by the ERA5 reanalysis, calculated from the bulk Richardson number \citep{seidel2012}.
-Plain lines: vertical (boundary layer) and linear (free atmosphere) fits to estimate the quantities in Fig.~\ref{Fig10_supp}.
+Plain lines: vertical (ABL) and linear (FA) fits to estimate the quantities in Fig.~\ref{Fig10_supp}.
 \textbf{b}: Examples of ill-processed vertical profiles at 3 different time steps (blue - 2/12/2013 - 2300 UTC, orange - 20/03/2017 - 0000 UTC, green - 14/07/2017 - 1400 UTC) at the South Namib station.
 \textbf{c}: Hourly distribution of ill-processed vertical profiles.
-\textbf{d}: Monthly distribution of ill-processed vertical profiles.
-\cg{These profiles are ill-processed because the temperature found at the boundary layer from the linear fit in the free-atm is smaller than the average one inside the boundary layer.
-This is an unstable situation, which does not allow to calculate the surface Froude number.}}
+\textbf{d}: Monthly distribution of ill-processed vertical profiles.}
   \label{Fig9_supp}

### 2022-01-04 Paper/main.tex
Commit before sending to Clement and Philippe

@@
 %
-At the largest scale, the feedback of mountains on the stratified part results in wave generation as well as significant wind disturbances, such as downslope winds in the lee side \citep{Duran1990}.
+At the largest scale, the feedback of mountains on the stratified free atmosphere results in wave generation as well as significant wind disturbances, such as downslope winds in the lee side \citep{Duran1990}.
 Inside the boundary layer, the interaction between a turbulent flow and hilly surfaces is for example key to the understanding ocean surface wind-driven waves~\citep{Sullivan2010}, or eolian bedforms in desert \citep{Courrech2015}.
@@
 In the case of aeolian bedforms, the measurements mentioned above were made on dunes a few meters high only.
-The resulting wind disturbances have then been shown to significantly impact the sediment pathways of the system \citep{Hesp2015}, but also to affect the collective behaviour of dune populations trough long-range interactions \citep{Bacik2020}.
+The resulting wind disturbances have then been shown to significantly impact the sediment pathways of the whole system \citep{Hesp2015}, but also to affect the collective behaviour of dune populations trough long-range interactions \citep{Bacik2020}.
 %

### 2022-01-29 Paper/main.tex

--- 2021-12-31 Paper/main.tex
@@
 \newcommand{\cg}{\textcolor{WildStrawberry}}
+\newcommand{\pc}{\textcolor{ForestGreen}}
+\newcommand{\new}{\textcolor{blue}}
+\newcommand{\todo}{\textcolor{red}}
 
+%---------------------------------------------------------------------------------------
 \begin{document}
 
-\title{Local wind regime induced by the presence of giant linear dunes.}
+\title{Local wind regime induced by giant linear dunes}
 
-\author{Cyril Gadal \and Pauline Delorme \and Clément Narteau \and
-        Giles Wiggs \and Matthew Baddock \and Joanna M. Nield \and Philippe Claudin }
+\author{Cyril Gadal \and Pauline Delorme \and Cl\'ement Narteau \and Giles F.S. Wiggs \and Matthew Baddock \and Joanna M. Nield \and Philippe Claudin}
 
-\institute{C. Gadal \at
-              Institut de Mécanique des Fluides de Toulouse (IMFT), Université de Toulouse, CNRS, INPT, UPS, Toulouse, France \\
-              \email{cyril.gadal@imft.fr}
+\institute{
+C. Gadal \at
+Institut de M\'ecanique des Fluides de Toulouse, Universit\'e de Toulouse Paul Sabatier, CNRS, Toulouse INP-ENSEEIHT, Toulouse, France. \email{cyril.gadal@imft.fr}
 \and
-           S. Author \at
-              second address
+P. Delorme \at
+School of Geography and Environmental Science, University of Southampton, Southampton, UK.
 \and
-            T. Author \at
-            third address
+C. Narteau \at
+Institut de Physique du Globe de Paris, Universit\'e de Paris, CNRS, Paris, France.
+\and
+G.S.F. Wiggs \at
+School of Geography and the Environment, University of Oxford, Oxford, UK.
+\and
+M. Baddock \at
+Geography and Environment, Loughborough University of Technology, Loughborough, UK.
+\and
+J.M. Nield \at
+School of Geography and Environmental Science, University of Southampton, Southampton, UK.
+\and
+P. Claudin \at
+Physique et M\'ecanique des Milieux H\'et\'erog\`enes, CNRS, ESPCI Paris, PSL Research University, Universit\'e de Paris, Sorbonne Universit\'e, Paris, France.  \email{philippe.claudin@espci.fr}
 }
@@
 \begin{abstract}
-
-  \keywords{Boundary layer \and Turbulent flow \and Sand dunes \and Fluide-structures interactions}
+Emergence and growth of sand dunes result from the interaction between the topography, which perturbs the wind flow, which itself controls the sediment transport.
+While these feedbacks are well studied at the scale of a single dune, the average effect of a periodic dune pattern on atmospheric flows remains poorly constrained due to a lack of data in major sand seas.
+Here, we compare field measurements of surface wind data to the predictions of the ERA5-Land climate reanalysis in four different places in Namibia, inside and north of the giant-dune field of the Namib sand sea.
+In the flat desert areas to the north, observations and predictions agree well despite the small amount of data in the assimilation process.
+This is also the case in the interdune areas of the sand sea, except for the weak winds blowing at night in a direction oblique to the orientation of the giant dunes.
+We quantify these similarities and differences and provide a physical understanding of the relevant aerodynamical regimes to relate them to the daily cycle of the turbulent atmospheric boundary layer over a dune pattern of given wavelength.
+We conclude by identifying the conditions under which the ERA5-Land reanalysis data can or can't be used to study dune morphodynamics.
+We also propose that, in multidirectional wind regimes, deflections of specific winds could explain the occurrence of secondary dune patterns with a different orientation than the primary structures between which they grow.
+\keywords{Atmospheric boundary layer \and Sand dunes \and Fluid-structure interactions}
 \end{abstract}
@@
 
+%---------------------------------------------------------------------------------------
 \section{Introduction}
 
-As a flow encounters an obstacle, different interactions can arise depending on the different time and length scales involved.
-In the case of atmospheric flows, this mainly depends on the part of the atmosphere, schematically composed of a turbulent boundary layer topped by a stably stratified one, with which the obstacle interacts \citep{Stull1988}.
-%
-At the largest scale, the feedback of mountains on the stratified part results in wave generation as well as significant wind disturbances, such as downslope winds in the lee side \citep{Duran1990}.
-Inside the boundary layer, the interaction between a turbulent flow and hilly surfaces is for example key to the understanding ocean surface wind-driven waves~\citep{Sullivan2010}, or eolian bedforms in desert \citep{Courrech2015}.
+The description of turbulent flows over a complex topography is a rich and active subject, that is relevant for a large variety of different environmental systems \citep{Finnigan2020}.
+For example, the flow over hills is of primary interest for wind power, meteorological and air pollution phenomena \citep{Taylor1987}.
+The properties of these flows are also key to the understanding of the formation of ocean surface wind-driven waves \citep{Sullivan2010}, dissolution bedforms \citep{Claudin2017}, or sedimentary ripples and dunes \citep{Charru2013, Courrech2015}.
+Importantly, the troposphere presents a vertical structure, with a lower convective boundary layer, of typical kilometer-scale thickness, capped by a stably stratified region \citep{Stull1988}.
+The largest topographic obstacles as mountains can therefore interact with this upper region, for example leading to internal wave generation or significant wind disturbances, such as downslope winds in lee sides \citep{Durran1990}.
 
-Looking at the flow close to the surface, two different components of the topography feedback can be isolated.
-First, the flow accelerates on the upwind slope, and slows down on the downwind one, with a maximum velocity slightly upwind of the crest~\citep{Jackson1975, Sykes1980, Hunt1988}.
-Several measurements are available, performed in subaqueous flumes~\citep{Zilker1977, Zilker1979, Frederick1988} and in eolian field conditions~\citep{Claudin2013, Lu2021}.
-Second, deflection occurs when the incident flow direction is not perpendicular to the ridge crest.
-While predicted to be small (less than $10^{\circ}$) in the linear regime~\citep{Gadal2019}, significant flow steering has been reported in the field on the downwind side of larger aspect ratios obstacles, such as mountain ranges~\citep{Kim2000, Lewis2008, Fernando2019} or sand dunes~\citep{Walker2009, Walker2013, Hesp2015, Smith2017}.
+Looking at the wind close to the surface, two topographic feedbacks on the flow, although related, can be commented separately.
+There is a first effect on the wind intensity: the flow accelerates on the upwind slope and slows down on the downwind one \citep{Baddock2011}, with a speed-up factor essentially proportional to the obstacle aspect ratio \pc{here a ref?}.
+Importantly, the velocity maximum is typically shifted upwind of the obstacle crest.
+This behaviour has been theoretically predicted by means of asymptotic analysis of a neutrally stratified boundary-layer flow over an obstacle of vanishing aspect ratio \citep{Jackson1975, Mason1979, Sykes1980, Hunt1988, Belcher1998}.
+Experiments in flumes \citep{Zilker1977, Zilker1979, Frederick1988, Poggi2007}, in wind tunnel \citep{Gong1989, Finnigan1990, Gong1996} and in field conditions \citep{Taylor1987a, Claudin2013, Fernando2019, Lu2021}, have also documented this effect.
+Interestingly, a similar behaviour exists for the pressure perturbation, but with a slight downwind shift for the pressure minimum \citep{Claudin2021}.
+The second effect is the flow deflection that occurs when the incident wind direction is not perpendicular to the ridge crest.
+While predicted to be small (less than $10^{\circ}$) in the linear regime valid for shallow topography \citep{Gadal2019}, significant flow steering has been reported in the field on the downwind side of steep enough obstacles, such as mountain ranges \citep{Kim2000, Lewis2008, Fernando2019} or well developed sand dunes \citep{Walker2009, Hesp2015, Walker2017, Smith2017}.
 
-In the case of aeolian bedforms, the measurements mentioned above were made on dunes a few meters high only.
-The resulting wind disturbances have then been shown to significantly impact the sediment pathways of the system \citep{Hesp2015}, but also to affect the collective behaviour of dune populations trough long-range interactions \citep{Bacik2020}.
-%
-However, eolian dunes coexists in sand seas at different scales, each representative of their different evolution stages, from emergence at twenty meters wavelengths to giant size, corresponding to kilometric wavelengths and heights of several dozen meters \citep{McKee1979}.
-Quantifying these two effects for giant dunes is then all the more important for several reasons.
-%
-First, larger wind disturbances are expected from larger obstacles.
-%
-Second, the flow--structure coupling is not straightforward due to vertical structure of the atmosphere with which the giant dunes interact due to their sizes \citep{andreotti2009}.
-%
-Finally, the morphodynamics of aeolian bedforms are strongly dependent on the wind regime~\citep{Livingstone1996}.
-Thus, small-scale bedforms are sensitive to wind disturbances induced by larger-scale bedforms.
-This has been demonstrated for impact ripples on dunes \citep{Howard1977,Hood2021}, but has never been studied for medium-scale dunes near giant dunes.
+For practical reasons, wind measurement over sand dunes have been performed over rather small bedforms, typically a few meters high only -- that is tens of meters in length.
+Giant dunes, with kilometer-scale wavelengths and heights of tens of meters, are more difficult to investigate although they provide, for several reasons, a choice configuration for the study of turbulent flows over a complex topography.
+First of all, one expects larger wind disturbances for larger obstacles.
+Second, their large size make them interact with the vertical structure of the atmosphere \citep{andreotti2009}.
+Also, they usually form large patterns in sand seas and thus behave as rather clean periodic perturbations, in contrast with isolated hills.
+Finally, because the morphodynamics of aeolian bedforms are strongly dependent on the local wind regime \citep{Livingstone1996}, one can expect to see the consequences of the wind disturbances on smaller dunes, in a similar way as what has been reported for the effect of dunes on impact ripples \citep{Howard1977,Hood2021}.
 
-In the literature, arid areas have been studied at the desert scale from climate reanalyses based on global atmospheric models \citep{Blumberg1996, Livingstone2010, Ashkenazy2012, Jolivet2021, Hu2021}, such as ERA-40, ERA-Interim or ERA-5 \citep{Uppala2005, Dee2011, Hersbach2020}.
-However, the spatial resolution (tens of kilometers) of these reanalyses implies average quantities that do not resolve the smaller scales, ranging from the study of individual dunes to the border of mountains \citep{Livingstone2010}.
-Lately, the release of ERA5-Land allow to push back this limitation by providing up to 70 years of hourly wind predictions at a 9~km spatial resolution \citep{munoz2021}.
-However, due to its recent nature, the applicability limitations of this dataset remain to be studied.
+Arid areas have been much studied at the desert scale from climate reanalyses based on global atmospheric models \citep{Blumberg1996, Livingstone2010, Ashkenazy2012, Jolivet2021, Hu2021}, such as ERA-40, ERA-Interim or ERA-5 \citep{Uppala2005, Dee2011, Hersbach2020}.
+However, the spatial resolution (tens of kilometers) of these reanalyses implies average quantities that do not resolve the smaller scales, ranging from individual dunes to the border of mountains \citep{Livingstone2010}.
+Lately, the release of ERA5-Land allows to push back this limitation by providing up to 70 years of hourly wind predictions at a $9$~km spatial resolution \citep{munoz2021}.
+However, its validity remains to be studied, especially in areas where data assimilation is low (such as deserts).
 
-Here, we compare local wind speeds and directions measured in four different places across the Namib desert to the regional predictions of the ERA5-Land climate reanalysis.
-When the measurement stations are surrounded by a relatively flat environment, we show that local measurements and regional predictions agree with each other.
-On the contrary, in the sand sea, we show that they differ due to the presence of giant dunes.
-Furthermore, we link the magnitude of these discrepancies to the circadian cycle of the atmospheric boundary layer.
-Finally, we draw implications for smaller-scale eolian bedforms.
+In this work, we compare local wind speeds and directions measured in four different places inside and north of the giant-dune field of the Namib sand sea to the regional predictions of the ERA5-Land climate reanalysis.
+When the meteorological stations are surrounded by a relatively flat environment, we show that local measurements and regional predictions agree well with each other.
+The find that the agreement is also good in the interdune areas of the sand sea, except for the weak winds blowing at night in a direction oblique to the orientation of the giant dunes (section 2).
+Furthermore, we are able to link the magnitude of these differences to the circadian cycle of the atmospheric boundary layer (section 3).
+Finally, we draw implications of the wind disturbances on smaller-scale dunes (section 4).
 
-\section{Wind regimes across the Namib Sand Sea}
-
-  We focus on four places across and nearby the Namib desert, highlighting different environments (see Fig.~\ref{Fig1}).
-  %
-  The Adamax station is located near the Adamax salt pan, in a highly vegetated area.
-The Huab station, on the coast at the outlet of the Huab river, is in an arid environment exhibiting $60$-m scale barchan dunes.
-While the surroundings of these two stations are relatively flat, this is not the case for the Deep Sea and South Namib stations.
-Both are in the interdune between tens of meters high giant linear dunes with kilometric wavelengths and superimposed patterns.
-In this section, we describe and compare winds from local measurements and climate reanalysis predictions.
-
-  \begin{figure}
+%---------------------------------------------
+\begin{figure}[t]
 \centering
 \includegraphics[scale=1]{Figures/Figure1.pdf}
-    \caption{Wind data used in this study \textbf{a}: Location of the studied sites.
-\textbf{b--e}: Satellite images of the studied sites (Google-Earth, Maxar Technologies, CNES/Airbus).
-In each subplot, the left and right wind roses represent the data from the ERA5Land climate reanalysis and the local wind stations, respectively.
+\caption{Wind data used in this study \textbf{a}: Location of the different sites in Namibia.
+\textbf{b-e}: Satellite images of these different environements (Google-Earth, Maxar Technologies, CNES/Airbus).
+In each subplot, the left and right wind roses represent the data from the ERA5-Land climate reanalysis and the local wind stations, respectively.
 Note that the bars show the direction towards which the wind blows.
@@
 \end{figure}
+%---------------------------------------------
 
-  \subsection{Datasets}
+%---------------------------------------------------------------------------------------
+\section{Wind regimes across the Namib Sand Sea}
+We have measured the wind regime in four different sites of Namibia, representative of various environments across and nearby the Namib desert (Fig.~\ref{Fig1}).
+The Adamax station is located near the Adamax salt pan, in a highly vegetated area.
+The Huab station, on the coast at the outlet of the Huab river, is in an arid place where $60$-m scale barchan dunes develop.
+These two stations are located in a relatively flat \pc{quantify how flat is flat ?} environment.
+In contrast, the Deep Sea and South Namib stations are located in the interdune between tens of meters high giant linear dunes with kilometer-scale wavelengths and superimposed patterns.
+In this section, we describe and compare winds from local measurements and climate reanalysis predictions.
 
-  Local winds are provided by measurement stations located in the four different places (see black dots in Fig.~\ref{Fig1}).
+%------------------------
+\subsection{Wind and elevation data}
+Local winds are measured at meteorological stations located in these four different places (black dots in Fig.~\ref{Fig1}).
 The wind strength and direction are sampled every 10 minutes by cup anemometers and wind vanes, at heights between $2$~m and $3$~m depending on the station.
-The available period of measurements ranges from 1 to 5 discontinuous years distributed between 2012 and 2020 (see Fig.~\ref{Fig1_supp}).
+The available period of measurements ranges from 1 to 5 discontinuous years distributed between 2012 and 2020 (Suppl.
+Fig.~\ref{Fig1_supp}).
 We checked that at least one complete seasonal cycle is available at each station.
-  %
 Regional winds are extracted at the same locations and periods from the ERA5-Land dataset, which is a replay at a smaller spatial resolution of ERA5, the latest climate reanalysis from the ECMWWF \citep{Hersbach2020, munoz2021}.
-It provides hourly predictions of the 10-m wind velocity and direction at a spatial resolution of $\sim 9~\textup{km}$ ($0.1^\circ\times0.1^\circ$).
+It provides hourly predictions of the 10-m wind velocity and direction at a spatial resolution of $0.1^\circ\times0.1^\circ$ ($\simeq 9~\textup{km}$ in Namibia).
 
-  For comparison, the local measurements are averaged into 1-hr bins centered on the temporal scale of the ERA5-Land estimates (see Fig.~\ref{Fig2_supp}).
-As the wind velocities of both datasets are provided at different heights, we convert them into shear velocities (see SI section 1), characteristic of the turbulent wind profile, which are then used together with the wind direction for further analysis.
-The resulting wind data are shown on the wind roses of Fig.~\ref{Fig1}(b--e).
+For comparison, the local measurements are averaged into 1-hr bins centered on the temporal scale of the ERA5-Land estimates (Suppl.
+Fig.~\ref{Fig2_supp}).
+As the wind velocities of both datasets are provided at different heights, we convert them into shear velocities (Suppl.
+Mat.
+section 1), characteristic of the turbulent wind profile, which are then used together with the wind direction for further analysis.
+Wind roses in Fig.~\ref{Fig1}(b--e) show the resulting wind data.
 
-  Finally, the dune properties are computed using autocorrelation on the 30-m Digital Elevation Models (DEMs) of the shuttle radar topography mission \citep{Farr2007}.
-For the South Namib and Deep Sea stations, we obtain respectively orientations of $85^\circ$ and $125^\circ$, wavelengths of $2.6~\textup{km}$ and $2.3~\textup{km}$ and amplitudes of $45~\textup{m}$ and $20~\textup{m}$ (see Fig.~\ref{Fig4_supp} for more details).
+Dune properties are computed using autocorrelation on the 30-m Digital Elevation Models (DEMs) of the shuttle radar topography mission \citep{Farr2007}.
+For the South Namib and Deep Sea stations, we obtain respectively orientations of $85^\circ$ and $125^\circ$ with respect to the North, wavelengths of $2.6~\textup{km}$ and $2.3~\textup{km}$ and amplitudes of $45~\textup{m}$ and $20~\textup{m}$ (Suppl.
+Fig.~\ref{Fig4_supp} for more details).
 
-  \subsection{Agreement between local and regional winds}
-
+%---------------------------------------------
 \begin{figure}
@@
 \includegraphics[scale=1]{Figures/Figure2.pdf}
-    \caption{Temporal comparison between the wind data coming from the Era5Land climate reanalysis (orange lines) and from the local measurements (blue lines).
-Color swathes indicate day (between 1000 UTC and 2200 UTC) and night (before 1000 UTC or after 2200 UTC) \textbf{a--b}: Huab station.
-\textbf{c--d}: Deep Sea station in winter.
-\textbf{e--f}: Deep Sea station in summer.}
+\caption{Temporal comparison between the wind data coming from the ERA5-Land climate reanalysis (orange lines) and from the local measurements (blue lines).
+Coloured swathes indicate day (between 1000 UTC and 2200 UTC) and night (before 1000 UTC or after 2200 UTC) \textbf{a-b}: Huab station.
+\textbf{c-d}: Deep Sea station in winter.
+\textbf{e-f}: Deep Sea station in summer.}
 \label{Fig2}
 \end{figure}
+%---------------------------------------------
 
+%------------------------
+\subsection{Comparison of local and regional winds}
+\label{section_data_feedback}
 The obtained wind regimes are shown in figure~\ref{Fig1}.
-In the Namib, the regional wind patterns are essentially controlled by the see breeze, resulting in strong northward components (sometimes slightly deviated by the large scale topography) present in all regional wind roses \citep{lancaster1985}.
-These daytime winds are dominant during the second-half of the year (Septembre-January).
-In winter, an additional easterly component can be recorded during the night, induced by the combination of katabatic winds forming on the mountains, and infrequent `berg' winds, which are responsible of the high wind velocities observed \citep{lancaster1984}.
-The frequency of these easterly components decreases from the inland to the coast, resulting in bidirectional wind regimes within the Namib Sand Sea and at the Adamax salt pan (Fig.~\ref{Fig1}b, \ref{Fig1}d and \ref{Fig1}e) and a unidirectional wind regime on the coast at the outlet of the Huab River (Fig.~\ref{Fig1}c).
+In the Namib, the regional wind patterns are essentially controlled by the see breeze, resulting in strong northward components (sometimes slightly deviated by the large scale topography \pc{mountain?}) present in all regional wind roses \citep{lancaster1985}.
+These daytime winds are dominant during the second-half of the year (September-January).
+In winter, an additional easterly component can be recorded during the night, induced by the combination of katabatic winds forming in the mountains, and infrequent `berg' winds, which are responsible of the high wind velocities observed \citep{lancaster1984}.
+The frequency of these easterly components decreases from the inland to the coast.
+As a result, bidirectional wind regimes within the Namib Sand Sea and at the Adamax salt pan (Fig.~\ref{Fig1}b,d,e) and a unidirectional wind regime on the coast at the outlet of the Huab River (Fig.~\ref{Fig1}c) are observed.
 
-  In the case of the Adamax and Huab stations, the wind roses from the regional predictions qualitatively match those corresponding to the local measurements.
-However, for the Deep Sea and South Namib stations, the measured wind roses exhibit additional components aligned with the giant dune orientation visible on the satellite images (Fig.~\ref{Fig1}c--d).
-  %
-  The time series of wind speed and direction show that this agreement in the case of Adamax and Huab stations is always verified (Fig.~\ref{Fig2}a--b) and Fig.~\ref{Fig5_supp}).
-In contrast, for the stations within the giant dune field, we observe that this agreement is limited to Septembre--January time periods (Fig.~\ref{Fig2}c--d).
+In the case of the Adamax and Huab stations, the time series of wind speed and direction from the regional predictions quantitatively match those corresponding to the local measurements (Fig.~\ref{Fig2}a,b) and Suppl.
+Fig.~\ref{Fig5_supp}).
+For the Deep Sea and South Namib stations within the giant dune field, we observe that this agreement is also good, but limited to the September-January time period (Fig.~\ref{Fig2}c,d).
+As a matter of fact, the measured wind roses exhibit additional wind components aligned with the giant dune orientation, as evidenced on the satellite images (Fig.~\ref{Fig1}c,d).
 
-  \subsection{Influence of the giant dunes on local wind regimes}
-  \label{section_data_feedback}
+More precisely, during the February-August period, the local and regional winds in the interdune match during daytime only, i.e when the southerly/southwesterly sea breeze dominates (Figs.~\ref{Fig2}(e,f) and \ref{Fig3}, Suppl.
+Fig.~\ref{Fig6_supp}).
+In the late afternoon and during the night, when the northwesterly `berg' and katabatic winds blow, measurements and predictions differ.
+In this case, the angular wind distribution of the local measurements exhibits two additional modes corresponding to reversing winds aligned with the giant dune orientation (purple frame in Fig.~\ref{Fig3}, Suppl.
+Figs.~\ref{Fig6_supp} and \ref{Fig7_supp}).
+This deviation is also associated with a global attenuation of the wind strength (Suppl.
+Fig.~\ref{Fig8_supp}).
+Remarkably, all these figures show that this wind reorientation and attenuation processes typically occur at low wind velocities, typically for $u_{*} \lesssim 0.1~\textrm{m}~\textrm{s}^{-1}$.
+For shear velocities larger than $\simeq 0.25~\textrm{m}~\textrm{s}^{-1}$, the wind reorientation is not significant.
+Finally, for intermediate shear velocities, moments of reorientation along the dune crest and others with no reorientation can both be observed (Suppl.
+Fig.~\ref{Fig7_supp}).
+\pc{CHECK the above lines cf.
+$\delta u$}
 
-  In the February--August period, when giant dunes are present, the local and regional winds match during daytime only, i.e when the southerly/southwesterly sea breeze dominates (see Fig.~\ref{Fig2}(e--f), Fig.~\ref{Fig3} and Fig.~\ref{Fig6_supp}).
-In the late afternoon and during the night, when the northwesterly `berg' and katabatic winds blow, measurements and predictions differ.
-In this case, the angular wind distribution of the local measurements exhibits two additional modes separated of $\simeq 180^\circ$, each corresponding to the giant dune alignement (see the purple frame in Fig.~\ref{Fig3} and Fig.~\ref{Fig6_supp}, as well as Fig.~\ref{Fig7_supp}).
-This deviation is also associated with a global attenuation of the wind strength (Fig.~\ref{Fig8_supp}).
-Remarkably, all these figures show that this process occurs for low wind velocities, typically for $u_{*} < 0.1~\textrm{m}~\textrm{s}^{-1}$.
-For shear velocities larger than $0.25~\textrm{m}~\textrm{s}^{-1}$, this wind reorientation does not occur.
-Finally, for intermediate shear velocities, both reorientation along the dune crest and no reorientation are observed (Fig.~\ref{Fig7_supp}).
-
+%---------------------------------------------
 \begin{figure}
@@
 \includegraphics[scale=1]{Figures/Figure3.pdf}
-    \caption{Distributions of wind direction at the Deep Sea Station for the Era5Land climate reanalysis (orange) and the local measurements (blue).
-In each subplot, both distributions are plotted from the same time steps, selected with constraints on the wind direction (columns) and/or wind velocity (rows) of the Era5Land dataset.
+\caption{Distributions of wind direction at the Deep Sea Station for the ERA5-Land climate reanalysis (orange) and the local measurements (blue).
+In each subplot, both distributions are plotted from the same time steps, selected with constraints on the wind direction (columns) and/or wind velocity (rows) of the ERA5-Land dataset.
 The gray vertical dashed lines indicate the dune orientation.
 The numbers at the top right give the percentage of time steps selected in each sub-range, as well as the percentage of them corresponding to the day (between 1000 UTC and 2200 UTC).
-The purple frame highlights the regime (small wind velocities, nocturnal summer wind) in which the data from both datasets differs.
-A similar figure can be obtained for the Deep Sea station (see Fig.~\ref{Fig6_supp}).}
+The purple frame highlights the regime (small wind velocities, nocturnal summer wind) in which the data from both datasets differ.
+A similar figure can be obtained for the Deep Sea station (Suppl.
+Fig.~\ref{Fig6_supp}).}
 \label{Fig3}
 \end{figure}
+%---------------------------------------------
 
-  \section{Influence of the circadian cycle of the atmospheric boundary layer}
+%---------------------------------------------------------------------------------------
+\section{Influence of wind speed and circadian cycle on the atmospheric boundary layer}
+The wind deflection induced by linear dunes has so far mainly been related to the angle between wind direction and crest orientation, with a maximum deflection for angles between $30^{\circ}$ and $70^{\circ}$ \citep{Walker2009, Hesp2015}.
+In the analysed data, the most deflected wind at both Deep Sean and South Namib stations is the most perpendicular to the giant dunes (\pc{Fig.~\ref{***}}).
+The incident wind direction then does not seem to be the only parameter controlling the wind deflection.
+In contrast, a different behaviour is observed between low and high wind velocities, suggesting a change in hydrodynamical regime.
+In this section, we discuss the relevant parameters associated with the dynamical mechanisms that govern the interactions between the atmospheric boundary layer flow and topographical obstacles.
+This analysis allows us to provide a physics-based interpretation the observations.
 
-  For linear ridges, dune-induced flow disturbances have mainly been related to the angle between wind direction and crest orientation, with a maximum for angles between $30^{\circ}$ and $70^{\circ}$~\citep{Walker2009, Hesp2015}.
-In our case, the most deflected wind for both stations is the most perpendicular, such that the incident wind direction does not seem to be the dominant parameter controlling the wind deflection.
-In contrast, a different behavior is observed between  low and high wind velocities, suggesting a change in hydrodynamical regime.
-
-  In the following, we discuss the relevant parameters leading to different hydrodynamical interactions with topographical obstacles, and interpret the data with respect to the corresponding physical mechanisms.
-
+%---------------------------------------------
 \begin{figure}
@@
 \includegraphics[scale=1]{Figures/Figure4.pdf}
-    \caption{\textbf{a--b}: Vertical profiles of the virtual potential temperature at 2 different time steps (day - 31/03/2017 - 1200 UTC, night - 21/03/2017 - 2200 UTC) at the Deep Sea station.
+\caption{\textbf{a-b}: Vertical profiles of the virtual potential temperature at 2 different time steps (day - 03/11/2015 - 1200 UTC, night - 01/13/2013 - 0900 UTC) at the Deep Sea station.
 Dots: data from the ERA5 reanalysis.
-Transparent dashed lines: boundary layer height given by the ERA5 reanalysis, calculated from the bulk Richardson number \citep{seidel2012}.
+Dashed lines: boundary layer height given by the ERA5 reanalysis (Suppl.
+Mat.
+section 2).
 Plain lines: vertical (boundary layer) and linear (free atmosphere) fits to estimate the stratification properties.
-\textbf{c--f}: Sketches representing the interaction between the giant dunes and the atmospheric flow for different meteorological conditions.
-\textbf{g}: Streamlines qualitatively representing the effect of low, medium and high flow confinement.
-For details on the streamline derivation, see Appendix~\ref{turbulent_wind_model}.}
+\textbf{c-f}: Sketches representing the interaction between the giant dunes and the atmospheric flow for different meteorological conditions.
+\textbf{g}: Streamlines over a sinusoidal topography $\xi(x,y)$ qualitatively representing the effect of low, medium and strong flow confinement, in relation to the above panels (see Appendix~1 for more details).
+\pc{add arrow with reference wind}}
 \label{Fig4}
 \end{figure}
+%---------------------------------------------
 
-  \subsection{Relevant non-dimensional parameters and physical considerations}
+%------------------------
+\subsection{Flow over a modulated bed}
 \label{theoretical_framework}
+Taking as a reference the turbulent flow over a flat bed, the general framework of this study is the understanding and the description of the flow response to a bed modulation.
+Without loss of generality, we can consider in this context an idealised bed elevation in the form of parallel sinusoidal ridges, with wavelength $\lambda$ (or wavenumber $k = 2\pi/\lambda$) and amplitude $\xi_0$, and where the reference flow direction makes a given angle with respect to the ridge crest \citep{Andreotti2012}.
+Part of this response, on which we focus here, is the flow deflection by the ridges.
+In a simplified way, it can be understood from the Bernoulli principle~\citep{Hesp2015}: as the flow approaches the ridge crest, the compression of the streamlines results in larger flow velocities, and thus lower pressures \citep{Rubin1987}.
+An incident flow oblique to the ridge is then deflected towards lower pressure zones, i.e towards the crest.
+Turbulent dissipation tends to increase this effect downstream, resulting in along the crest wind deflection in the lee side~\citep{Gadal2019}.
 
-  Flow deflection over ridges can be understood from the Bernoulli principle~\citep{Hesp2015}.
-As the flow approaches the ridge crest, the compression of the streamlines results in larger flow velocities, and thus lower pressures~\citep{Rubin1987}.
-An incident flow oblique to the ridge is then deflected towards lower pressure zones, i.e towards the crest.
-Turbulent dissipation at the bottom and non-linearities tends to increase this effect downstream, resulting in along the crest wind deflection in the lee side~\citep{Hesp2015, Gadal2019}.
+Flow confinement below a capping surface, which enhances streamline compression, has a strong effect on the hydrodynamic response and typically increases flow deflection.
+This is the case for bedforms forming in open channel flows such as rivers~\citep{Fourriere2010, Unsworth2018}.
+This is also relevant for aeolian dunes as they evolve in the turbulent atmospheric boundary layer (ABL) capped by the stratified free atmosphere (FA) \citep{andreotti2009}.
+Two main mechanisms, associated with dimensionless numbers, must then be considered (Fig.~\ref{Fig4}).
+First, topographic obstacles typically disturb the flow over a characteristic height similar to their length.
+As flow confinement is characterised by a thickness $H$, the interaction between the dunes and the wind in the ABL is well captured by the parameter $k H$.
+The height $H$ is directly related to the radiative fluxes at the Earth surface.
+It is typically on the order of a kilometre, but significantly varies with the circadian and seasonal cycles.
+Emerging and small dunes, with wavelengths in the range $20$ to $100$~m, are not affected by the confinement, corresponding to $k H \gg 1$.
+For giant dunes with kilometric wavelengths, however, their interaction with the FA is significant \citep{andreotti2009}.
+This translates into a parameter $kH$ in the range $0.02$--$5$, depending on the moment of the day and the season.
+A second important mechanism is associated with the existence of a thin intermediate so-called capping layer between the ABL and the FA.
+It is characterised by a density jump $\Delta\rho$, which controls the `rigidity' of this interface, i.e.
+how much its deformation affects streamline compression.
+This is usually quantified using the Froude number \citep{Vosper2004, Stull2006, Sheridan2006, Hunt2006, Jiang2014}:
+%
+\begin{equation}
+\mathcal{F} = \displaystyle\frac{U}{\sqrt{\displaystyle\frac{\Delta\rho}{\rho_{0}}gH}},
+\label{FroudeNumber}
+\end{equation}
+%
+where $U$ is the wind velocity at the top of the ABL and $\rho_{0}$ its average density.
+The intensity of the stratification, i.e.
+the amplitude of the gradient $\left | \partial_z \rho \right|$, also impact ability to deform the capping layer under the presence of an underlying obstacle, and thus affects the influence of flow confinement.
+This can be quantified using the internal Froude number~\citep{Vosper2004, Stull2006, Sheridan2006, Hunt2006, Jiang2014} $\mathcal{F}_{\textup{I}} = k U/N$, where $N = \sqrt{-g \partial_z \rho / \rho_{0}}$ is the the Brunt-V\"ais\"al\"a frequency \citep{Stull1988}.
+Both Froude numbers have in practice the same qualitative effect on flow confinement, and we shall restrict the main discussion to $\mathcal{F}$ only.
 
-  Another way to increase the flow deflection is its confinement below a capping surface, that results in further streamline compression.
-This is the case for bedforms forming in open channel flows such as rivers~\citep{Fourriere2010, Unsworth2018}, but also for eolian dunes.
-These dunes evolve in the turbulent atmospheric boundary layer (ABL), which is capped by a transitional layer separating it from the stratified atmosphere above (see Fig.~\ref{Fig4}).
-Two different mechanisms control the possibility of this additional streamline compression.
+This theoretical framework in mind, the smallest wind disturbances are expected to occur during the day, when the ABL depth is the largest and comparable to the dune wavelength ($k H \gtrsim 1$), which correspond to a weak confinement situation (Fig.~\ref{Fig4}c,d).
+On the contrary, large wind disturbances are expected to occur during the night, when the confinement is mainly induced by a shallow ABL (Fig.~\ref{Fig4}e).
+However, this strong confinement can be somewhat reduced in the case of strong winds, corresponding to large values of the Froude number (Fig.~\ref{Fig4}f).
+This is qualitative agreement with the transition from deflected to non-deflected winds related to low and high velocities observed in the data (Sec.~\ref{section_data_feedback}).
 
-  On one hand, it depends if the flow disturbance induced by the underlying topography reach the surface.
-As obstacles typically disturb flow over a characteristic height similar to their length, the potential of interaction between the dunes and the overlying surface is well captured by the parameter $k H$, where $k = 2\pi/\lambda$ is the wavenumber and $H$ the ABL depth.
-Note that $H$ is directly related to the radiative fluxes at the Earth surface, and thus varies with the circadian and seasonal cycles.
-Here, the giant dunes have kilometric wavelengths, such that $0.02 \lesssim k H \lesssim 5$, and they interact most of the time with the capping layer and the stratified free atmosphere (FA) above \citep{andreotti2009}.
-Interestingly, the limit of no-interactions between the topography and the boundary layer structure ($k H \gg 1$), in which the properties of the overlying atmospheric structure are irrelevant, is never reached here, in the case of giant dunes.
+%---------------------------------------------
+\begin{figure}
+\centering
+\includegraphics[scale=1]{Figures/Figure5.pdf}
+\caption{Regime diagrams of the wind deviation $\delta_{\theta}$ (\textbf{a}) and relative attenuation/amplification $\delta_{u}$ (\textbf{b}) in the space $(\mathcal{F}, \, kH)$, containing the data from both the Deep Sea and South Namib stations.
+Blue dashed lines empirically delimit the different regimes.
+The point density in each bin of the diagrams is shown in Suppl.
+Fig.~\ref{Fig11_supp}.
+Similar regime diagrams in the spaces $(\mathcal{F}_{\textup{I}}, \, kH)$ and $(\mathcal{F}_{\textup{I}}, \, \mathcal{F})$ are shown in Suppl.
+Fig.~\ref{Fig12_supp}.
+\pc{ticks on x-axis ?}}
+\label{Fig5}
+\end{figure}
+%---------------------------------------------
 
-  On the other hand, it depends on rigidity of the capping surface, as its deformation releases the confinement effect inducing streamline compression.
-This is typically quantified using the Froude number~\citep{Vosper2004, Stull2006, Sheridan2006, Hunt2006, Jiang2014}:
-  \begin{equation}
-        \mathcal{F} = \displaystyle\frac{U}{\sqrt{\displaystyle\frac{\Delta\rho}{\rho}gH}},
-  \end{equation}
-  where $U$ is the wind velocity at the top of the ABL, $\rho$ its average density, $\Delta\rho$ the density jump between the ABL and the FA.
-
-  % Note that the ability of the capping layer and stratification to accommodate a perturbation induced by the topography directly impacts the strength of this confinement effect (Fig.~\ref{Fig4}).
-This is typically quantified using surface and internal Froude numbers
-  % \citep{Vosper2004, Stull2006, Sheridan2006, Hunt2006, Jiang2014}:
-  % \begin{equation}
-  %       \mathcal{F} = \displaystyle\frac{U}{\sqrt{\displaystyle\frac{\Delta\rho}{\rho}gH}}, \, \mathcal{F}_{\textup{I}} = \displaystyle\frac{k U}{N},
-  % \end{equation}
-  % where $U$ is the wind velocity at the top of the ABL, $\rho$ its average density, $\Delta\rho$ the density jump between the ABL and the FA and $N$ is the Brunt-Väisälä frequency, characteristic of the stratification.
-
-  The smallest wind disturbances are expected during the day, when the ABL depth is comparable to the dune wavelength ($k H \gtrsim 1$) and for large wind velocities, which correspond to a weak confinement situation (Fig.~\ref{Fig4}d).
-On the contrary, large wind disturbances are expected to occur during the night, when the confinement is mainly induced by shallow ABL (Fig.~\ref{Fig4}e--f).
-Note that this strong confinement can be somewhat reduced in the case of strong winds (corresponding to large Froude numbers, see Fig.~\ref{Fig4}f), explaining the transition from deflected to non-deflected winds related to low and high velocities observed in the data (see section~\ref{section_data_feedback}).
-
+%------------------------
 \subsection{Flow regime diagrams}
-
-  To highlight these different regimes from our data, we compute wind disturbance diagrams in the space defined by the two relevant non-dimensional numbers presented above, $\left(k H,\, \mathcal{F}\right)$.
-Those are calculated from the time series of the geopotential, temperature and specific humidity vertical profiles available in the ERA5 climate reanalysis (see SI section 2).
-  %
-  Flow deviation is computed as the minimal angle between the wind orientations from the local measurements, and the regional predictions. The relative velocity modulation is computed as
+We can go one step further and analyse how our data quantitatively spread over the different regimes discussed above.
+For that purpose, one needs to compute $kH$ and $\mathcal{F}$ from the time series.
+$H$, $U$ and the other atmospheric parameters can be deduced from the various vertical profiles (temperature, humidity) available in the ERA5 climate reanalysis (Suppl.
+Mat.
+section 2).
+We quantify the flow deflection $\delta_{\theta}$ as the minimal angle between the wind orientations comparing the local measurements and the regional predictions.
+We also compute the relative velocity modulation as
 %
@@
 \end{equation}
+%
+These two quantities are represented as maps in the plane $(\mathcal{F}, \, kH)$ (Fig.~\ref{Fig5}a,b), and one can clearly identify different regions in these graphs.
+Small wind disturbances (small $\delta_{\theta}$ and $\delta_{\textup{u}}$) are located in the top-right part of the diagrams, corresponding to a regime with low-interaction as well as low-confinement ($k H$ and $\mathcal{F}$ large enough, Fig.~\ref{Fig4}d).
+Lower values of $k H$ (stronger interaction) or of Froude number (stronger confinement) both lead to an increase in wind disturbances, both in terms of orientation and velocity.
+Below a crossover value $k H \simeq 0.3$, wind disturbance is less sensitive to the $\mathcal{F}$-value, probably due to enhanced non-linear effects linked to strong flow modulation by the obstacle when confinement is strong \pc{more precise arguments?}.
+The Froude number also controls a transition from damped to amplified wind velocities in the interdune, with a crossover around $\mathcal{F} \simeq 0.4$ (Fig.~\ref{Fig5}b).
+Such an amplification is rather unexpected.
+Checking the occurrence of the corresponding data, it appears that they are mainly associated with the wind coming from the South, when it blows at the end of day, but in summer time only although it is also present in winter \pc{an additional suppl.
+fig.
+?}.
+This effect may be linked to a change in the flow behaviour in the lee side of the obstacle (lee waves, hydraulic jumps, rotors) but further measurements are clearly needed in order to assess the different possibilities \citep{baines1995, Vosper2004}.
 
-  When representing the two variables $\delta_{\theta}$ and $\delta_{\textup{u}}$ in this space, different regime emerges (Fig.~\ref{Fig5}).
-Small wind disturbances ($\delta_{\theta} \to 0$, $\delta_{\textup{u}} \to 0$) are located in the top-right part of the diagrams, corresponding to a regime mixing low-interaction and low-confinement ($k H$ and $\mathcal{F}$ large enough, Fig.~\ref{Fig4}d).
-  %
-  Lower values of $k H$ (stronger interaction) or Froude number (stronger confinement) then both lead to an increase in wind disturbances, both in terms of orientation and velocity.
-Below a threshold value of $k H \simeq 0.3$, wind disturbance occurs independently of the Froude number value, probably due to enhanced non-linear effects linked to strong flow modulation by the obstacle in this part of the diagram.
-The Froude number also controls a transition from damped to amplified wind velocities in the interdune, with a transition at $\mathcal{F} \simeq 0.4$ (Fig.~\ref{Fig5}b).
-This may be linked to a transition in the flow regime in the lee side of the obstacle (lee waves, hydraulic jumps, rotors) but further measurements are needed in order to assess this \citep{baines1995, Vosper2004}.
-
-  %
-  % Furthermore, this also seems to control a transition between from damped to amplified wind velocities within the interdune (Fig.~\ref{Fig5}b), for which we do not have an explanation.
-Note that the same interpretation can be done with the diagrams including the internal Froude number $\mathcal{F}_{\textup{I}}$, as shown by Fig.~\ref{Fig12_supp}.
-
-  \begin{figure}
-    \centering
-    \includegraphics[scale=1]{Figures/Figure5.pdf}
-    \caption{Regime diagrams of the wind deviation $\delta_{\theta}$ and relative attenuation/amplification $\delta_{u}$ in the space $(\mathcal{F}, \, kH)$, containing the data from both the Deep Sea and South Namib stations.
-Blue dashed lines empirically delimit the different regimes.
-The point density in each bin of the diagrams is shown in Fig.~\ref{Fig11_supp}.
-The regime diagrams in the spaces $(\mathcal{F}_{\textup{I}}, \, kH)$ and $(\mathcal{F}_{\textup{I}}, \, \mathcal{F})$ are shown in Fig.~\ref{Fig12_supp}.}
-    \label{Fig5}
-  \end{figure}
-
-  \subsection{On the influence of the stratification of the free atmosphere}
-
-  The presence of a stratification in the free atmosphere can also impact the flow confinement, depending on its ability to deform under the presence of an underlying obstacle.
-This can be quantified using the internal Froude number~\citep{Vosper2004, Stull2006, Sheridan2006, Hunt2006, Jiang2014}:
-  \begin{equation}
-    \mathcal{F}_{\textup{I}} = \displaystyle\frac{k U}{N},
-  \end{equation}
-  where $N = \sqrt{-(g/\rho)(\partial \rho/\partial z)}$ is characteristic of the stratification.
-Both Froude numbers have the same qualitative effect on flow confinement, as they quantify the rigidity of the overlying layers.
-This is confirmed by figure~\ref{Fig12_supp}, where we can also find the different regimes related to wind disturbances described previously for the Froude number $\mathcal{F}$.
-
-\section{Discussion}
-
+%---------------------------------------------
 \begin{figure}
@@
 \includegraphics[scale=1]{Figures/Figure6.pdf}
-    \caption{Implications for smaller scale patterns in (a) the South Namib and (b) Deep Sea.
-The ellipses indicates the diffe  rent types of elongating dunes, at large (plain) and small scale (dashed).
-The dune orientations are calculated using the model of \citet{Courrech2014} from the sand flux angular distributions, shown here for typical sand quartz grains of $180~\mu$m.
-The double blue and single red arrows correspond to the two possible dune growth mechanisms, bed instability and elongation, respectively.
-Likewise, plain arrows are calculated from the ERA5-Land data, and dashed arrows from the local measurements.
-Wedges show the uncertainty on the orientation calculation, and the arrows correspond to typical parameters found in the literature, i.e a grain diameter of $180~\mu$m and a flux-up ratio of 1.6.
-The black dots indicate the position of the measurement stations.
+\caption{Implications for smaller scale patterns in (\textbf{a}) the South Namib and (\textbf{b}) Deep Sea.
+The ellipses indicate the different types of elongating dunes, at large (plain line) and small (dashed line) scales.
+Dune orientation are predicted using the model of \citet{Courrech2014} from the sand flux angular distributions, shown here (roses and resultant transport direction) for typical values (grain size $180~\mu$m, flux-up ratio of $1.6$).
+Code for corresponding arrows: use of ERA5-Land data (plain line), use of local measurements (dashed line), prediction for the bed instability growth mechanism (blue), prediction for the elongation growth mechanism (red).
+Wedges show the uncertainty on the orientation calculation when parameters are varied.
+The black dots indicate the location of the meteorological stations with respect to the dunes.
 See Appendix 2 for additional details.}
@@
 \end{figure}
+%---------------------------------------------
 
- The comparison of local (direct measurements) and regional (climate reanalysis) wind data reveals the giant dunes feedback on the flow.
-In flat areas, the matching between measurements and prediction confirms the ability of the ERA5Land climate reanalysis to predict the wind flow down to scales $\sim 10~\textrm{km}$, i.e the grid model.
-When smaller scale topographies are present (giant dunes in our case), locally measured wind regimes can significantly differ from the regionally predicted ones.
-Furthermore, we link these disturbances induced by the dunes to their interaction with the lower part of the atmospheric vertical structure, and more specifically to its circadian variability.
-During the night, the presence of a shallow atmospheric boundary layer (ABL) induces a strong confinement of the flow, associated with large wind deviation and acceleration or deceleration.
-During the day, the capping layer is high enough to prevent its interaction with the giant dunes, resulting in a low confinement of the flow, and thus smaller wind disturbances.
-Interestingly, we also found that this effect could be counterbalanced by the presence of large wind velocities, capable of deforming the capping layer and/or the FA stratification, thus decreasing the confinement effect.
+%---------------------------------------------------------------------------------------
+\section{Discussion}
+The comparison of local (direct measurements) and regional (climate reanalysis) wind data reveals the giant dune feedback on the wind flow.
+In flat areas, the agreement between the two confirms the ability of the ERA5-Land climate reanalysis to predict the wind regime down to scales $\sim10$~km, i.e the model grid.
+When smaller scale topographies are present (giant dunes in our case), locally measured winds can significantly differ from the regionally predicted ones.
+This is the case when the disturbances induced by the dunes interact with the lower part of the ABL vertical structure, which presents circadian variations.
+During the day, when the capping layer is typically high, this interaction is small, and the ERA5-Land predictions are also quantitatively consistent with the local data.
+During the night, however, the presence of a shallow atmospheric boundary layer induces a strong confinement of the flow, and is associated with large wind deflection by the dunes.
+Importantly, we find that this effect can be counterbalanced for large wind velocities, which are capable of deforming the capping layer, thus decreasing the influence of the confinement.
 
- Simple linear models also suggest that larger wind disturbances occur under strong flow confinement such as described above \citet{andreotti2009, Andreotti2012}.
-However, they are unable to reproduce the magnitude of the observed deviations, probably due to the presence of hydrodynamical non-linear effects, all the more present in high confinement situations linked to strong flow modulations (see Fig.~\ref{Fig12_supp} and Appendix 1).
-They also predict different spatial flow structures such as lee waves and rotors~\citep{baines1995, Vosper2004}, which are likely to be complicated by these non-linearities, and which cannot be observed by our single point measurements.
-Measurements in different places along and across the ridge are then needed in order to properly map these flow structures, and allow further comparisons with models.
+The theoretical computation of the wind disturbances induced by sinusoidal ridges under flow confinement has been performed in the linear limit \citep{andreotti2009, Andreotti2012}, i.e.
+when the aspect ration of these ridges is small ($k\xi_0 \ll 1$).
+These models are able to qualitatively reproduce the observed wind deflection (Appendix 1, Suppl.
+Figs.~\ref{Fig12_supp} and \ref{Fig13_supp}), and thus provide the physical support for the hydrodynamic picture we propose here.
+They however cannot reach the magnitude of these observations, probably due to the presence of expected non-linearities in high confinement situations linked to strong flow modulations.
+Besides, these linear calculations only predict wind attenuation in the interdune, in contrast with the observed enhanced velocities associated with particular summer evening winds from the South.
+Some other models also predict different spatial flow structures such as lee waves and rotors~\citep{baines1995, Vosper2004}, which of course cannot be observed by our single point measurements.
+Data at different places along and across the ridges are needed to investigate and possibly map such flow structures, and for further comparisons with the models.
 
 This study highlights the interaction between giant dunes and the atmospheric boundary layer.
-It then supports the debated idea that the capping layer acts as a bounding surface limiting dune growth~\citep{andreotti2009}, as opposed to an unconstrained growth ever-slower with size \citep{Eastwood2011, gunn2021}.
-Once validated, this mechanism would then allow inference of the ABL depth from the giant bedforms spacing where measurements are not feasible or available, as performed by~\citet{Lorenz2010} on Titan.
+It then supports the debated idea that the capping layer acts as a bounding surface impacting the dune dynamics, and limiting dune growth in particular~\citep{andreotti2009}, as opposed to an unconstrained growth ever-slower with size \citep{Eastwood2011, gunn2021}.
+Interestingly, this mechanism would allow for the inference of the ABL depth from the giant bedforms spacing where measurements are not feasible or available, as e.g.
+performed by~\citet{Lorenz2010} on Titan.
 
- This interaction also have strong implications for smaller scales bedforms, as illustrated in Fig.~\ref{Fig6}.
-In the Namib Sand Sea, small linear dunes ($\sim 50$ m -wide) are present in the interdune between giant linear dunes ($\sim 2$ km -wide).
-While differences between larger and smaller scale dune patterns are observed ubiquitously, they are now largely attributed to the presence of two different dune growth mechanisms, leading to two different dune patterns (orientations and/or morphologies) for the same wind regime~\citep{Courrech2014, Runyon2017, lu2017, Song2019, Hu2021}.
-Here, coupling sediment transport and dune growth models, we show that the orientations of the small and giant linear dunes can be predicted from the same dune growth mechanism, using the locally measured and regionally predicted winds, respectively (red arrows in Fig.~\ref{Fig6}).
-The giant dune feedback on the flow described in this study then provides a mechanism for the existence of these small linear dunes elongating across the interdune, as yet unresolved.
-While further studies are needed, these dune type could provide additional strong constraints for the inference of local winds from bedforms, as currently performed on Mars using ripple orientations \citep{Liu2015, Hood2021}.
+This interaction also has important implications for smaller scales bedforms, as illustrated in Fig.~\ref{Fig6}.
+In the Namib Sand Sea, small linear dunes ($\sim50$~m -wide) are present in the interdune between giant linear dunes ($\sim2$~km -wide).
+Strangely enough, the small dunes do not have the same orientation as the large ones, and sometimes denotes as `crossing dunes'.
+While differences between large and small scale dune patterns are observed ubiquitously, they are now largely attributed to the presence of two different dune growth mechanisms, leading to two different dune patterns (orientations and/or morphologies) for the same wind regime~\citep{Courrech2014, Runyon2017, lu2017, Song2019, Gadal2020, Hu2021}.
+Here, however, we can get the orientations of the small and giant linear dunes from the same dune growth mechanism (elongating mode), respectively using the locally measured and regionally predicted winds (red arrows in Fig.~\ref{Fig6}).
+The feedback of the giant dunes on the wind described in this study then provides an explanation for the existence of these small linear dunes elongating across the interdune, as yet unresolved.
+Further studies are of course needed to confirm this hypothesis.
+These crossing dunes could provide additional constraints for the inference of local winds from bedforms, similarly to what is currently performed on Mars using ripple orientations \citep{Liu2015, Hood2021}.
 
+%---------------------------------------------------------------------------------------
 \begin{acknowledgements}
-[citing all grants ...]\\
-Finally, we would like to acknowledge the contributors of the following open-source python librairies, Matplotlib \citep{Hunter2007}, Numpy \citep{Harris2020} and Scipy \citep{Virtanen2020}, which provide an incredibly efficient ecosystem allowing scientific research in Python.
+We would like to acknowledge the contributors of the following open-source python librairies, Matplotlib \citep{Hunter2007}, Numpy \citep{Harris2020} and Scipy \citep{Virtanen2020}, which provide an incredibly efficient ecosystem allowing scientific research in Python.
+
+ERA5 and ERA5-Land datasets are publicly available at the Copernicus Climate Change Service (C3S) Climate Data Store.
+The locally measured wind data can be found at \cg{upload on public data repository}.
+The digital elevation models from the Shuttle Radar Topography Mission are publicly available from Nasa servers, and can be downloaded at https://dwtkns.com/srtm30m/.
+Fully documented codes used to analyse this study are available at https://github.com/Cgadal/GiantDunes (will be made public upon acceptance of this manuscript for publication).
+
+[citing all grants ...] \pc{TOAD}
 \end{acknowledgements}
 
-\section*{Appendix 1: ABL turbulent wind model}
+%---------------------------------------------------------------------------------------
+\section*{Appendix 1: Linear theory of wind response to topographic perturbation}
 \label{turbulent_wind_model}
 
-Following the work of \citet{Fourriere2010}, \citet{Andreotti2012} anc \citet{andreotti2009}, we briefly expose in this section the linear response of a turbulent flow to a small aspect ratio perturbation of the underlying topography.
-As this topography can be decomposed into several sinusoidal modes, we focus on a sinusoidal topography:
+Following the work of \citet{Fourriere2010}, \citet{Andreotti2012} and \citet{andreotti2009}, we briefly expose in this appendix the framework of the linear response of a turbulent flow to a topographic perturbation of small aspect ratio.
+As a general bed elevation can be decomposed into Fourier modes, we focus here on a sinusoidal topography:
+%
 \begin{equation}
@@
 \end{equation}
-which is also a good approximation for the giant dunes observed in the Deep Sea and South Namib Station (see Fig~\ref{Fig1} and Fig~\ref{Fig4_supp}).
-Here, $x$ and $y$ are the streamwise and spanwise coordinates, $k=2\pi/\lambda$ the wavenumber of the sinusoidal perturbation, and $\alpha$ its crest orientation, calculated with respect to the $y$--direction.
-
-The two components of the basal shear stress $\boldsymbol{\tau} = \rho u_{*}\boldsymbol{u}_{*}$, constant in a flat bottom situation, can then be written without loss of generality as:
+%
+which is also a good approximation for the giant dunes observed in the Deep Sea and South Namib Station (Fig.~\ref{Fig1} and Suppl.
+Fig.~\ref{Fig4_supp}).
+$x$ and $y$ are the streamwise and spanwise coordinates, $k=2\pi/\lambda$ the wavenumber of the sinusoidal perturbation, and $\alpha$ its crest orientation with respect to the $y$-direction.
+The two components of the basal shear stress $\boldsymbol{\tau} = \rho_{0} u_{*}\boldsymbol{u}_{*}$, constant in the flat bottom reference case, can then be generically written as:
+%
 \begin{align}
 \tau_{x} & = \tau_{0}\left(1 + k\xi_{0}\sqrt{\mathcal{A}_{x}^{2} + \mathcal{B}_{x}^{2}}\cos\left[k\left(\cos(\alpha)x + \sin(\alpha)y\right) + \phi_{x}\right]\right), \\
-  \tau_{y} & = \tau_{0}k\xi_{0}\sqrt{\mathcal{A}_{y}^{2} + \mathcal{B}_{y}^{2}}\cos\left[k\left(\cos(\alpha)x + \sin(\alpha)y\right) + \phi_{y}\right],
+\tau_{y} & = \tau_{0} \, k\xi_{0}\sqrt{\mathcal{A}_{y}^{2} + \mathcal{B}_{y}^{2}}\cos\left[k\left(\cos(\alpha)x + \sin(\alpha)y\right) + \phi_{y}\right],
 \end{align}
-where $\tau_{0}$ is the basal shear stress on a flat bed, and $\phi_{x, y} = \tan^{-1}\left(\mathcal{B}_{x, y}/\mathcal{A}_{x, y}\right)$.
-The in--phase and in--quadrature hydrodynamical coefficients $\mathcal{A}_{x, y}$ and $\mathcal{B}_{x, y}$ are functions of the flow conditions, i.e the bottom roughness, the vertical flow structure or the incident flow direction \citep{Fourriere2010, andreotti2009, Andreotti2012}.
+%
+where $\tau_{0}$ is the reference basal shear stress on a flat bed.
+We have defined the phase $\phi_{x, y} = \tan^{-1}\left(\mathcal{B}_{x, y}/\mathcal{A}_{x, y}\right)$ from the in-phase and in-quadrature hydrodynamical coefficients $\mathcal{A}_{x, y}$ and $\mathcal{B}_{x, y}$.
+They are functions of $k$ and of the flow conditions, i.e the bottom roughness, the vertical flow structure and the incident flow direction, and the theoretical framework developed in the above cited papers proposes methods to compute them in the linear regime.
 
-Following \citet{Andreotti2012}, the impact of the incident wind direction can be approximated by the following expressions:
+Following \citet{Andreotti2012}, the effect of the incident wind direction can be approximated by the following expressions:
+%
 \begin{align}
@@
 \end{align}
-where $\mathcal{A}_{0}$ and $\mathcal{B}_{0}$ are now two coefficients independent of the dune orientation $\alpha$.
-In the case of a fully turbulent boundary layer capped by a stratified atmosphere, they now only depend on $k H$, $k z_{0}$, $\mathcal{F}$ and $\mathcal{F}_{\textup{I}}$ \citet{andreotti2009}.
-In this study, we assume a constant hydrodynamic roughness $z_{0} \sim 1~\textup{mm}$, leading to a constant value of $k z_{0} \sim 10^{-6}$.
-Measured values of $z_{0}$ in the field indeed reports a variation of $z_{0}$ between 0.1 mm and 10 mm \citep{Sherman2008, Field2018}, but $\mathcal{A}_{0}$ and $\mathcal{B}_{0}$ does not vary much in the corresponding range of $k z_{0}$ \citep{Fourriere2010}.
 %
-Note that the linearity assumption of this theoretical framework requires $\left(\vert \tau \vert - \tau_{0}\right)/\tau_{0} \ll 1$, which is satisfied by $k\xi\sqrt{\mathcal{A}_{0}^{2} + \mathcal{B}_{0}^{2}} \ll 1$.
-In our case, the giant dune morphology gives $k\xi \simeq 0.1$, setting the upper bound of the coefficient modulus $\sqrt{\mathcal{A}_{0}^{2} + \mathcal{B}_{0}^{2}}$ to 10.
+where $\mathcal{A}_{0}$ and $\mathcal{B}_{0}$ are now two coefficients independent of the dune orientation $\alpha$, corresponding to the transverse case ($\alpha=0$).
+For a fully turbulent boundary layer capped by a stratified atmosphere, these coefficients depend on $k H$, $k z_{0}$, $\mathcal{F}$ and $\mathcal{F}_{\textup{I}}$ \citep{andreotti2009}.
+In this study, we assume a constant hydrodynamic roughness $z_{0} \simeq 1$~mm (Suppl.
+Mat.
+section 1).
+For the considered giant dunes, this leads to $k z_{0} \simeq 2 \cdot 10^{-6}$, as their wavelength is $\lambda \simeq 2.4$~km (or $k \simeq 2 \cdot 10^{-3}$~m$^{-1}$).
+Values of $z_{0}$ extracted from field data indeed typically fall between 0.1 mm and 10 mm \citep{Sherman2008, Field2018}.
+Importantly, $\mathcal{A}_{0}$ and $\mathcal{B}_{0}$ do not vary much in the corresponding range of $k z_{0}$ \citep{Fourriere2010}, and the results presented here are robust with respect to this choice.
 
-Additionlly, we also calculate the time series of the hydrodyamical coefficients from the time series of the non-dimensional numbers used in this study.
-The results, shown Fig.~\ref{Fig12_supp} under the similar form of the regime diagrams presented in Fig.~\ref{Fig5} and Fig.~\ref{Fig12_supp}, exhibit a qualitative matching with the observations presented in this study.
-%
-Small values of these coefficients ($\mathcal{A}_{0} \simeq 3.4$ and $\mathcal{B}_{0} \simeq 1$) are found in low confinement cases, i.e for $k H \gg 1$ (no interaction between the dunes and the capping layer) or for $k H \gg 1$ but large enough Froude numbers (reduced flow confinement due to the deformation of the olverlying capping layer and stratification).
-In contrast, larger values are obtained in high confinement cases (small $k H$ and Froude numbers).
-However, not that most of this part of the diagrams are outside of the linear limit discussed above ($\sqrt{\mathcal{A}_{0}^{2} + \mathcal{B}_{0}^{2}} \gtrsim 10$), which does not allow further quantitative comparison with the data.
+With capping a layer height and Froude numbers computed from the ERA5-Land time series, the corresponding $\mathcal{A}_{0}$ and $\mathcal{B}_{0}$ can be deduced, allowing to produce maps as displayed in Suppl.
+Fig.~\ref{Fig13_supp}.
+Interestingly, it shows similar regimes as in the diagrams of Fig.~\ref{Fig5} and Fig.~\ref{Fig12_supp}a,b, supporting the physics picture.
+However, the matching remains qualitative only.
+As a matter of fact, the linearity assumption of the theoretical framework requires $\left(\vert \tau \vert - \tau_{0}\right)/\tau_{0} \ll 1$, which translates into $k\xi\sqrt{\mathcal{A}_{0}^{2} + \mathcal{B}_{0}^{2}} \ll 1$.
+In our case, the giant dune morphology gives $k\xi_0 \simeq 0.1$, which means that one quits the regime of validity of the linear theory when the coefficient modulus $\sqrt{\mathcal{A}_{0}^{2} + \mathcal{B}_{0}^{2}}$ becomes larger than a few units.
+In accordance with the theoretical expectations, these coefficients present values on the order of unity ($\mathcal{A}_{0} \simeq 3$ and $\mathcal{B}_{0} \simeq 1$) in unconfined situations \citep{Claudin2013, Lu2021}.
+In contrast and as illustrated in Suppl.
+Fig.~\ref{Fig12_supp}a,b, larger values are predicted in case of strong confinement, which does not allow us to proceed to further quantitative comparison with the data.
 
+Finally, the linear model is also able to reproduce the enhancement of the flow deflection over the sinusoidal ridges when $\sqrt{\mathcal{A}_{0}^{2} + \mathcal{B}_{0}^{2}}$ is increased (Suppl.
+Fig.~\ref{Fig13_supp}).
+Here, using $k\xi_0 \simeq 0.1$ to be representative of the amplitude of the giant dunes at the Deep Sea station, the coefficient modulus is bounded to $10$.
+
+%---------------------------------------------------------------------------------------
 \section*{Appendix 2: Sediment transport and dune morphodynamics}
 
-Here, we briefly describe the sediment transport and dune morphodynamics theoretical framework leading to the prediction of sand fluxes and dune orientations from wind data.
+We summarise in this appendix the sediment transport and dune morphodynamics theoretical framework leading to the prediction of sand fluxes and dune orientations from wind data.
 
-\paragraph{Sediment transport}
-The prediction of sand fluxes from wind data has been a long standing issue in geomorphological studies~\citep{Fryberger79, Pearce2005, Sherman2012, Shen2019}.
-Based on laboratory studies in wind tunnels \citep{Rasmussen96, Iversen99, Creyssels2009, Ho2011}, as well as physical considerations \citep{Ungar1987, Andreotti2004bis, Duran2011, Pahtz2020}, it has been shown that the steady saturated sand flux over a flat sand bed depends linearly on the shear stress:
+%------------------------
+\paragraph{Sediment transport} ---
+The prediction of sand fluxes from wind data has been a long standing issue in aeolian geomorphological studies~\citep{Fryberger79, Pearce2005, Sherman2012, Shen2019}.
+Based on laboratory studies in wind tunnels \citep{Rasmussen96, Iversen99, Creyssels2009, Ho2011}, as well as physical considerations \citep{Ungar1987, Andreotti2004bis, Duran2011, Pahtz2020}, it has been shown that the steady saturated saltation flux over a flat sand bed depends linearly on the shear stress:
+%
 \begin{equation}
 \label{transport_law_quadratic}
-  \frac{q_{\textup{sat, t}}}{Q} = \Omega\sqrt{\Theta_{\textup{th}}}\left(\Theta_{t} - \Theta_{\textup{th}}\right),
+\frac{q_{\textup{sat}}}{Q} = \Omega\sqrt{\Theta_{\textup{th}}}\left(\Theta - \Theta_{\textup{th}}\right),
 \end{equation}
-where $\Omega$ is a proportionality constant, $Q = d\sqrt{(\rho_{\textup{s}} - \rho)gd/\rho}$ is a characteristic flux, $\Theta = \rho u_{*, t}^{2}/(\rho_{\textup{s}} - \rho)gd$ the Shields number, and $\Theta_{\textup{th}}$ its threshold value for incipient sediment transport.
-Here, $\rho_{\textup{s}} = 2.6~\textup{g}~\textup{cm}^{-3}$ and $d=180~\mu\textup{m}$ are the grain density and diameter, and $g$ is the gravitational acceleration.
+%
+where $\Omega$ is a proportionality constant, $Q = d\sqrt{(\rho_{\textup{s}} - \rho_{0})gd/\rho_{0}}$ is a characteristic flux, $\Theta = \rho_{0} u_{*}^{2}/(\rho_{\textup{s}} - \rho_{0})gd$ the Shields number, and $\Theta_{\textup{th}}$ its threshold value below which saltation vanishes.
+$\rho_{\textup{s}} = 2.6~\textup{g}~\textup{cm}^{-3}$ and $d=180~\mu\textup{m}$ are the grain density and diameter, and $g$ is the gravitational acceleration.
+The shear velocity, and consequently the Shields number as well as the sediment flux, are time dependent.
 
-Recently, \citet{Pahtz2020} suggested a quadratic dependency on the shear stress by taking into account grain--grain interactions within the transport layer, performing better at reproducing laboratory data at high wind velocities:
+Recently, \citet{Pahtz2020} suggested an additional quadratic term in Shields to account for grain-grain interactions within the transport layer at strong wind velocities:
+%
 \begin{equation}
 \label{transport_law_quartic}
-  \frac{q_{\textup{sat, t}}}{Q} = \frac{2\sqrt{\Theta_{\textup{th}}}}{\kappa\mu}\left(\Theta_{t} - \Theta_{\textup{th}}\right)\left(1 +\frac{C_{\textup{M}}}{\mu}\left[\Theta_{t} - \Theta_{\textup{th}}\right]\right),
+\frac{q_{\textup{sat, t}}}{Q} = \frac{2\sqrt{\Theta_{\textup{th}}}}{\kappa\mu}\left(\Theta - \Theta_{\textup{th}}\right)\left(1 +\frac{C_{\textup{M}}}{\mu}\left[\Theta - \Theta_{\textup{th}}\right]\right),
 \end{equation}
-where $\kappa = 0.4$ is the von Kármán constant, $C_{\rm M}=1.7$ a constant and $\mu$ a friction coefficient, taken to be the avalanche slope of the granular material, i.e $\sim 0.6$.
+%
+where $\kappa = 0.4$ is the von K\'arm\'an constant, $C_{\rm M} \simeq 1.7$ a constant and $\mu \simeq 0.6$ is a friction coefficient, taken to be the avalanche slope of the granular material.
 The fit of this law to the experimental data of \citet{Creyssels2009} and \citet{Ho2011} gives $\Theta_{\textup{th}} = 0.0035$.
-The sand flux angular distributions and the dune orientations in Fig.~\ref{Fig6} are calculated using this quartic law \eqref{transport_law_quartic}.
-However, we verified that using the quadratic law \eqref{transport_law_quadratic} instead did not change the predicted dune orientations by more than a few degrees.
+The fit of Eq.~\ref{transport_law_quadratic} on these same data similarly gives $\Omega \simeq 8$ and $\Theta_{\textup{th}} = 0.005$.
+The sand flux angular distributions and the dune orientations in Fig.~\ref{Fig6} are calculated using this law \eqref{transport_law_quartic}.
+We have checked that using the ordinary linear relationship \eqref{transport_law_quadratic} instead does not change the predicted dune orientations by more than a few degrees.
 
-\paragraph{Dune orientations}
+%------------------------
+\paragraph{Dune orientations} ---
+Dune orientations are predicted with the dimensional model of \citet{Courrech2014}, from the sand flux time series computed with the above transport law.
+Two orientations are possible depending on the mechanism dominating the dune growth: elongation or bed instability.
+The orientation $\alpha$ corresponding the bed instability is then the one that maximises the following growth rate \citep{Rubin1987}:
+%
+\begin{equation}
+\sigma \propto \frac{1}{H_{d} W_{d} T}\int_0^T  q_{\textup{crest}}\vert \sin\left(\theta - \alpha\right) \vert \, \textup{d}t,
+\end{equation}
+%
+where $\theta$ is the wind orientation measured with respect to the same reference as $\alpha$, and $H_{d}$ and $W_{d}$ are dimensional constants respectively representing the dune height and width.
+The integral runs over a time $T$, which must be representative of the characteristic period of the wind regime.
+The flux at the crest is expressed as:
+%
+\begin{equation}
+q_{\textup{crest}} = q_{\textup{sat}}\left[1 + \gamma\vert\sin\left(\theta - \alpha\right)\vert\right],
+\end{equation}
+%
+where the flux-up ratio $\gamma$ has been calibrated to $1.6$ using field studies, underwater laboratory experiments and numerical simulations.
+Predictions of the linear analysis of \citet{Gadal2019} give similar results.
 
-The dune orientations are predicted from the computed sand flux time series, using the dimensional model of \citet{Courrech2014}.
-Two orientations are possible depending on the mechanism dominating the dune growth: elongation or bed instability (the latter is also known as the rule of maximum gross bedform-normal transport from \citet{Rubin1987}).
+Similarly, the dune orientation corresponding to the elongation mechanism is the one that verifies:
+%
+\begin{equation}
+\tan(\alpha) = \frac{\langle q_{\textup{crest}}(\alpha) \boldsymbol{e}_{\theta}\rangle \cdot \boldsymbol{e}_{WE} }{ \langle q_{\textup{crest}}(\alpha) \boldsymbol{e}_{\theta} \rangle \cdot \boldsymbol{e}_{SN}},
+\end{equation}
+%
+where $\langle.\rangle$ denotes a vectorial time average.
+The unitary vectors $\boldsymbol{e}_{WE}$, $\boldsymbol{e}_{SN}$ and $\boldsymbol{e}_{\theta}$ are in the West-East, South-North and wind directions, respectively.
 
-The orientation $\alpha$ corresponding the bed instability is then the one that maximizes the following growth rate:
-\begin{equation}
-  \sigma \propto \frac{1}{H_{d} W_{d} T}\int_{t}  q_{\textup{crest}, t}\vert \sin\left(\theta_{t} - \alpha\right) \vert,
-\end{equation}
-where $H_{d}$ and $W_{d}$ are dimensional constants representing the dune height and width, respectively. The flux at the crest is expressed as:
-\begin{equation}
-  q_{\textup{crest}, t} = q_{\textup{sat}, t}\left[1 + \gamma\vert\sin\left(\theta_{t} - \alpha\right)\vert\right],
-\end{equation}
-where the flux-up ratio $\gamma$ has been calibrated to 1.6 using field studies, underwater laboratory experiments and numerical simulations.
-Similarly, the dune orientation corresponding to the elongation mechanism is the one that verifies:
-\begin{equation}
-  \tan(\alpha) = \frac{\langle q_{\textup{crest}, t}(\alpha) \boldsymbol{e}_{\theta_{t}}\rangle \cdot \boldsymbol{e}_{WE} }{ \langle q_{\textup{crest}, t}(\alpha) \boldsymbol{e}_{\theta_{t}} \rangle \cdot \boldsymbol{e}_{SN}},
-\end{equation}
-where $\langle.\rangle$ denotes a vectorial time average.
-The unitary vectors $\boldsymbol{e}_{WE}$, $\boldsymbol{e}_{SN}$ and $\boldsymbol{e}_{\theta_{t}}$ are in the West--East, South--North and wind direction, respectively.
-
-The resulting computed dune orientations, blue and red arrows in figure~\ref{Fig6}, are then depending on a certain number of parameters (grain properties, flux-up ratio), for which we took typical values for eolian desert on Earth.
-Due to the lack of measurements in the studied places, significant uncertainties can however be expected.
-We therefore run a sensibility test by calculating the dune orientations for grain diameters ranging from $100~\mu\textup{m}$ to $400~\mu\textup{m}$ and the speed-up ratio from $0.1$ to $10$ (wedges on figure~\ref{Fig6}).
+The resulting computed dune orientations, blue and red arrows in Fig.~\ref{Fig6}, then depend on a certain number of parameters (grain properties, flux-up ratio, etc.), for which we take typical values for aeolian sandy deserts.
+Due to the lack of measurements in the studied places, some uncertainties can be expected.
+We therefore run a sensibility test by calculating the dune orientations for grain diameters ranging from $100~\mu\textup{m}$ to $400~\mu\textup{m}$ and for a speed-up ratio between $0.1$ and $10$ (wedges in Fig.~\ref{Fig6}).
 
@@
 \bibliographystyle{spbasic_updated}
-\bibliography{biblio}
+\bibliography{main}
+
+%---------------------------------------------------------------------------------------
+%---------------------------------------------------------------------------------------
 
 \newpage
+
 \renewcommand{\thefigure}{S\arabic{figure}}
@@
 
-\section*{Supplementary Material for \textit{Boundary-Layer Meteorology} Sample Paper: Instructions for Authors}
+%---------------------------------------------------------------------------------------
+\begin{center}
+\textbf{\large
+Local wind regime induced by giant linear dunes \\
+--- Supplementary Material ---
+}
+\end{center}
 
-{\textbf{First Author* $\cdot$ Second Author $\cdot$ Third Author \\}}
-\\
-\text{*}Affiliation and email address for the corresponding author only (note that the corresponding author does not need to be the first author).
+\noindent
+\textbf{C. Gadal$^{*}$ $\cdot$ P. Delorme $\cdot$ C. Narteau $\cdot$ G.F.S. Wiggs $\cdot$ M. Baddock $\cdot$ J.M. Nield $\cdot$ P. Claudin}
+\\ \\
+$^{*}$ Institut de M\'ecanique des Fluides de Toulouse, Universit\'e de Toulouse Paul Sabatier, CNRS, Toulouse INP-ENSEEIHT, Toulouse, France.\\
+\texttt{cyril.gadal@imft.fr}
 
-\begin{figure}
+%---------------------------------------------------------------------------------------
+\section*{1. Shear velocity and calibration of the hydrodynamical roughness}
+\label{calib_z0}
+
+As the regionally predicted and locally measured velocities are available at different heights, we can not compare them directly.
+We then convert all velocities into shear velocities $u_{*}$, characteristic of the turbulent velocity profile \citep{Spalding1961, Stull1988}:
+%
+\begin{equation}
+\frac{u(z)}{u_{*}} = \frac{1}{\kappa}\ln\left(\frac{z}{z_{0}}\right),
+\end{equation}
+%
+where $z$ is the vertical coordinate, $\kappa = 0.4$ the von K\'arm\'an constant and $z_{0}$ the hydrodynamic roughness.
+Several field measurements of hydrodynamic roughnesses are available \pc{Here cite a few old/general papers on roughness}.
+In the absence of sediment transport, it scales with the geometric features of the bed \citep{Pelletier2016} \pc{too specific paper?
+other older papers?}.
+When transport occurs, it rather typically scales with the thickness of the transport layer.
+For aeolian saltation, this is controlled by the altitude of Bagnold's focal point \citep{Duran2011,Valance2015}, which depends on the wind velocity and grain properties \citep{Sherman2008, Zhang2016, Field2018}.
+Whether associated with geometric features or with sediment transport, its typical order of magnitude is the millimetre scale.
+
+We do not have precise velocity vertical profiles to be able to deduce an accurate value of $z_0$ in the various environments of the meteorological stations (vegetated, arid, sandy).
+Our approach is to rather select the hydrodynamic roughness which allows for the best possible matching between the regionally predicted and locally measured winds, i.e.
+minimising the relative difference $\delta$ between the wind vectors of both datasets:
+%
+\begin{equation}
+\label{metric_roughness}
+\delta = \frac{\sqrt{\langle\| \boldsymbol{u}_{*, \textrm{era}} - \boldsymbol{u}_{*, \textrm{station}} \|^{2}\rangle}}{\sqrt{ \langle \| \boldsymbol{u}_{*, \textrm{era}} \| \rangle \langle \| \boldsymbol{u}_{*, \textrm{station}} \| \rangle}} ,
+\end{equation}
+%
+where $\langle.\rangle$ denotes time average.
+This parameter is computed for values of $z_0$ in ERA5-Land analysis ranging from $10^{-5}$~m to $10^{-2}$~m for the four different stations.
+Note that for the Deep Sea and South Namib stations, where the giant dunes feedback presumably affect the wind, we take into account the non-deflected winds only in the calculation of $\delta$ (with a $15^\circ$ tolerance).
+
+As shown in Suppl.
+Fig.~\ref{Fig3_supp}, the minimum values of $\delta$ in the space ($z_{0}^{\textup{ERA5Land}}$, $z_{0}^{\textup{local}}$) form a line.
+We thus set the roughness in the ERA5-Land analysis to the typical value $z_0=10^{-3}$~m, and deduce the corresponding ones for the local stations.
+It leads to $2.7$, $0.8$, $0.1$ and $0.5$~mm for the Adamax, Deep Sea, Huab and South Namib stations, respectively.
+Importantly, this approach somewhat impacts the calculation of the shear velocities, but not that of the wind directions.
+As such, most of our conclusions are independent of such a choice.
+However, it may affect the magnitude of the wind velocity attenuation/amplification in flow confinement situations.
+
+%---------------------------------------------------------------------------------------
+\section*{2. Computation of the ABL characteristics}
+
+The estimation of the non-dimensional numbers associated with the ABL requires the computation of representative meteorological quantities.
+In arid areas, the vertical structure of the atmosphere can be approximated by a well mixed convective boundary layer of height $H$, topped by the stratified free atmosphere \citep{Stull1988, Shao2008}.
+In this context, one usually introduces the virtual potential temperature $T_{\textup{vp}}$, which is a constant $T_{0}$ inside the boundary layer, and increases linearly in the FA (Suppl.
+Fig.~\ref{Fig9_supp}a):
+%
+\begin{equation}
+T_{\textup{vp}}(z) =
+\begin{cases}
+T_{0} &\text{for $z \leq H$},\\
+T_{0}\left(1 + \displaystyle\frac{\Delta T_{\textup{vp}}}{T_{0}} + \frac{N^{2}}{g}(z - H)\right) &\text{for $ z \geq H$},
+\end{cases}
+\label{TvpVerticalProfile}
+\end{equation}
+%
+where $\Delta T_{\textup{vp}}$ is the temperature discontinuity at the capping layer and $N = \sqrt{g \partial_z T_{\textup{vp}} / T_{0}}$ is the Brunt-V\"ais\"al\"a frequency, characteristic of the stratification.
+Note that, under the usual Boussinesq approximation, temperature and air density variations are simply related by $\delta T_{\textup{vp}}/T_{0} \simeq - \delta\rho/\rho_{0}$ (see Suppl.
+Mat.
+of \citet{andreotti2009}), so that $N$ can equivalently be defined from the density gradient as next to Eq.~\ref{FroudeNumber}.
+
+The ERA5 dataset provides vertical profiles of the geopotential $\phi$, the actual temperature $T$ and the specific humidity $\eta$ at given pressure levels $P$.
+The vertical coordinate is then calculated as:
+%
+\begin{equation}
+z = \frac{\phi R_{\textup{t}}}{g R_{\textup{t}} - \phi},
+\end{equation}
+%
+where $R_{\textup{t}} = 6371229$~m is the reference Earth radius and $g = 9.81$~m s$^{-2}$ is the gravitational acceleration. One also computes the virtual potential temperature as:
+%
+\begin{equation}
+T_{\textup{vp}} = T\left[1 + \left(\frac{M_d}{M_w} - 1\right) \eta \right] \left(\frac{P_{0}}{P}\right)^{R/C_p},
+\end{equation}
+%
+where $P_{0} = 10^{5}$~Pa is the standard pressure, $R = 8.31$~J/K is the ideal gas constant, $C_p \simeq 29.1$~J/K is the air molar heat capacity, and $M_w = 0.018$~kg/Mol and $M_d = 0.029$~kg/Mol are the molecular masses of water and dry air respectively.
+The specific humidity is related to the vapour pressure $p_w$ as
+%
+\begin{equation}
+\eta = \frac{\frac{M_w}{M_d} p_w}{p - \left( 1 - \frac{M_w}{M_d} \right) p_w} \, .
+\end{equation}
+%
+
+The ERA5 dataset also provides an estimate of the ABL depth $H$, based on the behaviour of the Richardson vertical profile.
+This dimensionless number is defined as the ratio of buoyancy and flow shear terms, and can be expressed as $\textup{Ri} = N^2/(\partial_z u)^2$.
+It vanishes in the lower well-mixed layer where $T_{\textup{vp}}$ is constant, and increases in the stratified FA.
+Following the method and calibration of \citet{Vogelezang1996,seidel2012}, the value $\textup{Ri}(z) \simeq 0.25$ has been shown to be a good empirical criterion to give $z \simeq H$ within a precision varying from $50$\% for the shallower ABL (e.g.
+at night) to $20$\% for situations of stronger convection.
+
+Examples of vertical profiles of the virtual potential temperature deduced from ERA5 are shown in Suppl.
+Fig.~\ref{Fig9_supp}a.
+For each of them, an average temperature is computed below the ABL depth ($z<H$), and a linear function is fitted above, allowing us to extract the temperature jump $\Delta T_{\textup{vp}}$.
+Importantly, some profiles display a vertical structure that cannot be approximated by the simple form \eqref{TvpVerticalProfile} used here (Suppl.
+Fig.~\ref{Fig9_supp}b).
+In practice, we removed from the analysis all of those leading to the unphysical case $\Delta T_{\textup{vp}}<0$.
+We have noticed that these `ill-processed' profiles dominantly occur in winter and are evenly spread across the hours of the day.
+Importantly, they represent $\simeq12$\% of the data only (Suppl.
+Fig.~\ref{Fig9_supp}c,d), and we are thus confident that this data treatment does not affect our conclusions.
+
+%---------------------------------------------
+\begin{figure}[p]
 \centering
@@
 \end{figure}
+%---------------------------------------------
 
-\begin{figure}
+%---------------------------------------------
+\begin{figure}[p]
 \centering
 \includegraphics[scale=1]{Figures/Figure2_supp.pdf}
-  \caption{Comparison between raw local wind measurements, and centered averaged data over one hour for the South Namib station.
-\textbf{a}: wind direction.
-\textbf{b}: wind velocity at the measurement height, 2.6 m.}
+\caption{Comparison between raw local wind measurements, and hourly-averaged data for South Namib station. \textbf{a}: wind direction. \textbf{b}: wind velocity at height $2.6$~m.}
 \label{Fig2_supp}
 \end{figure}
+%---------------------------------------------
 
-\section*{1. Shear velocity and calibration of the hydrodynamical roughness}
-\label{calib_z0}
-
-For each station, the hydrodynamic roughness is calibrated by finding the one that minimizes the relative difference $\delta$ between the wind vectors of both datasets:
-\begin{equation}
-  \label{metric_roughness}
-      \delta = \frac{\sqrt{\langle\| \boldsymbol{u}_{*, \textrm{era}} - \boldsymbol{u}_{*, \textrm{station}} \|^{2}\rangle_{t}}}{\sqrt{ \langle \| \boldsymbol{u}_{*, \textrm{era}} \| \rangle_{t}\langle \| \boldsymbol{u}_{*, \textrm{station}} \| \rangle_{t}}}
-\end{equation}
-
-This $\delta$--parameter is computed for hydrodynamic roughness values ranging from $10^{-5}$~m to $10^{-2}$~m for the different stations.
-A shown by figure~\ref{Fig3_supp}, the minimum of $\delta$ in the space ($z_{0,\, \textup{Era}}$, $z_{0,\, \textup{in situ}}$) forms a line.
-We thus take the roughness of the Era5Land dataset as the typical value when sediment transport occurs, $10^{-3}$~m, corresponding to the thickness of the transport layer \citep{Duran2011}.
-It leads for the Adamax, Deep Sea, Huab and South Namib stations values of $2.7~\textup{mm}$, $0.76~\textup{mm}$, $0.12~\textup{mm}$ and $0.48~\textup{mm}$, respectively.
-
-The choice of the hydrodynamic roughness values only impacts the calculated shear velocities, but note the wind directions.
-As such, most of our conclusions are then independent of such a choice, and only the magnitude of the wind velocity attenuation in confined situation might be affected.
-
-\begin{figure}
+%---------------------------------------------
+\begin{figure}[p]
 \centering
 \includegraphics[scale=1]{Figures/Figure3_supp.pdf}
-  \caption{Calibration of the hydrodynamic roughnesses.
-The metric $\delta$ defined in \eqref{metric_roughness} is represented in colorscale as a function of the hydrodynamic roughnesses chosen for the Era5-Land and in situ datasets, for the Adamax (\textbf{a}), Deep Sea (\textbf{b}), Huab (\textbf{c}) and South Namib (\textbf{d}) Stations.
-The red dashed and plain lines shows the minima of $\delta$ and the identity line.
-The orange lines and dots highlight the chosen hydrodynamic roughnesses for the local datasets, by imposing $z_{0, \textup{\textup{ERA}}} = 1~\textup{mm}$, leading for each station to $2.7~\textup{mm}$, $0.76~\textup{mm}$, $0.12~\textup{mm}$ and $0.48~\textup{mm}$, respectively.}
+\caption{Calibration of hydrodynamic roughness.
+The parameter $\delta$ (Eq.~\ref{metric_roughness}) quantifying the difference between local and predicted winds is shown in colorscale as a function of the hydrodynamic roughnesses chosen for the ERA5-Land and for local winds, for the (\textbf{a}) Adamax, (\textbf{b}) Deep Sea, (\textbf{c}) Huab and (\textbf{d}) South Namib stations.
+The red dashed and plain lines shows the minima of $\delta$ and the identity line, respectively.
+The orange lines and dots highlight the chosen hydrodynamic roughnesses for the local winds deduced from setting $z_{0}^{\textup{ERA5Land}} = 1$~mm.}
 \label{Fig3_supp}
 \end{figure}
+%---------------------------------------------
 
-\begin{figure}
+%---------------------------------------------
+\begin{figure}[p]
 \centering
 \includegraphics[scale=1]{Figures/Figure4_supp.pdf}
-  \caption{Analysis of the DEMs of the Deep Sea (left column -- \textbf{a}, \textbf{b}, \textbf{c}) and South Namib (right column -- \textbf{d}, \textbf{e}, \textbf{f}) stations.
-\textbf{a--d}: Detrended topography (a second order polynomial is first fitted and then removed) .
-\textbf{b--e}: autocorrelation matrix shown in colorscale.
-The black line shows the detected orientation, and the red line the profile along which the wavelength is calculated, shown in \textbf{c--f}.
+\caption{Analysis of the DEMs of the Deep Sea (left column -- panels \textbf{a}, \textbf{b}, \textbf{c}) and South Namib (right column -- panels \textbf{d}, \textbf{e}, \textbf{f}) stations.
+\textbf{a-d}: Bed elevation detrended by a fitted second order polynomial base-line.
+\textbf{b-e}: Autocorrelation matrix shown in colorscale.
+The black line shows the detected dune orientation, and the red line represents the autocorrelation profile along which the dune wavelength is calculated, displayed in \textbf{c-f}.
 The blue lines and dots show the first peak of the autocorrelation profile, whose abscissa gives the characteristic wavelength of the dune pattern.}
@@
 \end{figure}
+%---------------------------------------------
 
-\begin{figure}
+%---------------------------------------------
+\begin{figure}[p]
 \centering
 \includegraphics[scale=1]{Figures/Figure5_supp.pdf}
-  \caption{Statistical agreement of the wind orientation (\textbf{a}) and velocity (\textbf{b}) between the Era5Land dataset and the local measurements for the Huab and Adamax stations.
-Note how the points are clustered around identity lines (dashed and black).}
+\caption{Statistical comparison of the wind orientation (\textbf{a}) and velocity (\textbf{b}) between the ERA5-Land dataset and the local measurements for the Huab and Adamax stations.
+Data point clustering around identity lines (dashed and black) provide evidence for agreement of the two sets.}
 \label{Fig5_supp}
 \end{figure}
+%---------------------------------------------
 
-\begin{figure}
+%---------------------------------------------
+\begin{figure}[p]
 \centering
 \includegraphics[scale=1]{Figures/Figure6_supp.pdf}
-  \caption{Distributions of wind direction at the South Namib Station for the Era5Land climate reanalysis (orange) and the in situ measurements (blue).
-In each subplot, both distributions are plotted from the same time steps, selected with constraints on the wind direction (columns) and/or wind velocity (rows) in the Era5Land dataset.
+\caption{Distributions of wind direction at the South Namib Station for the ERA5-Land climate reanalysis (orange) and the local measurements (blue) -- equivalent of Fig.~\ref{Fig3}.
+In each subplot, both distributions are plotted from the same time steps, selected with constraints on the wind direction (columns) and/or wind velocity (rows) in the ERA5-Land dataset.
 The gray vertical dashed lines indicate the dune orientation.
-The numbers at the top center give the percentage of time steps selected in each sub-range, as well as the percentage of them corresponding to the day (between 1000 UTC and 2200 UTC).
-The purple frame highlights the regime (small wind velocities, nocturnal summer wind) during which the wind data from both datasets differs.
-A similar figure can be obtained for the South Namib station (see Fig.~\ref{Fig3}).}
+The numbers at the top centre give the percentage of time steps selected in each sub-range, as well as the percentage of them corresponding to the day (between 1000 UTC and 2200 UTC).
+The purple frame highlights the regime (small wind velocities, nocturnal summer wind) during which the wind data from both datasets differ.}
 \label{Fig6_supp}
 \end{figure}
+%---------------------------------------------
 
-\begin{figure}
+%---------------------------------------------
+\begin{figure}[p]
 \centering
 \includegraphics[scale=1]{Figures/Figure7_supp.pdf}
-  \caption{Statistical comparison of the wind orientation between the Era5Land dataset and the local measurements for the South Namib and Deep Sea stations, for different velocity ranges.
+\caption{Statistical comparison of the wind orientation between the ERA5-Land dataset and the local measurements for the South Namib and Deep Sea stations, for different velocity ranges.
 \textbf{a}: $u_{*, \textup{\textup{ERA}}} < 0.1~\textup{m}~\textup{s}^{-1}$.
@@
 \textbf{c}: $u_{*, \textup{\textup{ERA}}} \geq 0.25~\textup{m}~\textup{s}^{-1}$.
-Note that the measured dune orientations are substracted to the wind orientation, which allows to plot both stations on the same graph.
-Black dashed lines indicate locally measured orientations aligned with the dune crests (here $0^\circ$, $180^\circ$ and $360^\circ$ -- \textbf{a}, \textbf{b}), as well as the identity lines (\textbf{b}, \textbf{c}).}
+The measured dune orientations are subtracted to the wind orientation, which allows us to plot both stations on the same graph.
+Black dashed lines indicate locally measured orientations aligned with the dune crests (here $0^\circ$, $180^\circ$ and $360^\circ$ -- panels \textbf{a}, \textbf{b}), as well as the identity lines (panels \textbf{b}, \textbf{c}).}
 \label{Fig7_supp}
 \end{figure}
+%---------------------------------------------
 
-\begin{figure}
+%---------------------------------------------
+\begin{figure}[p]
 \centering
 \includegraphics[scale=1]{Figures/Figure8_supp.pdf}
-  \caption{Statistical comparison of the wind velocity between the Era5Land dataset and the local measurements for the South Namib and Deep Sea stations.
+\caption{Statistical comparison of the wind velocity between the ERA5-Land dataset and the local measurements for the South Namib and Deep Sea stations.
 \textbf{a}: Nocturnal summer easterly wind.
@@
 Black dashed lines are identity lines.
-The angle ranges used to select diurnal and nocturnal summer winds are those taken in Fig.~\ref{Fig3} and Fig.~\ref{Fig6_supp}.}
+The angle ranges used to select diurnal and nocturnal summer winds are the same as those in Fig.~\ref{Fig3} and Suppl.
+Fig.~\ref{Fig6_supp}.}
 \label{Fig8_supp}
 \end{figure}
+%---------------------------------------------
 
-\section*{2. Extraction of the ABL properties}
-
-In order to estimate the relevant non-dimensional numbers, one need to estimate in addition to the wind and dune properties some parameters of the ABL.
-The Era5 dataset provides a direct bulk estimate of the ABL depth $H$ from a bulk Richardson number calculation, as well as vertical profiles of the geopotential $\phi$, temperature $T$ and specific humidity $e_{\textup{w}}$ at given pressure levels $P$.
-From these quantities, the virtual potential temperature, which takes into account the vertical pressure and humidity changes, can be calculated as:
-\begin{equation}
-  T_{\textup{vp}} = T\left(1 + \left[R_{\textup{M}} - 1\right]e_{\textup{w}}\right)\left(\frac{P_{0}}{P}\right)^{P_{\textup{c}}(1 - 0.24e_{\textup{w}})},
-\end{equation}
-where $P_{0} = 10^{5}~\textup{Pa}$ is the standard pressure, $P_{\textup{c}} = 0.2854$ the Poisson coefficient for dry air and $R_{\textup{M}}=1.61$ is the ratio between the melocular masses of dry air and water.
-The vertical coordinates are calculated as:
-\begin{equation}
-  z = \frac{\phi R_{\textup{t}}}{g R_{\textup{t}} - \phi},
-\end{equation}
-where $R_{\textup{t}} = 6356766~\textup{m}$ is the average Earth radius, and $g = 9.81~\textup{m}~\textup{s}^{-2}$ the gravitational acceleration.
-
-Example of obtained vertical profiles of the virtual potential temperature are shown in Fig.~\ref{Fig9_supp}.
-On each of them, an average is computed below the ABL depth given by the Era5 dataset, and a linear function is fitted above.
-
-Under the Boussinesq approximation, the temperature variations are assumed to induce most of those of the density, leading to $\Delta\rho/\rho \simeq \Delta T_{\textup{vp}}/T_{\textup{vp}}$.
-Here, $T_{\textup{vp}}/T_{\textup{vp}}$ is the relative virtual potential temperature jump at the capping, directly measured on the vertical profiles.
-
-Following \citet{Tritton2012}, the relative density jump at the capping layer
-
-\begin{figure}
+%---------------------------------------------
+\begin{figure}[p]
 \centering
 \includegraphics[scale=1]{Figures/Figure9_supp.pdf}
-  \caption{\textbf{a}: Vertical profiles of the virtual potential temperature at 3 different time steps (blue - 29/11/2012 - 1100 UTC, orange - 21/03/2017 - 1200 UTC, green - 21/03/2017 - 2000 UTC) at the South Namib station.
+\caption{\textbf{a}: Vertical profiles of the virtual potential temperature at three different times (blue: 29/11/2012 - 1100 UTC, orange: 21/03/2017 - 1200 UTC, green: 21/03/2017 - 2000 UTC) at the South Namib station.
 Dots: data from the ERA5 reanalysis.
-Transparent dashed lines: boundary layer height given by the ERA5 reanalysis, calculated from the bulk Richardson number \citep{seidel2012}.
-Plain lines: vertical (boundary layer) and linear (free atmosphere) fits to estimate the quantities in Fig.~\ref{Fig10_supp}.
-\textbf{b}: Examples of ill-processed vertical profiles at 3 different time steps (blue - 2/12/2013 - 2300 UTC, orange - 20/03/2017 - 0000 UTC, green - 14/07/2017 - 1400 UTC) at the South Namib station.
+Dashed lines: boundary layer height given by the ERA5 reanalysis.
+Plain lines: vertical (ABL) and linear (FA) fits to estimate the quantities displayed in Suppl.
+Fig.~\ref{Fig10_supp}.
+\textbf{b}: Examples of ill-processed vertical profiles at three different times (blue: 2/12/2013 - 2300 UTC, orange: 20/03/2017 - 0000 UTC, green: 14/07/2017 - 1400 UTC) at the South Namib station.
 \textbf{c}: Hourly distribution of ill-processed vertical profiles.
-\textbf{d}: Monthly distribution of ill-processed vertical profiles.
-\cg{These profiles are ill-processed because the temperature found at the boundary layer from the linear fit in the free-atm is smaller than the average one inside the boundary layer.
-This is an unstable situation, which does not allow to calculate the surface Froude number.}}
+\textbf{d}: Monthly distribution of ill-processed vertical profiles.}
 \label{Fig9_supp}
 \end{figure}
+%---------------------------------------------
 
-\begin{figure}
+%---------------------------------------------
+\begin{figure}[p]
 \centering
 \includegraphics{Figures/Figure10_supp.pdf}
-  \caption{Distributions of the meteorological parameters resulting from the processing of the Era5-Land data for the South Namib (blue) and the Deep Sea (orange) stations.}
+\caption{Distributions of the meteorological parameters resulting from the processing of the ERA5-Land data for the South Namib (blue) and the Deep Sea (orange) stations.
+\pc{axis label : panel c -velocity at (height) H}.
+\pc{Is this for all day/night/seasons ?}}
 \label{Fig10_supp}
 \end{figure}
+%---------------------------------------------
 
-\begin{figure}
+%---------------------------------------------
+\begin{figure}[p]
 \centering
 \includegraphics[scale=1]{Figures/Figure11_supp.pdf}
-  \caption{Non-dimensional parameters distributions. For the marginal distributions, the orange correspond to the South Namib station, and the blue to the Deep Sea station.}
+\caption{Non-dimensional parameters distributions. For the marginal \pc{???} distributions, the orange correspond to the South Namib station, and the blue to the Deep Sea station.}
 \label{Fig11_supp}
 \end{figure}
+%---------------------------------------------
 
-\begin{figure}
+%---------------------------------------------
+\begin{figure}[p]
 \centering
@@
 Blue dashed lines empirically delimit the different regimes.
-The point density in each bin of the diagrams is shown in Fig.~\ref{Fig11_supp}.
-The regime diagrams in the space $(\mathcal{F}, \, kH)$ are shown in Fig.~\ref{Fig5} of the main article.}
+The point density in each bin of the diagrams is shown in Suppl.
+Fig.~\ref{Fig11_supp}.
+The similar regime diagrams in the space $(\mathcal{F}, \, kH)$ are shown in Fig.~\ref{Fig5}.}
 \label{Fig12_supp}
 \end{figure}
+%---------------------------------------------
 
-\begin{figure}
+%---------------------------------------------
+\begin{figure}[p]
 \centering
 \includegraphics[scale=1]{Figures/Figure13_supp.pdf}
-  \caption{Physical interpretation of the flow disturbance.
-(a) and (b) Magnitude of the disturbance induced by a sinusoidal topography calculated from the time series of the non-dimensional numbers presented in Figures~\ref{Fig4} and \ref{Fig5} using the linear model of \citet{andreotti2009}.
-(c) Shear velocity streamlines represented in the case of the Deep Sea station, for increasing values of $\sqrt{\mathcal{A}_{0}^{2} + \mathcal{B}_{0}^{2}}$.
-From the upper to the lower streamline, values of $\left(kH,\, Fr_{\textup{surface}},\, Fr_{\textup{internal}},\, \mathcal{A}_{0},\, \mathcal{B}_{0},\, \sqrt{\mathcal{A}_{0}^{2} + \mathcal{B}_{0}^{2}}\right)$ are (1.9, 0.6, 1.5, 3.4, 1.0, 3.5), (1.5, 0.3, 0.4, 4.8, 1.4, 5.0), (0.1, 3.5, 1.0, 8.6, 0.1, 8.6), (0.5, 0.05, 0.04, 9.6, 2.5, 9.9).}
+\caption{Computation of the flow disturbance with the linear model of \citet{andreotti2009}.
+(a) and (b) Magnitude of the hydrodynamic coefficients $\mathcal{A}_0$ and $\mathcal{B}_0$, calculated from the values of the non-dimensional numbers corresponding to the ERA5-Land time series presented in Figs.~\ref{Fig4} and \ref{Fig5}.
+(c) Shear velocity streamlines over sinusoidal ridges of amplitude $k\xi_0 = 0.1$ and for increasing values of $\sqrt{\mathcal{A}_{0}^{2} + \mathcal{B}_{0}^{2}}$.
+From the upper to the lower streamline, values of $\left(kH,\, \mathcal{F},\, \mathcal{F}_{\rm I},\, \mathcal{A}_{0},\, \mathcal{B}_{0},\, \sqrt{\mathcal{A}_{0}^{2} + \mathcal{B}_{0}^{2}}\right)$ are (1.9, 0.6, 1.5, 3.4, 1.0, 3.5), (1.5, 0.3, 0.4, 4.8, 1.4, 5.0), (0.1, 3.5, 1.0, 8.6, 0.1, 8.6), (0.5, 0.05, 0.04, 9.6, 2.5, 9.9).}
 \label{Fig13_supp}
 \end{figure}
+%---------------------------------------------
 

### 2022-02-28 Paper/main.tex
Starting modifications after review by collaborators:
- Expanding Fig2 and adding its double in Supp as Fig4_supp
- Renaming figures and references with respect to previous change
- Changing in situ to local in figures
- Chaning 'Deep Sea' to 'North Namib' in text and figures

@@
 These two stations are located in a relatively flat \pc{quantify how flat is flat ?} environment.
-In contrast, the Deep Sea and South Namib stations are located in the interdune between tens of meters high giant linear dunes with kilometer-scale wavelengths and superimposed patterns.
+In contrast, the North Namib and South Namib stations are located in the interdune between tens of meters high giant linear dunes with kilometer-scale wavelengths and superimposed patterns.
 In this section, we describe and compare winds from local measurements and climate reanalysis predictions.
@@
 Dune properties are computed using autocorrelation on the 30-m Digital Elevation Models (DEMs) of the shuttle radar topography mission \citep{Farr2007}.
-For the South Namib and Deep Sea stations, we obtain respectively orientations of $85^\circ$ and $125^\circ$ with respect to the North, wavelengths of $2.6~\textup{km}$ and $2.3~\textup{km}$ and amplitudes of $45~\textup{m}$ and $20~\textup{m}$ (Suppl.
-Fig.~\ref{Fig4_supp} for more details).
+For the South Namib and North Namib stations, we obtain respectively orientations of $85^\circ$ and $125^\circ$ with respect to the North, wavelengths of $2.6~\textup{km}$ and $2.3~\textup{km}$ and amplitudes of $45~\textup{m}$ and $20~\textup{m}$ (Suppl.
+Fig.~\ref{Fig5_supp} for more details).
 
@@
 \caption{Temporal comparison between the wind data coming from the ERA5-Land climate reanalysis (orange lines) and from the local measurements (blue lines).
-Coloured swathes indicate day (between 1000 UTC and 2200 UTC) and night (before 1000 UTC or after 2200 UTC) \textbf{a-b}: Huab station.
-\textbf{c-d}: Deep Sea station in winter.
-\textbf{e-f}: Deep Sea station in summer.}
+Coloured swathes indicate day (between 1000 UTC and 2200 UTC) and night (before 1000 UTC or after 2200 UTC) \textbf{a-b}: Adamax station in summer.
+\textbf{b-c}: Adamax station in winter.
+\textbf{d-e}: North Namib station in summer.
+\textbf{f-g}: North Namib station in winter.
+Time series of the two other stations are shown in Suppl.
+Fig.~\ref{Fig4_supp}.}
 \label{Fig2}
@@
 
-In the case of the Adamax and Huab stations, the time series of wind speed and direction from the regional predictions quantitatively match those corresponding to the local measurements (Fig.~\ref{Fig2}a,b) and Suppl.
-Fig.~\ref{Fig5_supp}).
-For the Deep Sea and South Namib stations within the giant dune field, we observe that this agreement is also good, but limited to the September-January time period (Fig.~\ref{Fig2}c,d).
+In the case of the Adamax and Huab stations, the time series of wind speed and direction from the regional predictions quantitatively match those corresponding to the local measurements (Fig.~\ref{Fig2}a--d) and Suppl.
+Fig.~\ref{Fig5_supp}a--d,~\ref{Fig6_supp}).
+For the North Namib and South Namib stations within the giant dune field, we observe that this agreement is also good, but limited to the September-January time period (Fig.~\ref{Fig2}e--h and and Suppl.
+Fig.~\ref{Fig5_supp}e--h).
 As a matter of fact, the measured wind roses exhibit additional wind components aligned with the giant dune orientation, as evidenced on the satellite images (Fig.~\ref{Fig1}c,d).
 
-More precisely, during the February-August period, the local and regional winds in the interdune match during daytime only, i.e when the southerly/southwesterly sea breeze dominates (Figs.~\ref{Fig2}(e,f) and \ref{Fig3}, Suppl.
-Fig.~\ref{Fig6_supp}).
+More precisely, during the February-August period, the local and regional winds in the interdune match during daytime only, i.e when the southerly/southwesterly sea breeze dominates (Figs.~\ref{Fig2}e,f and \ref{Fig3}, Suppl.
+Fig.~\ref{Fig7_supp}).
 In the late afternoon and during the night, when the northwesterly `berg' and katabatic winds blow, measurements and predictions differ.
 In this case, the angular wind distribution of the local measurements exhibits two additional modes corresponding to reversing winds aligned with the giant dune orientation (purple frame in Fig.~\ref{Fig3}, Suppl.
-Figs.~\ref{Fig6_supp} and \ref{Fig7_supp}).
+Figs.~\ref{Fig7_supp} and \ref{Fig8_supp}).
 This deviation is also associated with a global attenuation of the wind strength (Suppl.
-Fig.~\ref{Fig8_supp}).
+Fig.~\ref{Fig9_supp}).
 Remarkably, all these figures show that this wind reorientation and attenuation processes typically occur at low wind velocities, typically for $u_{*} \lesssim 0.1~\textrm{m}~\textrm{s}^{-1}$.
@@
 Finally, for intermediate shear velocities, moments of reorientation along the dune crest and others with no reorientation can both be observed (Suppl.
-Fig.~\ref{Fig7_supp}).
+Fig.~\ref{Fig8_supp}).
 \pc{CHECK the above lines cf.
@@
 \includegraphics[scale=1]{Figures/Figure3.pdf}
-\caption{Distributions of wind direction at the Deep Sea Station for the ERA5-Land climate reanalysis (orange) and the local measurements (blue).
+\caption{Distributions of wind direction at the North Namib Station for the ERA5-Land climate reanalysis (orange) and the local measurements (blue).
 In each subplot, both distributions are plotted from the same time steps, selected with constraints on the wind direction (columns) and/or wind velocity (rows) of the ERA5-Land dataset.
@@
 The purple frame highlights the regime (small wind velocities, nocturnal summer wind) in which the data from both datasets differ.
-A similar figure can be obtained for the Deep Sea station (Suppl.
-Fig.~\ref{Fig6_supp}).}
+A similar figure can be obtained for the North Namib station (Suppl.
+Fig.~\ref{Fig7_supp}).}
 \label{Fig3}
@@
 The wind deflection induced by linear dunes has so far mainly been related to the angle between wind direction and crest orientation, with a maximum deflection for angles between $30^{\circ}$ and $70^{\circ}$ \citep{Walker2009, Hesp2015}.
-In the analysed data, the most deflected wind at both Deep Sean and South Namib stations is the most perpendicular to the giant dunes (\pc{Fig.~\ref{***}}).
+In the analysed data, the most deflected wind at both North Namibn and South Namib stations is the most perpendicular to the giant dunes (\pc{Fig.~\ref{***}}).
 The incident wind direction then does not seem to be the only parameter controlling the wind deflection.
@@
 \includegraphics[scale=1]{Figures/Figure4.pdf}
-\caption{\textbf{a-b}: Vertical profiles of the virtual potential temperature at 2 different time steps (day - 03/11/2015 - 1200 UTC, night - 01/13/2013 - 0900 UTC) at the Deep Sea station.
+\caption{\textbf{a-b}: Vertical profiles of the virtual potential temperature at 2 different time steps (day - 03/11/2015 - 1200 UTC, night - 01/13/2013 - 0900 UTC) at the North Namib station.
 Dots: data from the ERA5 reanalysis.
@@
 \includegraphics[scale=1]{Figures/Figure5.pdf}
-\caption{Regime diagrams of the wind deviation $\delta_{\theta}$ (\textbf{a}) and relative attenuation/amplification $\delta_{u}$ (\textbf{b}) in the space $(\mathcal{F}, \, kH)$, containing the data from both the Deep Sea and South Namib stations.
+\caption{Regime diagrams of the wind deviation $\delta_{\theta}$ (\textbf{a}) and relative attenuation/amplification $\delta_{u}$ (\textbf{b}) in the space $(\mathcal{F}, \, kH)$, containing the data from both the North Namib and South Namib stations.
 Blue dashed lines empirically delimit the different regimes.
 The point density in each bin of the diagrams is shown in Suppl.
-Fig.~\ref{Fig11_supp}.
+Fig.~\ref{Fig12_supp}.
 Similar regime diagrams in the spaces $(\mathcal{F}_{\textup{I}}, \, kH)$ and $(\mathcal{F}_{\textup{I}}, \, \mathcal{F})$ are shown in Suppl.
-Fig.~\ref{Fig12_supp}.
+Fig.~\ref{Fig13_supp}.
 \pc{ticks on x-axis ?}}
@@
 \includegraphics[scale=1]{Figures/Figure6.pdf}
-\caption{Implications for smaller scale patterns in (\textbf{a}) the South Namib and (\textbf{b}) Deep Sea.
+\caption{Implications for smaller scale patterns in (\textbf{a}) the South Namib and (\textbf{b}) North Namib.
 The ellipses indicate the different types of elongating dunes, at large (plain line) and small (dashed line) scales.
@@
 These models are able to qualitatively reproduce the observed wind deflection (Appendix 1, Suppl.
-Figs.~\ref{Fig12_supp} and \ref{Fig13_supp}), and thus provide the physical support for the hydrodynamic picture we propose here.
+Figs.~\ref{Fig13_supp} and \ref{Fig14_supp}), and thus provide the physical support for the hydrodynamic picture we propose here.
 They however cannot reach the magnitude of these observations, probably due to the presence of expected non-linearities in high confinement situations linked to strong flow modulations.
@@
 %
-which is also a good approximation for the giant dunes observed in the Deep Sea and South Namib Station (Fig.~\ref{Fig1} and Suppl.
+which is also a good approximation for the giant dunes observed in the North Namib and South Namib Station (Fig.~\ref{Fig1} and Suppl.
 Fig.~\ref{Fig4_supp}).
@@
 Fig.~\ref{Fig13_supp}).
-Here, using $k\xi_0 \simeq 0.1$ to be representative of the amplitude of the giant dunes at the Deep Sea station, the coefficient modulus is bounded to $10$.
+Here, using $k\xi_0 \simeq 0.1$ to be representative of the amplitude of the giant dunes at the North Namib station, the coefficient modulus is bounded to $10$.
 
@@
 This parameter is computed for values of $z_0$ in ERA5-Land analysis ranging from $10^{-5}$~m to $10^{-2}$~m for the four different stations.
-Note that for the Deep Sea and South Namib stations, where the giant dunes feedback presumably affect the wind, we take into account the non-deflected winds only in the calculation of $\delta$ (with a $15^\circ$ tolerance).
+Note that for the North Namib and South Namib stations, where the giant dunes feedback presumably affect the wind, we take into account the non-deflected winds only in the calculation of $\delta$ (with a $15^\circ$ tolerance).
 
@@
 We thus set the roughness in the ERA5-Land analysis to the typical value $z_0=10^{-3}$~m, and deduce the corresponding ones for the local stations.
-It leads to $2.7$, $0.8$, $0.1$ and $0.5$~mm for the Adamax, Deep Sea, Huab and South Namib stations, respectively.
+It leads to $2.7$, $0.8$, $0.1$ and $0.5$~mm for the Adamax, North Namib, Huab and South Namib stations, respectively.
 Importantly, this approach somewhat impacts the calculation of the shear velocities, but not that of the wind directions.
@@
 \caption{Calibration of hydrodynamic roughness.
-The parameter $\delta$ (Eq.~\ref{metric_roughness}) quantifying the difference between local and predicted winds is shown in colorscale as a function of the hydrodynamic roughnesses chosen for the ERA5-Land and for local winds, for the (\textbf{a}) Adamax, (\textbf{b}) Deep Sea, (\textbf{c}) Huab and (\textbf{d}) South Namib stations.
+The parameter $\delta$ (Eq.~\ref{metric_roughness}) quantifying the difference between local and predicted winds is shown in colorscale as a function of the hydrodynamic roughnesses chosen for the ERA5-Land and for local winds, for the (\textbf{a}) Adamax, (\textbf{b}) North Namib, (\textbf{c}) Huab and (\textbf{d}) South Namib stations.
 The red dashed and plain lines shows the minima of $\delta$ and the identity line, respectively.
@@
   \includegraphics[scale=1]{Figures/Figure4_supp.pdf}
-\caption{Analysis of the DEMs of the Deep Sea (left column -- panels \textbf{a}, \textbf{b}, \textbf{c}) and South Namib (right column -- panels \textbf{d}, \textbf{e}, \textbf{f}) stations.
+  \caption{Analysis of the DEMs of the North Namib (left column -- panels \textbf{a}, \textbf{b}, \textbf{c}) and South Namib (right column -- panels \textbf{d}, \textbf{e}, \textbf{f}) stations.
 \textbf{a-d}: Bed elevation detrended by a fitted second order polynomial base-line.
@@
 %---------------------------------------------
-\begin{figure}[p]
+\begin{figure}
 \centering
 \includegraphics[scale=1]{Figures/Figure5_supp.pdf}
-\caption{Statistical comparison of the wind orientation (\textbf{a}) and velocity (\textbf{b}) between the ERA5-Land dataset and the local measurements for the Huab and Adamax stations.
-Data point clustering around identity lines (dashed and black) provide evidence for agreement of the two sets.}
+\caption{Temporal comparison between the wind data coming from the ERA5-Land climate reanalysis (orange lines) and from the local measurements (blue lines).
+Coloured swathes indicate day (between 1000 UTC and 2200 UTC) and night (before 1000 UTC or after 2200 UTC) \textbf{a-b}: Huab station in summer.
+\textbf{b-c}: Huab station in winter.
+\textbf{d-e}: South Namib station in summer.
+\textbf{f-g}: South Namib station in winter.}
 \label{Fig5_supp}
@@
 \includegraphics[scale=1]{Figures/Figure6_supp.pdf}
-\caption{Distributions of wind direction at the South Namib Station for the ERA5-Land climate reanalysis (orange) and the local measurements (blue) -- equivalent of Fig.~\ref{Fig3}.
-In each subplot, both distributions are plotted from the same time steps, selected with constraints on the wind direction (columns) and/or wind velocity (rows) in the ERA5-Land dataset.
-The gray vertical dashed lines indicate the dune orientation.
-The numbers at the top centre give the percentage of time steps selected in each sub-range, as well as the percentage of them corresponding to the day (between 1000 UTC and 2200 UTC).
-The purple frame highlights the regime (small wind velocities, nocturnal summer wind) during which the wind data from both datasets differ.}
+\caption{Statistical comparison of the wind orientation (\textbf{a}) and velocity (\textbf{b}) between the ERA5-Land dataset and the local measurements for the Huab and Adamax stations.
+Data point clustering around identity lines (dashed and black) provide evidence for agreement of the two sets.}
 \label{Fig6_supp}
@@
 \includegraphics[scale=1]{Figures/Figure7_supp.pdf}
-\caption{Statistical comparison of the wind orientation between the ERA5-Land dataset and the local measurements for the South Namib and Deep Sea stations, for different velocity ranges.
-\textbf{a}: $u_{*, \textup{\textup{ERA}}} < 0.1~\textup{m}~\textup{s}^{-1}$.
-\textbf{b}: $0.1 < u_{*, \textup{\textup{ERA}}} \leq 0.25~\textup{m}~\textup{s}^{-1}$.
-\textbf{c}: $u_{*, \textup{\textup{ERA}}} \geq 0.25~\textup{m}~\textup{s}^{-1}$.
-The measured dune orientations are subtracted to the wind orientation, which allows us to plot both stations on the same graph.
-Black dashed lines indicate locally measured orientations aligned with the dune crests (here $0^\circ$, $180^\circ$ and $360^\circ$ -- panels \textbf{a}, \textbf{b}), as well as the identity lines (panels \textbf{b}, \textbf{c}).}
+\caption{Distributions of wind direction at the South Namib Station for the ERA5-Land climate reanalysis (orange) and the local measurements (blue) -- equivalent of Fig.~\ref{Fig3}.
+In each subplot, both distributions are plotted from the same time steps, selected with constraints on the wind direction (columns) and/or wind velocity (rows) in the ERA5-Land dataset.
+The gray vertical dashed lines indicate the dune orientation.
+The numbers at the top centre give the percentage of time steps selected in each sub-range, as well as the percentage of them corresponding to the day (between 1000 UTC and 2200 UTC).
+The purple frame highlights the regime (small wind velocities, nocturnal summer wind) during which the wind data from both datasets differ.}
 \label{Fig7_supp}
@@
 \includegraphics[scale=1]{Figures/Figure8_supp.pdf}
-\caption{Statistical comparison of the wind velocity between the ERA5-Land dataset and the local measurements for the South Namib and Deep Sea stations.
-\textbf{a}: Nocturnal summer easterly wind.
-\textbf{b}: Diurnal southerly wind.
-Black dashed lines are identity lines.
-The angle ranges used to select diurnal and nocturnal summer winds are the same as those in Fig.~\ref{Fig3} and Suppl.
-Fig.~\ref{Fig6_supp}.}
+\caption{Statistical comparison of the wind orientation between the ERA5-Land dataset and the local measurements for the South Namib and North Namib stations, for different velocity ranges.
+\textbf{a}: $u_{*, \textup{\textup{ERA}}} < 0.1~\textup{m}~\textup{s}^{-1}$.
+\textbf{b}: $0.1 < u_{*, \textup{\textup{ERA}}} \leq 0.25~\textup{m}~\textup{s}^{-1}$.
+\textbf{c}: $u_{*, \textup{\textup{ERA}}} \geq 0.25~\textup{m}~\textup{s}^{-1}$.
+The measured dune orientations are subtracted to the wind orientation, which allows us to plot both stations on the same graph.
+Black dashed lines indicate locally measured orientations aligned with the dune crests (here $0^\circ$, $180^\circ$ and $360^\circ$ -- panels \textbf{a}, \textbf{b}), as well as the identity lines (panels \textbf{b}, \textbf{c}).}
 \label{Fig8_supp}
@@
 \includegraphics[scale=1]{Figures/Figure9_supp.pdf}
+\caption{Statistical comparison of the wind velocity between the ERA5-Land dataset and the local measurements for the South Namib and North Namib stations.
+\textbf{a}: Nocturnal summer easterly wind.
+\textbf{b}: Diurnal southerly wind.
+Black dashed lines are identity lines.
+The angle ranges used to select diurnal and nocturnal summer winds are the same as those in Fig.~\ref{Fig3} and Suppl.
+Fig.~\ref{Fig6_supp}.}
+\label{Fig9_supp}
+\end{figure}
+%---------------------------------------------
+
+%---------------------------------------------
+\begin{figure}[p]
+\centering
+\includegraphics[scale=1]{Figures/Figure10_supp.pdf}
 \caption{\textbf{a}: Vertical profiles of the virtual potential temperature at three different times (blue: 29/11/2012 - 1100 UTC, orange: 21/03/2017 - 1200 UTC, green: 21/03/2017 - 2000 UTC) at the South Namib station.
@@
 \textbf{d}: Monthly distribution of ill-processed vertical profiles.}
-\label{Fig9_supp}
-\end{figure}
-%---------------------------------------------
-
-%---------------------------------------------
-\begin{figure}[p]
-\centering
-\includegraphics{Figures/Figure10_supp.pdf}
-\caption{Distributions of the meteorological parameters resulting from the processing of the ERA5-Land data for the South Namib (blue) and the Deep Sea (orange) stations.
-\pc{axis label : panel c -velocity at (height) H}.
-\pc{Is this for all day/night/seasons ?}}
 \label{Fig10_supp}
@@
 \centering
-\includegraphics[scale=1]{Figures/Figure11_supp.pdf}
-\caption{Non-dimensional parameters distributions. For the marginal \pc{???} distributions, the orange correspond to the South Namib station, and the blue to the Deep Sea station.}
+\includegraphics{Figures/Figure11_supp.pdf}
+\caption{Distributions of the meteorological parameters resulting from the processing of the ERA5-Land data for the South Namib (blue) and the North Namib (orange) stations.
+\pc{axis label : panel c -velocity at (height) H}.
+\pc{Is this for all day/night/seasons ?}}
 \label{Fig11_supp}
@@
 \includegraphics[scale=1]{Figures/Figure12_supp.pdf}
-\caption{Regime diagrams of the wind deviation $\delta_{\theta}$ and relative attenuation/amplification $\delta_{u}$ in the spaces $(\mathcal{F}_{\textup{I}}, \, kH)$ and $(\mathcal{F}_{\textup{I}}, \, \mathcal{F})$, containing the data from both the Deep Sea and South Namib stations.
-Blue dashed lines empirically delimit the different regimes.
-The point density in each bin of the diagrams is shown in Suppl.
-Fig.~\ref{Fig11_supp}.
-The similar regime diagrams in the space $(\mathcal{F}, \, kH)$ are shown in Fig.~\ref{Fig5}.}
+\caption{Non-dimensional parameters distributions. For the marginal \pc{???} distributions, the orange correspond to the South Namib station, and the blue to the North Namib station.}
 \label{Fig12_supp}
@@
 \includegraphics[scale=1]{Figures/Figure13_supp.pdf}
+\caption{Regime diagrams of the wind deviation $\delta_{\theta}$ and relative attenuation/amplification $\delta_{u}$ in the spaces $(\mathcal{F}_{\textup{I}}, \, kH)$ and $(\mathcal{F}_{\textup{I}}, \, \mathcal{F})$, containing the data from both the North Namib and South Namib stations.
+Blue dashed lines empirically delimit the different regimes.
+The point density in each bin of the diagrams is shown in Suppl.
+Fig.~\ref{Fig11_supp}.
+The similar regime diagrams in the space $(\mathcal{F}, \, kH)$ are shown in Fig.~\ref{Fig5}.}
+\label{Fig13_supp}
+\end{figure}
+%---------------------------------------------
+
+%---------------------------------------------
+\begin{figure}[p]
+\centering
+\includegraphics[scale=1]{Figures/Figure14_supp.pdf}
 \caption{Computation of the flow disturbance with the linear model of \citet{andreotti2009}.
@@
 From the upper to the lower streamline, values of $\left(kH,\, \mathcal{F},\, \mathcal{F}_{\rm I},\, \mathcal{A}_{0},\, \mathcal{B}_{0},\, \sqrt{\mathcal{A}_{0}^{2} + \mathcal{B}_{0}^{2}}\right)$ are (1.9, 0.6, 1.5, 3.4, 1.0, 3.5), (1.5, 0.3, 0.4, 4.8, 1.4, 5.0), (0.1, 3.5, 1.0, 8.6, 0.1, 8.6), (0.5, 0.05, 0.04, 9.6, 2.5, 9.9).}
-\label{Fig13_supp}
+\label{Fig14_supp}
 \end{figure}

### 2022-02-28 Paper_figure/Figure02.py

@@ -9,11 +9,14 @@ import numpy as np
 import matplotlib.pyplot as plt
 import matplotlib.dates as mdates
 import calendar
+import locale
 from datetime import datetime, timedelta
 import sys
 import os
 sys.path.append('../')
 import python_codes.theme as theme
+
+locale.setlocale(locale.LC_ALL, 'en_US.utf8')
 
 
 def tick_formatter(ax, fmt='%d'):
@@ -36,21 +39,23 @@ Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=Tru
 Stations = sorted(Data.keys())
 
 # Figure properties
-variables = ['U', 'Orientation']
-label_var = {'U': r'Velocity, $u_{*}~[\textup{m}~\textup{s}^{-1}]$', 'Orientation': r'Orientation, $\theta~[^\circ]$'}
-labels = [(r'\textbf{a}', r'\textbf{b}'), (r'\textbf{c}', r'\textbf{d}'), (r'\textbf{e}', r'\textbf{f}')]
-row_labels = ['Huab station', 'Deep Sea station -- summer', 'Deep Sea station -- winter']
-years = [2018, 2017, 2017]
-months = [2, 12, 6]
-days = [(11, 14), (5, 8), (1, 4)]
+variables = ['U_star', 'Orientation']
+label_var = {'U_star': r'Velocity, $u_{*}~[\textup{m}~\textup{s}^{-1}]$', 'Orientation': r'Orientation, $\theta~[^\circ]$'}
+labels = [(r'\textbf{a}', r'\textbf{b}'), (r'\textbf{c}', r'\textbf{d}'),
+          (r'\textbf{e}', r'\textbf{f}'), (r'\textbf{g}', r'\textbf{h}')]
+row_labels = ['Adamax station -- summer', 'Adamax station -- winter', 'North Namib station -- summer',
+              'North Namib station -- winter']
+years = [2015, 2016, 2017, 2017]
+months = [11, 8, 12, 6]
+days = [(10, 14), (17, 20), (5, 8), (1, 4)]
 month_calendar = {index: month for index, month in enumerate(calendar.month_name) if month}
 
 
-stations_plot = ['Huab_Station', 'Deep_Sea_Station', 'Deep_Sea_Station']
+stations_plot = ['Adamax_Station', 'Adamax_Station', 'Deep_Sea_Station', 'Deep_Sea_Station']
 
 # #### Figure
[…]

### 2022-02-28 Paper_figure/Supplementary_Figures/Figure01_supp.py

@@ -64,10 +64,10 @@ for station in Stations:
 
 plt.xlabel('time [years]')
 plt.gca().set_yticks(centers)
-plt.gca().set_yticklabels([station[:-8].replace('_', ' ') for station in Stations])
+plt.gca().set_yticklabels([station[:-8].replace('_', ' ') if station != 'Deep_Sea_Station' else 'North Namib' for station in Stations])
 ptch_Era5Land = mpatches.Patch(color=colors[0], label='Era5Land/Era5')
-ptch_InSitu = mpatches.Patch(color=colors[1], label='In situ')
-plt.legend(handles=[ptch_Era5Land, ptch_InSitu], loc='lower left')
+ptch_InSitu = mpatches.Patch(color=colors[1], label='local measurements')
+plt.legend(handles=[ptch_Era5Land, ptch_InSitu], loc='lower left', bbox_to_anchor=(0, 0.1))
 
 
 plt.savefig(os.path.join(path_savefig, 'Figure1_supp.pdf'))

### 2022-02-28 Paper_figure/Supplementary_Figures/Figure03_supp.py

@@ -49,7 +49,7 @@ plt.ylim([Data_roughness['z0_insitu_vals'].min(), Data_roughness['z0_insitu_vals
 plt.gca().set_xscale('log')
 plt.gca().set_yscale('log')
 fig.supxlabel(r'Hydrodynamic roughness - ERA, $z_{0}^{\textup{Era5Land}}$ [m]')
-fig.supylabel(r'Hydrodynamic roughness - in situ, $z_{0}^{\textup{local}}$ [m]')
+fig.supylabel(r'Hydrodynamic roughness - local, $z_{0}^{\textup{local}}$ [m]')
 
 # colorbar
 sm = plt.cm.ScalarMappable(cmap='viridis', norm=norm)

### 2022-02-28 Paper_figure/Supplementary_Figures/Figure05_supp.py

@@ -6,58 +6,87 @@ Figure 5 -- SI
 """
 
 import numpy as np
-import os
 import matplotlib.pyplot as plt
+import matplotlib.dates as mdates
+import calendar
+import locale
+from datetime import datetime, timedelta
 import sys
+import os
 sys.path.append('../../')
 import python_codes.theme as theme
-from python_codes.plot_functions import plot_scatter_surrounded
+
+locale.setlocale(locale.LC_ALL, 'en_US.utf8')
 
 
+def tick_formatter(ax, fmt='%d'):
+    myFmt = mdates.DateFormatter(fmt)
+    ax.xaxis.set_major_formatter(myFmt)
+    ticklabels = ax.get_xticklabels()
+    ticklabels[0].set_ha('left')
+
+
+# Loading figure theme
 theme.load_style()
 
 # paths
 path_savefig = '../../Paper/Figures'
 path_outputdata = '../../static/data/processed_data/'
 
+# Loading wind data
 Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
-labels = [r'\textbf{a}', r'\textbf{b}']
+Stations = sorted(Data.keys())
 
-# preparing data
-Stations_ref = ['Adamax_Station', 'Huab_Station']
-#
-Theta_ERA = np.concatenate([Data[station]['Orientation_era'] for station in Stations_ref])
-Theta_Station = np.concatenate([Data[station]['Orientation_insitu'] for station in Stations_ref])
-#
-U_ERA = np.concatenate([Data[station]['U_star_era'] for station in Stations_ref])
-U_Station = np.concatenate([Data[station]['U_star_insitu'] for station in Stations_ref])
+# Figure properties
+variables = ['U_star', 'Orientation']
+label_var = {'U_star': r'Velocity, $u_{*}~[\textup{m}~\textup{s}^{-1}]$', 'Orientation': r'Orientation, $\theta~[^\circ]$'}
+labels = [(r'\textbf{a}', r'\textbf{b}'), (r'\textbf{c}', r'\textbf{d}'),
+          (r'\textbf{e}', r'\textbf{f}'), (r'\textbf{g}', r'\textbf{h}')]
+row_labels = ['Huab station -- summer', 'Huab station -- winter', 'South Namib station -- summer',
+              'South Namib station -- winter']
[…]

### 2022-02-28 Paper_figure/Supplementary_Figures/Figure06_supp.py

@@ -6,76 +6,58 @@ Figure 6 -- SI
 """
 
 import numpy as np
+import os
 import matplotlib.pyplot as plt
 import sys
-import os
 sys.path.append('../../')
 import python_codes.theme as theme
-from python_codes.plot_functions import make_nice_histogram
+from python_codes.plot_functions import plot_scatter_surrounded
 
-# Loading figure theme
+
 theme.load_style()
 
 # paths
 path_savefig = '../../Paper/Figures'
 path_outputdata = '../../static/data/processed_data/'
 
-# Loading wind data
 Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
-Stations = sorted(Data.keys())
+labels = [r'\textbf{a}', r'\textbf{b}']
 
-# Figure properties
-station = 'South_Namib_Station'
+# preparing data
+Stations_ref = ['Adamax_Station', 'Huab_Station']
 #
-theta_bins = [[0, 140], [150, 260]]
-velocity_bins = [[0.05, 0.2], [0.2, 10]]
-Data_pattern = np.load(os.path.join(path_outputdata, 'Data_DEM.npy'), allow_pickle=True).item()[station]
-icon = [r'\faSun', r'\faMoon']
+Theta_ERA = np.concatenate([Data[station]['Orientation_era'] for station in Stations_ref])
+Theta_Station = np.concatenate([Data[station]['Orientation_insitu'] for station in Stations_ref])
+#
+U_ERA = np.concatenate([Data[station]['U_star_era'] for station in Stations_ref])
+U_Station = np.concatenate([Data[station]['U_star_insitu'] for station in Stations_ref])
 
-color_ax = 'purple'
 
-# ################ Figure
-fig, axarr = plt.subplots(3, 3, figsize=(theme.fig_width, 0.925*theme.fig_width), constrained_layout=True, sharex=True)
-for i in range(3):  # Loop over velocites
-    if i < 2:
-        mask_U = (Data[station]['U_star_era'] >= velocity_bins[i][0]) & (Data[station]['U_star_era'] <= velocity_bins[i][1])
-        label_u = r'$u_{*} < ' + str(velocity_bins[i][1]) + '$' if i == 0 else r'$u_{*} > ' + str(velocity_bins[i][0]) + '$'
-    else:
[…]

### 2022-02-28 Paper_figure/Supplementary_Figures/Figure10_supp.py

@@ -6,64 +6,106 @@ Figure 10 -- SI
 """
 
 import numpy as np
+import os
 import matplotlib.pyplot as plt
 import sys
-import os
 sys.path.append('../../')
 import python_codes.theme as theme
-from python_codes.meteo_analysis import mu
 from python_codes.plot_functions import make_nice_histogram
 
 
-# Loading figure theme
+def plot_vertical_profile(ax, height, Virtual_potential_temperature, grad_free_atm, theta_free_atm, blh, theta_ground, Hmax_fit, color='tab:blue', label=None):
+    Hfit = np.linspace(blh, Hmax_fit, 100)
+    #
+    line = ax.vlines(theta_ground, 0, blh/1e3, color=color, label=label, zorder=-3)
+    ax.axhline(blh/1e3, alpha=0.5, color=color, ls='--')
+    ax.plot(np.poly1d([grad_free_atm, theta_free_atm])(Hfit), Hfit/1e3, color=line.get_color(), zorder=-2)
+    ax.plot(Virtual_potential_temperature, height/1e3, '.', color=line.get_color(), zorder=-1)
+    # ax.scatter(theta_ground, blh/1e3, s=30, facecolors=line.get_color(), edgecolors='k', linewidth=2, zorder=0)
+
+
 theme.load_style()
 
 # paths
 path_savefig = '../../Paper/Figures'
 path_outputdata = '../../static/data/processed_data/'
 
-# ##### Loading meteo data
+# Loading data
 Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
 
-# ## histograms parameters
-Stations = ['South_Namib_Station', 'Deep_Sea_Station']
-g = 9.81  # [m/s2] gravitational constant
-z0_era = 1e-3  # [m] hydrodynamic roughness
-#
 labels = [r'\textbf{a}', r'\textbf{b}', r'\textbf{c}', r'\textbf{d}']
-nbins = 80
 
-# #### Figure
-fig, axarr = plt.subplots(2, 2, figsize=(theme.fig_width, 0.9*theme.fig_width),
-                          constrained_layout=True, sharey=True)
+# ## vertical profiles parameters
+station = 'Deep_Sea_Station'
+time_steps_bad = [10856, 30266, 33463]
+time_steps_good = [2012, 30302, 30310]
+colors = ['tab:blue', 'tab:orange', 'tab:green']
[…]

### 2022-02-28 Paper_figure/Supplementary_Figures/Figure11_supp.py

@@ -6,123 +6,64 @@ Figure 11 -- SI
 """
 
 import numpy as np
-import os
 import matplotlib.pyplot as plt
-from matplotlib.colors import LogNorm
 import sys
+import os
 sys.path.append('../../')
 import python_codes.theme as theme
+from python_codes.meteo_analysis import mu
 from python_codes.plot_functions import make_nice_histogram
 
+
+# Loading figure theme
 theme.load_style()
 
 # paths
 path_savefig = '../../Paper/Figures'
 path_outputdata = '../../static/data/processed_data/'
 
-# Loading data
+# ##### Loading meteo data
 Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
+
+# ## histograms parameters
 Stations = ['South_Namib_Station', 'Deep_Sea_Station']
-
-numbers = {key: np.concatenate([Data[station][key] for station in Stations]) for key in ('Froude', 'kH', 'kLB')}
-mask = ~np.isnan(numbers['Froude'])
-ad_hoc_quantity = np.concatenate([Data[station]['U_star_era'] for station in Stations])
-
-# Figure properties
-couples = [('Froude', 'kLB'), ('Froude', 'kH'), ('kLB', 'kH')]
-lims = {'Froude': (5.8e-3, 450), 'kLB': (0.009, 7.5), 'kH': (2.2e-2, 10.8)}
-# ax_labels = {'kH': r'$kH$', 'Froude': r'$\mathcal{F} =  U/\sqrt{(\Delta\rho/\rho) g H}$',
-# 'kLB': r'$\mathcal{F}_{\textup{I}} =  kU/N$'}
-ax_labels = {'kH': r'$kH$', 'Froude': r'$\mathcal{F}$',
-             'kLB': r'$\mathcal{F}_{\textup{I}}$'}
-norm = LogNorm(vmin=1, vmax=1.5e3)
+g = 9.81  # [m/s2] gravitational constant
+z0_era = 1e-3  # [m] hydrodynamic roughness
+#
+labels = [r'\textbf{a}', r'\textbf{b}', r'\textbf{c}', r'\textbf{d}']
+nbins = 80
 
 # #### Figure
-fig, axarr = plt.subplots(4, 3, figsize=(theme.fig_width, 0.95*theme.fig_width),
-                          # constrained_layout=True,
-                          gridspec_kw={'height_ratios': [0.2, 1.3, 2, 2],
[…]

### 2022-02-28 Paper_figure/Supplementary_Figures/Figure14_supp.py

@@ -1,0 +1,133 @@
+"""
+============
+Figure 13 -- SI
+============
+
+"""
+
+import numpy as np
+import matplotlib.pyplot as plt
+import sys
+import os
+sys.path.append('../../')
+import python_codes.theme as theme
+from python_codes.general import cosd, sind
+from python_codes.plot_functions import plot_regime_diagram
+from python_codes.linear_theory import Cisaillement_basal_rotated_wind
+
+
+def topo(x, y, alpha, k, xi):
+    return xi*np.cos(k*(cosd(alpha)*x + sind(alpha)*y))
+
+
+# Loading figure theme
+theme.load_style()
+
+# paths
+path_savefig = '../../Paper/Figures'
+path_outputdata = '../../static/data/processed_data/'
+
+# ## regime diagram properties
+# data
+Stations = ['South_Namib_Station', 'Deep_Sea_Station']
+Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
+numbers = {key: np.concatenate([Data[station][key] for station in Stations]) for key in ('Froude', 'kH', 'kLB')}
+
+# Time series hydrodynamic coefficients
+Hydro_coeffs_time = np.load(os.path.join(path_outputdata, 'time_series_hydro_coeffs.npy'), allow_pickle=True).item()
+modulus = np.linalg.norm(np.concatenate([Hydro_coeffs_time[station] for station in Stations], axis=1), axis=0)
+
+#
+couples = [('Froude', 'kH'), ('kLB', 'kH')]
+labels = [r'\textbf{a}', r'\textbf{b}', r'\textbf{c}']
+#
+ax_labels = {'Froude': r'$\mathcal{F} =  U/\sqrt{(\Delta\rho/\rho_{0}) g H}$', 'kH': '$k H$', 'kLB': r'$\mathcal{F}_{\textup{I}} = k U/N$'}
+lims = {'Froude': (5.8e-3, 450), 'kLB': (0.009, 7.5), 'kH': (2.2e-2, 10.8)}
+#
+regime_line_color = 'tab:blue'
+cbar_labels = [r'$\delta_{\theta}$ [deg.]', r'$\delta_{u}$']
+
+mask = ~np.isnan(numbers['Froude'])
+
+# ## streamline parameters
+station = Stations[1]
[…]

### 2022-02-28 Paper/main.tex
Entering Giles 'white' comments for abstract and introduction:
- Ignoring 'Arid areas'
- chossing something else for 'except for some weak winds blowing at
  night'

Keep inluding Giles comments and modifications:

- White comments until section3 (until page 9 included)
- changing ref to SI to 'Online Ressource ...' according to BLM template
- changing station names to 'North Sand Sea' and 'South Sand Sea' in
  text and figures

Changing title to Giles suggestion, and correction Mat affiliation

updating latexdiff before sending to Philippe

Phillipe corrections afetr Giles remark

Adding arro

--- 2022-01-29 Paper/main.tex
@@
 
-\title{Local wind regime induced by giant linear dunes}
+% \title{Local wind regime induced by giant linear dunes}
+\title{Local wind regime induced by giant linear dunes: comparison of  ERA5 re-analysis with surface measurements}
 
@@
 \and
-G.S.F. Wiggs \at
+G. Wiggs \at
 School of Geography and the Environment, University of Oxford, Oxford, UK.
@@
 M. Baddock \at
-Geography and Environment, Loughborough University of Technology, Loughborough, UK.
+Geography and Environment, Loughborough University, Loughborough, UK.
 \and
@@
 \begin{abstract}
-Emergence and growth of sand dunes result from the interaction between the topography, which perturbs the wind flow, which itself controls the sediment transport.
-While these feedbacks are well studied at the scale of a single dune, the average effect of a periodic dune pattern on atmospheric flows remains poorly constrained due to a lack of data in major sand seas.
-Here, we compare field measurements of surface wind data to the predictions of the ERA5-Land climate reanalysis in four different places in Namibia, inside and north of the giant-dune field of the Namib sand sea.
-In the flat desert areas to the north, observations and predictions agree well despite the small amount of data in the assimilation process.
-This is also the case in the interdune areas of the sand sea, except for the weak winds blowing at night in a direction oblique to the orientation of the giant dunes.
+Emergence and growth of sand dunes results from the dynamic interaction between topography, wind flow and sediment transport.
+While feedbacks between these variables are well studied at the scale of a single dune, the average effect of a periodic dune pattern on atmospheric flows remains poorly constrained due to a lack of data in major sand seas.
+Here, we compare field measurements of surface wind data to the predictions of the ERA5-Land climate reanalysis at four locations in Namibia, including within the giant-dune field of the Namib sand sea.
+In the desert plains to the north of the sand sea, observations and predictions agree well.
+This is also the case in the interdune areas of the sand sea, except for the weak winds blowing at night, which exhibit additional components aligned with the giant dune orientation, which are not predicted by the ERA5-Land reanalysis.
 We quantify these similarities and differences and provide a physical understanding of the relevant aerodynamical regimes to relate them to the daily cycle of the turbulent atmospheric boundary layer over a dune pattern of given wavelength.
-We conclude by identifying the conditions under which the ERA5-Land reanalysis data can or can't be used to study dune morphodynamics.
-We also propose that, in multidirectional wind regimes, deflections of specific winds could explain the occurrence of secondary dune patterns with a different orientation than the primary structures between which they grow.
+We conclude by identifying the conditions under which the ERA5-Land reanalysis data can reliably be used to study dune morphodynamics.
+We also propose that, in multidirectional wind regimes, deflections of specific winds by giant dunes could explain the occurrence of secondary dune patterns with a different orientation to the primary structures between which they develop.
 \keywords{Atmospheric boundary layer \and Sand dunes \and Fluid-structure interactions}
@@
 Importantly, the troposphere presents a vertical structure, with a lower convective boundary layer, of typical kilometer-scale thickness, capped by a stably stratified region \citep{Stull1988}.
-The largest topographic obstacles as mountains can therefore interact with this upper region, for example leading to internal wave generation or significant wind disturbances, such as downslope winds in lee sides \citep{Durran1990}.
+The largest topographic obstacles, such as mountains can therefore interact with this upper region, and lead to internal wave generation or significant wind disturbances, such as lee-side downslope winds \citep{Durran1990}.
 
-Looking at the wind close to the surface, two topographic feedbacks on the flow, although related, can be commented separately.
-There is a first effect on the wind intensity: the flow accelerates on the upwind slope and slows down on the downwind one \citep{Baddock2011}, with a speed-up factor essentially proportional to the obstacle aspect ratio \pc{here a ref?}.
+Focusing on the wind close to the surface, two related topographic feedbacks on the windflow over dunes can be commented on separately.
+First is the effect on wind speed with documented flow acceleration on upwind slopes \citep{Weaver2011} and deceleration on downwind slopes \citep{Baddock2007}, with a speed-up factor essentially proportional to the obstacle aspect ratio \citep{Jackson1975}.
 Importantly, the velocity maximum is typically shifted upwind of the obstacle crest.
 This behaviour has been theoretically predicted by means of asymptotic analysis of a neutrally stratified boundary-layer flow over an obstacle of vanishing aspect ratio \citep{Jackson1975, Mason1979, Sykes1980, Hunt1988, Belcher1998}.
-Experiments in flumes \citep{Zilker1977, Zilker1979, Frederick1988, Poggi2007}, in wind tunnel \citep{Gong1989, Finnigan1990, Gong1996} and in field conditions \citep{Taylor1987a, Claudin2013, Fernando2019, Lu2021}, have also documented this effect.
+Experiments in flumes \citep{Zilker1977, Zilker1979, Frederick1988, Poggi2007}, in wind tunnels \citep{Gong1989, Finnigan1990, Gong1996} and in field conditions \citep{Taylor1987a, Claudin2013, Fernando2019, Lu2021, Bristow2022}, have also documented this effect.
 Interestingly, a similar behaviour exists for the pressure perturbation, but with a slight downwind shift for the pressure minimum \citep{Claudin2021}.
-The second effect is the flow deflection that occurs when the incident wind direction is not perpendicular to the ridge crest.
-While predicted to be small (less than $10^{\circ}$) in the linear regime valid for shallow topography \citep{Gadal2019}, significant flow steering has been reported in the field on the downwind side of steep enough obstacles, such as mountain ranges \citep{Kim2000, Lewis2008, Fernando2019} or well developed sand dunes \citep{Walker2009, Hesp2015, Walker2017, Smith2017}.
+The second effect, much less studied, is the flow deflection that occurs when the incident wind direction is not perpendicular to the ridge crest.
+While predicted to be small (less than $10^{\circ}$) in the linear regime valid for shallow topography \citep{Gadal2019}, significant flow steering has been reported in the field on the downwind side of steep enough obstacles, such as mountain ranges \citep{Kim2000, Lewis2008, Fernando2019}, well-developed sand dunes \citep{Walker2009, Hesp2015, Walker2017, Smith2017}, and valley topographies \citep{Wiggs2002}.
 
-For practical reasons, wind measurement over sand dunes have been performed over rather small bedforms, typically a few meters high only -- that is tens of meters in length.
-Giant dunes, with kilometer-scale wavelengths and heights of tens of meters, are more difficult to investigate although they provide, for several reasons, a choice configuration for the study of turbulent flows over a complex topography.
-First of all, one expects larger wind disturbances for larger obstacles.
-Second, their large size make them interact with the vertical structure of the atmosphere \citep{andreotti2009}.
-Also, they usually form large patterns in sand seas and thus behave as rather clean periodic perturbations, in contrast with isolated hills.
-Finally, because the morphodynamics of aeolian bedforms are strongly dependent on the local wind regime \citep{Livingstone1996}, one can expect to see the consequences of the wind disturbances on smaller dunes, in a similar way as what has been reported for the effect of dunes on impact ripples \citep{Howard1977,Hood2021}.
+For practical reasons, wind measurement over sand dunes has been performed over rather small bedforms, typically a few meters high only (corresponding to tens of meters long) \citep{Wiggs1996}.
+Giant dunes, with kilometer-scale wavelengths and heights of tens of meters, are more difficult to investigate although for several reasons they provide a choice configuration for the study of turbulent flows over a complex topography.
+First, one expects larger wind disturbances for larger obstacles.
+Secondly, their large size makes them interact with the vertical structure of the atmosphere \citep{Andreotti2009}.
+Third, they usually form large patterns in sand seas and thus behave as rather clean periodic perturbations, in contrast with isolated dunes.
+Finally, because the morphodynamics of aeolian bedforms are strongly dependent on the local wind regime \citep{Livingstone2019}, one can expect to see the consequences of windflow disturbance by large dunes on neighbouring small dunes, in a similar manner to that reported for the effect of dunes on impact ripples \citep{Howard1977,Hood2021}.
 
-Arid areas have been much studied at the desert scale from climate reanalyses based on global atmospheric models \citep{Blumberg1996, Livingstone2010, Ashkenazy2012, Jolivet2021, Hu2021}, such as ERA-40, ERA-Interim or ERA-5 \citep{Uppala2005, Dee2011, Hersbach2020}.
+Atmospheric flows have been much studied at the desert-scale with climate reanalyses based on global atmospheric models \citep{Blumberg1996, Livingstone2010, Ashkenazy2012, Jolivet2021, Hu2021}, such as ERA-40, ERA-Interim or ERA-5 \citep{Uppala2005, Dee2011, Hersbach2020}.
 However, the spatial resolution (tens of kilometers) of these reanalyses implies average quantities that do not resolve the smaller scales, ranging from individual dunes to the border of mountains \citep{Livingstone2010}.
-Lately, the release of ERA5-Land allows to push back this limitation by providing up to 70 years of hourly wind predictions at a $9$~km spatial resolution \citep{munoz2021}.
-However, its validity remains to be studied, especially in areas where data assimilation is low (such as deserts).
+Recently, the release of ERA5-Land has resolved this limitation by providing up to 70 years of hourly wind predictions at a $9$~km spatial resolution \citep{munoz2021}.
+However, its validity remains to be studied, especially in remote desert areas where data assimilation is very low.
 
-In this work, we compare local wind speeds and directions measured in four different places inside and north of the giant-dune field of the Namib sand sea to the regional predictions of the ERA5-Land climate reanalysis.
-When the meteorological stations are surrounded by a relatively flat environment, we show that local measurements and regional predictions agree well with each other.
-The find that the agreement is also good in the interdune areas of the sand sea, except for the weak winds blowing at night in a direction oblique to the orientation of the giant dunes (section 2).
-Furthermore, we are able to link the magnitude of these differences to the circadian cycle of the atmospheric boundary layer (section 3).
-Finally, we draw implications of the wind disturbances on smaller-scale dunes (section 4).
+In this work, we compare local wind speeds and directions measured by meteorological stations at four different locations inside and north of the giant-dune field of the Namib sand sea to the regional predictions of the ERA5-Land climate reanalysis.
+Where the meteorological stations are surrounded by a relatively flat environment, we show that local measurements and regional predictions agree well.
+The agreement is also good in the interdune areas of the sand sea, except for some weak winds blowing at night, which exhibit an additional component aligned with the giant dune orientation, which are not predicted by the ERA5Land reanalysis (section 2).
+Further, we are able to link the magnitude of these differences to the circadian cycle of the atmospheric boundary layer (section 3).
+Finally, we draw implications for the wind disturbances on smaller-scale dunes (section 4), suggesting a possible origin for crossing dunes.
 
@@
 \caption{Wind data used in this study \textbf{a}: Location of the different sites in Namibia.
-\textbf{b-e}: Satellite images of these different environements (Google-Earth, Maxar Technologies, CNES/Airbus).
+\textbf{b--e}: Satellite images of these different environements (Google-Earth, Maxar Technologies, CNES/Airbus).
 In each subplot, the left and right wind roses represent the data from the ERA5-Land climate reanalysis and the local wind stations, respectively.
@@
 \section{Wind regimes across the Namib Sand Sea}
-We have measured the wind regime in four different sites of Namibia, representative of various environments across and nearby the Namib desert (Fig.~\ref{Fig1}).
-The Adamax station is located near the Adamax salt pan, in a highly vegetated area.
-The Huab station, on the coast at the outlet of the Huab river, is in an arid place where $60$-m scale barchan dunes develop.
-These two stations are located in a relatively flat \pc{quantify how flat is flat ?} environment.
-In contrast, the Deep Sea and South Namib stations are located in the interdune between tens of meters high giant linear dunes with kilometer-scale wavelengths and superimposed patterns.
+We measured the wind regime at four different locations in Namibia, representative of various arid environments across the Namib desert (Fig.~\ref{Fig1}, Online Resource Fig.~\ref{Fig1_supp}).
+The Adamax station was located at the Adamax waterhole to the west of Etosha Pan in northern Namibia, in a sparsely vegetated area.
+The Huab station was near the coast on a hyper-arid flat gravel plain lying above the ephemeral Huab river.
+Here, 5--6~m high barchan dunes develop from the sediment blowing out of the river valley.
+These two stations were located in a relatively flat environments.
+In contrast, the North Sand Sea and South Sand Sea stations were located in the interdunes between 50--100~m high linear dunes with kilometer-scale wavelengths and superimposed patterns.
 In this section, we describe and compare winds from local measurements and climate reanalysis predictions.
@@
 \subsection{Wind and elevation data}
-Local winds are measured at meteorological stations located in these four different places (black dots in Fig.~\ref{Fig1}).
-The wind strength and direction are sampled every 10 minutes by cup anemometers and wind vanes, at heights between $2$~m and $3$~m depending on the station.
-The available period of measurements ranges from 1 to 5 discontinuous years distributed between 2012 and 2020 (Suppl.
-Fig.~\ref{Fig1_supp}).
-We checked that at least one complete seasonal cycle is available at each station.
-Regional winds are extracted at the same locations and periods from the ERA5-Land dataset, which is a replay at a smaller spatial resolution of ERA5, the latest climate reanalysis from the ECMWWF \citep{Hersbach2020, munoz2021}.
-It provides hourly predictions of the 10-m wind velocity and direction at a spatial resolution of $0.1^\circ\times0.1^\circ$ ($\simeq 9~\textup{km}$ in Namibia).
+At each meteorological station (Fig.~\ref{Fig1}), wind speed and direction were sampled every 10 minutes using cup anemometers and wind vanes at heights between 2~m and 3~m depending on the station.
+The available period of measurements at each station ranged from 1 to 5 discontinuous years distributed between 2012 and 2020 (Online Resource Fig.~\ref{Fig2_supp}).
+We checked that at least one complete seasonal cycle was available for each station.
+Regional winds were extracted at the same locations and periods from the ERA5-Land dataset, which is a replay at a smaller spatial resolution of ERA5, the latest climate reanalysis from the ECMWWF \citep{Hersbach2020, munoz2021}.
+This dataset provided hourly predictions of the 10-m wind velocity and direction at a spatial resolution of $0.1^\circ\times0.1^\circ$ ($\simeq 9~\textup{km}$ in Namibia).
 
-For comparison, the local measurements are averaged into 1-hr bins centered on the temporal scale of the ERA5-Land estimates (Suppl.
-Fig.~\ref{Fig2_supp}).
-As the wind velocities of both datasets are provided at different heights, we convert them into shear velocities (Suppl.
-Mat.
-section 1), characteristic of the turbulent wind profile, which are then used together with the wind direction for further analysis.
+To enable direct comparison, the local wind measurements were averaged into 1-hr bins centered on the temporal scale of the ERA5-Land estimates (Online Resource Fig.~\ref{Fig3_supp}).
+As the wind velocities of both datasets were provided at different heights, we converted them into shear velocities $u_{*}$ (Online Resource section 1), characteristic of the turbulent wind profile.
 Wind roses in Fig.~\ref{Fig1}(b--e) show the resulting wind data.
 
-Dune properties are computed using autocorrelation on the 30-m Digital Elevation Models (DEMs) of the shuttle radar topography mission \citep{Farr2007}.
-For the South Namib and Deep Sea stations, we obtain respectively orientations of $85^\circ$ and $125^\circ$ with respect to the North, wavelengths of $2.6~\textup{km}$ and $2.3~\textup{km}$ and amplitudes of $45~\textup{m}$ and $20~\textup{m}$ (Suppl.
-Fig.~\ref{Fig4_supp} for more details).
+Dune properties were computed using autocorrelation on the 30-m Digital Elevation Models (DEMs) of the shuttle radar topography mission \citep{Farr2007}.
+For the South Sand Sea and North Sand Sea stations, we obtain, respectively, orientations of $85^\circ$ and $125^\circ$ with respect to the North, wavelengths of $2.6~\textup{km}$ and $2.3~\textup{km}$ and amplitudes (or half-heights) of $45~\textup{m}$ and $20~\textup{m}$ (Online Resource Fig.~\ref{Fig5_supp} for more details).
 
@@
 \caption{Temporal comparison between the wind data coming from the ERA5-Land climate reanalysis (orange lines) and from the local measurements (blue lines).
-Coloured swathes indicate day (between 1000 UTC and 2200 UTC) and night (before 1000 UTC or after 2200 UTC) \textbf{a-b}: Huab station.
-\textbf{c-d}: Deep Sea station in winter.
-\textbf{e-f}: Deep Sea station in summer.}
+Coloured swathes indicate day (between 1000 UTC and 2200 UTC) and night (before 1000 UTC or after 2200 UTC) \textbf{a--b}: Adamax station in summer.
+\textbf{b--c}: Adamax station in winter.
+\textbf{d--e}: North Sand Sea station in summer.
+\textbf{f--g}: North Sand Sea station in winter.
+Time series of the two other stations are shown in Online Resource Fig.~\ref{Fig6_supp}.}
 \label{Fig2}
@@
 \label{section_data_feedback}
-The obtained wind regimes are shown in figure~\ref{Fig1}.
-In the Namib, the regional wind patterns are essentially controlled by the see breeze, resulting in strong northward components (sometimes slightly deviated by the large scale topography \pc{mountain?}) present in all regional wind roses \citep{lancaster1985}.
-These daytime winds are dominant during the second-half of the year (September-January).
-In winter, an additional easterly component can be recorded during the night, induced by the combination of katabatic winds forming in the mountains, and infrequent `berg' winds, which are responsible of the high wind velocities observed \citep{lancaster1984}.
-The frequency of these easterly components decreases from the inland to the coast.
+The obtained measured and predicted wind regimes are shown in figure~\ref{Fig1}.
+In the Namib, the regional wind patterns are essentially controlled by the see breeze, resulting in strong northward components (sometimes slightly deviated by the large scale topography) present in all regional wind roses \citep{lancaster1985}.
+These daytime winds are dominant during the period October-March (Fig.~\ref{Fig2}f and Online Resource Fig.~\ref{Fig6_supp}f).
+During April-September, an additional easterly component can be recorded during the night, induced by the combination of katabatic winds forming in the mountains, and infrequent `berg' winds, which are responsible for the high wind velocities observed \citep{lancaster1984}.
+The frequency of these easterly components decreases from inland to the coast.
 As a result, bidirectional wind regimes within the Namib Sand Sea and at the Adamax salt pan (Fig.~\ref{Fig1}b,d,e) and a unidirectional wind regime on the coast at the outlet of the Huab River (Fig.~\ref{Fig1}c) are observed.
 
-In the case of the Adamax and Huab stations, the time series of wind speed and direction from the regional predictions quantitatively match those corresponding to the local measurements (Fig.~\ref{Fig2}a,b) and Suppl.
-Fig.~\ref{Fig5_supp}).
-For the Deep Sea and South Namib stations within the giant dune field, we observe that this agreement is also good, but limited to the September-January time period (Fig.~\ref{Fig2}c,d).
+In the case of the Adamax and Huab stations, the time series of wind speed and direction from the regional predictions quantitatively match those corresponding to the local measurements (Fig.~\ref{Fig2}a--d) and Online Resource Fig.~\ref{Fig6_supp}a--d,~\ref{Fig7_supp}).
+For the North Sand Sea and South Sand Sea stations within the giant dune field, we observe that this agreement is also good, but limited to the October-March time period (Fig.~\ref{Fig2}e--h and and Online Resource Fig.~\ref{Fig6_supp}e--h).
 As a matter of fact, the measured wind roses exhibit additional wind components aligned with the giant dune orientation, as evidenced on the satellite images (Fig.~\ref{Fig1}c,d).
 
-More precisely, during the February-August period, the local and regional winds in the interdune match during daytime only, i.e when the southerly/southwesterly sea breeze dominates (Figs.~\ref{Fig2}(e,f) and \ref{Fig3}, Suppl.
-Fig.~\ref{Fig6_supp}).
+More precisely, during the April-September period, the local and regional winds in the interdune match during daytime only, i.e when the southerly/southwesterly sea breeze dominates (Figs.~\ref{Fig2}e,f and \ref{Fig3}, Online Resource Fig.~\ref{Fig8_supp}).
 In the late afternoon and during the night, when the northwesterly `berg' and katabatic winds blow, measurements and predictions differ.
-In this case, the angular wind distribution of the local measurements exhibits two additional modes corresponding to reversing winds aligned with the giant dune orientation (purple frame in Fig.~\ref{Fig3}, Suppl.
-Figs.~\ref{Fig6_supp} and \ref{Fig7_supp}).
-This deviation is also associated with a global attenuation of the wind strength (Suppl.
-Fig.~\ref{Fig8_supp}).
-Remarkably, all these figures show that this wind reorientation and attenuation processes typically occur at low wind velocities, typically for $u_{*} \lesssim 0.1~\textrm{m}~\textrm{s}^{-1}$.
-For shear velocities larger than $\simeq 0.25~\textrm{m}~\textrm{s}^{-1}$, the wind reorientation is not significant.
-Finally, for intermediate shear velocities, moments of reorientation along the dune crest and others with no reorientation can both be observed (Suppl.
-Fig.~\ref{Fig7_supp}).
-\pc{CHECK the above lines cf.
-$\delta u$}
+In this case, the angular wind distribution of the local measurements exhibits two additional modes corresponding to reversing winds aligned with the giant dune orientation (purple frame in Fig.~\ref{Fig3}, Online Resource Figs.~\ref{Fig8_supp} and \ref{Fig9_supp}).
+This deviation is also associated with a general attenuation of the wind strength (Online Resource Fig.~\ref{Fig10_supp}).
+Remarkably, all these figures show that these wind reorientation and attenuation processes occur only at low wind velocities, typically for $u_{*} \lesssim 0.1~\textrm{m}~\textrm{s}^{-1}$.
+For shear velocities larger than $\simeq 0.25~\textrm{m}~\textrm{s}^{-1}$, the wind reorientation is not apparent.
+Finally, for intermediate shear velocities, both situations of wind flow reoriented along the dune crest and not reoriented can be successively observed (Online Resource Fig.~\ref{Fig9_supp}).
 
@@
 \includegraphics[scale=1]{Figures/Figure3.pdf}
-\caption{Distributions of wind direction at the Deep Sea Station for the ERA5-Land climate reanalysis (orange) and the local measurements (blue).
+\caption{Distributions of wind direction at the North Sand Sea Station for the ERA5-Land climate reanalysis (orange) and the local measurements (blue).
 In each subplot, both distributions are plotted from the same time steps, selected with constraints on the wind direction (columns) and/or wind velocity (rows) of the ERA5-Land dataset.
-The gray vertical dashed lines indicate the dune orientation.
-The numbers at the top right give the percentage of time steps selected in each sub-range, as well as the percentage of them corresponding to the day (between 1000 UTC and 2200 UTC).
-The purple frame highlights the regime (small wind velocities, nocturnal summer wind) in which the data from both datasets differ.
-A similar figure can be obtained for the Deep Sea station (Suppl.
-Fig.~\ref{Fig6_supp}).}
+The grey vertical dashed lines indicate the dune orientation.
+The numbers at the top right give the percentage of time steps selected in each sub-range, as well as the percentage corresponding to the daytime (between 1000 UTC and 2200 UTC).
+The purple frame highlights the regime (low wind velocities, nocturnal easterly wind) in which the data from both datasets differ.
+A similar figure can be obtained for the North Sand Sea station (Online Resource Fig.~\ref{Fig8_supp}).}
 \label{Fig3}
@@
 \section{Influence of wind speed and circadian cycle on the atmospheric boundary layer}
-The wind deflection induced by linear dunes has so far mainly been related to the angle between wind direction and crest orientation, with a maximum deflection for angles between $30^{\circ}$ and $70^{\circ}$ \citep{Walker2009, Hesp2015}.
-In the analysed data, the most deflected wind at both Deep Sean and South Namib stations is the most perpendicular to the giant dunes (\pc{Fig.~\ref{***}}).
-The incident wind direction then does not seem to be the only parameter controlling the wind deflection.
-In contrast, a different behaviour is observed between low and high wind velocities, suggesting a change in hydrodynamical regime.
-In this section, we discuss the relevant parameters associated with the dynamical mechanisms that govern the interactions between the atmospheric boundary layer flow and topographical obstacles.
-This analysis allows us to provide a physics-based interpretation the observations.
+The wind deflection induced by linear dunes has previously been related to the incident angle between wind direction and crest orientation, with a maximum deflection evident for incident angles between $30^{\circ}$ and $70^{\circ}$ \citep{Walker2009, Hesp2015}.
+In the data analysed here, the most deflected wind at both the North and South Sand Sea stations is seen to be where the incident angle is perpendicular to the giant dunes (Figs.~\ref{Fig1} and \ref{Fig3}, Online Resource Fig.~\ref{Fig8_supp}).
+It therefore appears that the incident wind angle is not the dominant control on maximum wind deflection.
+Further, and as shown in Figure~\ref{Fig3}, winds of high and low velocities show contrasting behaviour in characteristics of deflection.
+This suggests a change in hydrodynamical regime between the winds.
+In this section, we discuss the relevant parameters associated with the dynamical mechanisms that govern the interactions between the atmospheric boundary layer flow and giant dune topographies.
+This analysis allows us to provide a physics-based interpretation of our measured wind data.
 
@@
 \includegraphics[scale=1]{Figures/Figure4.pdf}
-\caption{\textbf{a-b}: Vertical profiles of the virtual potential temperature at 2 different time steps (day - 03/11/2015 - 1200 UTC, night - 01/13/2013 - 0900 UTC) at the Deep Sea station.
+\caption{\textbf{a--b}: Vertical profiles of the virtual potential temperature at 2 different time steps (day - 03/11/2015 - 1200 UTC, night - 01/13/2013 - 0900 UTC) at the North Sand Sea station.
 Dots: data from the ERA5 reanalysis.
-Dashed lines: boundary layer height given by the ERA5 reanalysis (Suppl.
-Mat.
-section 2).
+Dashed lines: boundary layer height given by the ERA5 reanalysis (Online Resource section 2).
 Plain lines: vertical (boundary layer) and linear (free atmosphere) fits to estimate the stratification properties.
-\textbf{c-f}: Sketches representing the interaction between the giant dunes and the atmospheric flow for different meteorological conditions.
+\textbf{c--f}: Sketches representing the interaction between the giant dunes and the atmospheric flow for different meteorological conditions.
 \textbf{g}: Streamlines over a sinusoidal topography $\xi(x,y)$ qualitatively representing the effect of low, medium and strong flow confinement, in relation to the above panels (see Appendix~1 for more details).
-\pc{add arrow with reference wind}}
+The blue arrow indicates the undisturbed wind direction.}
 \label{Fig4}
@@
 \label{theoretical_framework}
-Taking as a reference the turbulent flow over a flat bed, the general framework of this study is the understanding and the description of the flow response to a bed modulation.
-Without loss of generality, we can consider in this context an idealised bed elevation in the form of parallel sinusoidal ridges, with wavelength $\lambda$ (or wavenumber $k = 2\pi/\lambda$) and amplitude $\xi_0$, and where the reference flow direction makes a given angle with respect to the ridge crest \citep{Andreotti2012}.
+Taking as a reference the turbulent flow over a flat bed, the general framework of our study is understanding and describing the flow response to a bed modulation (e.g.
+a giant dune).
+Without loss of generality, we can consider in this context an idealised bed elevation in the form of parallel sinusoidal ridges, with wavelength $\lambda$ (or wavenumber $k = 2\pi/\lambda$) and amplitude $\xi_0$, and where the reference flow direction makes a given incident angle with respect to the ridge crest \citep{Andreotti2012}.
 Part of this response, on which we focus here, is the flow deflection by the ridges.
@@
 An incident flow oblique to the ridge is then deflected towards lower pressure zones, i.e towards the crest.
-Turbulent dissipation tends to increase this effect downstream, resulting in along the crest wind deflection in the lee side~\citep{Gadal2019}.
+Turbulent dissipation tends to increase this effect downstream, resulting in wind deflection along the crest in the lee side~\citep{Gadal2019}.
 
@@
 This is the case for bedforms forming in open channel flows such as rivers~\citep{Fourriere2010, Unsworth2018}.
-This is also relevant for aeolian dunes as they evolve in the turbulent atmospheric boundary layer (ABL) capped by the stratified free atmosphere (FA) \citep{andreotti2009}.
-Two main mechanisms, associated with dimensionless numbers, must then be considered (Fig.~\ref{Fig4}).
+This is also relevant for aeolian dunes as they evolve in the turbulent atmospheric boundary layer (ABL) capped by the stratified free atmosphere (FA) \citep{Andreotti2009}.
+Two main mechanisms, associated with dimensionless numbers must then be considered (Fig.~\ref{Fig4}).
 First, topographic obstacles typically disturb the flow over a characteristic height similar to their length.
@@
 Emerging and small dunes, with wavelengths in the range $20$ to $100$~m, are not affected by the confinement, corresponding to $k H \gg 1$.
-For giant dunes with kilometric wavelengths, however, their interaction with the FA is significant \citep{andreotti2009}.
+For giant dunes with kilometric wavelengths, however, their interaction with the FA is significant \citep{Andreotti2009}.
 This translates into a parameter $kH$ in the range $0.02$--$5$, depending on the moment of the day and the season.
@@
 The intensity of the stratification, i.e.
-the amplitude of the gradient $\left | \partial_z \rho \right|$, also impact ability to deform the capping layer under the presence of an underlying obstacle, and thus affects the influence of flow confinement.
+the amplitude of the gradient $\left | \partial_z \rho \right|$, also impacts its ability to deform the capping layer under the presence of an underlying obstacle, and thus affects the influence of flow confinement.
 This can be quantified using the internal Froude number~\citep{Vosper2004, Stull2006, Sheridan2006, Hunt2006, Jiang2014} $\mathcal{F}_{\textup{I}} = k U/N$, where $N = \sqrt{-g \partial_z \rho / \rho_{0}}$ is the the Brunt-V\"ais\"al\"a frequency \citep{Stull1988}.
@@
 
-This theoretical framework in mind, the smallest wind disturbances are expected to occur during the day, when the ABL depth is the largest and comparable to the dune wavelength ($k H \gtrsim 1$), which correspond to a weak confinement situation (Fig.~\ref{Fig4}c,d).
-On the contrary, large wind disturbances are expected to occur during the night, when the confinement is mainly induced by a shallow ABL (Fig.~\ref{Fig4}e).
+With this theoretical framework in mind, and in the context of the measured wind data in the North and South Sand Sea stations, the smallest wind disturbances are expected to occur during the day, when the ABL depth is the largest and comparable to the dune wavelength ($k H \gtrsim 1$), which corresponds to a weak confinement situation (Fig.~\ref{Fig4}c,d).
+In contrast, large wind disturbances are expected to occur during the night, when the confinement is mainly induced by a shallow ABL (Fig.~\ref{Fig4}e).
 However, this strong confinement can be somewhat reduced in the case of strong winds, corresponding to large values of the Froude number (Fig.~\ref{Fig4}f).
-This is qualitative agreement with the transition from deflected to non-deflected winds related to low and high velocities observed in the data (Sec.~\ref{section_data_feedback}).
+This is in qualitative agreement with the transition from deflected to non-deflected winds related to low and high velocities observed in our data (Sec.~\ref{section_data_feedback}).
 
@@
 \includegraphics[scale=1]{Figures/Figure5.pdf}
-\caption{Regime diagrams of the wind deviation $\delta_{\theta}$ (\textbf{a}) and relative attenuation/amplification $\delta_{u}$ (\textbf{b}) in the space $(\mathcal{F}, \, kH)$, containing the data from both the Deep Sea and South Namib stations.
+\caption{Regime diagrams of the wind deviation $\delta_{\theta}$ (\textbf{a}) and relative attenuation/amplification $\delta_{u}$ (\textbf{b}) in the space $(\mathcal{F}, \, kH)$, containing the data from both the North Sand Sea and South Sand Sea stations.
 Blue dashed lines empirically delimit the different regimes.
-The point density in each bin of the diagrams is shown in Suppl.
-Fig.~\ref{Fig11_supp}.
-Similar regime diagrams in the spaces $(\mathcal{F}_{\textup{I}}, \, kH)$ and $(\mathcal{F}_{\textup{I}}, \, \mathcal{F})$ are shown in Suppl.
-Fig.~\ref{Fig12_supp}.
-\pc{ticks on x-axis ?}}
+The point density in each bin of the diagrams is shown in Online Resource Fig.~\ref{Fig13_supp} -- 95\% of the data occur in the range $-1 < \delta u < 1$.
+Similar regime diagrams in the spaces $(\mathcal{F}_{\textup{I}}, \, kH)$ and $(\mathcal{F}_{\textup{I}}, \, \mathcal{F})$ are shown in Online Resource Fig.~\ref{Fig14_supp}.}
 \label{Fig5}
@@
 %------------------------
-\subsection{Flow regime diagrams}
+\subsection{Data distribution in the flow regimes}
 We can go one step further and analyse how our data quantitatively spread over the different regimes discussed above.
 For that purpose, one needs to compute $kH$ and $\mathcal{F}$ from the time series.
-$H$, $U$ and the other atmospheric parameters can be deduced from the various vertical profiles (temperature, humidity) available in the ERA5 climate reanalysis (Suppl.
-Mat.
-section 2).
+$H$, $U$ and the other atmospheric parameters can be deduced from the various vertical profiles (temperature, humidity) available in the ERA5 climate reanalysis (Online Resource section 2).
 We quantify the flow deflection $\delta_{\theta}$ as the minimal angle between the wind orientations comparing the local measurements and the regional predictions.
@@
 Lower values of $k H$ (stronger interaction) or of Froude number (stronger confinement) both lead to an increase in wind disturbances, both in terms of orientation and velocity.
-Below a crossover value $k H \simeq 0.3$, wind disturbance is less sensitive to the $\mathcal{F}$-value, probably due to enhanced non-linear effects linked to strong flow modulation by the obstacle when confinement is strong \pc{more precise arguments?}.
+Below a crossover value $k H \simeq 0.3$, wind disturbance is less sensitive to the $\mathcal{F}$-value.
+This is probably due to enhanced non-linear effects linked to flow modulation by the obstacle when confinement is strong.
 The Froude number also controls a transition from damped to amplified wind velocities in the interdune, with a crossover around $\mathcal{F} \simeq 0.4$ (Fig.~\ref{Fig5}b).
 Such an amplification is rather unexpected.
-Checking the occurrence of the corresponding data, it appears that they are mainly associated with the wind coming from the South, when it blows at the end of day, but in summer time only although it is also present in winter \pc{an additional suppl.
-fig.
-?}.
-This effect may be linked to a change in the flow behaviour in the lee side of the obstacle (lee waves, hydraulic jumps, rotors) but further measurements are clearly needed in order to assess the different possibilities \citep{baines1995, Vosper2004}.
+Checking the occurrence of the corresponding data, it appears that these amplifications are associated with the southerly sea breeze, and occur dominantly during the October-March period, when the other easterly wind is not present (Online Resource Fig.~\ref{Fig15_supp}a--b).
+Furthermore, they occur less frequently during the afternoon, and more frequently at the end of the day (Online Resource Fig.~\ref{Fig15_supp}c).
+This effect may be linked to a change in the flow behaviour in the lee side of the obstacles but further measurements are clearly needed in order to assess the different possibilities \citep{baines1995, Vosper2004}.
 
@@
 \includegraphics[scale=1]{Figures/Figure6.pdf}
-\caption{Implications for smaller scale patterns in (\textbf{a}) the South Namib and (\textbf{b}) Deep Sea.
+\caption{Implications for smaller scale patterns in (\textbf{a}) the South Sand Sea and (\textbf{b}) North Sand Sea.
 The ellipses indicate the different types of elongating dunes, at large (plain line) and small (dashed line) scales.
@@
 %---------------------------------------------------------------------------------------
-\section{Discussion}
-The comparison of local (direct measurements) and regional (climate reanalysis) wind data reveals the giant dune feedback on the wind flow.
+\section{Discussion and conclusion}
+The feedback of the giant dunes on the wind flow has important implications for smaller scales bedforms.
+As illustrated in Fig.~\ref{Fig6}, small linear dunes ($\sim50$~m -wide) are often present in the 1--2~km interdune between giant linear dunes in the Namib Sand Sea.
+These smaller dunes do not exhibit the same orientation as the large ones, and are sometimes considered to be `crossing dunes'.
+Whilst differences between large and small scale dune patterns are observed ubiquitously, they are largely attributed to the presence of two different dune growth mechanisms, leading to two different dune patterns (orientations and/or morphologies) for the same wind regime~\citep{Courrech2014, Runyon2017, lu2017, Song2019, Gadal2020, Hu2021}.
+Here, however, our arguments enable the development of differing orientations for the small and giant linear dunes whilst also imposing the same dune growth mechanism (elongating mode).
+Figure~\ref{Fig6} shows how the orientations for the small and giant dunes can be derived from the locally measured and regionally predicted winds respectively (red arrows in Fig.~\ref{Fig6}).
+The feedback of the giant dunes on the wind described in this study thus provides an explanation for the existence of these small linear dunes elongating across the interdune, a dynamic which has remained unresolved to date.
+Further studies are of course needed to confirm this hypothesis.
+These crossing dunes could provide additional constraints for the inference of local winds from bedforms, similarly to what is currently performed on Mars using ripple orientations \citep{Liu2015, Hood2021}.
+
+Also, this study highlights the interaction between giant dunes and the atmospheric boundary layer.
+It thus supports the still debated idea that the capping layer acts as a bounding surface impacting the dune dynamics, and limiting dune growth in particular~\citep{Andreotti2009}, in contrast to an unconstrained growth ever-slower with size \citep{Eastwood2011, gunn2021}.
+Interestingly, this mechanism would allow for the inference of the ABL depth from the giant bedforms spacing where measurements are not feasible or available, as e.g.
+performed by~\citet{Lorenz2010} on Titan.
+
+To sum up conditions under which the ERA5-Land reanalysis data can reliably be used to study dune morphodynamics, we summarise the comparison of local (direct measurements) and regional (climate reanalysis) wind data.
 In flat areas, the agreement between the two confirms the ability of the ERA5-Land climate reanalysis to predict the wind regime down to scales $\sim10$~km, i.e the model grid.
@@
 
-The theoretical computation of the wind disturbances induced by sinusoidal ridges under flow confinement has been performed in the linear limit \citep{andreotti2009, Andreotti2012}, i.e.
+The theoretical computation of the wind disturbances induced by sinusoidal ridges under flow confinement has been performed in the linear limit \citep{Andreotti2009, Andreotti2012}, i.e.
 when the aspect ration of these ridges is small ($k\xi_0 \ll 1$).
-These models are able to qualitatively reproduce the observed wind deflection (Appendix 1, Suppl.
-Figs.~\ref{Fig12_supp} and \ref{Fig13_supp}), and thus provide the physical support for the hydrodynamic picture we propose here.
-They however cannot reach the magnitude of these observations, probably due to the presence of expected non-linearities in high confinement situations linked to strong flow modulations.
-Besides, these linear calculations only predict wind attenuation in the interdune, in contrast with the observed enhanced velocities associated with particular summer evening winds from the South.
-Some other models also predict different spatial flow structures such as lee waves and rotors~\citep{baines1995, Vosper2004}, which of course cannot be observed by our single point measurements.
+These models are able to qualitatively reproduce the observed wind deflection (Appendix 1, Online Resource Figs.~\ref{Fig14_supp} and \ref{Fig16_supp}), and thus provide the physical support for the interpretation we propose here based on hydrodynamic regimes.
+However, these models cannot quantitatively predict the magnitude of these observations, probably due to the presence of expected non-linearities in high confinement situations linked to strong flow modulations.
+Besides, these linear calculations only predict wind attenuation in the interdune, in contrast with the observed enhanced velocities associated with particular evening winds from the South during the period October-March (Online Resource Fig.~\ref{Fig15_supp}).
+Some other models predict different spatial flow structures in response to a modulated topography, such as lee waves and rotors \citep{baines1995, Vosper2004}.
+Of course they cannot be observed by our measurements located at a single point in the interdune.
 Data at different places along and across the ridges are needed to investigate and possibly map such flow structures, and for further comparisons with the models.
-
-This study highlights the interaction between giant dunes and the atmospheric boundary layer.
-It then supports the debated idea that the capping layer acts as a bounding surface impacting the dune dynamics, and limiting dune growth in particular~\citep{andreotti2009}, as opposed to an unconstrained growth ever-slower with size \citep{Eastwood2011, gunn2021}.
-Interestingly, this mechanism would allow for the inference of the ABL depth from the giant bedforms spacing where measurements are not feasible or available, as e.g.
-performed by~\citet{Lorenz2010} on Titan.
-
-This interaction also has important implications for smaller scales bedforms, as illustrated in Fig.~\ref{Fig6}.
-In the Namib Sand Sea, small linear dunes ($\sim50$~m -wide) are present in the interdune between giant linear dunes ($\sim2$~km -wide).
-Strangely enough, the small dunes do not have the same orientation as the large ones, and sometimes denotes as `crossing dunes'.
-While differences between large and small scale dune patterns are observed ubiquitously, they are now largely attributed to the presence of two different dune growth mechanisms, leading to two different dune patterns (orientations and/or morphologies) for the same wind regime~\citep{Courrech2014, Runyon2017, lu2017, Song2019, Gadal2020, Hu2021}.
-Here, however, we can get the orientations of the small and giant linear dunes from the same dune growth mechanism (elongating mode), respectively using the locally measured and regionally predicted winds (red arrows in Fig.~\ref{Fig6}).
-The feedback of the giant dunes on the wind described in this study then provides an explanation for the existence of these small linear dunes elongating across the interdune, as yet unresolved.
-Further studies are of course needed to confirm this hypothesis.
-These crossing dunes could provide additional constraints for the inference of local winds from bedforms, similarly to what is currently performed on Mars using ripple orientations \citep{Liu2015, Hood2021}.
 
@@
 
-[citing all grants ...] \pc{TOAD}
+[citing all relevant grants ...] \pc{TOAD}
 \end{acknowledgements}
@@
 
-Following the work of \citet{Fourriere2010}, \citet{Andreotti2012} and \citet{andreotti2009}, we briefly expose in this appendix the framework of the linear response of a turbulent flow to a topographic perturbation of small aspect ratio.
+Following the work of \citet{Fourriere2010}, \citet{Andreotti2012} and \citet{Andreotti2009}, we briefly describe in this appendix the framework for the linear response of a turbulent flow to a topographic perturbation of small aspect ratio.
 As a general bed elevation can be decomposed into Fourier modes, we focus here on a sinusoidal topography:
@@
 \begin{equation}
-\xi = \xi_{0}\cos\left[k\left(\cos(\alpha)x + \sin(\alpha)y\right)\right],
+\xi = \xi_{0}\cos\left[k\left(\cos(\alpha)y - \sin(\alpha)x\right)\right],
 \end{equation}
 %
-which is also a good approximation for the giant dunes observed in the Deep Sea and South Namib Station (Fig.~\ref{Fig1} and Suppl.
-Fig.~\ref{Fig4_supp}).
-$x$ and $y$ are the streamwise and spanwise coordinates, $k=2\pi/\lambda$ the wavenumber of the sinusoidal perturbation, and $\alpha$ its crest orientation with respect to the $y$-direction.
+which is also a good approximation for the giant dunes observed in the North Sand Sea and South Sand Sea Station (Fig.~\ref{Fig1} and Online Resource Fig.~\ref{Fig5_supp}).
+Here, $x$ and $y$ are the streamwise and spanwise coordinates, $k=2\pi/\lambda$ the wavenumber of the sinusoidal perturbation, $\alpha$ its crest orientation with respect to the $x$-direction (anticlockwise) and $\xi_{0}$ its amplitude.
 The two components of the basal shear stress $\boldsymbol{\tau} = \rho_{0} u_{*}\boldsymbol{u}_{*}$, constant in the flat bottom reference case, can then be generically written as:
@@
 \begin{align}
-\tau_{x} & = \tau_{0}\left(1 + k\xi_{0}\sqrt{\mathcal{A}_{x}^{2} + \mathcal{B}_{x}^{2}}\cos\left[k\left(\cos(\alpha)x + \sin(\alpha)y\right) + \phi_{x}\right]\right), \\
-\tau_{y} & = \tau_{0} \, k\xi_{0}\sqrt{\mathcal{A}_{y}^{2} + \mathcal{B}_{y}^{2}}\cos\left[k\left(\cos(\alpha)x + \sin(\alpha)y\right) + \phi_{y}\right],
+\tau_{x} & = \tau_{0}\left(1 + k\xi_{0}\sqrt{\mathcal{A}_{x}^{2} + \mathcal{B}_{x}^{2}}\cos\left[k\left(\cos(\alpha)y - \sin(\alpha)x\right) + \phi_{x}\right]\right), \\
+\tau_{y} & = \tau_{0} \, k\xi_{0}\sqrt{\mathcal{A}_{y}^{2} + \mathcal{B}_{y}^{2}}\cos\left[k\left(\cos(\alpha)y - \sin(\alpha)x\right) + \phi_{y}\right],
 \end{align}
@@
 \begin{align}
-  \mathcal{A}_{x} & = \mathcal{A}_{0}\cos^{2}\alpha, \\
-  \mathcal{B}_{x} & = \mathcal{B}_{0}\cos^{2}\alpha, \\
-  \mathcal{A}_{y} & = \displaystyle\frac{1}{2}\mathcal{A}_{0}\cos\alpha \sin\alpha, \\
-  \mathcal{B}_{y} & = \displaystyle\frac{1}{2}\mathcal{B}_{0}\cos\alpha \sin\alpha,
+  \mathcal{A}_{x} & = \mathcal{A}_{0}\sin^{2}\alpha, \\
+  \mathcal{B}_{x} & = \mathcal{B}_{0}\sin^{2}\alpha, \\
+  \mathcal{A}_{y} & = -\displaystyle\frac{1}{2}\mathcal{A}_{0}\cos\alpha \sin\alpha, \\
+  \mathcal{B}_{y} & = -\displaystyle\frac{1}{2}\mathcal{B}_{0}\cos\alpha \sin\alpha,
 \end{align}
 %
-where $\mathcal{A}_{0}$ and $\mathcal{B}_{0}$ are now two coefficients independent of the dune orientation $\alpha$, corresponding to the transverse case ($\alpha=0$).
-For a fully turbulent boundary layer capped by a stratified atmosphere, these coefficients depend on $k H$, $k z_{0}$, $\mathcal{F}$ and $\mathcal{F}_{\textup{I}}$ \citep{andreotti2009}.
-In this study, we assume a constant hydrodynamic roughness $z_{0} \simeq 1$~mm (Suppl.
-Mat.
-section 1).
+where $\mathcal{A}_{0}$ and $\mathcal{B}_{0}$ are now two coefficients independent of the dune orientation $\alpha$, corresponding to the transverse case ($\alpha=90$).
+For a fully turbulent boundary layer capped by a stratified atmosphere, these coefficients depend on $k H$, $k z_{0}$, $\mathcal{F}$ and $\mathcal{F}_{\textup{I}}$ \citep{Andreotti2009}.
+In this study, we assume a constant hydrodynamic roughness $z_{0} \simeq 1$~mm (Online Resource section 1).
 For the considered giant dunes, this leads to $k z_{0} \simeq 2 \cdot 10^{-6}$, as their wavelength is $\lambda \simeq 2.4$~km (or $k \simeq 2 \cdot 10^{-3}$~m$^{-1}$).
@@
 
-With capping a layer height and Froude numbers computed from the ERA5-Land time series, the corresponding $\mathcal{A}_{0}$ and $\mathcal{B}_{0}$ can be deduced, allowing to produce maps as displayed in Suppl.
-Fig.~\ref{Fig13_supp}.
-Interestingly, it shows similar regimes as in the diagrams of Fig.~\ref{Fig5} and Fig.~\ref{Fig12_supp}a,b, supporting the physics picture.
+With capping layer height and Froude numbers computed from the ERA5-Land time series, the corresponding $\mathcal{A}_{0}$ and $\mathcal{B}_{0}$ can be deduced, allowing to produce maps as displayed in Online Resource Fig.~\ref{Fig16_supp}.
+Interestingly, it shows similar regimes as in the diagrams of Fig.~\ref{Fig5} and Online Resource Fig.~\ref{Fig14_supp}a,b, supporting the physics picture.
 However, the matching remains qualitative only.
@@
 In accordance with the theoretical expectations, these coefficients present values on the order of unity ($\mathcal{A}_{0} \simeq 3$ and $\mathcal{B}_{0} \simeq 1$) in unconfined situations \citep{Claudin2013, Lu2021}.
-In contrast and as illustrated in Suppl.
-Fig.~\ref{Fig12_supp}a,b, larger values are predicted in case of strong confinement, which does not allow us to proceed to further quantitative comparison with the data.
+In contrast and as illustrated in Online Resource Fig.~\ref{Fig16_supp}a,b, larger values are predicted in case of strong confinement, which does not allow us to proceed to further quantitative comparison with the data.
 
-Finally, the linear model is also able to reproduce the enhancement of the flow deflection over the sinusoidal ridges when $\sqrt{\mathcal{A}_{0}^{2} + \mathcal{B}_{0}^{2}}$ is increased (Suppl.
-Fig.~\ref{Fig13_supp}).
-Here, using $k\xi_0 \simeq 0.1$ to be representative of the amplitude of the giant dunes at the Deep Sea station, the coefficient modulus is bounded to $10$.
+Finally, the linear model is also able to reproduce the enhancement of the flow deflection over the sinusoidal ridges when $\sqrt{\mathcal{A}_{0}^{2} + \mathcal{B}_{0}^{2}}$ is increased (Online Resource Fig.~\ref{Fig16_supp}).
+Here, using $k\xi_0 \simeq 0.1$ to be representative of the amplitude of the giant dunes at the North Sand Sea station, the coefficient modulus is bounded to $10$.
 
@@
 where the flux-up ratio $\gamma$ has been calibrated to $1.6$ using field studies, underwater laboratory experiments and numerical simulations.
-Predictions of the linear analysis of \citet{Gadal2019} give similar results.
+Predictions of the linear analysis of \citet{Gadal2019} and \citet{Delorme2020} give similar results.
 
@@
 Due to the lack of measurements in the studied places, some uncertainties can be expected.
-We therefore run a sensibility test by calculating the dune orientations for grain diameters ranging from $100~\mu\textup{m}$ to $400~\mu\textup{m}$ and for a speed-up ratio between $0.1$ and $10$ (wedges in Fig.~\ref{Fig6}).
+We therefore run a sensitivity test by calculating the dune orientations for grain diameters ranging from $100~\mu\textup{m}$ to $400~\mu\textup{m}$ and for a speed-up ratio between $0.1$ and $10$ (wedges in Fig.~\ref{Fig6}).
 
@@
 As the regionally predicted and locally measured velocities are available at different heights, we can not compare them directly.
-We then convert all velocities into shear velocities $u_{*}$, characteristic of the turbulent velocity profile \citep{Spalding1961, Stull1988}:
+We therefore convert all velocities into shear velocities $u_{*}$, characteristic of the turbulent velocity profile \citep{Spalding1961, Stull1988}:
 %
@@
 where $z$ is the vertical coordinate, $\kappa = 0.4$ the von K\'arm\'an constant and $z_{0}$ the hydrodynamic roughness.
-Several field measurements of hydrodynamic roughnesses are available \pc{Here cite a few old/general papers on roughness}.
-In the absence of sediment transport, it scales with the geometric features of the bed \citep{Pelletier2016} \pc{too specific paper?
-other older papers?}.
-When transport occurs, it rather typically scales with the thickness of the transport layer.
-For aeolian saltation, this is controlled by the altitude of Bagnold's focal point \citep{Duran2011,Valance2015}, which depends on the wind velocity and grain properties \citep{Sherman2008, Zhang2016, Field2018}.
-Whether associated with geometric features or with sediment transport, its typical order of magnitude is the millimetre scale.
+Several measurements of hydrodynamic roughnesses are available \citep{Raupach1992,Bauer1992,Brown2008,Nield2014}.
+In the absence of sediment transport, it is governed by the geometric features of the bed \citep{Flack2010,Pelletier2016}.
+When aeolian saltation occurs, it is rather controlled by the altitude of Bagnold's focal point \citep{Duran2011,Valance2015}, which depends on the wind velocity and grain properties \citep{Sherman2008, Zhang2016, Field2018}.
+Whether associated with geometric features or with sediment transport, its typical order of magnitude is the millimetre scale on sandy surfaces.
 
@@
 This parameter is computed for values of $z_0$ in ERA5-Land analysis ranging from $10^{-5}$~m to $10^{-2}$~m for the four different stations.
-Note that for the Deep Sea and South Namib stations, where the giant dunes feedback presumably affect the wind, we take into account the non-deflected winds only in the calculation of $\delta$ (with a $15^\circ$ tolerance).
+Note that for the North Sand Sea and South Sand Sea stations, where the giant dunes feedback presumably affect the wind, we take into account the non-deflected winds only in the calculation of $\delta$ (with a $15^\circ$ tolerance).
 
-As shown in Suppl.
-Fig.~\ref{Fig3_supp}, the minimum values of $\delta$ in the space ($z_{0}^{\textup{ERA5Land}}$, $z_{0}^{\textup{local}}$) form a line.
+As shown in Online Resource Fig.~\ref{Fig4_supp}, the minimum values of $\delta$ in the space ($z_{0}^{\textup{ERA5Land}}$, $z_{0}^{\textup{local}}$) form a line.
 We thus set the roughness in the ERA5-Land analysis to the typical value $z_0=10^{-3}$~m, and deduce the corresponding ones for the local stations.
-It leads to $2.7$, $0.8$, $0.1$ and $0.5$~mm for the Adamax, Deep Sea, Huab and South Namib stations, respectively.
+It leads to $2.7$, $0.8$, $0.1$ and $0.5$~mm for the Adamax, North Sand Sea, Huab and South Sand Sea stations, respectively.
 Importantly, this approach somewhat impacts the calculation of the shear velocities, but not that of the wind directions.
@@
 In arid areas, the vertical structure of the atmosphere can be approximated by a well mixed convective boundary layer of height $H$, topped by the stratified free atmosphere \citep{Stull1988, Shao2008}.
-In this context, one usually introduces the virtual potential temperature $T_{\textup{vp}}$, which is a constant $T_{0}$ inside the boundary layer, and increases linearly in the FA (Suppl.
-Fig.~\ref{Fig9_supp}a):
+In this context, one usually introduces the virtual potential temperature $T_{\textup{vp}}$, which is a constant $T_{0}$ inside the boundary layer, and increases linearly in the FA (Online Resource Fig.~\ref{Fig11_supp}a):
 %
@@
 where $\Delta T_{\textup{vp}}$ is the temperature discontinuity at the capping layer and $N = \sqrt{g \partial_z T_{\textup{vp}} / T_{0}}$ is the Brunt-V\"ais\"al\"a frequency, characteristic of the stratification.
-Note that, under the usual Boussinesq approximation, temperature and air density variations are simply related by $\delta T_{\textup{vp}}/T_{0} \simeq - \delta\rho/\rho_{0}$ (see Suppl.
-Mat.
-of \citet{andreotti2009}), so that $N$ can equivalently be defined from the density gradient as next to Eq.~\ref{FroudeNumber}.
+Note that, under the usual Boussinesq approximation, temperature and air density variations are simply related by $\delta T_{\textup{vp}}/T_{0} \simeq - \delta\rho/\rho_{0}$ (see Online Resource of \citet{Andreotti2009}), so that $N$ can equivalently be defined from the density gradient as next to Eq.~\ref{FroudeNumber}.
 
@@
 
-Examples of vertical profiles of the virtual potential temperature deduced from ERA5 are shown in Suppl.
-Fig.~\ref{Fig9_supp}a.
+Examples of vertical profiles of the virtual potential temperature deduced from ERA5 are shown in Online Resource Fig.~\ref{Fig11_supp}a.
 For each of them, an average temperature is computed below the ABL depth ($z<H$), and a linear function is fitted above, allowing us to extract the temperature jump $\Delta T_{\textup{vp}}$.
-Importantly, some profiles display a vertical structure that cannot be approximated by the simple form \eqref{TvpVerticalProfile} used here (Suppl.
-Fig.~\ref{Fig9_supp}b).
+Importantly, some profiles display a vertical structure that cannot be approximated by the simple form \eqref{TvpVerticalProfile} used here (Online Resource Fig.~\ref{Fig11_supp}b).
 In practice, we removed from the analysis all of those leading to the unphysical case $\Delta T_{\textup{vp}}<0$.
 We have noticed that these `ill-processed' profiles dominantly occur in winter and are evenly spread across the hours of the day.
-Importantly, they represent $\simeq12$\% of the data only (Suppl.
-Fig.~\ref{Fig9_supp}c,d), and we are thus confident that this data treatment does not affect our conclusions.
+Importantly, they represent $\simeq12$\% of the data only (Online Resource Fig.~\ref{Fig11_supp}c,d), and we are thus confident that this data treatment does not affect our conclusions.
 
@@
 \includegraphics[scale=1]{Figures/Figure1_supp.pdf}
-\caption{Gant chart representing the valid time steps for the two data sets, for all stations.}
+\caption{Photographs of the meteorological stations. \textbf{a}: South Sand Sea station. \textbf{b--c}: North Sand Sea station. \textbf{d}: Huab station. \textbf{e} Adamax station.}
 \label{Fig1_supp}
@@
 \includegraphics[scale=1]{Figures/Figure2_supp.pdf}
-\caption{Comparison between raw local wind measurements, and hourly-averaged data for South Namib station. \textbf{a}: wind direction. \textbf{b}: wind velocity at height $2.6$~m.}
+\caption{Gant chart representing the valid time steps for the two data sets, for all stations.}
 \label{Fig2_supp}
@@
 \includegraphics[scale=1]{Figures/Figure3_supp.pdf}
-\caption{Calibration of hydrodynamic roughness.
-The parameter $\delta$ (Eq.~\ref{metric_roughness}) quantifying the difference between local and predicted winds is shown in colorscale as a function of the hydrodynamic roughnesses chosen for the ERA5-Land and for local winds, for the (\textbf{a}) Adamax, (\textbf{b}) Deep Sea, (\textbf{c}) Huab and (\textbf{d}) South Namib stations.
-The red dashed and plain lines shows the minima of $\delta$ and the identity line, respectively.
-The orange lines and dots highlight the chosen hydrodynamic roughnesses for the local winds deduced from setting $z_{0}^{\textup{ERA5Land}} = 1$~mm.}
+\caption{Comparison between raw local wind measurements, and hourly-averaged data for South Sand Sea station. \textbf{a}: wind direction. \textbf{b}: wind velocity at height $2.6$~m.}
 \label{Fig3_supp}
@@
 \includegraphics[scale=1]{Figures/Figure4_supp.pdf}
-\caption{Analysis of the DEMs of the Deep Sea (left column -- panels \textbf{a}, \textbf{b}, \textbf{c}) and South Namib (right column -- panels \textbf{d}, \textbf{e}, \textbf{f}) stations.
-\textbf{a-d}: Bed elevation detrended by a fitted second order polynomial base-line.
-\textbf{b-e}: Autocorrelation matrix shown in colorscale.
-The black line shows the detected dune orientation, and the red line represents the autocorrelation profile along which the dune wavelength is calculated, displayed in \textbf{c-f}.
-The blue lines and dots show the first peak of the autocorrelation profile, whose abscissa gives the characteristic wavelength of the dune pattern.}
+\caption{Calibration of hydrodynamic roughness.
+The parameter $\delta$ (Eq.~\ref{metric_roughness}) quantifying the difference between local and predicted winds is shown in colorscale as a function of the hydrodynamic roughnesses chosen for the ERA5-Land and for local winds, for the (\textbf{a}) Adamax, (\textbf{b}) North Sand Sea, (\textbf{c}) Huab and (\textbf{d}) South Sand Sea stations.
+The red dashed and plain lines shows the minima of $\delta$ and the identity line, respectively.
+The orange lines and dots highlight the chosen hydrodynamic roughnesses for the local winds deduced from setting $z_{0}^{\textup{ERA5Land}} = 1$~mm.}
 \label{Fig4_supp}
@@
   \includegraphics[scale=1]{Figures/Figure5_supp.pdf}
-\caption{Statistical comparison of the wind orientation (\textbf{a}) and velocity (\textbf{b}) between the ERA5-Land dataset and the local measurements for the Huab and Adamax stations.
-Data point clustering around identity lines (dashed and black) provide evidence for agreement of the two sets.}
+  \caption{Analysis of the DEMs of the North Sand Sea (left column -- panels \textbf{a}, \textbf{b}, \textbf{c}) and South Sand Sea (right column -- panels \textbf{d}, \textbf{e}, \textbf{f}) stations.
+\textbf{a--d}: Bed elevation detrended by a fitted second order polynomial base-line.
+\textbf{b--e}: Autocorrelation matrix shown in colorscale.
+The black line shows the detected dune orientation, and the red line represents the autocorrelation profile along which the dune wavelength is calculated, displayed in \textbf{c--f}.
+The blue lines and dots show the first peak of the autocorrelation profile, whose abscissa gives the characteristic wavelength of the dune pattern.}
   \label{Fig5_supp}
@@
 %---------------------------------------------
-\begin{figure}[p]
+\begin{figure}
 \centering
 \includegraphics[scale=1]{Figures/Figure6_supp.pdf}
-\caption{Distributions of wind direction at the South Namib Station for the ERA5-Land climate reanalysis (orange) and the local measurements (blue) -- equivalent of Fig.~\ref{Fig3}.
-In each subplot, both distributions are plotted from the same time steps, selected with constraints on the wind direction (columns) and/or wind velocity (rows) in the ERA5-Land dataset.
-The gray vertical dashed lines indicate the dune orientation.
-The numbers at the top centre give the percentage of time steps selected in each sub-range, as well as the percentage of them corresponding to the day (between 1000 UTC and 2200 UTC).
-The purple frame highlights the regime (small wind velocities, nocturnal summer wind) during which the wind data from both datasets differ.}
+\caption{Temporal comparison between the wind data coming from the ERA5-Land climate reanalysis (orange lines) and from the local measurements (blue lines).
+Coloured swathes indicate day (between 1000 UTC and 2200 UTC) and night (before 1000 UTC or after 2200 UTC) \textbf{a--b}: Huab station in summer.
+\textbf{b--c}: Huab station in winter.
+\textbf{d--e}: South Sand Sea station in summer.
+\textbf{f--g}: South Sand Sea station in winter.}
 \label{Fig6_supp}
@@
 \includegraphics[scale=1]{Figures/Figure7_supp.pdf}
-\caption{Statistical comparison of the wind orientation between the ERA5-Land dataset and the local measurements for the South Namib and Deep Sea stations, for different velocity ranges.
-\textbf{a}: $u_{*, \textup{\textup{ERA}}} < 0.1~\textup{m}~\textup{s}^{-1}$.
-\textbf{b}: $0.1 < u_{*, \textup{\textup{ERA}}} \leq 0.25~\textup{m}~\textup{s}^{-1}$.
-\textbf{c}: $u_{*, \textup{\textup{ERA}}} \geq 0.25~\textup{m}~\textup{s}^{-1}$.
-The measured dune orientations are subtracted to the wind orientation, which allows us to plot both stations on the same graph.
-Black dashed lines indicate locally measured orientations aligned with the dune crests (here $0^\circ$, $180^\circ$ and $360^\circ$ -- panels \textbf{a}, \textbf{b}), as well as the identity lines (panels \textbf{b}, \textbf{c}).}
+\caption{Statistical comparison of the wind orientation (\textbf{a}) and velocity (\textbf{b}) between the ERA5-Land dataset and the local measurements for the Huab and Adamax stations.
+Data point clustering around identity lines (dashed and black) provide evidence for agreement of the two sets.}
 \label{Fig7_supp}
@@
 \includegraphics[scale=1]{Figures/Figure8_supp.pdf}
-\caption{Statistical comparison of the wind velocity between the ERA5-Land dataset and the local measurements for the South Namib and Deep Sea stations.
-\textbf{a}: Nocturnal summer easterly wind.
-\textbf{b}: Diurnal southerly wind.
-Black dashed lines are identity lines.
-The angle ranges used to select diurnal and nocturnal summer winds are the same as those in Fig.~\ref{Fig3} and Suppl.
-Fig.~\ref{Fig6_supp}.}
+\caption{Distributions of wind direction at the South Sand Sea Station for the ERA5-Land climate reanalysis (orange) and the local measurements (blue) -- equivalent of Fig.~\ref{Fig3}.
+In each subplot, both distributions are plotted from the same time steps, selected with constraints on the wind direction (columns) and/or wind velocity (rows) in the ERA5-Land dataset.
+The gray vertical dashed lines indicate the dune orientation.
+The numbers at the top centre give the percentage of time steps selected in each sub-range, as well as the percentage of them corresponding to the day (between 1000 UTC and 2200 UTC).
+The purple frame highlights the regime (small wind velocities, nocturnal summer wind) during which the wind data from both datasets differ.}
 \label{Fig8_supp}
@@
 \includegraphics[scale=1]{Figures/Figure9_supp.pdf}
-\caption{\textbf{a}: Vertical profiles of the virtual potential temperature at three different times (blue: 29/11/2012 - 1100 UTC, orange: 21/03/2017 - 1200 UTC, green: 21/03/2017 - 2000 UTC) at the South Namib station.
-Dots: data from the ERA5 reanalysis.
-Dashed lines: boundary layer height given by the ERA5 reanalysis.
-Plain lines: vertical (ABL) and linear (FA) fits to estimate the quantities displayed in Suppl.
-Fig.~\ref{Fig10_supp}.
-\textbf{b}: Examples of ill-processed vertical profiles at three different times (blue: 2/12/2013 - 2300 UTC, orange: 20/03/2017 - 0000 UTC, green: 14/07/2017 - 1400 UTC) at the South Namib station.
-\textbf{c}: Hourly distribution of ill-processed vertical profiles.
-\textbf{d}: Monthly distribution of ill-processed vertical profiles.}
+\caption{Statistical comparison of the wind orientation between the ERA5-Land dataset and the local measurements for the South Sand Sea and North Sand Sea stations, for different velocity ranges.
+\textbf{a}: $u_{*, \textup{\textup{ERA}}} < 0.1~\textup{m}~\textup{s}^{-1}$.
+\textbf{b}: $0.1 < u_{*, \textup{\textup{ERA}}} \leq 0.25~\textup{m}~\textup{s}^{-1}$.
+\textbf{c}: $u_{*, \textup{\textup{ERA}}} \geq 0.25~\textup{m}~\textup{s}^{-1}$.
+The measured dune orientations are subtracted to the wind orientation, which allows us to plot both stations on the same graph.
+Black dashed lines indicate locally measured orientations aligned with the dune crests (here $0^\circ$, $180^\circ$ and $360^\circ$ -- panels \textbf{a}, \textbf{b}), as well as the identity lines (panels \textbf{b}, \textbf{c}).}
 \label{Fig9_supp}
@@
 \centering
-\includegraphics{Figures/Figure10_supp.pdf}
-\caption{Distributions of the meteorological parameters resulting from the processing of the ERA5-Land data for the South Namib (blue) and the Deep Sea (orange) stations.
-\pc{axis label : panel c -velocity at (height) H}.
-\pc{Is this for all day/night/seasons ?}}
+\includegraphics[scale=1]{Figures/Figure10_supp.pdf}
+\caption{Statistical comparison of the wind velocity between the ERA5-Land dataset and the local measurements for the South Sand Sea and North Sand Sea stations.
+\textbf{a}: Nocturnal summer easterly wind.
+\textbf{b}: Diurnal southerly wind.
+Black dashed lines are identity lines.
+The angle ranges used to select diurnal and nocturnal summer winds are the same as those in Fig.~\ref{Fig3} and Online Resource Fig.~\ref{Fig8_supp}.}
 \label{Fig10_supp}
@@
 \includegraphics[scale=1]{Figures/Figure11_supp.pdf}
-\caption{Non-dimensional parameters distributions. For the marginal \pc{???} distributions, the orange correspond to the South Namib station, and the blue to the Deep Sea station.}
+\caption{\textbf{a}: Vertical profiles of the virtual potential temperature at three different times (blue: 29/11/2012 - 1100 UTC, orange: 21/03/2017 - 1200 UTC, green: 21/03/2017 - 2000 UTC) at the South Sand Sea station.
+Dots: data from the ERA5 reanalysis.
+Dashed lines: boundary layer height given by the ERA5 reanalysis.
+Plain lines: vertical (ABL) and linear (FA) fits to estimate the quantities displayed in Online Resource Fig.~\ref{Fig12_supp}.
+\textbf{b}: Examples of ill-processed vertical profiles at three different times (blue: 2/12/2013 - 2300 UTC, orange: 20/03/2017 - 0000 UTC, green: 14/07/2017 - 1400 UTC) at the South Sand Sea station.
+\textbf{c}: Hourly distribution of ill-processed vertical profiles.
+\textbf{d}: Monthly distribution of ill-processed vertical profiles.}
 \label{Fig11_supp}
@@
 \centering
-\includegraphics[scale=1]{Figures/Figure12_supp.pdf}
-\caption{Regime diagrams of the wind deviation $\delta_{\theta}$ and relative attenuation/amplification $\delta_{u}$ in the spaces $(\mathcal{F}_{\textup{I}}, \, kH)$ and $(\mathcal{F}_{\textup{I}}, \, \mathcal{F})$, containing the data from both the Deep Sea and South Namib stations.
-Blue dashed lines empirically delimit the different regimes.
-The point density in each bin of the diagrams is shown in Suppl.
-Fig.~\ref{Fig11_supp}.
-The similar regime diagrams in the space $(\mathcal{F}, \, kH)$ are shown in Fig.~\ref{Fig5}.}
+\includegraphics{Figures/Figure12_supp.pdf}
+\caption{Distributions of the meteorological parameters resulting from the processing of the ERA5-Land data for the South Sand Sea (blue) and the North Sand Sea (orange) stations.}
 \label{Fig12_supp}
@@
 \includegraphics[scale=1]{Figures/Figure13_supp.pdf}
-\caption{Computation of the flow disturbance with the linear model of \citet{andreotti2009}.
-(a) and (b) Magnitude of the hydrodynamic coefficients $\mathcal{A}_0$ and $\mathcal{B}_0$, calculated from the values of the non-dimensional numbers corresponding to the ERA5-Land time series presented in Figs.~\ref{Fig4} and \ref{Fig5}.
-(c) Shear velocity streamlines over sinusoidal ridges of amplitude $k\xi_0 = 0.1$ and for increasing values of $\sqrt{\mathcal{A}_{0}^{2} + \mathcal{B}_{0}^{2}}$.
-From the upper to the lower streamline, values of $\left(kH,\, \mathcal{F},\, \mathcal{F}_{\rm I},\, \mathcal{A}_{0},\, \mathcal{B}_{0},\, \sqrt{\mathcal{A}_{0}^{2} + \mathcal{B}_{0}^{2}}\right)$ are (1.9, 0.6, 1.5, 3.4, 1.0, 3.5), (1.5, 0.3, 0.4, 4.8, 1.4, 5.0), (0.1, 3.5, 1.0, 8.6, 0.1, 8.6), (0.5, 0.05, 0.04, 9.6, 2.5, 9.9).}
+\caption{Non-dimensional parameters distributions. For the marginal distributions, the orange correspond to the South Sand Sea station, and the blue to the North Sand Sea station.}
 \label{Fig13_supp}
@@
 
+%---------------------------------------------
+\begin{figure}[p]
+\centering
+\includegraphics[scale=1]{Figures/Figure14_supp.pdf}
+\caption{Regime diagrams of the wind deviation $\delta_{\theta}$ and relative attenuation/amplification $\delta_{u}$ in the spaces $(\mathcal{F}_{\textup{I}}, \, kH)$ and $(\mathcal{F}_{\textup{I}}, \, \mathcal{F})$, containing the data from both the North Sand Sea and South Sand Sea stations.
+Blue dashed lines empirically delimit the different regimes.
+The point density in each bin of the diagrams is shown in Online Resource Fig.~\ref{Fig13_supp} -- 95\% of the data occur in the range $-1 < \delta u < 1$.
+The similar regime diagrams in the space $(\mathcal{F}, \, kH)$ are shown in Fig.~\ref{Fig5}.}
+\label{Fig14_supp}
+\end{figure}
+%---------------------------------------------
+
+%---------------------------------------------
+\begin{figure}[p]
+\centering
+\includegraphics[scale=1]{Figures/Figure15_supp.pdf}
+\caption{Normalized distributions of amplified velocities for the North sand Sea (blue: $\delta_u < 0$, orange: $\delta_u < -0.5$).
+\textbf{a}: Angular distributions.
+\textbf{b}: Monthly distributions.
+\textbf{c}: Hourly distributions.}
+\label{Fig15_supp}
+\end{figure}
+%---------------------------------------------
+
+%---------------------------------------------
+\begin{figure}[p]
+\centering
+\includegraphics[scale=1]{Figures/Figure16_supp.pdf}
+\caption{Computation of the flow disturbance with the linear model of \citet{Andreotti2009}.
+\textbf{a} and \textbf{b} Magnitude of the hydrodynamic coefficients $\mathcal{A}_0$ and $\mathcal{B}_0$, calculated from the values of the non-dimensional numbers corresponding to the ERA5-Land time series presented in Figs.~\ref{Fig4} and \ref{Fig5}.
+\textbf{c} Shear velocity streamlines over sinusoidal ridges of amplitude $k\xi_0 = 0.1$ and for increasing values of $\sqrt{\mathcal{A}_{0}^{2} + \mathcal{B}_{0}^{2}}$.
+From the upper to the lower streamline, values of $\left(kH,\, \mathcal{F},\, \mathcal{F}_{\rm I},\, \mathcal{A}_{0},\, \mathcal{B}_{0},\, \sqrt{\mathcal{A}_{0}^{2} + \mathcal{B}_{0}^{2}}\right)$ are (1.9, 0.6, 1.5, 3.4, 1.0, 3.5), (1.5, 0.3, 0.4, 4.8, 1.4, 5.0), (0.1, 3.5, 1.0, 8.6, 0.1, 8.6), (0.5, 0.05, 0.04, 9.6, 2.5, 9.9).}
+\label{Fig16_supp}
+\end{figure}
+%---------------------------------------------
+
 \end{document}

### 2022-02-28 Paper_figure/Figure02.py

@@ -43,8 +43,8 @@ variables = ['U_star', 'Orientation']
 label_var = {'U_star': r'Velocity, $u_{*}~[\textup{m}~\textup{s}^{-1}]$', 'Orientation': r'Orientation, $\theta~[^\circ]$'}
 labels = [(r'\textbf{a}', r'\textbf{b}'), (r'\textbf{c}', r'\textbf{d}'),
           (r'\textbf{e}', r'\textbf{f}'), (r'\textbf{g}', r'\textbf{h}')]
-row_labels = ['Adamax station -- summer', 'Adamax station -- winter', 'North Namib station -- summer',
-              'North Namib station -- winter']
+row_labels = ['Adamax -- summer', 'Adamax -- winter', 'North Sand Sea -- summer',
+              'North Sand Sea -- winter']
 years = [2015, 2016, 2017, 2017]
 months = [11, 8, 12, 6]
 days = [(10, 14), (17, 20), (5, 8), (1, 4)]

### 2022-02-28 Paper_figure/Figure04.py

@@ -7,6 +7,7 @@ Figure 4
 
 import numpy as np
 import matplotlib.pyplot as plt
+from matplotlib.patches import FancyArrowPatch
 import sys
 import os
 sys.path.append('../')
@@ -188,6 +189,13 @@ for i, (theta, A0, B0, c) in enumerate(zip(Theta_list, A0_list, B0_list, colors)
     # ax.quiver(X[skip], Y[skip], TAU[0][skip], TAU[1][skip], color='grey')
     # strm = ax.streamplot(X, Y, TAU[0], TAU[1], color=np.sqrt(TAU[0]**2 + TAU[1]**2), cmap='inferno', density=50, start_points=[[4, 5-0.5*i]])
     strm = ax.streamplot(X, Y, ustar*np.cos(theta), ustar*np.sin(theta), density=50, start_points=[[4, 3-0.75*i]], color='k')
+
+tail = np.array([10.5, 0.36])
+length = 4
+head = tail + np.array([cosd(Theta_list[0]), sind(Theta_list[0])])*length
+
+arrow = FancyArrowPatch(tail, head, mutation_scale=20, facecolor='lightblue')
+ax.add_patch(arrow)
 
 ax.text(-6.57, 1, ' ' + labels[3] + ' ', bbox=bbox)
 ax.text(-6.57, 0, ' ' + labels[2] + ', ' + labels[5] + ' ', bbox=bbox)

### 2022-02-28 Paper_figure/Figure05.py

@@ -44,10 +44,11 @@ delta_u = (U_era - U_insitu)/U_era
 lims = {'Froude': (5.8e-3, 450), 'kLB': (0.009, 7.5), 'kH': (2.2e-2, 10.8)}
 cmaps = [theme.cmap_delta_theta, theme.cmap_delta_u]
 norms = [mpcolors.Normalize(vmin=0, vmax=99),
-         mpcolors.TwoSlopeNorm(vmin=-3, vcenter=0, vmax=3)]
+         mpcolors.TwoSlopeNorm(vmin=-3, vcenter=0, vmax=1)]
 cbar_labels = [r'$\delta_{\theta}$ [deg.]', r'$\delta_{u}$']
 quantities = [delta_angle, delta_u]
 labels = [r'\textbf{a}', r'\textbf{b}']
+cbticks = [[0, 25, 50, 75], [-3, -1.5, 0, 0.5, 1]]
 
 mask = ~np.isnan(numbers['Froude'])
 log_counts_max = np.log10(2230)
@@ -61,7 +62,9 @@ xlabel = r'$\mathcal{F} =  U/\sqrt{(\Delta\rho/\rho_{0}) g H}$'
 fig, axarr = plt.subplots(1, 2, figsize=(theme.fig_width, 0.375*theme.fig_height_max),
                           constrained_layout=True)
 
-for i, (ax, quantity, cmap, norm) in enumerate(zip(axarr.flatten(), quantities, cmaps, norms)):
+for i, (ax, quantity, cmap, norm, cbtick) in enumerate(zip(axarr.flatten(),
+                                                           quantities, cmaps, norms,
+                                                           cbticks)):
     ylabel = '$k H$' if i == 0 else None
     #
     vars = [numbers[var1][mask], numbers[var2][mask]]
@@ -73,14 +76,14 @@ for i, (ax, quantity, cmap, norm) in enumerate(zip(axarr.flatten(), quantities, 
     a = plot_regime_diagram(ax, quantity[mask], vars, lims_list, xlabel, ylabel, bins=bins, norm=norm, cmap=cmap, type='binned')
     #
     ax.text(0.04, 0.94, labels[i], transform=ax.transAxes, ha='left', va='center')
-
+    #
     # regime lines
     ax.axvline(0.4, color=regime_line_color, linestyle='--', lw=2)
     ax.axhline(0.32, color=regime_line_color, linestyle='--', lw=2)
-
+    #
[…]

### 2022-02-28 Paper_figure/Supplementary_Figures/Figure02_supp.py

@@ -1,65 +1,76 @@
 """
-============
-Figure 2 -- SI
-============
+============================
+Figure 2 -- Online Resource
+============================
 
 """
 
-import numpy as np
 import os
-from datetime import datetime
-import matplotlib.dates as mdates
-import matplotlib.pyplot as plt
 import sys
+import glob
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.patches as mpatches
 sys.path.append('../../')
 import python_codes.theme as theme
-
+from datetime import timedelta
+#
 theme.load_style()
+
+
+def make_range_broken_barh(time, dt):
+    diff = np.diff(time)
+    t_diff = np.concatenate((time[1:][diff > dt_threshold], time[:-1][diff > dt_threshold]))
+    t_diff = np.insert(t_diff, [0, t_diff.size], [time[0], time[-1]])
+    t_diff = np.array(sorted(t_diff))
+    return [(tstart, tspan) for tstart, tspan in zip(t_diff[::2], np.diff(t_diff)[::2])]
+
 
 # paths
 path_savefig = '../../Paper/Figures'
-path_outputdata = '../../static/data/processed_data'
-path_inputdata = '../../static/data/raw_data'
-
-# figure parameters
-station = 'South_Namib_Station'
-tmin, tmax = datetime(2017, 6, 3), datetime(2017, 6, 10)
-Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
-
-# Loading and recomputing some raw data
-path_insitu = os.path.join(path_inputdata, 'measured_wind_data/in_situ_wind_data_' + station + '.npy')
-Data_insitu = np.load(path_insitu, allow_pickle=True).item()
-#
-t_insitu = Data_insitu['time']
-U_insitu = Data_insitu['velocity']
-# putting angles in trigo. ref.
-Orientation_insitu = (270 - Data_insitu['direction']) % 360
+path_inputdata = '../../static/data/raw_data/'
 
 
-# ### Figure
-fig, axarr = plt.subplots(2, 1, figsize=(theme.fig_width, 0.85*theme.fig_width),
-                          constrained_layout=True, sharex=True)
+Stations = ['Adamax_Station', 'Huab_Station', 'Deep_Sea_Station', 'South_Namib_Station']
[…]

### 2022-02-28 Paper_figure/Supplementary_Figures/Figure03_supp.py

@@ -1,14 +1,15 @@
 """
-============
-Figure 3 -- SI
-============
+============================
+Figure 3 -- Online Resource
+============================
 
 """
 
 import numpy as np
 import os
+from datetime import datetime
+import matplotlib.dates as mdates
 import matplotlib.pyplot as plt
-from matplotlib.colors import Normalize
 import sys
 sys.path.append('../../')
 import python_codes.theme as theme
@@ -17,43 +18,48 @@ theme.load_style()
 
 # paths
 path_savefig = '../../Paper/Figures'
-path_outputdata = '../../static/data/processed_data/'
+path_outputdata = '../../static/data/processed_data'
+path_inputdata = '../../static/data/raw_data'
 
+# figure parameters
+station = 'South_Namib_Station'
+tmin, tmax = datetime(2017, 6, 3), datetime(2017, 6, 10)
 Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
-Data_roughness = np.load(os.path.join(path_outputdata, 'Data_calib_roughness.npy'),
-                         allow_pickle=True).item()
 
-labels = [r'\textbf{a}', r'\textbf{b}', r'\textbf{c}', r'\textbf{d}']
-norm = Normalize(vmin=0.3, vmax=1, clip=True)
-
-# #### Figure
-fig, axrr = plt.subplots(2, 2, figsize=(theme.fig_width, 1.1*theme.fig_width),
-                         sharex=True, sharey=True, constrained_layout=True)
+# Loading and recomputing some raw data
+path_insitu = os.path.join(path_inputdata, 'measured_wind_data/in_situ_wind_data_' + station + '.npy')
+Data_insitu = np.load(path_insitu, allow_pickle=True).item()
 #
-for i, (ax, p, metric, station, label) in enumerate(zip(axrr.flatten(), Data_roughness['Pvals'],
-                                                        Data_roughness['Metrics'], Data_roughness['Stations'], labels)):
-    cs = ax.contourf(Data_roughness['z0_era_vals'], Data_roughness['z0_insitu_vals'], metric, levels=50, norm=norm)
-    for c in cs.collections:
-        c.set_edgecolor("face")
[…]

### 2022-02-28 Paper_figure/Supplementary_Figures/Figure04_supp.py

@@ -1,18 +1,17 @@
 """
-============
-Figure 4 -- SI
-============
+============================
+Figure 4 -- Online Resource
+============================
 
 """
 
 import numpy as np
 import os
 import matplotlib.pyplot as plt
+from matplotlib.colors import Normalize
 import sys
-from types import SimpleNamespace
 sys.path.append('../../')
 import python_codes.theme as theme
-from python_codes.general import cosd, sind
 
 theme.load_style()
 
@@ -20,59 +19,41 @@ theme.load_style()
 path_savefig = '../../Paper/Figures'
 path_outputdata = '../../static/data/processed_data/'
 
-Data_DEM = np.load(os.path.join(path_outputdata, 'Data_DEM.npy'),
-                   allow_pickle=True).item()
+Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
+Data_roughness = np.load(os.path.join(path_outputdata, 'Data_calib_roughness.npy'),
+                         allow_pickle=True).item()
 
-labels = [(r'\textbf{a}', r'\textbf{d}'), (r'\textbf{b}', r'\textbf{e}'), (r'\textbf{c}', r'\textbf{f}')]
+labels = [r'\textbf{a}', r'\textbf{b}', r'\textbf{c}', r'\textbf{d}']
+norm = Normalize(vmin=0.3, vmax=1, clip=True)
 
-fig, axrr = plt.subplots(3, 2, figsize=(theme.fig_width, 0.75*theme.fig_height_max),
-                         constrained_layout=True, gridspec_kw={'width_ratios': (0.9, 1)})
-for i, station in enumerate(Data_DEM.keys()):
-    # loading into namespace from data dictionnary to shorten call
-    n = SimpleNamespace(**Data_DEM[station])
-    # ax0: Topo
-    cs = axrr[0, i].contourf(n.lon, n.lat, n.topo, levels=50)
+# #### Figure
+fig, axrr = plt.subplots(2, 2, figsize=(theme.fig_width, 1.1*theme.fig_width),
+                         sharex=True, sharey=True, constrained_layout=True)
+#
+for i, (ax, p, metric, station, label) in enumerate(zip(axrr.flatten(), Data_roughness['Pvals'],
[…]

### 2022-02-28 Paper_figure/Supplementary_Figures/Figure06_supp.py

@@ -1,63 +1,92 @@
 """
-============
-Figure 6 -- SI
-============
+============================
+Figure 6 -- Online Resource
+============================
 
 """
 
 import numpy as np
-import os
 import matplotlib.pyplot as plt
+import matplotlib.dates as mdates
+import calendar
+import locale
+from datetime import datetime, timedelta
 import sys
+import os
 sys.path.append('../../')
 import python_codes.theme as theme
-from python_codes.plot_functions import plot_scatter_surrounded
+
+locale.setlocale(locale.LC_ALL, 'en_US.utf8')
 
 
+def tick_formatter(ax, fmt='%d'):
+    myFmt = mdates.DateFormatter(fmt)
+    ax.xaxis.set_major_formatter(myFmt)
+    ticklabels = ax.get_xticklabels()
+    ticklabels[0].set_ha('left')
+
+
+# Loading figure theme
 theme.load_style()
 
 # paths
 path_savefig = '../../Paper/Figures'
 path_outputdata = '../../static/data/processed_data/'
 
+# Loading wind data
 Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
-labels = [r'\textbf{a}', r'\textbf{b}']
+Stations = sorted(Data.keys())
 
-# preparing data
-Stations_ref = ['Adamax_Station', 'Huab_Station']
-#
-Theta_ERA = np.concatenate([Data[station]['Orientation_era'] for station in Stations_ref])
-Theta_Station = np.concatenate([Data[station]['Orientation_insitu'] for station in Stations_ref])
-#
-U_ERA = np.concatenate([Data[station]['U_star_era'] for station in Stations_ref])
-U_Station = np.concatenate([Data[station]['U_star_insitu'] for station in Stations_ref])
+# Figure properties
+variables = ['U_star', 'Orientation']
+label_var = {'U_star': r'Velocity, $u_{*}~[\textup{m}~\textup{s}^{-1}]$', 'Orientation': r'Orientation, $\theta~[^\circ]$'}
+labels = [(r'\textbf{a}', r'\textbf{b}'), (r'\textbf{c}', r'\textbf{d}'),
+          (r'\textbf{e}', r'\textbf{f}'), (r'\textbf{g}', r'\textbf{h}')]
[…]

### 2022-02-28 Paper_figure/Supplementary_Figures/Figure09_supp.py

@@ -1,7 +1,7 @@
 """
-============
-Figure 9 -- SI
-============
+============================
+Figure 9 -- Online Resource
+============================
 
 """
 
@@ -12,6 +12,7 @@ import sys
 sys.path.append('../../')
 import python_codes.theme as theme
 from python_codes.plot_functions import plot_scatter_surrounded
+
 
 theme.load_style()
 
@@ -27,45 +28,53 @@ labels = [r'\textbf{a}', r'\textbf{b}']
 # preparing data
 Stations = ['South_Namib_Station', 'Deep_Sea_Station']
 
-Nocturnal_wind = {'South_Namib_Station': [150, 260], 'Deep_Sea_Station': [150, 230]}
+Dune_orientations = [np.load(os.path.join(path_outputdata, 'Data_DEM.npy'), allow_pickle=True).item()[station]['orientation']
+                     for station in Stations]
+
+velocity_thresholds = [0.1, 0.25]
 
 # variables
-x1 = np.concatenate([Data[station]['U_star_era'][(Data[station]['Orientation_era'] > Nocturnal_wind[station][0]) & (Data[station]['Orientation_era'] < Nocturnal_wind[station][1])]
+x1 = np.concatenate([Data[station]['Orientation_era'][Data[station]['U_star_era'] < velocity_thresholds[0]] - Dune_orientations[Stations.index(station)]
                      for station in Stations])
-y1 = np.concatenate([Data[station]['U_star_insitu'][(Data[station]['Orientation_era'] > Nocturnal_wind[station][0]) & (Data[station]['Orientation_era'] < Nocturnal_wind[station][1])]
-                     for station in Stations])
-t1 = np.concatenate([Data[station]['time'][(Data[station]['Orientation_era'] > Nocturnal_wind[station][0]) & (Data[station]['Orientation_era'] < Nocturnal_wind[station][1])]
+y1 = np.concatenate([Data[station]['Orientation_insitu'][Data[station]['U_star_era'] < velocity_thresholds[0]] - Dune_orientations[Stations.index(station)]
                      for station in Stations])
 
-hours = [i.hour for i in t1]
-#
[…]

### 2022-02-28 Paper_figure/Supplementary_Figures/Figure11_supp.py

@@ -1,69 +1,110 @@
 """
-============
-Figure 11 -- SI
-============
+============================
+Figure 11 -- Online Resource
+============================
 
 """
 
 import numpy as np
+import os
 import matplotlib.pyplot as plt
 import sys
-import os
 sys.path.append('../../')
 import python_codes.theme as theme
-from python_codes.meteo_analysis import mu
 from python_codes.plot_functions import make_nice_histogram
 
 
-# Loading figure theme
+def plot_vertical_profile(ax, height, Virtual_potential_temperature, grad_free_atm, theta_free_atm, blh, theta_ground, Hmax_fit, color='tab:blue', label=None):
+    Hfit = np.linspace(blh, Hmax_fit, 100)
+    #
+    line = ax.vlines(theta_ground, 0, blh/1e3, color=color, label=label, zorder=-3)
+    ax.axhline(blh/1e3, alpha=0.5, color=color, ls='--')
+    ax.plot(np.poly1d([grad_free_atm, theta_free_atm])(Hfit), Hfit/1e3, color=line.get_color(), zorder=-2)
+    ax.plot(Virtual_potential_temperature, height/1e3, '.', color=line.get_color(), zorder=-1)
+    # ax.scatter(theta_ground, blh/1e3, s=30, facecolors=line.get_color(), edgecolors='k', linewidth=2, zorder=0)
+
+
 theme.load_style()
 
 # paths
 path_savefig = '../../Paper/Figures'
 path_outputdata = '../../static/data/processed_data/'
 
-# ##### Loading meteo data
+# Loading data
 Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
 
-# ## histograms parameters
-Stations = ['South_Namib_Station', 'Deep_Sea_Station']
-g = 9.81  # [m/s2] gravitational constant
-z0_era = 1e-3  # [m] hydrodynamic roughness
-#
 labels = [r'\textbf{a}', r'\textbf{b}', r'\textbf{c}', r'\textbf{d}']
-nbins = 80
 
-# #### Figure
-fig, axarr = plt.subplots(2, 2, figsize=(theme.fig_width, 0.9*theme.fig_width),
-                          constrained_layout=True, sharey=True)
+# ## vertical profiles parameters
+station = 'Deep_Sea_Station'
[…]

### 2022-02-28 Paper_figure/Supplementary_Figures/Figure12_supp.py

@@ -1,128 +1,69 @@
 """
-============
-Figure 12 -- SI
-============
+============================
+Figure 12 -- Online Resource
+============================
 
 """
 
 import numpy as np
-import os
 import matplotlib.pyplot as plt
-from matplotlib.colors import LogNorm
 import sys
+import os
 sys.path.append('../../')
 import python_codes.theme as theme
+from python_codes.meteo_analysis import mu
 from python_codes.plot_functions import make_nice_histogram
 
+
+# Loading figure theme
 theme.load_style()
 
 # paths
 path_savefig = '../../Paper/Figures'
 path_outputdata = '../../static/data/processed_data/'
 
-# Loading data
+# ##### Loading meteo data
 Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
+
+# ## histograms parameters
 Stations = ['South_Namib_Station', 'Deep_Sea_Station']
-
-numbers = {key: np.concatenate([Data[station][key] for station in Stations]) for key in ('Froude', 'kH', 'kLB')}
-mask = ~np.isnan(numbers['Froude'])
-ad_hoc_quantity = np.concatenate([Data[station]['U_star_era'] for station in Stations])
-
-# Figure properties
-couples = [('Froude', 'kLB'), ('Froude', 'kH'), ('kLB', 'kH')]
-lims = {'Froude': (5.8e-3, 450), 'kLB': (0.009, 7.5), 'kH': (2.2e-2, 10.8)}
-# ax_labels = {'kH': r'$kH$', 'Froude': r'$\mathcal{F} =  U/\sqrt{(\Delta\rho/\rho) g H}$',
-# 'kLB': r'$\mathcal{F}_{\textup{I}} =  kU/N$'}
-ax_labels = {'kH': r'$kH$', 'Froude': r'$\mathcal{F}$',
-             'kLB': r'$\mathcal{F}_{\textup{I}}$'}
-norm = LogNorm(vmin=1, vmax=1.5e3)
+g = 9.81  # [m/s2] gravitational constant
+z0_era = 1e-3  # [m] hydrodynamic roughness
+#
+labels = [r'\textbf{a}', r'\textbf{b}', r'\textbf{c}', r'\textbf{d}']
+nbins = 80
 
 # #### Figure
-fig, axarr = plt.subplots(4, 3, figsize=(theme.fig_width, 0.95*theme.fig_width),
[…]

### 2022-02-28 Paper_figure/Supplementary_Figures/Figure13_supp.py

@@ -1,97 +1,128 @@
 """
-============
-Figure 13 -- SI
-============
+============================
+Figure 13 -- Online Resource
+============================
 
 """
 
 import numpy as np
-import matplotlib.pyplot as plt
-import matplotlib.colors as mpcolors
-import sys
 import os
+import matplotlib.pyplot as plt
+from matplotlib.colors import LogNorm
+import sys
 sys.path.append('../../')
 import python_codes.theme as theme
-from python_codes.general import smallestSignedAngleBetween, find_mode_distribution
-from python_codes.plot_functions import plot_regime_diagram
+from python_codes.plot_functions import make_nice_histogram
 
-# Loading figure theme
 theme.load_style()
 
 # paths
 path_savefig = '../../Paper/Figures'
 path_outputdata = '../../static/data/processed_data/'
 
-# ##### Loading meteo data
+# Loading data
 Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
 Stations = ['South_Namib_Station', 'Deep_Sea_Station']
 
-# #### Computing quantities
-
-Orientation_era = np.concatenate([Data[station]['Orientation_era'] for station in Stations])
-Orientation_insitu = np.concatenate([Data[station]['Orientation_insitu'] for station in Stations])
-U_era = np.concatenate([Data[station]['U_star_era'] for station in Stations])
-U_insitu = np.concatenate([Data[station]['U_star_insitu'] for station in Stations])
 numbers = {key: np.concatenate([Data[station][key] for station in Stations]) for key in ('Froude', 'kH', 'kLB')}
-#
-Delta = smallestSignedAngleBetween(Orientation_era, Orientation_insitu)
-mode_delta = np.array([find_mode_distribution(Delta, i) for i in np.arange(150, 350)]).mean()
-delta_angle = np.abs(Delta)
-delta_u = (U_era - U_insitu)/U_era
-
-# #### Figure parameters
-
-lims = {'Froude': (5.8e-3, 450), 'kLB': (0.009, 7.5), 'kH': (2.2e-2, 10.8)}
[…]

### 2022-02-28 python_codes/theme.py

@@ -3,7 +3,9 @@ import matplotlib.pyplot as plt
 import matplotlib.colors as mpcolors
 import numpy as np
 import cmocean
+import cmasher as cmr
 import os
+from matplotlib.colors import ListedColormap
 
 
 def truncate_colormap(cmap, minval=0.0, maxval=1.0, n=100):
@@ -53,7 +55,11 @@ markersize = 5
 markersize_small = 2
 
 # #### colormaps
-cmap_delta_u = cmocean.cm.curl_r
+# cmap_delta_u = plt.cm.twilight
+cmap_delta_u = truncate_colormap(cmr.emergency_s_r, minval=0.05, maxval=1.0, n=256)
+# cmap_delta_u = truncate_colormap(cmr.seasons_s, minval=0.05, maxval=1.0, n=256)
+# cmap_delta_u = truncate_colormap(cmr.iceburn, minval=0, maxval=1.0, n=256)
+# cmap_delta_u = cmr.emergency_s
 cmap_delta_theta = 'plasma'
 cmap_topo = cmocean.cm.turbid_r
 cmap_wind = truncate_colormap(plt.cm.viridis, 0.15, 1, 256)

### 2022-03-11 docs/Paper_figure/Figure02.py
Adding License and readmde file

@@ -9,11 +9,14 @@ import numpy as np
 import matplotlib.pyplot as plt
 import matplotlib.dates as mdates
 import calendar
+import locale
 from datetime import datetime, timedelta
 import sys
 import os
 sys.path.append('../')
 import python_codes.theme as theme
+
+locale.setlocale(locale.LC_ALL, 'en_US.utf8')
 
 
 def tick_formatter(ax, fmt='%d'):
@@ -29,28 +32,30 @@ theme.load_style()
 # paths
 path_imgs = '../static/images/'
 path_savefig = '../Paper/Figures'
-path_outputdata = '../static/processed_data'
+path_outputdata = '../static/data/processed_data'
 
 # Loading wind data
 Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
 Stations = sorted(Data.keys())
 
 # Figure properties
-variables = ['U', 'Orientation']
-label_var = {'U': r'Velocity, $u_{*}~[\textup{m}~\textup{s}^{-1}]$', 'Orientation': r'Orientation, $\theta~[^\circ]$'}
-labels = [(r'\textbf{a}', r'\textbf{b}'), (r'\textbf{c}', r'\textbf{d}'), (r'\textbf{e}', r'\textbf{f}')]
-row_labels = ['Huab station', 'Deep Sea station -- summer', 'Deep Sea station -- winter']
-years = [2018, 2017, 2017]
-months = [2, 12, 6]
-days = [(11, 14), (5, 8), (1, 4)]
+variables = ['U_star', 'Orientation']
+label_var = {'U_star': r'Velocity, $u_{*}~[\textup{m}~\textup{s}^{-1}]$', 'Orientation': r'Orientation, $\theta~[^\circ]$'}
+labels = [(r'\textbf{a}', r'\textbf{b}'), (r'\textbf{c}', r'\textbf{d}'),
+          (r'\textbf{e}', r'\textbf{f}'), (r'\textbf{g}', r'\textbf{h}')]
+row_labels = ['Adamax -- summer', 'Adamax -- winter', 'North Sand Sea -- summer',
+              'North Sand Sea -- winter']
+years = [2015, 2016, 2017, 2017]
+months = [11, 8, 12, 6]
+days = [(10, 14), (17, 20), (5, 8), (1, 4)]
 month_calendar = {index: month for index, month in enumerate(calendar.month_name) if month}
 
 
[…]

### 2022-03-11 docs/Paper_figure/Figure04.py

@@ -7,6 +7,7 @@ Figure 4
 
 import numpy as np
 import matplotlib.pyplot as plt
+from matplotlib.patches import FancyArrowPatch
 import sys
 import os
 sys.path.append('../')
@@ -44,7 +45,7 @@ theme.load_style()
 # paths
 path_imgs = '../static/images/'
 path_savefig = '../Paper/Figures'
-path_outputdata = '../static/processed_data'
+path_outputdata = '../static/data/processed_data'
 
 # ##### Loading meteo data
 Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
@@ -188,6 +189,13 @@ for i, (theta, A0, B0, c) in enumerate(zip(Theta_list, A0_list, B0_list, colors)
     # ax.quiver(X[skip], Y[skip], TAU[0][skip], TAU[1][skip], color='grey')
     # strm = ax.streamplot(X, Y, TAU[0], TAU[1], color=np.sqrt(TAU[0]**2 + TAU[1]**2), cmap='inferno', density=50, start_points=[[4, 5-0.5*i]])
     strm = ax.streamplot(X, Y, ustar*np.cos(theta), ustar*np.sin(theta), density=50, start_points=[[4, 3-0.75*i]], color='k')
+
+tail = np.array([10.5, 0.36])
+length = 4
+head = tail + np.array([cosd(Theta_list[0]), sind(Theta_list[0])])*length
+
+arrow = FancyArrowPatch(tail, head, mutation_scale=20, facecolor='lightblue')
+ax.add_patch(arrow)
 
 ax.text(-6.57, 1, ' ' + labels[3] + ' ', bbox=bbox)
 ax.text(-6.57, 0, ' ' + labels[2] + ', ' + labels[5] + ' ', bbox=bbox)

### 2022-03-11 docs/Paper_figure/Figure05.py

@@ -21,7 +21,7 @@ theme.load_style()
 # paths
 path_imgs = '../static/images/'
 path_savefig = '../Paper/Figures'
-path_outputdata = '../static/processed_data'
+path_outputdata = '../static/data/processed_data'
 
 # ##### Loading meteo data
 Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
@@ -44,10 +44,11 @@ delta_u = (U_era - U_insitu)/U_era
 lims = {'Froude': (5.8e-3, 450), 'kLB': (0.009, 7.5), 'kH': (2.2e-2, 10.8)}
 cmaps = [theme.cmap_delta_theta, theme.cmap_delta_u]
 norms = [mpcolors.Normalize(vmin=0, vmax=99),
-         mpcolors.TwoSlopeNorm(vmin=-3, vcenter=0, vmax=3)]
+         mpcolors.TwoSlopeNorm(vmin=-3, vcenter=0, vmax=1)]
 cbar_labels = [r'$\delta_{\theta}$ [deg.]', r'$\delta_{u}$']
 quantities = [delta_angle, delta_u]
 labels = [r'\textbf{a}', r'\textbf{b}']
+cbticks = [[0, 25, 50, 75], [-3, -1.5, 0, 0.5, 1]]
 
 mask = ~np.isnan(numbers['Froude'])
 log_counts_max = np.log10(2230)
@@ -61,7 +62,9 @@ xlabel = r'$\mathcal{F} =  U/\sqrt{(\Delta\rho/\rho_{0}) g H}$'
 fig, axarr = plt.subplots(1, 2, figsize=(theme.fig_width, 0.375*theme.fig_height_max),
                           constrained_layout=True)
 
-for i, (ax, quantity, cmap, norm) in enumerate(zip(axarr.flatten(), quantities, cmaps, norms)):
+for i, (ax, quantity, cmap, norm, cbtick) in enumerate(zip(axarr.flatten(),
+                                                           quantities, cmaps, norms,
+                                                           cbticks)):
     ylabel = '$k H$' if i == 0 else None
     #
     vars = [numbers[var1][mask], numbers[var2][mask]]
@@ -73,14 +76,14 @@ for i, (ax, quantity, cmap, norm) in enumerate(zip(axarr.flatten(), quantities, 
[…]

### 2022-03-11 docs/Paper_figure/Figure06.py

@@ -55,7 +55,7 @@ theme.load_style()
 # paths
 path_imgs = '../static/images/'
 path_savefig = '../Paper/Figures'
-path_outputdata = '../static/processed_data'
+path_outputdata = '../static/data/processed_data'
 
 # ##### Loading meteo data
 Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
@@ -157,7 +157,7 @@ for ax, station, label in zip(axarr.flatten(), Stations, labels):
     anchor = [1-pad_bord_x-size_rose_x, 1-(pad_bord_x+size_rose_x)*ax_aspect, size_rose_x, size_rose_x*ax_aspect]
     RDD, RDP = Vector_average(Angles, PDF[station][0, index, 0, :])
     subax = ax.inset_axes(bounds=anchor, transform=ax.transAxes)
-    a = plot_flux_rose(Angles, PDF[station][0, index, 0, :], subax, fig, cmap=theme.flux_color, edgecolor='k', linewidth=0.5, label='in situ', props=props)
+    a = plot_flux_rose(Angles, PDF[station][0, index, 0, :], subax, fig, cmap=theme.flux_color, edgecolor='k', linewidth=0.5, label='local', props=props)
     plot_arrow(a, (RDD*np.pi/180, 0), (RDD*np.pi/180, 0.85*a.get_rmax()),
                arrowprops=dict(arrowstyle="<|-", shrinkA=0, shrinkB=0, color='saddlebrown', mutation_scale=10, ls='--'))
     a.grid(linewidth=0.4, color='k', linestyle='--')

### 2022-03-11 docs/Paper_figure/Supplementary_Figures/Figure02_supp.py

@@ -1,65 +1,76 @@
 """
-============
-Figure 2 -- SI
-============
+============================
+Figure 2 -- Online Resource
+============================
 
 """
 
-import numpy as np
 import os
-from datetime import datetime
-import matplotlib.dates as mdates
-import matplotlib.pyplot as plt
 import sys
+import glob
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.patches as mpatches
 sys.path.append('../../')
 import python_codes.theme as theme
-
+from datetime import timedelta
+#
 theme.load_style()
+
+
+def make_range_broken_barh(time, dt):
+    diff = np.diff(time)
+    t_diff = np.concatenate((time[1:][diff > dt_threshold], time[:-1][diff > dt_threshold]))
+    t_diff = np.insert(t_diff, [0, t_diff.size], [time[0], time[-1]])
+    t_diff = np.array(sorted(t_diff))
+    return [(tstart, tspan) for tstart, tspan in zip(t_diff[::2], np.diff(t_diff)[::2])]
+
 
 # paths
 path_savefig = '../../Paper/Figures'
-path_outputdata = '../../static/processed_data'
-path_inputdata = '../../static/raw_data'
-
-# figure parameters
-station = 'South_Namib_Station'
-tmin, tmax = datetime(2017, 6, 3), datetime(2017, 6, 10)
-Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
-
-# Loading and recomputing some raw data
-path_insitu = os.path.join(path_inputdata, 'measured_wind_data/in_situ_wind_data_' + station + '.npy')
-Data_insitu = np.load(path_insitu, allow_pickle=True).item()
-#
-t_insitu = Data_insitu['time']
-U_insitu = Data_insitu['velocity']
-# putting angles in trigo. ref.
-Orientation_insitu = (270 - Data_insitu['direction']) % 360
+path_inputdata = '../../static/data/raw_data/'
 
 
-# ### Figure
-fig, axarr = plt.subplots(2, 1, figsize=(theme.fig_width, 0.85*theme.fig_width),
-                          constrained_layout=True, sharex=True)
+Stations = ['Adamax_Station', 'Huab_Station', 'Deep_Sea_Station', 'South_Namib_Station']
+labels = ['Adamax', 'Huab', 'North Sand Sea', 'South Sand Sea']
 
[…]

### 2022-03-11 docs/Paper_figure/Supplementary_Figures/Figure03_supp.py

@@ -1,14 +1,15 @@
 """
-============
-Figure 3 -- SI
-============
+============================
+Figure 3 -- Online Resource
+============================
 
 """
 
 import numpy as np
 import os
+from datetime import datetime
+import matplotlib.dates as mdates
 import matplotlib.pyplot as plt
-from matplotlib.colors import Normalize
 import sys
 sys.path.append('../../')
 import python_codes.theme as theme
@@ -17,43 +18,48 @@ theme.load_style()
 
 # paths
 path_savefig = '../../Paper/Figures'
-path_outputdata = '../../static/processed_data/'
+path_outputdata = '../../static/data/processed_data'
+path_inputdata = '../../static/data/raw_data'
 
+# figure parameters
+station = 'South_Namib_Station'
+tmin, tmax = datetime(2017, 6, 3), datetime(2017, 6, 10)
 Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
-Data_roughness = np.load(os.path.join(path_outputdata, 'Data_calib_roughness.npy'),
-                         allow_pickle=True).item()
 
-labels = [r'\textbf{a}', r'\textbf{b}', r'\textbf{c}', r'\textbf{d}']
-norm = Normalize(vmin=0.3, vmax=1, clip=True)
-
-# #### Figure
-fig, axrr = plt.subplots(2, 2, figsize=(theme.fig_width, 1.1*theme.fig_width),
-                         sharex=True, sharey=True, constrained_layout=True)
+# Loading and recomputing some raw data
+path_insitu = os.path.join(path_inputdata, 'measured_wind_data/in_situ_wind_data_' + station + '.npy')
+Data_insitu = np.load(path_insitu, allow_pickle=True).item()
 #
-for i, (ax, p, metric, station, label) in enumerate(zip(axrr.flatten(), Data_roughness['Pvals'],
-                                                        Data_roughness['Metrics'], Data_roughness['Stations'], labels)):
-    cs = ax.contourf(Data_roughness['z0_era_vals'], Data_roughness['z0_insitu_vals'], metric, levels=50, norm=norm)
-    for c in cs.collections:
-        c.set_edgecolor("face")
[…]

### 2022-03-11 docs/Paper_figure/Supplementary_Figures/Figure08_supp.py

@@ -1,71 +1,81 @@
 """
-============
-Figure 8 -- SI
-============
+============================
+Figure 8 -- Online Resource
+============================
 
 """
 
 import numpy as np
-import os
 import matplotlib.pyplot as plt
 import sys
+import os
 sys.path.append('../../')
 import python_codes.theme as theme
-from python_codes.plot_functions import plot_scatter_surrounded
+from python_codes.plot_functions import make_nice_histogram
 
+# Loading figure theme
 theme.load_style()
 
 # paths
 path_savefig = '../../Paper/Figures'
-path_outputdata = '../../static/processed_data/'
+path_outputdata = '../../static/data/processed_data/'
 
-# Loading data
+# Loading wind data
 Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
+Stations = sorted(Data.keys())
 
-labels = [r'\textbf{a}', r'\textbf{b}']
-
-# preparing data
-Stations = ['South_Namib_Station', 'Deep_Sea_Station']
-
-Nocturnal_wind = {'South_Namib_Station': [150, 260], 'Deep_Sea_Station': [150, 230]}
-
-# variables
-x1 = np.concatenate([Data[station]['U_star_era'][(Data[station]['Orientation_era'] > Nocturnal_wind[station][0]) & (Data[station]['Orientation_era'] < Nocturnal_wind[station][1])]
-                     for station in Stations])
-y1 = np.concatenate([Data[station]['U_star_insitu'][(Data[station]['Orientation_era'] > Nocturnal_wind[station][0]) & (Data[station]['Orientation_era'] < Nocturnal_wind[station][1])]
-                     for station in Stations])
-t1 = np.concatenate([Data[station]['time'][(Data[station]['Orientation_era'] > Nocturnal_wind[station][0]) & (Data[station]['Orientation_era'] < Nocturnal_wind[station][1])]
-                     for station in Stations])
-
-hours = [i.hour for i in t1]
+# Figure properties
+station = 'South_Namib_Station'
 #
[…]

### 2022-03-11 docs/Paper_figure/Supplementary_Figures/Figure09_supp.py

@@ -1,7 +1,7 @@
 """
-============
-Figure 9 -- SI
-============
+============================
+Figure 9 -- Online Resource
+============================
 
 """
 
@@ -11,98 +11,70 @@ import matplotlib.pyplot as plt
 import sys
 sys.path.append('../../')
 import python_codes.theme as theme
-from python_codes.plot_functions import make_nice_histogram
-
-
-def plot_vertical_profile(ax, height, Virtual_potential_temperature, grad_free_atm, theta_free_atm, blh, theta_ground, Hmax_fit, color='tab:blue', label=None):
-    Hfit = np.linspace(blh, Hmax_fit, 100)
-    #
-    line = ax.vlines(theta_ground, 0, blh/1e3, color=color, label=label, zorder=-3)
-    ax.axhline(blh/1e3, alpha=0.5, color=color, ls='--')
-    ax.plot(np.poly1d([grad_free_atm, theta_free_atm])(Hfit), Hfit/1e3, color=line.get_color(), zorder=-2)
-    ax.plot(Virtual_potential_temperature, height/1e3, '.', color=line.get_color(), zorder=-1)
-    # ax.scatter(theta_ground, blh/1e3, s=30, facecolors=line.get_color(), edgecolors='k', linewidth=2, zorder=0)
+from python_codes.plot_functions import plot_scatter_surrounded
 
 
 theme.load_style()
 
 # paths
 path_savefig = '../../Paper/Figures'
-path_outputdata = '../../static/processed_data/'
+path_outputdata = '../../static/data/processed_data/'
 
 # Loading data
 Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
 
-labels = [r'\textbf{a}', r'\textbf{b}', r'\textbf{c}', r'\textbf{d}']
+labels = [r'\textbf{a}', r'\textbf{b}']
 
-# ## vertical profiles parameters
-station = 'Deep_Sea_Station'
-time_steps_bad = [10856, 30266, 33463]
-time_steps_good = [2012, 30302, 30310]
-colors = ['tab:blue', 'tab:orange', 'tab:green']
-Hmax_fit = 10000  # [m]
-
-# ## Distribution parameters
+# preparing data
 Stations = ['South_Namib_Station', 'Deep_Sea_Station']
 
-fig, axarr = plt.subplots(2, 2, figsize=(theme.fig_width, 1*theme.fig_width),
[…]

### 2022-03-11 docs/Paper_figure/Supplementary_Figures/Figure12_supp.py

@@ -1,97 +1,69 @@
 """
-============
-Figure 12 -- SI
-============
+============================
+Figure 12 -- Online Resource
+============================
 
 """
 
 import numpy as np
 import matplotlib.pyplot as plt
-import matplotlib.colors as mpcolors
 import sys
 import os
 sys.path.append('../../')
 import python_codes.theme as theme
-from python_codes.general import smallestSignedAngleBetween, find_mode_distribution
-from python_codes.plot_functions import plot_regime_diagram
+from python_codes.meteo_analysis import mu
+from python_codes.plot_functions import make_nice_histogram
+
 
 # Loading figure theme
 theme.load_style()
 
 # paths
 path_savefig = '../../Paper/Figures'
-path_outputdata = '../../static/processed_data/'
+path_outputdata = '../../static/data/processed_data/'
 
 # ##### Loading meteo data
 Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
+
+# ## histograms parameters
 Stations = ['South_Namib_Station', 'Deep_Sea_Station']
-
-# #### Computing quantities
-
-Orientation_era = np.concatenate([Data[station]['Orientation_era'] for station in Stations])
-Orientation_insitu = np.concatenate([Data[station]['Orientation_insitu'] for station in Stations])
-U_era = np.concatenate([Data[station]['U_star_era'] for station in Stations])
-U_insitu = np.concatenate([Data[station]['U_star_insitu'] for station in Stations])
-numbers = {key: np.concatenate([Data[station][key] for station in Stations]) for key in ('Froude', 'kH', 'kLB')}
+g = 9.81  # [m/s2] gravitational constant
+z0_era = 1e-3  # [m] hydrodynamic roughness
 #
-Delta = smallestSignedAngleBetween(Orientation_era, Orientation_insitu)
-mode_delta = np.array([find_mode_distribution(Delta, i) for i in np.arange(150, 350)]).mean()
-delta_angle = np.abs(Delta)
-delta_u = (U_era - U_insitu)/U_era
-
-# #### Figure parameters
-
[…]

### 2022-03-11 docs/Paper_figure/Supplementary_Figures/Figure14_supp.py

@@ -1,0 +1,95 @@
+"""
+============================
+Figure 14 -- Online Resource
+============================
+
+"""
+
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.colors as mpcolors
+import sys
+import os
+sys.path.append('../../')
+import python_codes.theme as theme
+from python_codes.general import smallestSignedAngleBetween, find_mode_distribution
+from python_codes.plot_functions import plot_regime_diagram
+
+# Loading figure theme
+theme.load_style()
+
+# paths
+path_savefig = '../../Paper/Figures'
+path_outputdata = '../../static/data/processed_data/'
+
+# ##### Loading meteo data
+Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
+Stations = ['South_Namib_Station', 'Deep_Sea_Station']
+
+# #### Computing quantities
+
+Orientation_era = np.concatenate([Data[station]['Orientation_era'] for station in Stations])
+Orientation_insitu = np.concatenate([Data[station]['Orientation_insitu'] for station in Stations])
+U_era = np.concatenate([Data[station]['U_star_era'] for station in Stations])
+U_insitu = np.concatenate([Data[station]['U_star_insitu'] for station in Stations])
+numbers = {key: np.concatenate([Data[station][key] for station in Stations]) for key in ('Froude', 'kH', 'kLB')}
+#
+Delta = smallestSignedAngleBetween(Orientation_era, Orientation_insitu)
+mode_delta = np.array([find_mode_distribution(Delta, i) for i in np.arange(150, 350)]).mean()
+delta_angle = np.abs(Delta)
+delta_u = (U_era - U_insitu)/U_era
+
+# #### Figure parameters
+
+lims = {'Froude': (5.8e-3, 450), 'kLB': (0.009, 7.5), 'kH': (2.2e-2, 10.8)}
+cmaps = [theme.cmap_delta_theta, theme.cmap_delta_u]
+norms = [mpcolors.Normalize(vmin=0, vmax=99),
+         mpcolors.TwoSlopeNorm(vmin=-3, vcenter=0, vmax=1)]
+cbar_labels = [r'$\delta_{\theta}$ [deg.]', r'$\delta_{u}$']
[…]

### 2022-04-02 Paper/main.tex
Adding License and readmde file
Modifying figures
Attention: Re-updating repo at first submission state (loss of data -- crash system)

@@
 
-\usepackage[dvipsnames]{xcolor}
-\newcommand{\cg}{\textcolor{WildStrawberry}}
-\newcommand{\pc}{\textcolor{ForestGreen}}
-\newcommand{\new}{\textcolor{blue}}
-\newcommand{\todo}{\textcolor{red}}
-
 %---------------------------------------------------------------------------------------
@@
 
-% \title{Local wind regime induced by giant linear dunes}
-\title{Local wind regime induced by giant linear dunes: comparison of  ERA5 re-analysis with surface measurements}
+\title{Local wind regime induced by giant linear dunes: comparison of ERA5-Land re-analysis with surface measurements}
 
@@
 P. Delorme \at
-School of Geography and Environmental Science, University of Southampton, Southampton, UK.
+Energy and Environment Institute, University of Hull, Hull, UK.
 \and
@@
 P. Claudin \at
-Physique et M\'ecanique des Milieux H\'et\'erog\`enes, CNRS, ESPCI Paris, PSL Research University, Universit\'e de Paris, Sorbonne Universit\'e, Paris, France.  \email{philippe.claudin@espci.fr}
+Physique et M\'ecanique des Milieux H\'et\'erog\`enes, CNRS, ESPCI Paris, PSL Research University, Universit\'e de Paris, Sorbonne Universit\'e, Paris, France.
 }
@@
 
-The description of turbulent flows over a complex topography is a rich and active subject, that is relevant for a large variety of different environmental systems \citep{Finnigan2020}.
+The description of turbulent flows over complex topography is relevant for a large variety of different environmental systems \citep{Finnigan2020}.
 For example, the flow over hills is of primary interest for wind power, meteorological and air pollution phenomena \citep{Taylor1987}.
-The properties of these flows are also key to the understanding of the formation of ocean surface wind-driven waves \citep{Sullivan2010}, dissolution bedforms \citep{Claudin2017}, or sedimentary ripples and dunes \citep{Charru2013, Courrech2015}.
+The properties of these flows are also key to the understanding of geophysical phenomena, including the formation of wind-driven waves on the ocean surface \citep{Sullivan2010}, dissolution bedforms \citep{Claudin2017}, or sedimentary ripples and dunes \citep{Charru2013, Courrech2015}.
 Importantly, the troposphere presents a vertical structure, with a lower convective boundary layer, of typical kilometer-scale thickness, capped by a stably stratified region \citep{Stull1988}.
-The largest topographic obstacles, such as mountains can therefore interact with this upper region, and lead to internal wave generation or significant wind disturbances, such as lee-side downslope winds \citep{Durran1990}.
+The largest topographic obstacles, such as mountains, can therefore interact with this upper region and lead to internal wave generation or significant wind disturbances, such as lee-side downslope winds \citep{Durran1990}.
 
 Focusing on the wind close to the surface, two related topographic feedbacks on the windflow over dunes can be commented on separately.
-First is the effect on wind speed with documented flow acceleration on upwind slopes \citep{Weaver2011} and deceleration on downwind slopes \citep{Baddock2007}, with a speed-up factor essentially proportional to the obstacle aspect ratio \citep{Jackson1975}.
+First is the effect on wind speed, with documented flow acceleration on upwind slopes \citep{Weaver2011} and deceleration on downwind slopes \citep{Baddock2007}, where the speed-up factor is essentially proportional to the obstacle aspect ratio \citep{Jackson1975}.
 Importantly, the velocity maximum is typically shifted upwind of the obstacle crest.
 This behaviour has been theoretically predicted by means of asymptotic analysis of a neutrally stratified boundary-layer flow over an obstacle of vanishing aspect ratio \citep{Jackson1975, Mason1979, Sykes1980, Hunt1988, Belcher1998}.
-Experiments in flumes \citep{Zilker1977, Zilker1979, Frederick1988, Poggi2007}, in wind tunnels \citep{Gong1989, Finnigan1990, Gong1996} and in field conditions \citep{Taylor1987a, Claudin2013, Fernando2019, Lu2021, Bristow2022}, have also documented this effect.
+Experiments in flumes \citep{Zilker1977, Zilker1979, Frederick1988, Poggi2007, Bristow2022}, in wind tunnels \citep{Gong1989, Finnigan1990, Gong1996} and in field conditions \citep{Taylor1987a, Claudin2013, Fernando2019, Lu2021}, have also documented this effect.
 Interestingly, a similar behaviour exists for the pressure perturbation, but with a slight downwind shift for the pressure minimum \citep{Claudin2021}.
 The second effect, much less studied, is the flow deflection that occurs when the incident wind direction is not perpendicular to the ridge crest.
-While predicted to be small (less than $10^{\circ}$) in the linear regime valid for shallow topography \citep{Gadal2019}, significant flow steering has been reported in the field on the downwind side of steep enough obstacles, such as mountain ranges \citep{Kim2000, Lewis2008, Fernando2019}, well-developed sand dunes \citep{Walker2009, Hesp2015, Walker2017, Smith2017}, and valley topographies \citep{Wiggs2002}.
+While predicted to be small (less than $10^{\circ}$) in the linear regime valid for shallow topography \citep{Gadal2019}, significant flow steering has been reported in the field on the downwind side of steep enough obstacles, such as mountain ranges \citep{Kim2000, Lewis2008, Fernando2019}, well-developed sand dunes \citep{Walker2009, Hesp2015, Walker2017, Smith2017, deWinter2020}, and valley topographies \citep{Wiggs2002, Garvey2005}.
 
-For practical reasons, wind measurement over sand dunes has been performed over rather small bedforms, typically a few meters high only (corresponding to tens of meters long) \citep{Wiggs1996}.
+For practical reasons, wind measurement over sand dunes has been performed over small bedforms, typically a few meters high (corresponding to tens of meters long) (e.g.~\citet{Lancaster1996}, \citet{MckennaNeuman1997}, \citet{Sauermann2003}, \citet{Andreotti2002}, \citet{Walker2002}, \citet{Weaver2011}).
 Giant dunes, with kilometer-scale wavelengths and heights of tens of meters, are more difficult to investigate although for several reasons they provide a choice configuration for the study of turbulent flows over a complex topography.
@@
 Third, they usually form large patterns in sand seas and thus behave as rather clean periodic perturbations, in contrast with isolated dunes.
-Finally, because the morphodynamics of aeolian bedforms are strongly dependent on the local wind regime \citep{Livingstone2019}, one can expect to see the consequences of windflow disturbance by large dunes on neighbouring small dunes, in a similar manner to that reported for the effect of dunes on impact ripples \citep{Howard1977,Hood2021}.
+Finally, because the morphodynamics of aeolian bedforms are strongly dependent on the local wind regime \citep{Livingstone2019}, one can expect to see the consequences of windflow disturbance by large dunes on neighbouring small dunes.
+A similar effect is observed on the properties of impact ripple patterns due to the presence of dunes \citep{Howard1977,Hood2021}.
 
 Atmospheric flows have been much studied at the desert-scale with climate reanalyses based on global atmospheric models \citep{Blumberg1996, Livingstone2010, Ashkenazy2012, Jolivet2021, Hu2021}, such as ERA-40, ERA-Interim or ERA-5 \citep{Uppala2005, Dee2011, Hersbach2020}.
-However, the spatial resolution (tens of kilometers) of these reanalyses implies average quantities that do not resolve the smaller scales, ranging from individual dunes to the border of mountains \citep{Livingstone2010}.
+However, the spatial resolution (tens of kilometers) of these reanalyses implies average quantities that do not resolve the smaller scales of interest, which range from individual dunes to small mountains \citep{Livingstone2010}.
 Recently, the release of ERA5-Land has resolved this limitation by providing up to 70 years of hourly wind predictions at a $9$~km spatial resolution \citep{munoz2021}.
-However, its validity remains to be studied, especially in remote desert areas where data assimilation is very low.
+However, its validity remains to be studied, especially in remote desert areas where assimilation of measured data is very low.
 
@@
 Where the meteorological stations are surrounded by a relatively flat environment, we show that local measurements and regional predictions agree well.
-The agreement is also good in the interdune areas of the sand sea, except for some weak winds blowing at night, which exhibit an additional component aligned with the giant dune orientation, which are not predicted by the ERA5Land reanalysis (section 2).
+The agreement is also good in the interdune areas of the sand sea, except for some weak winds blowing at night, which exhibit an additional component aligned with the giant dune orientation.
+These winds are not predicted by the ERA5-Land reanalysis (section 2).
 Further, we are able to link the magnitude of these differences to the circadian cycle of the atmospheric boundary layer (section 3).
@@
 We measured the wind regime at four different locations in Namibia, representative of various arid environments across the Namib desert (Fig.~\ref{Fig1}, Online Resource Fig.~\ref{Fig1_supp}).
-The Adamax station was located at the Adamax waterhole to the west of Etosha Pan in northern Namibia, in a sparsely vegetated area.
-The Huab station was near the coast on a hyper-arid flat gravel plain lying above the ephemeral Huab river.
-Here, 5--6~m high barchan dunes develop from the sediment blowing out of the river valley.
-These two stations were located in a relatively flat environments.
-In contrast, the North Sand Sea and South Sand Sea stations were located in the interdunes between 50--100~m high linear dunes with kilometer-scale wavelengths and superimposed patterns.
+The Etosha West station was located at the Adamax waterhole to the west of Etosha Pan in northern Namibia, in a sparsely vegetated area.
+The Huab station was near the coast on a hyper-arid flat gravel plain lying north the ephemeral Huab river.
+Here, barchan dunes up to a few meters in height develop from the sediment blowing out of the river valley \citep{Nield2017, Hesp1998}.
+These two stations were both located in relatively flat environments.
+In contrast, the North Sand Sea and South Sand Sea stations were located in the interdunes between linear dunes with kilometer-scale wavelengths, hectometer-scale heights and superimposed patterns.
 In this section, we describe and compare winds from local measurements and climate reanalysis predictions.
@@
 \subsection{Wind and elevation data}
-At each meteorological station (Fig.~\ref{Fig1}), wind speed and direction were sampled every 10 minutes using cup anemometers and wind vanes at heights between 2~m and 3~m depending on the station.
+At each meteorological station (Fig.~\ref{Fig1}), wind speed and direction were sampled every 10 minutes using cup anemometers (Vector Instruments A100-LK) and wind vanes (Vector Instruments W200-P) at heights which varied between 2~m and 3~m depending on the station.
 The available period of measurements at each station ranged from 1 to 5 discontinuous years distributed between 2012 and 2020 (Online Resource Fig.~\ref{Fig2_supp}).
@@
 Dune properties were computed using autocorrelation on the 30-m Digital Elevation Models (DEMs) of the shuttle radar topography mission \citep{Farr2007}.
-For the South Sand Sea and North Sand Sea stations, we obtain, respectively, orientations of $85^\circ$ and $125^\circ$ with respect to the North, wavelengths of $2.6~\textup{km}$ and $2.3~\textup{km}$ and amplitudes (or half-heights) of $45~\textup{m}$ and $20~\textup{m}$ (Online Resource Fig.~\ref{Fig5_supp} for more details).
+For the North and South Sand Sea stations, we obtain, respectively, orientations of $85^\circ$ and $125^\circ$ with respect to the North, wavelengths of $2.6~\textup{km}$ and $2.3~\textup{km}$ and amplitudes (or half-heights) of $45~\textup{m}$ and $20~\textup{m}$ (Online Resource Fig.~\ref{Fig5_supp} for more details).
+This agrees with direct measurements made on site.
 
@@
 \caption{Temporal comparison between the wind data coming from the ERA5-Land climate reanalysis (orange lines) and from the local measurements (blue lines).
-Coloured swathes indicate day (between 1000 UTC and 2200 UTC) and night (before 1000 UTC or after 2200 UTC) \textbf{a--b}: Adamax station in summer.
-\textbf{b--c}: Adamax station in winter.
+Coloured swathes indicate day (between 1000 UTC and 2200 UTC) and night (before 1000 UTC or after 2200 UTC) \textbf{a--b}: Etosha West station in summer.
+\textbf{b--c}: Etosha West station in winter.
 \textbf{d--e}: North Sand Sea station in summer.
@@
 \label{section_data_feedback}
-The obtained measured and predicted wind regimes are shown in figure~\ref{Fig1}.
-In the Namib, the regional wind patterns are essentially controlled by the see breeze, resulting in strong northward components (sometimes slightly deviated by the large scale topography) present in all regional wind roses \citep{lancaster1985}.
+The measured and predicted wind regimes are shown in Fig.~\ref{Fig1}.
+In the Namib, the regional wind patterns are essentially controlled by the sea breeze, resulting in strong northward components (sometimes slightly deviated by the large scale topography) present in all regional wind roses \citep{lancaster1985}.
 These daytime winds are dominant during the period October-March (Fig.~\ref{Fig2}f and Online Resource Fig.~\ref{Fig6_supp}f).
-During April-September, an additional easterly component can be recorded during the night, induced by the combination of katabatic winds forming in the mountains, and infrequent `berg' winds, which are responsible for the high wind velocities observed \citep{lancaster1984}.
+During April-September, an additional (and often nocturnal) easterly component can also be recorded, induced by the combination of katabatic winds forming in the mountains, and infrequent `berg' winds, which are responsible for the high wind velocities observed \citep{lancaster1984}.
 The frequency of these easterly components decreases from inland to the coast.
-As a result, bidirectional wind regimes within the Namib Sand Sea and at the Adamax salt pan (Fig.~\ref{Fig1}b,d,e) and a unidirectional wind regime on the coast at the outlet of the Huab River (Fig.~\ref{Fig1}c) are observed.
+As a result, bidirectional wind regimes within the Namib Sand Sea and at the west Etosha site (Fig.~\ref{Fig1}b,d,e) and a unidirectional wind regime on the coast at the outlet of the Huab River (Fig.~\ref{Fig1}c) are observed.
 
-In the case of the Adamax and Huab stations, the time series of wind speed and direction from the regional predictions quantitatively match those corresponding to the local measurements (Fig.~\ref{Fig2}a--d) and Online Resource Fig.~\ref{Fig6_supp}a--d,~\ref{Fig7_supp}).
+In the case of the Etosha West and Huab stations, the time series of wind speed and direction from the regional predictions quantitatively match those corresponding to the local measurements (Fig.~\ref{Fig2}a--d) and Online Resource Figs.~\ref{Fig6_supp}a--d,~\ref{Fig7_supp},~\ref{Fig9_supp}).
 For the North Sand Sea and South Sand Sea stations within the giant dune field, we observe that this agreement is also good, but limited to the October-March time period (Fig.~\ref{Fig2}e--h and and Online Resource Fig.~\ref{Fig6_supp}e--h).
-As a matter of fact, the measured wind roses exhibit additional wind components aligned with the giant dune orientation, as evidenced on the satellite images (Fig.~\ref{Fig1}c,d).
+However, the field-measured wind roses exhibit additional wind components aligned with the giant dune orientation, as evidenced on the satellite images (Fig.~\ref{Fig1}c,d).
 
 More precisely, during the April-September period, the local and regional winds in the interdune match during daytime only, i.e when the southerly/southwesterly sea breeze dominates (Figs.~\ref{Fig2}e,f and \ref{Fig3}, Online Resource Fig.~\ref{Fig8_supp}).
-In the late afternoon and during the night, when the northwesterly `berg' and katabatic winds blow, measurements and predictions differ.
-In this case, the angular wind distribution of the local measurements exhibits two additional modes corresponding to reversing winds aligned with the giant dune orientation (purple frame in Fig.~\ref{Fig3}, Online Resource Figs.~\ref{Fig8_supp} and \ref{Fig9_supp}).
-This deviation is also associated with a general attenuation of the wind strength (Online Resource Fig.~\ref{Fig10_supp}).
-Remarkably, all these figures show that these wind reorientation and attenuation processes occur only at low wind velocities, typically for $u_{*} \lesssim 0.1~\textrm{m}~\textrm{s}^{-1}$.
-For shear velocities larger than $\simeq 0.25~\textrm{m}~\textrm{s}^{-1}$, the wind reorientation is not apparent.
-Finally, for intermediate shear velocities, both situations of wind flow reoriented along the dune crest and not reoriented can be successively observed (Online Resource Fig.~\ref{Fig9_supp}).
+In the late afternoon and during the night, when the easterly `berg' and katabatic winds blow, measurements and predictions differ.
+In this case, the angular wind distribution of the local measurements exhibits two additional modes corresponding to reversing winds aligned with the giant dune orientation (purple frame in Fig.~\ref{Fig3}, Online Resource Figs.~\ref{Fig8_supp} and \ref{Fig10_supp}).
+This deviation is also associated with a general attenuation of the wind strength (Online Resource Fig.~\ref{Fig11_supp}).
+Remarkably, all these figures show that these wind reorientation and attenuation processes occur only at low velocities of the regional wind, typically for $u_{*, \, \textup{ERA}} \lesssim 0.2~\textrm{m}~\textrm{s}^{-1}$.
+For shear velocities larger than $u_{*, \, \textup{ERA}} \simeq 0.3~\textrm{m}~\textrm{s}^{-1}$, the wind reorientation is not apparent.
+Finally, for intermediate shear velocities, both situations of wind flow reoriented along the dune crest and not reoriented can be successively observed (Online Resource Fig.~\ref{Fig10_supp}).
+Importantly, these values are not precise thresholds, but indicative of a crossover between regimes, whose physical interpretation is discussed in the next section.
 
@@
 In the data analysed here, the most deflected wind at both the North and South Sand Sea stations is seen to be where the incident angle is perpendicular to the giant dunes (Figs.~\ref{Fig1} and \ref{Fig3}, Online Resource Fig.~\ref{Fig8_supp}).
-It therefore appears that the incident wind angle is not the dominant control on maximum wind deflection.
-Further, and as shown in Figure~\ref{Fig3}, winds of high and low velocities show contrasting behaviour in characteristics of deflection.
+It therefore appears that in our case, the incident wind angle is not the dominant control on maximum wind deflection.
+Further, and as shown in Fig.~\ref{Fig3}, winds of high and low velocities show contrasting behaviour in characteristics of deflection.
 This suggests a change in hydrodynamical regime between the winds.
@@
 In contrast, large wind disturbances are expected to occur during the night, when the confinement is mainly induced by a shallow ABL (Fig.~\ref{Fig4}e).
-However, this strong confinement can be somewhat reduced in the case of strong winds, corresponding to large values of the Froude number (Fig.~\ref{Fig4}f).
+However, this strong confinement can be somewhat reduced in the case of strong winds, corresponding to large values of the Froude number and a less `rigid' interface (Fig.~\ref{Fig4}f).
 This is in qualitative agreement with the transition from deflected to non-deflected winds related to low and high velocities observed in our data (Sec.~\ref{section_data_feedback}).
@@
 \caption{Regime diagrams of the wind deviation $\delta_{\theta}$ (\textbf{a}) and relative attenuation/amplification $\delta_{u}$ (\textbf{b}) in the space $(\mathcal{F}, \, kH)$, containing the data from both the North Sand Sea and South Sand Sea stations.
-Blue dashed lines empirically delimit the different regimes.
-The point density in each bin of the diagrams is shown in Online Resource Fig.~\ref{Fig13_supp} -- 95\% of the data occur in the range $-1 < \delta u < 1$.
-Similar regime diagrams in the spaces $(\mathcal{F}_{\textup{I}}, \, kH)$ and $(\mathcal{F}_{\textup{I}}, \, \mathcal{F})$ are shown in Online Resource Fig.~\ref{Fig14_supp}.}
+Green dashed lines empirically delimit the different regimes.
+The point density in each bin of the diagrams is shown in Online Resource Fig.~\ref{Fig14_supp} -- 95\% of the data occur in the range $-1 < \delta u < 1$.
+Similar regime diagrams in the spaces $(\mathcal{F}_{\textup{I}}, \, kH)$ and $(\mathcal{F}_{\textup{I}}, \, \mathcal{F})$ are shown in Online Resource Fig.~\ref{Fig15_supp}.}
 \label{Fig5}
@@
 Such an amplification is rather unexpected.
-Checking the occurrence of the corresponding data, it appears that these amplifications are associated with the southerly sea breeze, and occur dominantly during the October-March period, when the other easterly wind is not present (Online Resource Fig.~\ref{Fig15_supp}a--b).
-Furthermore, they occur less frequently during the afternoon, and more frequently at the end of the day (Online Resource Fig.~\ref{Fig15_supp}c).
-This effect may be linked to a change in the flow behaviour in the lee side of the obstacles but further measurements are clearly needed in order to assess the different possibilities \citep{baines1995, Vosper2004}.
+Checking the occurrence of the corresponding data, it appears that these amplifications are associated with the southerly sea breeze, and occur dominantly during the October-March period, when the other easterly wind is not present (Online Resource Fig.~\ref{Fig16_supp}a--b).
+Furthermore, they occur less frequently during the afternoon, and more frequently at the end of the day (Online Resource Fig.~\ref{Fig16_supp}c).
+This effect may be linked to a change in the flow behaviour in the lee side of the obstacles but further measurements are needed in order to assess the different possibilities \citep{baines1995, Vosper2004}.
 
@@
 The feedback of the giant dunes on the wind flow has important implications for smaller scales bedforms.
-As illustrated in Fig.~\ref{Fig6}, small linear dunes ($\sim50$~m -wide) are often present in the 1--2~km interdune between giant linear dunes in the Namib Sand Sea.
-These smaller dunes do not exhibit the same orientation as the large ones, and are sometimes considered to be `crossing dunes'.
+As illustrated in Fig.~\ref{Fig6}, small linear dunes ($\sim50$~m wide) are often present in the 1--2~km interdune between giant linear dunes in the Namib Sand Sea \citep{Livingstone2010}.
+These smaller dunes do not exhibit the same orientation as the large ones, and are sometimes named `crossing dunes'.
 Whilst differences between large and small scale dune patterns are observed ubiquitously, they are largely attributed to the presence of two different dune growth mechanisms, leading to two different dune patterns (orientations and/or morphologies) for the same wind regime~\citep{Courrech2014, Runyon2017, lu2017, Song2019, Gadal2020, Hu2021}.
@@
 Figure~\ref{Fig6} shows how the orientations for the small and giant dunes can be derived from the locally measured and regionally predicted winds respectively (red arrows in Fig.~\ref{Fig6}).
-The feedback of the giant dunes on the wind described in this study thus provides an explanation for the existence of these small linear dunes elongating across the interdune, a dynamic which has remained unresolved to date.
-Further studies are of course needed to confirm this hypothesis.
-These crossing dunes could provide additional constraints for the inference of local winds from bedforms, similarly to what is currently performed on Mars using ripple orientations \citep{Liu2015, Hood2021}.
+These predictions require a specification for the threshold of eolian sand transport.
+Importantly, its value expressed as a shear velocity $u_{\textup{th}} \simeq 0.15~\textup{m}\textup{s}^{-1}$ is reached in the deflected wind regime already.
+The feedback of the giant dunes on the wind described in this study thus provides a potential explanation for the existence of these small linear dunes elongating across the interdune, a dynamic which has remained unresolved to date.
+These crossing dunes could provide additional constraints for the inference of local winds from bedforms, similarly to that currently performed on Mars using ripple orientations \citep{Liu2015, Hood2021}.
+Further work is needed to investigate these processes in more detail, including measurements of sediment transport and flow on the top of dunes.
 
-Also, this study highlights the interaction between giant dunes and the atmospheric boundary layer.
-It thus supports the still debated idea that the capping layer acts as a bounding surface impacting the dune dynamics, and limiting dune growth in particular~\citep{Andreotti2009}, in contrast to an unconstrained growth ever-slower with size \citep{Eastwood2011, gunn2021}.
-Interestingly, this mechanism would allow for the inference of the ABL depth from the giant bedforms spacing where measurements are not feasible or available, as e.g.
-performed by~\citet{Lorenz2010} on Titan.
+This study presents the evidence that wind flow patterns around giant dunes are influenced by the atmospheric boundary layer, particularly during nocturnal conditions.
+It leaves open the debate as to whether the size of giant dunes is limited by the depth of this layer \citep{Andreotti2009}, in contrast to an unconstrained dune growth, ever-slower with size \citep{Gunn2021b}.
+More field evidence is needed from additional dune fields, but this mechanism would then allow for the inference of the ABL depth from giant bedform wavelengths where measurements are not feasible or available, such as Titan \citep{Lorenz2010}.
 
-To sum up conditions under which the ERA5-Land reanalysis data can reliably be used to study dune morphodynamics, we summarise the comparison of local (direct measurements) and regional (climate reanalysis) wind data.
+To conclude on conditions under which the ERA5-Land reanalysis data can reliably be used to study dune morphodynamics, we summarise the comparison of local (direct measurements) and regional (climate reanalysis) wind data.
 In flat areas, the agreement between the two confirms the ability of the ERA5-Land climate reanalysis to predict the wind regime down to scales $\sim10$~km, i.e the model grid.
@@
 when the aspect ration of these ridges is small ($k\xi_0 \ll 1$).
-These models are able to qualitatively reproduce the observed wind deflection (Appendix 1, Online Resource Figs.~\ref{Fig14_supp} and \ref{Fig16_supp}), and thus provide the physical support for the interpretation we propose here based on hydrodynamic regimes.
+These models are able to qualitatively reproduce the observed wind deflection (Appendix 1, Online Resource Figs.~\ref{Fig15_supp} and \ref{Fig17_supp}), and thus provide the physical support for the interpretation we propose here based on hydrodynamic regimes.
 However, these models cannot quantitatively predict the magnitude of these observations, probably due to the presence of expected non-linearities in high confinement situations linked to strong flow modulations.
-Besides, these linear calculations only predict wind attenuation in the interdune, in contrast with the observed enhanced velocities associated with particular evening winds from the South during the period October-March (Online Resource Fig.~\ref{Fig15_supp}).
+Besides, these linear calculations only predict wind attenuation in the interdune, in contrast with the observed enhanced velocities associated with particular evening winds from the South during the period October-March (Online Resource Fig.~\ref{Fig16_supp}).
 Some other models predict different spatial flow structures in response to a modulated topography, such as lee waves and rotors \citep{baines1995, Vosper2004}.
-Of course they cannot be observed by our measurements located at a single point in the interdune.
+However, our measurements are located at a single point in the interdune, so we are unable to explore these types of responses.
 Data at different places along and across the ridges are needed to investigate and possibly map such flow structures, and for further comparisons with the models.
@@
 
-ERA5 and ERA5-Land datasets are publicly available at the Copernicus Climate Change Service (C3S) Climate Data Store.
+All data used in this study can be found in \citet{DataPaper}.
+Note that it contains modified Copernicus Climate Change Service Information (2021).
+Neither the European Commission nor ECMWF is responsible for any use that may be made of the Copernicus Information or Data it
+contains. Fully documented codes used to analyse this study are available at https://github.com/Cgadal/GiantDunes (will be made public upon acceptance of this manuscript for publication).
+
+% ERA5 and ERA5-Land datasets are publicly available at the Copernicus Climate Change Service (C3S) Climate Data Store. As such, It contains modified Copernicus Climate Change
+% Service Information (2021). Neither the European Commission nor ECMWF is
+% responsible for any use that may be made of the Copernicus Information or Data it
+% contains.
 The locally measured wind data can be found at \cg{upload on public data repository}.
@@
 
-[citing all relevant grants ...] \pc{TOAD}
+Multiple grants have supported the collection of wind data through visits to the four sites between 2013 and 2020 (John Fell Oxford University Press (OUP) Research Fund (121/474); National Geographic (CP-029R-17); Natural Environment Research Council UK (NE/R010196/1 and NE/H021841/1 NSFGEO-NERC); Southampton Marine and Maritime Institute SMMI EPSRC-GCRF UK), along with research permits (1978/2014, 2140/2016, 2304/2017, 2308/2017, RPIV00022018, RPIV0052018, RPIV00230218).
+The authors are very grateful for support from Etosha National Park (especially Shyane Kötting, Boas Erckie, Pierre du Preez, Claudine Cloete, Immanuel Kapofi, Wilferd Versfeld, and Werner Kilian), Gobabeb Namib Research Institute (Gillian Maggs-Kölling and Eugene Marais), The Skeleton Coast National Park (Joshua Kazeurua).
+Various researchers and desert enthusiasts have assisted with instruments and the logistics of expeditions, especially Mary Seely for expert guidance at the North Sand Sea site.
+
+Finally, we acknowledge financial support from the Laboratoire d’Excellence UnivEarthS Grant ANR-10-LABX-0023, the Initiative d’Excellence Universite de Paris Grant ANR-18-IDEX-0001, the French National Research Agency Grants ANR-17-CE01-0014/SONO and the National Science Center of Poland Grant 2016/23/B/ST10/01700.
+
 \end{acknowledgements}
@@
 
-With capping layer height and Froude numbers computed from the ERA5-Land time series, the corresponding $\mathcal{A}_{0}$ and $\mathcal{B}_{0}$ can be deduced, allowing to produce maps as displayed in Online Resource Fig.~\ref{Fig16_supp}.
-Interestingly, it shows similar regimes as in the diagrams of Fig.~\ref{Fig5} and Online Resource Fig.~\ref{Fig14_supp}a,b, supporting the physics picture.
-However, the matching remains qualitative only.
-As a matter of fact, the linearity assumption of the theoretical framework requires $\left(\vert \tau \vert - \tau_{0}\right)/\tau_{0} \ll 1$, which translates into $k\xi\sqrt{\mathcal{A}_{0}^{2} + \mathcal{B}_{0}^{2}} \ll 1$.
+With capping layer height and Froude numbers computed from the ERA5-Land time series, the corresponding $\mathcal{A}_{0}$ and $\mathcal{B}_{0}$ can be deduced, as displayed in Online Resource Fig.~\ref{Fig17_supp}.
+Interestingly, it shows similar regimes as in the diagrams of Fig.~\ref{Fig5} and Online Resource Fig.~\ref{Fig15_supp}a,b, supporting the underlying physics.
+However, the agreement is qualitative only.
+Further, the linearity assumption of the theoretical framework requires $\left(\vert \tau \vert - \tau_{0}\right)/\tau_{0} \ll 1$, which translates into $k\xi\sqrt{\mathcal{A}_{0}^{2} + \mathcal{B}_{0}^{2}} \ll 1$.
 In our case, the giant dune morphology gives $k\xi_0 \simeq 0.1$, which means that one quits the regime of validity of the linear theory when the coefficient modulus $\sqrt{\mathcal{A}_{0}^{2} + \mathcal{B}_{0}^{2}}$ becomes larger than a few units.
 In accordance with the theoretical expectations, these coefficients present values on the order of unity ($\mathcal{A}_{0} \simeq 3$ and $\mathcal{B}_{0} \simeq 1$) in unconfined situations \citep{Claudin2013, Lu2021}.
-In contrast and as illustrated in Online Resource Fig.~\ref{Fig16_supp}a,b, larger values are predicted in case of strong confinement, which does not allow us to proceed to further quantitative comparison with the data.
+In contrast and as illustrated in Online Resource Fig.~\ref{Fig17_supp}a,b, larger values are predicted in case of strong confinement, which does not allow us to proceed to further quantitative comparison with the data.
 
-Finally, the linear model is also able to reproduce the enhancement of the flow deflection over the sinusoidal ridges when $\sqrt{\mathcal{A}_{0}^{2} + \mathcal{B}_{0}^{2}}$ is increased (Online Resource Fig.~\ref{Fig16_supp}).
+Finally, the linear model is also able to reproduce the enhancement of the flow deflection over the sinusoidal ridges when $\sqrt{\mathcal{A}_{0}^{2} + \mathcal{B}_{0}^{2}}$ is increased (Online Resource Fig.~\ref{Fig17_supp}).
 Here, using $k\xi_0 \simeq 0.1$ to be representative of the amplitude of the giant dunes at the North Sand Sea station, the coefficient modulus is bounded to $10$.
@@
 We thus set the roughness in the ERA5-Land analysis to the typical value $z_0=10^{-3}$~m, and deduce the corresponding ones for the local stations.
-It leads to $2.7$, $0.8$, $0.1$ and $0.5$~mm for the Adamax, North Sand Sea, Huab and South Sand Sea stations, respectively.
+It leads to $2.7$, $0.8$, $0.1$ and $0.5$~mm for the Etosha West, North Sand Sea, Huab and South Sand Sea stations, respectively.
 Importantly, this approach somewhat impacts the calculation of the shear velocities, but not that of the wind directions.
@@
 In arid areas, the vertical structure of the atmosphere can be approximated by a well mixed convective boundary layer of height $H$, topped by the stratified free atmosphere \citep{Stull1988, Shao2008}.
-In this context, one usually introduces the virtual potential temperature $T_{\textup{vp}}$, which is a constant $T_{0}$ inside the boundary layer, and increases linearly in the FA (Online Resource Fig.~\ref{Fig11_supp}a):
+In this context, one usually introduces the virtual potential temperature $T_{\textup{vp}}$, which is a constant $T_{0}$ inside the boundary layer, and increases linearly in the FA (Online Resource Fig.~\ref{Fig12_supp}a):
 %
@@
 
-Examples of vertical profiles of the virtual potential temperature deduced from ERA5 are shown in Online Resource Fig.~\ref{Fig11_supp}a.
+Examples of vertical profiles of the virtual potential temperature deduced from ERA5 are shown in Online Resource Fig.~\ref{Fig12_supp}a.
 For each of them, an average temperature is computed below the ABL depth ($z<H$), and a linear function is fitted above, allowing us to extract the temperature jump $\Delta T_{\textup{vp}}$.
-Importantly, some profiles display a vertical structure that cannot be approximated by the simple form \eqref{TvpVerticalProfile} used here (Online Resource Fig.~\ref{Fig11_supp}b).
+Importantly, some profiles display a vertical structure that cannot be approximated by the simple form \eqref{TvpVerticalProfile} used here (Online Resource Fig.~\ref{Fig12_supp}b).
 In practice, we removed from the analysis all of those leading to the unphysical case $\Delta T_{\textup{vp}}<0$.
 We have noticed that these `ill-processed' profiles dominantly occur in winter and are evenly spread across the hours of the day.
-Importantly, they represent $\simeq12$\% of the data only (Online Resource Fig.~\ref{Fig11_supp}c,d), and we are thus confident that this data treatment does not affect our conclusions.
+Importantly, they represent $\simeq12$\% of the data only (Online Resource Fig.~\ref{Fig12_supp}c,d), and we are thus confident that this data treatment does not affect our conclusions.
 
@@
 \includegraphics[scale=1]{Figures/Figure1_supp.pdf}
-\caption{Photographs of the meteorological stations. \textbf{a}: South Sand Sea station. \textbf{b--c}: North Sand Sea station. \textbf{d}: Huab station. \textbf{e} Adamax station.}
+\caption{Photographs of the meteorological stations. \textbf{a}: South Sand Sea station. \textbf{b--c}: North Sand Sea station. \textbf{d}: Huab station. \textbf{e} Etosha West station.}
 \label{Fig1_supp}
@@
 \caption{Calibration of hydrodynamic roughness.
-The parameter $\delta$ (Eq.~\ref{metric_roughness}) quantifying the difference between local and predicted winds is shown in colorscale as a function of the hydrodynamic roughnesses chosen for the ERA5-Land and for local winds, for the (\textbf{a}) Adamax, (\textbf{b}) North Sand Sea, (\textbf{c}) Huab and (\textbf{d}) South Sand Sea stations.
+The parameter $\delta$ (Eq.~\ref{metric_roughness}) quantifying the difference between local and predicted winds is shown in colorscale as a function of the hydrodynamic roughnesses chosen for the ERA5-Land and for local winds, for the (\textbf{a}) Etosha West, (\textbf{b}) North Sand Sea, (\textbf{c}) Huab and (\textbf{d}) South Sand Sea stations.
 The red dashed and plain lines shows the minima of $\delta$ and the identity line, respectively.
@@
 \includegraphics[scale=1]{Figures/Figure7_supp.pdf}
-\caption{Statistical comparison of the wind orientation (\textbf{a}) and velocity (\textbf{b}) between the ERA5-Land dataset and the local measurements for the Huab and Adamax stations.
+\caption{Statistical comparison of the wind orientation (\textbf{a}) and velocity (\textbf{b}) between the ERA5-Land dataset and the local measurements for the Huab and Etosha West stations.
 Data point clustering around identity lines (dashed and black) provide evidence for agreement of the two sets.}
@@
 \includegraphics[scale=1]{Figures/Figure9_supp.pdf}
-\caption{Statistical comparison of the wind orientation between the ERA5-Land dataset and the local measurements for the South Sand Sea and North Sand Sea stations, for different velocity ranges.
-\textbf{a}: $u_{*, \textup{\textup{ERA}}} < 0.1~\textup{m}~\textup{s}^{-1}$.
-\textbf{b}: $0.1 < u_{*, \textup{\textup{ERA}}} \leq 0.25~\textup{m}~\textup{s}^{-1}$.
-\textbf{c}: $u_{*, \textup{\textup{ERA}}} \geq 0.25~\textup{m}~\textup{s}^{-1}$.
-The measured dune orientations are subtracted to the wind orientation, which allows us to plot both stations on the same graph.
-Black dashed lines indicate locally measured orientations aligned with the dune crests (here $0^\circ$, $180^\circ$ and $360^\circ$ -- panels \textbf{a}, \textbf{b}), as well as the identity lines (panels \textbf{b}, \textbf{c}).}
+\caption{Distributions of wind direction at the Huab and Etosha West stations for the ERA5-Land climate reanalysis (orange) and the local measurements (blue).
+In each subplot, both distributions are plotted from the same time steps, selected with constraints on the wind wind velocity (rows) in the ERA5-Land dataset.
+The numbers at the top centre give the percentage of time steps selected in each sub-range, as well as the percentage of them corresponding to the day (between 1000 UTC and 2200 UTC).
+Compared to the North and South Namib stations (Fig.~\ref{Fig3} and Fig.~\ref{Fig8_supp}), histograms match for high and low velocities.}
 \label{Fig9_supp}
@@
 \includegraphics[scale=1]{Figures/Figure10_supp.pdf}
-\caption{Statistical comparison of the wind velocity between the ERA5-Land dataset and the local measurements for the South Sand Sea and North Sand Sea stations.
-\textbf{a}: Nocturnal summer easterly wind.
-\textbf{b}: Diurnal southerly wind.
-Black dashed lines are identity lines.
-The angle ranges used to select diurnal and nocturnal summer winds are the same as those in Fig.~\ref{Fig3} and Online Resource Fig.~\ref{Fig8_supp}.}
+\caption{Statistical comparison of the wind orientation between the ERA5-Land dataset and the local measurements for the South Sand Sea and North Sand Sea stations, for different velocity ranges.
+\textbf{a}: $u_{*, \textup{\textup{ERA}}} < 0.1~\textup{m}~\textup{s}^{-1}$.
+\textbf{b}: $0.1 < u_{*, \textup{\textup{ERA}}} \leq 0.25~\textup{m}~\textup{s}^{-1}$.
+\textbf{c}: $u_{*, \textup{\textup{ERA}}} \geq 0.25~\textup{m}~\textup{s}^{-1}$.
+The measured dune orientations are subtracted to the wind orientation, which allows us to plot both stations on the same graph.
+Black dashed lines indicate locally measured orientations aligned with the dune crests (here $0^\circ$, $180^\circ$ and $360^\circ$ -- panels \textbf{a}, \textbf{b}), as well as the identity lines (panels \textbf{b}, \textbf{c}).}
 \label{Fig10_supp}
@@
 \includegraphics[scale=1]{Figures/Figure11_supp.pdf}
-\caption{\textbf{a}: Vertical profiles of the virtual potential temperature at three different times (blue: 29/11/2012 - 1100 UTC, orange: 21/03/2017 - 1200 UTC, green: 21/03/2017 - 2000 UTC) at the South Sand Sea station.
-Dots: data from the ERA5 reanalysis.
-Dashed lines: boundary layer height given by the ERA5 reanalysis.
-Plain lines: vertical (ABL) and linear (FA) fits to estimate the quantities displayed in Online Resource Fig.~\ref{Fig12_supp}.
-\textbf{b}: Examples of ill-processed vertical profiles at three different times (blue: 2/12/2013 - 2300 UTC, orange: 20/03/2017 - 0000 UTC, green: 14/07/2017 - 1400 UTC) at the South Sand Sea station.
-\textbf{c}: Hourly distribution of ill-processed vertical profiles.
-\textbf{d}: Monthly distribution of ill-processed vertical profiles.}
+\caption{Statistical comparison of the wind velocity between the ERA5-Land dataset and the local measurements for the South Sand Sea and North Sand Sea stations.
+\textbf{a}: Nocturnal summer easterly wind.
+\textbf{b}: Diurnal southerly wind.
+Black dashed lines are identity lines.
+The angle ranges used to select diurnal and nocturnal summer winds are the same as those in Fig.~\ref{Fig3} and Online Resource Fig.~\ref{Fig8_supp}.}
 \label{Fig11_supp}
@@
 \centering
-\includegraphics{Figures/Figure12_supp.pdf}
-\caption{Distributions of the meteorological parameters resulting from the processing of the ERA5-Land data for the South Sand Sea (blue) and the North Sand Sea (orange) stations.}
+\includegraphics[scale=1]{Figures/Figure12_supp.pdf}
+\caption{\textbf{a}: Vertical profiles of the virtual potential temperature at three different times (blue: 29/11/2012 - 1100 UTC, orange: 21/03/2017 - 1200 UTC, green: 21/03/2017 - 2000 UTC) at the South Sand Sea station.
+Dots: data from the ERA5 reanalysis.
+Dashed lines: boundary layer height given by the ERA5 reanalysis.
+Plain lines: vertical (ABL) and linear (FA) fits to estimate the quantities displayed in Online Resource Fig.~\ref{Fig13_supp}.
+\textbf{b}: Examples of ill-processed vertical profiles at three different times (blue: 2/12/2013 - 2300 UTC, orange: 20/03/2017 - 0000 UTC, green: 14/07/2017 - 1400 UTC) at the South Sand Sea station.
+\textbf{c}: Hourly distribution of ill-processed vertical profiles.
+\textbf{d}: Monthly distribution of ill-processed vertical profiles.}
 \label{Fig12_supp}
@@
 \includegraphics[scale=1]{Figures/Figure13_supp.pdf}
-\caption{Non-dimensional parameters distributions. For the marginal distributions, the orange correspond to the South Sand Sea station, and the blue to the North Sand Sea station.}
+\caption{Distributions of the meteorological parameters resulting from the processing of the ERA5-Land data for the South Sand Sea (blue) and the North Sand Sea (orange) stations.}
 \label{Fig13_supp}
@@
 \includegraphics[scale=1]{Figures/Figure14_supp.pdf}
-\caption{Regime diagrams of the wind deviation $\delta_{\theta}$ and relative attenuation/amplification $\delta_{u}$ in the spaces $(\mathcal{F}_{\textup{I}}, \, kH)$ and $(\mathcal{F}_{\textup{I}}, \, \mathcal{F})$, containing the data from both the North Sand Sea and South Sand Sea stations.
-Blue dashed lines empirically delimit the different regimes.
-The point density in each bin of the diagrams is shown in Online Resource Fig.~\ref{Fig13_supp} -- 95\% of the data occur in the range $-1 < \delta u < 1$.
-The similar regime diagrams in the space $(\mathcal{F}, \, kH)$ are shown in Fig.~\ref{Fig5}.}
+\caption{Non-dimensional parameters distributions. For the marginal distributions, the orange correspond to the South Sand Sea station, and the blue to the North Sand Sea station.}
 \label{Fig14_supp}
@@
 \includegraphics[scale=1]{Figures/Figure15_supp.pdf}
-\caption{Normalized distributions of amplified velocities for the North sand Sea (blue: $\delta_u < 0$, orange: $\delta_u < -0.5$).
-\textbf{a}: Angular distributions.
-\textbf{b}: Monthly distributions.
-\textbf{c}: Hourly distributions.}
+\caption{Regime diagrams of the wind deviation $\delta_{\theta}$ and relative attenuation/amplification $\delta_{u}$ in the spaces $(\mathcal{F}_{\textup{I}}, \, kH)$ and $(\mathcal{F}_{\textup{I}}, \, \mathcal{F})$, containing the data from both the North Sand Sea and South Sand Sea stations.
+Green dashed lines empirically delimit the different regimes.
+The point density in each bin of the diagrams is shown in Online Resource Fig.~\ref{Fig14_supp} -- 95\% of the data occur in the range $-1 < \delta u < 1$.
+The similar regime diagrams in the space $(\mathcal{F}, \, kH)$ are shown in Fig.~\ref{Fig5}.}
 \label{Fig15_supp}
@@
 \includegraphics[scale=1]{Figures/Figure16_supp.pdf}
+\caption{Normalized distributions of amplified velocities for the North sand Sea (blue: $\delta_u < 0$, orange: $\delta_u < -0.5$).
+\textbf{a}: Angular distributions.
+\textbf{b}: Monthly distributions.
+\textbf{c}: Hourly distributions.}
+\label{Fig16_supp}
+\end{figure}
+%---------------------------------------------
+
+%---------------------------------------------
+\begin{figure}[p]
+\centering
+\includegraphics[scale=1]{Figures/Figure17_supp.pdf}
 \caption{Computation of the flow disturbance with the linear model of \citet{Andreotti2009}.
@@
 From the upper to the lower streamline, values of $\left(kH,\, \mathcal{F},\, \mathcal{F}_{\rm I},\, \mathcal{A}_{0},\, \mathcal{B}_{0},\, \sqrt{\mathcal{A}_{0}^{2} + \mathcal{B}_{0}^{2}}\right)$ are (1.9, 0.6, 1.5, 3.4, 1.0, 3.5), (1.5, 0.3, 0.4, 4.8, 1.4, 5.0), (0.1, 3.5, 1.0, 8.6, 0.1, 8.6), (0.5, 0.05, 0.04, 9.6, 2.5, 9.9).}
-\label{Fig16_supp}
+\label{Fig17_supp}
 \end{figure}

### 2022-04-02 Paper_figure/Figure05.py

@@ -52,8 +52,6 @@ cbticks = [[0, 25, 50, 75], [-3, -1.5, 0, 0.5, 1]]
 
 mask = ~np.isnan(numbers['Froude'])
 log_counts_max = np.log10(2230)
-regime_line_color = 'tab:blue'
-
 
 var1, var2 = 'Froude', 'kH'
 xlabel = r'$\mathcal{F} =  U/\sqrt{(\Delta\rho/\rho_{0}) g H}$'
@@ -78,8 +76,8 @@ for i, (ax, quantity, cmap, norm, cbtick) in enumerate(zip(axarr.flatten(),
     ax.text(0.04, 0.94, labels[i], transform=ax.transAxes, ha='left', va='center')
     #
     # regime lines
-    ax.axvline(0.4, color=regime_line_color, linestyle='--', lw=2)
-    ax.axhline(0.32, color=regime_line_color, linestyle='--', lw=2)
+    ax.axvline(0.4, color=theme.regime_line_color, linestyle='--', lw=2)
+    ax.axhline(0.32, color=theme.regime_line_color, linestyle='--', lw=2)
     #
     # colorbar
     sm = plt.cm.ScalarMappable(cmap=cmap, norm=norm)

### 2022-04-02 Paper_figure/Supplementary_Figures/Figure02_supp.py

@@ -32,7 +32,7 @@ path_inputdata = '../../static/data/raw_data/'
 
 
 Stations = ['Adamax_Station', 'Huab_Station', 'Deep_Sea_Station', 'South_Namib_Station']
-labels = ['Adamax', 'Huab', 'North Sand Sea', 'South Sand Sea']
+labels = ['Etosha West', 'Huab', 'North Sand Sea', 'South Sand Sea']
 
 directory_types = ['ERA5Land', 'measured_wind_data']
 

### 2022-04-02 Paper_figure/Supplementary_Figures/Figure09_supp.py

@@ -6,75 +6,68 @@ Figure 9 -- Online Resource
 """
 
 import numpy as np
-import os
 import matplotlib.pyplot as plt
 import sys
+import os
 sys.path.append('../../')
 import python_codes.theme as theme
-from python_codes.plot_functions import plot_scatter_surrounded
+from python_codes.plot_functions import make_nice_histogram
 
-
+# Loading figure theme
 theme.load_style()
 
 # paths
 path_savefig = '../../Paper/Figures'
 path_outputdata = '../../static/data/processed_data/'
 
-# Loading data
-Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
+# Loading wind data
+Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'),
+               allow_pickle=True).item()
+Stations = ['Huab_Station', 'Adamax_Station']
 
-labels = [r'\textbf{a}', r'\textbf{b}']
-
-# preparing data
-Stations = ['South_Namib_Station', 'Deep_Sea_Station']
-
-Dune_orientations = [np.load(os.path.join(path_outputdata, 'Data_DEM.npy'), allow_pickle=True).item()[station]['orientation']
-                     for station in Stations]
-
-velocity_thresholds = [0.1, 0.25]
-
-# variables
-x1 = np.concatenate([Data[station]['Orientation_era'][Data[station]['U_star_era'] < velocity_thresholds[0]] - Dune_orientations[Stations.index(station)]
-                     for station in Stations])
-y1 = np.concatenate([Data[station]['Orientation_insitu'][Data[station]['U_star_era'] < velocity_thresholds[0]] - Dune_orientations[Stations.index(station)]
-                     for station in Stations])
-
-x2 = np.concatenate([Data[station]['Orientation_era'][(Data[station]['U_star_era'] >= velocity_thresholds[0]) & (Data[station]['U_star_era'] < velocity_thresholds[1])] - Dune_orientations[Stations.index(station)]
-                     for station in Stations])
-y2 = np.concatenate([Data[station]['Orientation_insitu'][(Data[station]['U_star_era'] >= velocity_thresholds[0]) & (Data[station]['U_star_era'] < velocity_thresholds[1])] - Dune_orientations[Stations.index(station)]
[…]

### 2022-04-02 Paper_figure/Supplementary_Figures/Figure15_supp.py

@@ -1,18 +1,19 @@
 """
 ============================
-Figure 15 -- Online Resource
+Figure 15-- Online Resource
 ============================
 
 """
 
 import numpy as np
 import matplotlib.pyplot as plt
+import matplotlib.colors as mpcolors
 import sys
 import os
 sys.path.append('../../')
 import python_codes.theme as theme
-from python_codes.plot_functions import make_nice_histogram
-
+from python_codes.general import smallestSignedAngleBetween, find_mode_distribution
+from python_codes.plot_functions import plot_regime_diagram
 
 # Loading figure theme
 theme.load_style()
@@ -23,55 +24,71 @@ path_outputdata = '../../static/data/processed_data/'
 
 # ##### Loading meteo data
 Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
-# Stations = ['South_Namib_Station', 'Deep_Sea_Station']
-Stations = ['Deep_Sea_Station']
+Stations = ['South_Namib_Station', 'Deep_Sea_Station']
 
 # #### Computing quantities
+
 Orientation_era = np.concatenate([Data[station]['Orientation_era'] for station in Stations])
+Orientation_insitu = np.concatenate([Data[station]['Orientation_insitu'] for station in Stations])
 U_era = np.concatenate([Data[station]['U_star_era'] for station in Stations])
 U_insitu = np.concatenate([Data[station]['U_star_insitu'] for station in Stations])
-time = np.concatenate([Data[station]['time'] for station in Stations])
-month = np.array([i.month for i in time])
-hour = np.array([i.hour for i in time])
+numbers = {key: np.concatenate([Data[station][key] for station in Stations]) for key in ('Froude', 'kH', 'kLB')}
 #
+Delta = smallestSignedAngleBetween(Orientation_era, Orientation_insitu)
+mode_delta = np.array([find_mode_distribution(Delta, i) for i in np.arange(150, 350)]).mean()
+delta_angle = np.abs(Delta)
 delta_u = (U_era - U_insitu)/U_era
[…]

### 2022-05-22 Paper/main.tex
Re-arranging figure for review:
- the former fig1 is now split in two
- fig 3 -- 6 integrate material that was in the supplementary before
- On the fig 3 and 5, possible numbers that quantify the discrepancy
between the time series as asked by Ryan
- a new figure at the end of the supp to show that our choice of z0 has
little impact on the results (not referenced in the text yet)

New figures main manuscript

Re-formatting supplementary figure names

Adding new figure to supp (delta_u as a function of z0)

New figure set: updating captions and referencing through text

Last mofis on new figure

@@
 \includegraphics[scale=1]{Figures/Figure1.pdf}
-\caption{Wind data used in this study \textbf{a}: Location of the different sites in Namibia.
-\textbf{b--e}: Satellite images of these different environements (Google-Earth, Maxar Technologies, CNES/Airbus).
-In each subplot, the left and right wind roses represent the data from the ERA5-Land climate reanalysis and the local wind stations, respectively.
-Note that the bars show the direction towards which the wind blows.
-The black dots show the location of local wind stations.}
+\caption{Studied field sites. \textbf{a}: Location of the different sites in Namibia. \textbf{b--e}: Photographs of the meteorological stations.}
 \label{Fig1}
+\end{figure}
+%---------------------------------------------
+
+%---------------------------------------------
+\begin{figure}[t]
+\centering
+\includegraphics[scale=1]{Figures/Figure2.pdf}
+\caption{Wind data used in this study.
+Satellite images of these different environements (Google-Earth, Maxar Technologies, CNES/Airbus) are shown on the left.
+The black dots show the location of the wind measurements stations.
+On the center and on the right, wind roses representing the data from the ERA5-Land climate reanalysis and the local wind stations are respectively shown.
+Note that the bars show the direction towards which the wind blows.}
+\label{Fig2}
 \end{figure}
@@
 \section{Wind regimes across the Namib Sand Sea}
-We measured the wind regime at four different locations in Namibia, representative of various arid environments across the Namib desert (Fig.~\ref{Fig1}, Online Resource Fig.~\ref{Fig1_supp}).
+We measured the wind regime at four different locations in Namibia, representative of various arid environments across the Namib desert (Fig.~\ref{Fig1}, Fig.~\ref{Fig2}).
 The Etosha West station was located at the Adamax waterhole to the west of Etosha Pan in northern Namibia, in a sparsely vegetated area.
@@
 At each meteorological station (Fig.~\ref{Fig1}), wind speed and direction were sampled every 10 minutes using cup anemometers (Vector Instruments A100-LK) and wind vanes (Vector Instruments W200-P) at heights which varied between 2~m and 3~m depending on the station.
-The available period of measurements at each station ranged from 1 to 5 discontinuous years distributed between 2012 and 2020 (Online Resource Fig.~\ref{Fig2_supp}).
+The available period of measurements at each station ranged from 1 to 5 discontinuous years distributed between 2012 and 2020 (Online Resource Fig.~\ref{Fig1_supp}).
 We checked that at least one complete seasonal cycle was available for each station.
@@
 
-To enable direct comparison, the local wind measurements were averaged into 1-hr bins centered on the temporal scale of the ERA5-Land estimates (Online Resource Fig.~\ref{Fig3_supp}).
+To enable direct comparison, the local wind measurements were averaged into 1-hr bins centered on the temporal scale of the ERA5-Land estimates (Online Resource Fig.~\ref{Fig2_supp}).
 As the wind velocities of both datasets were provided at different heights, we converted them into shear velocities $u_{*}$ (Online Resource section 1), characteristic of the turbulent wind profile.
-Wind roses in Fig.~\ref{Fig1}(b--e) show the resulting wind data.
+Wind roses in Fig.~\ref{Fig2} show the resulting wind data.
 
 Dune properties were computed using autocorrelation on the 30-m Digital Elevation Models (DEMs) of the shuttle radar topography mission \citep{Farr2007}.
-For the North and South Sand Sea stations, we obtain, respectively, orientations of $85^\circ$ and $125^\circ$ with respect to the North, wavelengths of $2.6~\textup{km}$ and $2.3~\textup{km}$ and amplitudes (or half-heights) of $45~\textup{m}$ and $20~\textup{m}$ (Online Resource Fig.~\ref{Fig5_supp} for more details).
+For the North and South Sand Sea stations, we obtain, respectively, orientations of $85^\circ$ and $125^\circ$ with respect to the North, wavelengths of $2.6~\textup{km}$ and $2.3~\textup{km}$ and amplitudes (or half-heights) of $45~\textup{m}$ and $20~\textup{m}$ (Online Resource Fig.~\ref{Fig4_supp} for more details).
 This agrees with direct measurements made on site.
@@
 \centering
-\includegraphics[scale=1]{Figures/Figure2.pdf}
+\includegraphics[scale=1]{Figures/Figure3.pdf}
 \caption{Temporal comparison between the wind data coming from the ERA5-Land climate reanalysis (orange lines) and from the local measurements (blue lines).
@@
 \textbf{b--c}: Etosha West station in winter.
-\textbf{d--e}: North Sand Sea station in summer.
-\textbf{f--g}: North Sand Sea station in winter.
-Time series of the two other stations are shown in Online Resource Fig.~\ref{Fig6_supp}.}
-\label{Fig2}
+\textbf{d--e}: Huab station in summer.
+\textbf{f--g}: Huab station in winter.
+Time series of the two other stations are shown in Fig.~\ref{Fig5}.}
+\label{Fig3}
+\end{figure}
+%---------------------------------------------
+
+%---------------------------------------------
+\begin{figure}
+\centering
+\includegraphics[scale=1]{Figures/Figure4.pdf}
+\caption{Distributions of wind direction at the Huab and Etosha West stations for the
+ERA5-Land climate reanalysis (orange) and the local measurements (blue). In each subplot,
+both distributions are plotted from the same time steps, selected with constraints on the
+wind wind velocity (rows) in the ERA5-Land dataset. The numbers at the top centre give
+the percentage of time steps selected in each sub-range, as well as the percentage of them
+corresponding to the day (between 1000 UTC and 2200 UTC).}
+\label{Fig4}
 \end{figure}
@@
 \label{section_data_feedback}
-The measured and predicted wind regimes are shown in Fig.~\ref{Fig1}.
+The measured and predicted wind regimes are shown in Fig.~\ref{Fig2}.
 In the Namib, the regional wind patterns are essentially controlled by the sea breeze, resulting in strong northward components (sometimes slightly deviated by the large scale topography) present in all regional wind roses \citep{lancaster1985}.
-These daytime winds are dominant during the period October-March (Fig.~\ref{Fig2}f and Online Resource Fig.~\ref{Fig6_supp}f).
+These daytime winds are dominant during the period October-March (Fig.~\ref{Fig3}f and Online Resource Fig.~\ref{Fig4}f).
 During April-September, an additional (and often nocturnal) easterly component can also be recorded, induced by the combination of katabatic winds forming in the mountains, and infrequent `berg' winds, which are responsible for the high wind velocities observed \citep{lancaster1984}.
 The frequency of these easterly components decreases from inland to the coast.
-As a result, bidirectional wind regimes within the Namib Sand Sea and at the west Etosha site (Fig.~\ref{Fig1}b,d,e) and a unidirectional wind regime on the coast at the outlet of the Huab River (Fig.~\ref{Fig1}c) are observed.
+As a result, bidirectional wind regimes within the Namib Sand Sea and at the west Etosha site (Fig.~\ref{Fig2}a,c,d) and a unidirectional wind regime on the coast at the outlet of the Huab River (Fig.~\ref{Fig2}b) are observed.
 
-In the case of the Etosha West and Huab stations, the time series of wind speed and direction from the regional predictions quantitatively match those corresponding to the local measurements (Fig.~\ref{Fig2}a--d) and Online Resource Figs.~\ref{Fig6_supp}a--d,~\ref{Fig7_supp},~\ref{Fig9_supp}).
-For the North Sand Sea and South Sand Sea stations within the giant dune field, we observe that this agreement is also good, but limited to the October-March time period (Fig.~\ref{Fig2}e--h and and Online Resource Fig.~\ref{Fig6_supp}e--h).
-However, the field-measured wind roses exhibit additional wind components aligned with the giant dune orientation, as evidenced on the satellite images (Fig.~\ref{Fig1}c,d).
+In the case of the Etosha West and Huab stations, the time series of wind speed and direction from the regional predictions quantitatively match those corresponding to the local measurements (Figs.~\ref{Fig3}, ~\ref{Fig4} and Online Resource Fig.~\ref{Fig5_supp}).
+For the North Sand Sea and South Sand Sea stations within the giant dune field, we observe that this agreement is also good, but limited to the October-March time period (Fig.~\ref{Fig4}a, b, e, f).
+However, the field-measured wind roses exhibit additional wind components aligned with the giant dune orientation, as evidenced on the satellite images (Fig.~\ref{Fig2}c,d).
 
-More precisely, during the April-September period, the local and regional winds in the interdune match during daytime only, i.e when the southerly/southwesterly sea breeze dominates (Figs.~\ref{Fig2}e,f and \ref{Fig3}, Online Resource Fig.~\ref{Fig8_supp}).
+More precisely, during the April-September period, the local and regional winds in the interdune match during daytime only, i.e when the southerly/southwesterly sea breeze dominates (Figs.~\ref{Fig5}c,d,g,h and \ref{Fig6}).
 In the late afternoon and during the night, when the easterly `berg' and katabatic winds blow, measurements and predictions differ.
-In this case, the angular wind distribution of the local measurements exhibits two additional modes corresponding to reversing winds aligned with the giant dune orientation (purple frame in Fig.~\ref{Fig3}, Online Resource Figs.~\ref{Fig8_supp} and \ref{Fig10_supp}).
-This deviation is also associated with a general attenuation of the wind strength (Online Resource Fig.~\ref{Fig11_supp}).
+In this case, the angular wind distribution of the local measurements exhibits two additional modes corresponding to reversing winds aligned with the giant dune orientation (purple frame in Fig.~\ref{Fig6}, Online Resource Fig.~\ref{Fig6_supp}).
+This deviation is also associated with a general attenuation of the wind strength (Online Resource Fig.~\ref{Fig7_supp}).
 Remarkably, all these figures show that these wind reorientation and attenuation processes occur only at low velocities of the regional wind, typically for $u_{*, \, \textup{ERA}} \lesssim 0.2~\textrm{m}~\textrm{s}^{-1}$.
 For shear velocities larger than $u_{*, \, \textup{ERA}} \simeq 0.3~\textrm{m}~\textrm{s}^{-1}$, the wind reorientation is not apparent.
-Finally, for intermediate shear velocities, both situations of wind flow reoriented along the dune crest and not reoriented can be successively observed (Online Resource Fig.~\ref{Fig10_supp}).
+Finally, for intermediate shear velocities, both situations of wind flow reoriented along the dune crest and not reoriented can be successively observed (Online Resource Fig.~\ref{Fig6_supp}).
 Importantly, these values are not precise thresholds, but indicative of a crossover between regimes, whose physical interpretation is discussed in the next section.
@@
 \centering
-\includegraphics[scale=1]{Figures/Figure3.pdf}
-\caption{Distributions of wind direction at the North Sand Sea Station for the ERA5-Land climate reanalysis (orange) and the local measurements (blue).
+\includegraphics[scale=1]{Figures/Figure5.pdf}
+\caption{Temporal comparison between the wind data coming from the ERA5-Land climate reanalysis (orange lines) and from the local measurements (blue lines).
+Coloured swathes indicate day (between 1000 UTC and 2200 UTC) and night (before 1000 UTC or after 2200 UTC) \textbf{a--b}: North Sand Sea station in summer.
+\textbf{b--c}: North Sand Sea station in winter.
+\textbf{d--e}: South Sand Sea in summer.
+\textbf{f--g}: South Sand Sea station in winter.
+Time series of the two other stations are shown in Fig.~\ref{Fig3}.}
+\label{Fig5}
+\end{figure}
+%---------------------------------------------
+
+%---------------------------------------------
+\begin{figure}
+\centering
+\includegraphics[scale=1]{Figures/Figure6.pdf}
+\caption{Distributions of wind direction at the North Sand Sea (\textbf{a}) and South Sand Sea (\textbf{b}) stations for the ERA5-Land climate reanalysis (orange) and the local measurements (blue).
 In each subplot, both distributions are plotted from the same time steps, selected with constraints on the wind direction (columns) and/or wind velocity (rows) of the ERA5-Land dataset.
@@
 The numbers at the top right give the percentage of time steps selected in each sub-range, as well as the percentage corresponding to the daytime (between 1000 UTC and 2200 UTC).
-The purple frame highlights the regime (low wind velocities, nocturnal easterly wind) in which the data from both datasets differ.
-A similar figure can be obtained for the North Sand Sea station (Online Resource Fig.~\ref{Fig8_supp}).}
-\label{Fig3}
+Contrary to the Huab and Etosha West stations (Fig.~\ref{Fig4}), histograms does not match low wind velocities.
+More specifically, the purple frame highlights the regime (low wind velocities, nocturnal easterly wind) in which the data from both datasets differ.}
+\label{Fig6}
 \end{figure}
@@
 The wind deflection induced by linear dunes has previously been related to the incident angle between wind direction and crest orientation, with a maximum deflection evident for incident angles between $30^{\circ}$ and $70^{\circ}$ \citep{Walker2009, Hesp2015}.
-In the data analysed here, the most deflected wind at both the North and South Sand Sea stations is seen to be where the incident angle is perpendicular to the giant dunes (Figs.~\ref{Fig1} and \ref{Fig3}, Online Resource Fig.~\ref{Fig8_supp}).
+In the data analysed here, the most deflected wind at both the North and South Sand Sea stations is seen to be where the incident angle is perpendicular to the giant dunes (Figs.~\ref{Fig2} and \ref{Fig6}).
 It therefore appears that in our case, the incident wind angle is not the dominant control on maximum wind deflection.
-Further, and as shown in Fig.~\ref{Fig3}, winds of high and low velocities show contrasting behaviour in characteristics of deflection.
+Further, and as shown in Fig.~\ref{Fig6}, winds of high and low velocities show contrasting behaviour in characteristics of deflection.
 This suggests a change in hydrodynamical regime between the winds.
@@
 \centering
-\includegraphics[scale=1]{Figures/Figure4.pdf}
+\includegraphics[scale=1]{Figures/Figure7.pdf}
 \caption{\textbf{a--b}: Vertical profiles of the virtual potential temperature at 2 different time steps (day - 03/11/2015 - 1200 UTC, night - 01/13/2013 - 0900 UTC) at the North Sand Sea station.
@@
 The blue arrow indicates the undisturbed wind direction.}
-\label{Fig4}
+\label{Fig7}
 \end{figure}
@@
 This is also relevant for aeolian dunes as they evolve in the turbulent atmospheric boundary layer (ABL) capped by the stratified free atmosphere (FA) \citep{Andreotti2009}.
-Two main mechanisms, associated with dimensionless numbers must then be considered (Fig.~\ref{Fig4}).
+Two main mechanisms, associated with dimensionless numbers must then be considered (Fig.~\ref{Fig7}).
 First, topographic obstacles typically disturb the flow over a characteristic height similar to their length.
@@
 
-With this theoretical framework in mind, and in the context of the measured wind data in the North and South Sand Sea stations, the smallest wind disturbances are expected to occur during the day, when the ABL depth is the largest and comparable to the dune wavelength ($k H \gtrsim 1$), which corresponds to a weak confinement situation (Fig.~\ref{Fig4}c,d).
-In contrast, large wind disturbances are expected to occur during the night, when the confinement is mainly induced by a shallow ABL (Fig.~\ref{Fig4}e).
-However, this strong confinement can be somewhat reduced in the case of strong winds, corresponding to large values of the Froude number and a less `rigid' interface (Fig.~\ref{Fig4}f).
+With this theoretical framework in mind, and in the context of the measured wind data in the North and South Sand Sea stations, the smallest wind disturbances are expected to occur during the day, when the ABL depth is the largest and comparable to the dune wavelength ($k H \gtrsim 1$), which corresponds to a weak confinement situation (Fig.~\ref{Fig7}c,d).
+In contrast, large wind disturbances are expected to occur during the night, when the confinement is mainly induced by a shallow ABL (Fig.~\ref{Fig7}e).
+However, this strong confinement can be somewhat reduced in the case of strong winds, corresponding to large values of the Froude number and a less `rigid' interface (Fig.~\ref{Fig7}f).
 This is in qualitative agreement with the transition from deflected to non-deflected winds related to low and high velocities observed in our data (Sec.~\ref{section_data_feedback}).
@@
 \centering
-\includegraphics[scale=1]{Figures/Figure5.pdf}
+\includegraphics[scale=1]{Figures/Figure8.pdf}
 \caption{Regime diagrams of the wind deviation $\delta_{\theta}$ (\textbf{a}) and relative attenuation/amplification $\delta_{u}$ (\textbf{b}) in the space $(\mathcal{F}, \, kH)$, containing the data from both the North Sand Sea and South Sand Sea stations.
 Green dashed lines empirically delimit the different regimes.
-The point density in each bin of the diagrams is shown in Online Resource Fig.~\ref{Fig14_supp} -- 95\% of the data occur in the range $-1 < \delta u < 1$.
-Similar regime diagrams in the spaces $(\mathcal{F}_{\textup{I}}, \, kH)$ and $(\mathcal{F}_{\textup{I}}, \, \mathcal{F})$ are shown in Online Resource Fig.~\ref{Fig15_supp}.}
-\label{Fig5}
+The point density in each bin of the diagrams is shown in Online Resource Fig.~\ref{Fig10_supp} -- 95\% of the data occur in the range $-1 < \delta u < 1$.
+Similar regime diagrams in the spaces $(\mathcal{F}_{\textup{I}}, \, kH)$ and $(\mathcal{F}_{\textup{I}}, \, \mathcal{F})$ are shown in Online Resource Fig.~\ref{Fig11_supp}.}
+\label{Fig8}
 \end{figure}
@@
 %
-These two quantities are represented as maps in the plane $(\mathcal{F}, \, kH)$ (Fig.~\ref{Fig5}a,b), and one can clearly identify different regions in these graphs.
-Small wind disturbances (small $\delta_{\theta}$ and $\delta_{\textup{u}}$) are located in the top-right part of the diagrams, corresponding to a regime with low-interaction as well as low-confinement ($k H$ and $\mathcal{F}$ large enough, Fig.~\ref{Fig4}d).
+These two quantities are represented as maps in the plane $(\mathcal{F}, \, kH)$ (Fig.~\ref{Fig8}a,b), and one can clearly identify different regions in these graphs.
+Small wind disturbances (small $\delta_{\theta}$ and $\delta_{\textup{u}}$) are located in the top-right part of the diagrams, corresponding to a regime with low-interaction as well as low-confinement ($k H$ and $\mathcal{F}$ large enough, Fig.~\ref{Fig7}d).
 Lower values of $k H$ (stronger interaction) or of Froude number (stronger confinement) both lead to an increase in wind disturbances, both in terms of orientation and velocity.
@@
 This is probably due to enhanced non-linear effects linked to flow modulation by the obstacle when confinement is strong.
-The Froude number also controls a transition from damped to amplified wind velocities in the interdune, with a crossover around $\mathcal{F} \simeq 0.4$ (Fig.~\ref{Fig5}b).
+The Froude number also controls a transition from damped to amplified wind velocities in the interdune, with a crossover around $\mathcal{F} \simeq 0.4$ (Fig.~\ref{Fig8}b).
 Such an amplification is rather unexpected.
-Checking the occurrence of the corresponding data, it appears that these amplifications are associated with the southerly sea breeze, and occur dominantly during the October-March period, when the other easterly wind is not present (Online Resource Fig.~\ref{Fig16_supp}a--b).
-Furthermore, they occur less frequently during the afternoon, and more frequently at the end of the day (Online Resource Fig.~\ref{Fig16_supp}c).
+Checking the occurrence of the corresponding data, it appears that these amplifications are associated with the southerly sea breeze, and occur dominantly during the October-March period, when the other easterly wind is not present (Online Resource Fig.~\ref{Fig12_supp}a--b).
+Furthermore, they occur less frequently during the afternoon, and more frequently at the end of the day (Online Resource Fig.~\ref{Fig12_supp}c).
 This effect may be linked to a change in the flow behaviour in the lee side of the obstacles but further measurements are needed in order to assess the different possibilities \citep{baines1995, Vosper2004}.
@@
 \centering
-\includegraphics[scale=1]{Figures/Figure6.pdf}
+\includegraphics[scale=1]{Figures/Figure9.pdf}
 \caption{Implications for smaller scale patterns in (\textbf{a}) the South Sand Sea and (\textbf{b}) North Sand Sea.
@@
 See Appendix 2 for additional details.}
-\label{Fig6}
+\label{Fig9}
 \end{figure}
@@
 The feedback of the giant dunes on the wind flow has important implications for smaller scales bedforms.
-As illustrated in Fig.~\ref{Fig6}, small linear dunes ($\sim50$~m wide) are often present in the 1--2~km interdune between giant linear dunes in the Namib Sand Sea \citep{Livingstone2010}.
+As illustrated in Fig.~\ref{Fig9}, small linear dunes ($\sim50$~m wide) are often present in the 1--2~km interdune between giant linear dunes in the Namib Sand Sea \citep{Livingstone2010}.
 These smaller dunes do not exhibit the same orientation as the large ones, and are sometimes named `crossing dunes'.
@@
 Here, however, our arguments enable the development of differing orientations for the small and giant linear dunes whilst also imposing the same dune growth mechanism (elongating mode).
-Figure~\ref{Fig6} shows how the orientations for the small and giant dunes can be derived from the locally measured and regionally predicted winds respectively (red arrows in Fig.~\ref{Fig6}).
+Figure~\ref{Fig9} shows how the orientations for the small and giant dunes can be derived from the locally measured and regionally predicted winds respectively (red arrows in Fig.~\ref{Fig9}).
 These predictions require a specification for the threshold of eolian sand transport.
@@
 when the aspect ration of these ridges is small ($k\xi_0 \ll 1$).
-These models are able to qualitatively reproduce the observed wind deflection (Appendix 1, Online Resource Figs.~\ref{Fig15_supp} and \ref{Fig17_supp}), and thus provide the physical support for the interpretation we propose here based on hydrodynamic regimes.
+These models are able to qualitatively reproduce the observed wind deflection (Appendix 1, Online Resource Figs.~\ref{Fig11_supp} and \ref{Fig13_supp}), and thus provide the physical support for the interpretation we propose here based on hydrodynamic regimes.
 However, these models cannot quantitatively predict the magnitude of these observations, probably due to the presence of expected non-linearities in high confinement situations linked to strong flow modulations.
-Besides, these linear calculations only predict wind attenuation in the interdune, in contrast with the observed enhanced velocities associated with particular evening winds from the South during the period October-March (Online Resource Fig.~\ref{Fig16_supp}).
+Besides, these linear calculations only predict wind attenuation in the interdune, in contrast with the observed enhanced velocities associated with particular evening winds from the South during the period October-March (Online Resource Fig.~\ref{Fig12_supp}).
 Some other models predict different spatial flow structures in response to a modulated topography, such as lee waves and rotors \citep{baines1995, Vosper2004}.
@@
 Note that it contains modified Copernicus Climate Change Service Information (2021).
-Neither the European Commission nor ECMWF is responsible for any use that may be made of the Copernicus Information or Data it
-contains. Fully documented codes used to analyse this study are available at https://github.com/Cgadal/GiantDunes (will be made public upon acceptance of this manuscript for publication).
+Neither the European Commission nor ECMWF is responsible for any use that may be made of the Copernicus Information or Data it contains.
+Fully documented codes used to analyse this study are available at https://github.com/Cgadal/GiantDunes (will be made public upon acceptance of this manuscript for publication).
 
@@
 %
-which is also a good approximation for the giant dunes observed in the North Sand Sea and South Sand Sea Station (Fig.~\ref{Fig1} and Online Resource Fig.~\ref{Fig5_supp}).
+which is also a good approximation for the giant dunes observed in the North Sand Sea and South Sand Sea Station (Fig.~\ref{Fig2} and Online Resource Fig.~\ref{Fig4_supp}).
 Here, $x$ and $y$ are the streamwise and spanwise coordinates, $k=2\pi/\lambda$ the wavenumber of the sinusoidal perturbation, $\alpha$ its crest orientation with respect to the $x$-direction (anticlockwise) and $\xi_{0}$ its amplitude.
@@
 
-With capping layer height and Froude numbers computed from the ERA5-Land time series, the corresponding $\mathcal{A}_{0}$ and $\mathcal{B}_{0}$ can be deduced, as displayed in Online Resource Fig.~\ref{Fig17_supp}.
-Interestingly, it shows similar regimes as in the diagrams of Fig.~\ref{Fig5} and Online Resource Fig.~\ref{Fig15_supp}a,b, supporting the underlying physics.
+With capping layer height and Froude numbers computed from the ERA5-Land time series, the corresponding $\mathcal{A}_{0}$ and $\mathcal{B}_{0}$ can be deduced, as displayed in Online Resource Fig.~\ref{Fig13_supp}.
+Interestingly, it shows similar regimes as in the diagrams of Fig.~\ref{Fig8} and Online Resource Fig.~\ref{Fig11_supp}a,b, supporting the underlying physics.
 However, the agreement is qualitative only.
@@
 In accordance with the theoretical expectations, these coefficients present values on the order of unity ($\mathcal{A}_{0} \simeq 3$ and $\mathcal{B}_{0} \simeq 1$) in unconfined situations \citep{Claudin2013, Lu2021}.
-In contrast and as illustrated in Online Resource Fig.~\ref{Fig17_supp}a,b, larger values are predicted in case of strong confinement, which does not allow us to proceed to further quantitative comparison with the data.
+In contrast and as illustrated in Online Resource Fig.~\ref{Fig13_supp}a,b, larger values are predicted in case of strong confinement, which does not allow us to proceed to further quantitative comparison with the data.
 
-Finally, the linear model is also able to reproduce the enhancement of the flow deflection over the sinusoidal ridges when $\sqrt{\mathcal{A}_{0}^{2} + \mathcal{B}_{0}^{2}}$ is increased (Online Resource Fig.~\ref{Fig17_supp}).
+Finally, the linear model is also able to reproduce the enhancement of the flow deflection over the sinusoidal ridges when $\sqrt{\mathcal{A}_{0}^{2} + \mathcal{B}_{0}^{2}}$ is increased (Online Resource Fig.~\ref{Fig13_supp}).
 Here, using $k\xi_0 \simeq 0.1$ to be representative of the amplitude of the giant dunes at the North Sand Sea station, the coefficient modulus is bounded to $10$.
@@
 The fit of Eq.~\ref{transport_law_quadratic} on these same data similarly gives $\Omega \simeq 8$ and $\Theta_{\textup{th}} = 0.005$.
-The sand flux angular distributions and the dune orientations in Fig.~\ref{Fig6} are calculated using this law \eqref{transport_law_quartic}.
+The sand flux angular distributions and the dune orientations in Fig.~\ref{Fig9} are calculated using this law \eqref{transport_law_quartic}.
 We have checked that using the ordinary linear relationship \eqref{transport_law_quadratic} instead does not change the predicted dune orientations by more than a few degrees.
@@
 
-The resulting computed dune orientations, blue and red arrows in Fig.~\ref{Fig6}, then depend on a certain number of parameters (grain properties, flux-up ratio, etc.), for which we take typical values for aeolian sandy deserts.
+The resulting computed dune orientations, blue and red arrows in Fig.~\ref{Fig9}, then depend on a certain number of parameters (grain properties, flux-up ratio, etc.), for which we take typical values for aeolian sandy deserts.
 Due to the lack of measurements in the studied places, some uncertainties can be expected.
-We therefore run a sensitivity test by calculating the dune orientations for grain diameters ranging from $100~\mu\textup{m}$ to $400~\mu\textup{m}$ and for a speed-up ratio between $0.1$ and $10$ (wedges in Fig.~\ref{Fig6}).
+We therefore run a sensitivity test by calculating the dune orientations for grain diameters ranging from $100~\mu\textup{m}$ to $400~\mu\textup{m}$ and for a speed-up ratio between $0.1$ and $10$ (wedges in Fig.~\ref{Fig9}).
 
@@
 
-As shown in Online Resource Fig.~\ref{Fig4_supp}, the minimum values of $\delta$ in the space ($z_{0}^{\textup{ERA5Land}}$, $z_{0}^{\textup{local}}$) form a line.
+As shown in Online Resource Fig.~\ref{Fig3_supp}, the minimum values of $\delta$ in the space ($z_{0}^{\textup{ERA5Land}}$, $z_{0}^{\textup{local}}$) form a line.
 We thus set the roughness in the ERA5-Land analysis to the typical value $z_0=10^{-3}$~m, and deduce the corresponding ones for the local stations.
@@
 In arid areas, the vertical structure of the atmosphere can be approximated by a well mixed convective boundary layer of height $H$, topped by the stratified free atmosphere \citep{Stull1988, Shao2008}.
-In this context, one usually introduces the virtual potential temperature $T_{\textup{vp}}$, which is a constant $T_{0}$ inside the boundary layer, and increases linearly in the FA (Online Resource Fig.~\ref{Fig12_supp}a):
+In this context, one usually introduces the virtual potential temperature $T_{\textup{vp}}$, which is a constant $T_{0}$ inside the boundary layer, and increases linearly in the FA (Online Resource Fig.~\ref{Fig8_supp}a):
 %
@@
 where $\Delta T_{\textup{vp}}$ is the temperature discontinuity at the capping layer and $N = \sqrt{g \partial_z T_{\textup{vp}} / T_{0}}$ is the Brunt-V\"ais\"al\"a frequency, characteristic of the stratification.
-Note that, under the usual Boussinesq approximation, temperature and air density variations are simply related by $\delta T_{\textup{vp}}/T_{0} \simeq - \delta\rho/\rho_{0}$ (see Online Resource of \citet{Andreotti2009}), so that $N$ can equivalently be defined from the density gradient as next to Eq.~\ref{FroudeNumber}.
+Note that, under the usual Boussinesq approximation, temperature and air density variations are simply related by $\delta T_{\textup{vp}}/T_{0} \simeq - \delta\rho/\rho_{0}$ (see Online Resource of \citet{Andreotti2009}), so that $N$ can equivalently be defined from the density gradient as next to \eqref{FroudeNumber}.
 
@@
 
-Examples of vertical profiles of the virtual potential temperature deduced from ERA5 are shown in Online Resource Fig.~\ref{Fig12_supp}a.
+Examples of vertical profiles of the virtual potential temperature deduced from ERA5 are shown in Online Resource Fig.~\ref{Fig8_supp}a.
 For each of them, an average temperature is computed below the ABL depth ($z<H$), and a linear function is fitted above, allowing us to extract the temperature jump $\Delta T_{\textup{vp}}$.
-Importantly, some profiles display a vertical structure that cannot be approximated by the simple form \eqref{TvpVerticalProfile} used here (Online Resource Fig.~\ref{Fig12_supp}b).
+Importantly, some profiles display a vertical structure that cannot be approximated by the simple form \eqref{TvpVerticalProfile} used here (Online Resource Fig.~\ref{Fig8_supp}b).
 In practice, we removed from the analysis all of those leading to the unphysical case $\Delta T_{\textup{vp}}<0$.
 We have noticed that these `ill-processed' profiles dominantly occur in winter and are evenly spread across the hours of the day.
-Importantly, they represent $\simeq12$\% of the data only (Online Resource Fig.~\ref{Fig12_supp}c,d), and we are thus confident that this data treatment does not affect our conclusions.
+Importantly, they represent $\simeq12$\% of the data only (Online Resource Fig.~\ref{Fig8_supp}c,d), and we are thus confident that this data treatment does not affect our conclusions.
 
@@
 \includegraphics[scale=1]{Figures/Figure1_supp.pdf}
-\caption{Photographs of the meteorological stations. \textbf{a}: South Sand Sea station. \textbf{b--c}: North Sand Sea station. \textbf{d}: Huab station. \textbf{e} Etosha West station.}
+\caption{Gant chart representing the valid time steps for the two data sets, for all stations.}
 \label{Fig1_supp}
@@
 \includegraphics[scale=1]{Figures/Figure2_supp.pdf}
-\caption{Gant chart representing the valid time steps for the two data sets, for all stations.}
+\caption{Comparison between raw local wind measurements, and hourly-averaged data for South Sand Sea station. \textbf{a}: wind direction. \textbf{b}: wind velocity at height $2.6$~m.}
 \label{Fig2_supp}
@@
 \includegraphics[scale=1]{Figures/Figure3_supp.pdf}
-\caption{Comparison between raw local wind measurements, and hourly-averaged data for South Sand Sea station. \textbf{a}: wind direction. \textbf{b}: wind velocity at height $2.6$~m.}
+\caption{Calibration of hydrodynamic roughness.
+The parameter $\delta$ \eqref{metric_roughness} quantifying the difference between local and predicted winds is shown in colorscale as a function of the hydrodynamic roughnesses chosen for the ERA5-Land and for local winds, for the (\textbf{a}) Etosha West, (\textbf{b}) North Sand Sea, (\textbf{c}) Huab and (\textbf{d}) South Sand Sea stations.
+The red dashed and plain lines shows the minima of $\delta$ and the identity line, respectively.
+The orange lines and dots highlight the chosen hydrodynamic roughnesses for the local winds deduced from setting $z_{0}^{\textup{ERA5Land}} = 1$~mm.}
 \label{Fig3_supp}
@@
   \includegraphics[scale=1]{Figures/Figure4_supp.pdf}
-\caption{Calibration of hydrodynamic roughness.
-The parameter $\delta$ (Eq.~\ref{metric_roughness}) quantifying the difference between local and predicted winds is shown in colorscale as a function of the hydrodynamic roughnesses chosen for the ERA5-Land and for local winds, for the (\textbf{a}) Etosha West, (\textbf{b}) North Sand Sea, (\textbf{c}) Huab and (\textbf{d}) South Sand Sea stations.
-The red dashed and plain lines shows the minima of $\delta$ and the identity line, respectively.
-The orange lines and dots highlight the chosen hydrodynamic roughnesses for the local winds deduced from setting $z_{0}^{\textup{ERA5Land}} = 1$~mm.}
+  \caption{Analysis of the DEMs of the North Sand Sea (left column -- panels \textbf{a}, \textbf{b}, \textbf{c}) and South Sand Sea (right column -- panels \textbf{d}, \textbf{e}, \textbf{f}) stations.
+\textbf{a--d}: Bed elevation detrended by a fitted second order polynomial base-line.
+\textbf{b--e}: Autocorrelation matrix shown in colorscale.
+The black line shows the detected dune orientation, and the red line represents the autocorrelation profile along which the dune wavelength is calculated, displayed in \textbf{c--f}.
+The blue lines and dots show the first peak of the autocorrelation profile, whose abscissa gives the characteristic wavelength of the dune pattern.}
   \label{Fig4_supp}
@@
 \includegraphics[scale=1]{Figures/Figure5_supp.pdf}
-  \caption{Analysis of the DEMs of the North Sand Sea (left column -- panels \textbf{a}, \textbf{b}, \textbf{c}) and South Sand Sea (right column -- panels \textbf{d}, \textbf{e}, \textbf{f}) stations.
-\textbf{a--d}: Bed elevation detrended by a fitted second order polynomial base-line.
-\textbf{b--e}: Autocorrelation matrix shown in colorscale.
-The black line shows the detected dune orientation, and the red line represents the autocorrelation profile along which the dune wavelength is calculated, displayed in \textbf{c--f}.
-The blue lines and dots show the first peak of the autocorrelation profile, whose abscissa gives the characteristic wavelength of the dune pattern.}
+\caption{Statistical comparison of the wind orientation (\textbf{a}) and velocity (\textbf{b}) between the ERA5-Land dataset and the local measurements for the Huab and Etosha West stations.
+Data point clustering around identity lines (dashed and black) provide evidence for agreement of the two sets.}
 \label{Fig5_supp}
@@
 %---------------------------------------------
-\begin{figure}
+\begin{figure}[p]
 \centering
 \includegraphics[scale=1]{Figures/Figure6_supp.pdf}
-\caption{Temporal comparison between the wind data coming from the ERA5-Land climate reanalysis (orange lines) and from the local measurements (blue lines).
-Coloured swathes indicate day (between 1000 UTC and 2200 UTC) and night (before 1000 UTC or after 2200 UTC) \textbf{a--b}: Huab station in summer.
-\textbf{b--c}: Huab station in winter.
-\textbf{d--e}: South Sand Sea station in summer.
-\textbf{f--g}: South Sand Sea station in winter.}
+\caption{Statistical comparison of the wind orientation between the ERA5-Land dataset and the local measurements for the South Sand Sea and North Sand Sea stations, for different velocity ranges.
+\textbf{a}: $u_{*, \textup{\textup{ERA}}} < 0.1~\textup{m}~\textup{s}^{-1}$.
+\textbf{b}: $0.1 < u_{*, \textup{\textup{ERA}}} \leq 0.25~\textup{m}~\textup{s}^{-1}$.
+\textbf{c}: $u_{*, \textup{\textup{ERA}}} \geq 0.25~\textup{m}~\textup{s}^{-1}$.
+The measured dune orientations are subtracted to the wind orientation, which allows us to plot both stations on the same graph.
+Black dashed lines indicate locally measured orientations aligned with the dune crests (here $0^\circ$, $180^\circ$ and $360^\circ$ -- panels \textbf{a}, \textbf{b}), as well as the identity lines (panels \textbf{b}, \textbf{c}).}
 \label{Fig6_supp}
@@
 \includegraphics[scale=1]{Figures/Figure7_supp.pdf}
-\caption{Statistical comparison of the wind orientation (\textbf{a}) and velocity (\textbf{b}) between the ERA5-Land dataset and the local measurements for the Huab and Etosha West stations.
-Data point clustering around identity lines (dashed and black) provide evidence for agreement of the two sets.}
+\caption{Statistical comparison of the wind velocity between the ERA5-Land dataset and the local measurements for the South Sand Sea and North Sand Sea stations.
+\textbf{a}: Nocturnal summer easterly wind.
+\textbf{b}: Diurnal southerly wind.
+Black dashed lines are identity lines.
+The angle ranges used to select diurnal and nocturnal summer winds are the same as those in Figs.~\ref{Fig4} and Figs.~\ref{Fig6} of the main article.}
 \label{Fig7_supp}
@@
 \includegraphics[scale=1]{Figures/Figure8_supp.pdf}
-\caption{Distributions of wind direction at the South Sand Sea Station for the ERA5-Land climate reanalysis (orange) and the local measurements (blue) -- equivalent of Fig.~\ref{Fig3}.
-In each subplot, both distributions are plotted from the same time steps, selected with constraints on the wind direction (columns) and/or wind velocity (rows) in the ERA5-Land dataset.
-The gray vertical dashed lines indicate the dune orientation.
-The numbers at the top centre give the percentage of time steps selected in each sub-range, as well as the percentage of them corresponding to the day (between 1000 UTC and 2200 UTC).
-The purple frame highlights the regime (small wind velocities, nocturnal summer wind) during which the wind data from both datasets differ.}
+\caption{\textbf{a}: Vertical profiles of the virtual potential temperature at three different times (blue: 29/11/2012 - 1100 UTC, orange: 21/03/2017 - 1200 UTC, green: 21/03/2017 - 2000 UTC) at the South Sand Sea station.
+Dots: data from the ERA5 reanalysis.
+Dashed lines: boundary layer height given by the ERA5 reanalysis.
+Plain lines: vertical (ABL) and linear (FA) fits to estimate the quantities displayed in Online Resource Fig.~\ref{Fig9_supp}.
+\textbf{b}: Examples of ill-processed vertical profiles at three different times (blue: 2/12/2013 - 2300 UTC, orange: 20/03/2017 - 0000 UTC, green: 14/07/2017 - 1400 UTC) at the South Sand Sea station.
+\textbf{c}: Hourly distribution of ill-processed vertical profiles.
+\textbf{d}: Monthly distribution of ill-processed vertical profiles.}
 \label{Fig8_supp}
@@
 \includegraphics[scale=1]{Figures/Figure9_supp.pdf}
-\caption{Distributions of wind direction at the Huab and Etosha West stations for the ERA5-Land climate reanalysis (orange) and the local measurements (blue).
-In each subplot, both distributions are plotted from the same time steps, selected with constraints on the wind wind velocity (rows) in the ERA5-Land dataset.
-The numbers at the top centre give the percentage of time steps selected in each sub-range, as well as the percentage of them corresponding to the day (between 1000 UTC and 2200 UTC).
-Compared to the North and South Namib stations (Fig.~\ref{Fig3} and Fig.~\ref{Fig8_supp}), histograms match for high and low velocities.}
+\caption{Distributions of the meteorological parameters resulting from the processing of the ERA5-Land data for the South Sand Sea (blue) and the North Sand Sea (orange) stations.}
 \label{Fig9_supp}
@@
 \includegraphics[scale=1]{Figures/Figure10_supp.pdf}
-\caption{Statistical comparison of the wind orientation between the ERA5-Land dataset and the local measurements for the South Sand Sea and North Sand Sea stations, for different velocity ranges.
-\textbf{a}: $u_{*, \textup{\textup{ERA}}} < 0.1~\textup{m}~\textup{s}^{-1}$.
-\textbf{b}: $0.1 < u_{*, \textup{\textup{ERA}}} \leq 0.25~\textup{m}~\textup{s}^{-1}$.
-\textbf{c}: $u_{*, \textup{\textup{ERA}}} \geq 0.25~\textup{m}~\textup{s}^{-1}$.
-The measured dune orientations are subtracted to the wind orientation, which allows us to plot both stations on the same graph.
-Black dashed lines indicate locally measured orientations aligned with the dune crests (here $0^\circ$, $180^\circ$ and $360^\circ$ -- panels \textbf{a}, \textbf{b}), as well as the identity lines (panels \textbf{b}, \textbf{c}).}
+\caption{Non-dimensional parameters distributions. For the marginal distributions, the orange correspond to the South Sand Sea station, and the blue to the North Sand Sea station.}
 \label{Fig10_supp}
@@
 \includegraphics[scale=1]{Figures/Figure11_supp.pdf}
-\caption{Statistical comparison of the wind velocity between the ERA5-Land dataset and the local measurements for the South Sand Sea and North Sand Sea stations.
-\textbf{a}: Nocturnal summer easterly wind.
-\textbf{b}: Diurnal southerly wind.
-Black dashed lines are identity lines.
-The angle ranges used to select diurnal and nocturnal summer winds are the same as those in Fig.~\ref{Fig3} and Online Resource Fig.~\ref{Fig8_supp}.}
+\caption{Regime diagrams of the wind deviation $\delta_{\theta}$ and relative attenuation/amplification $\delta_{u}$ in the spaces $(\mathcal{F}_{\textup{I}}, \, kH)$ and $(\mathcal{F}_{\textup{I}}, \, \mathcal{F})$, containing the data from both the North Sand Sea and South Sand Sea stations.
+Green dashed lines empirically delimit the different regimes.
+The point density in each bin of the diagrams is shown in Online Resource Fig.~\ref{Fig10_supp} -- 95\% of the data occur in the range $-1 < \delta u < 1$.
+The similar regime diagrams in the space $(\mathcal{F}, \, kH)$ are shown in Fig.~\ref{Fig8}.}
 \label{Fig11_supp}
@@
 \includegraphics[scale=1]{Figures/Figure12_supp.pdf}
-\caption{\textbf{a}: Vertical profiles of the virtual potential temperature at three different times (blue: 29/11/2012 - 1100 UTC, orange: 21/03/2017 - 1200 UTC, green: 21/03/2017 - 2000 UTC) at the South Sand Sea station.
-Dots: data from the ERA5 reanalysis.
-Dashed lines: boundary layer height given by the ERA5 reanalysis.
-Plain lines: vertical (ABL) and linear (FA) fits to estimate the quantities displayed in Online Resource Fig.~\ref{Fig13_supp}.
-\textbf{b}: Examples of ill-processed vertical profiles at three different times (blue: 2/12/2013 - 2300 UTC, orange: 20/03/2017 - 0000 UTC, green: 14/07/2017 - 1400 UTC) at the South Sand Sea station.
-\textbf{c}: Hourly distribution of ill-processed vertical profiles.
-\textbf{d}: Monthly distribution of ill-processed vertical profiles.}
+\caption{Normalized distributions of amplified velocities for the North sand Sea (blue: $\delta_u < 0$, orange: $\delta_u < -0.5$).
+\textbf{a}: Angular distributions.
+\textbf{b}: Monthly distributions.
+\textbf{c}: Hourly distributions.}
 \label{Fig12_supp}
@@
 \includegraphics[scale=1]{Figures/Figure13_supp.pdf}
-\caption{Distributions of the meteorological parameters resulting from the processing of the ERA5-Land data for the South Sand Sea (blue) and the North Sand Sea (orange) stations.}
+\caption{Computation of the flow disturbance with the linear model of \citet{Andreotti2009}.
+\textbf{a}--\textbf{b}: Magnitude of the hydrodynamic coefficients $\mathcal{A}_0$ and $\mathcal{B}_0$, calculated from the time series of the non-dimensional numbers corresponding to the ERA5-Land wind data and ERA5 data on vertical pressure levels.
+\textbf{c} Shear velocity streamlines over sinusoidal ridges of amplitude $k\xi_0 = 0.1$ and for increasing values of $\sqrt{\mathcal{A}_{0}^{2} + \mathcal{B}_{0}^{2}}$.
+From the upper to the lower streamline, values of $\left(kH,\, \mathcal{F},\, \mathcal{F}_{\rm I},\, \mathcal{A}_{0},\, \mathcal{B}_{0},\, \sqrt{\mathcal{A}_{0}^{2} + \mathcal{B}_{0}^{2}}\right)$ are (1.9, 0.6, 1.5, 3.4, 1.0, 3.5), (1.5, 0.3, 0.4, 4.8, 1.4, 5.0), (0.1, 3.5, 1.0, 8.6, 0.1, 8.6), (0.5, 0.05, 0.04, 9.6, 2.5, 9.9).}
 \label{Fig13_supp}
@@
 \includegraphics[scale=1]{Figures/Figure14_supp.pdf}
-\caption{Non-dimensional parameters distributions. For the marginal distributions, the orange correspond to the South Sand Sea station, and the blue to the North Sand Sea station.}
+\caption{Time series of the relative velocity disturbance $\delta_{u}$ corresponding to Fig.~\ref{Fig5}, for different values of the hydroydnamic roughnesses.
+\textbf{a}: North Sand Sea – summer, \textbf{b}: North Sand Sea – winter, \textbf{d}: South Sand Sea – summer, \textbf{e}: South Sand Sea – winter.
+Note that $\delta_{\theta}$ is independant of the choice of $z_{0}^{\textup{Era5Land}}$ and $z_{0}^{\textup{local}}$.}
 \label{Fig14_supp}
@@
 
-%---------------------------------------------
-\begin{figure}[p]
-\centering
-\includegraphics[scale=1]{Figures/Figure15_supp.pdf}
-\caption{Regime diagrams of the wind deviation $\delta_{\theta}$ and relative attenuation/amplification $\delta_{u}$ in the spaces $(\mathcal{F}_{\textup{I}}, \, kH)$ and $(\mathcal{F}_{\textup{I}}, \, \mathcal{F})$, containing the data from both the North Sand Sea and South Sand Sea stations.
-Green dashed lines empirically delimit the different regimes.
-The point density in each bin of the diagrams is shown in Online Resource Fig.~\ref{Fig14_supp} -- 95\% of the data occur in the range $-1 < \delta u < 1$.
-The similar regime diagrams in the space $(\mathcal{F}, \, kH)$ are shown in Fig.~\ref{Fig5}.}
-\label{Fig15_supp}
-\end{figure}
-%---------------------------------------------
-
-%---------------------------------------------
-\begin{figure}[p]
-\centering
-\includegraphics[scale=1]{Figures/Figure16_supp.pdf}
-\caption{Normalized distributions of amplified velocities for the North sand Sea (blue: $\delta_u < 0$, orange: $\delta_u < -0.5$).
-\textbf{a}: Angular distributions.
-\textbf{b}: Monthly distributions.
-\textbf{c}: Hourly distributions.}
-\label{Fig16_supp}
-\end{figure}
-%---------------------------------------------
-
-%---------------------------------------------
-\begin{figure}[p]
-\centering
-\includegraphics[scale=1]{Figures/Figure17_supp.pdf}
-\caption{Computation of the flow disturbance with the linear model of \citet{Andreotti2009}.
-\textbf{a} and \textbf{b} Magnitude of the hydrodynamic coefficients $\mathcal{A}_0$ and $\mathcal{B}_0$, calculated from the values of the non-dimensional numbers corresponding to the ERA5-Land time series presented in Figs.~\ref{Fig4} and \ref{Fig5}.
-\textbf{c} Shear velocity streamlines over sinusoidal ridges of amplitude $k\xi_0 = 0.1$ and for increasing values of $\sqrt{\mathcal{A}_{0}^{2} + \mathcal{B}_{0}^{2}}$.
-From the upper to the lower streamline, values of $\left(kH,\, \mathcal{F},\, \mathcal{F}_{\rm I},\, \mathcal{A}_{0},\, \mathcal{B}_{0},\, \sqrt{\mathcal{A}_{0}^{2} + \mathcal{B}_{0}^{2}}\right)$ are (1.9, 0.6, 1.5, 3.4, 1.0, 3.5), (1.5, 0.3, 0.4, 4.8, 1.4, 5.0), (0.1, 3.5, 1.0, 8.6, 0.1, 8.6), (0.5, 0.05, 0.04, 9.6, 2.5, 9.9).}
-\label{Fig17_supp}
-\end{figure}
-%---------------------------------------------
-
 \end{document}

### 2022-05-22 Paper_figure/Figure02.py

@@ -1,30 +1,22 @@
 """
+
 ============
 Figure 2
 ============
 
 """
 
-import numpy as np
-import matplotlib.pyplot as plt
-import matplotlib.dates as mdates
-import calendar
-import locale
-from datetime import datetime, timedelta
-import sys
 import os
+import sys
+import numpy as np
+from PIL import Image
+import matplotlib.pyplot as plt
+from matplotlib.patches import Rectangle
+from matplotlib.colors import BoundaryNorm
+from mpl_toolkits.axes_grid1.anchored_artists import AnchoredSizeBar
 sys.path.append('../')
 import python_codes.theme as theme
-
-locale.setlocale(locale.LC_ALL, 'en_US.utf8')
-
-
-def tick_formatter(ax, fmt='%d'):
-    myFmt = mdates.DateFormatter(fmt)
-    ax.xaxis.set_major_formatter(myFmt)
-    ticklabels = ax.get_xticklabels()
-    ticklabels[0].set_ha('left')
-
+from python_codes.plot_functions import plot_wind_rose
 
 # Loading figure theme
 theme.load_style()
@@ -38,56 +30,75 @@ path_outputdata = '../static/data/processed_data'
 Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
 Stations = sorted(Data.keys())
 
-# Figure properties
-variables = ['U_star', 'Orientation']
-label_var = {'U_star': r'Velocity, $u_{*}~[\textup{m}~\textup{s}^{-1}]$', 'Orientation': r'Orientation, $\theta~[^\circ]$'}
-labels = [(r'\textbf{a}', r'\textbf{b}'), (r'\textbf{c}', r'\textbf{d}'),
-          (r'\textbf{e}', r'\textbf{f}'), (r'\textbf{g}', r'\textbf{h}')]
-row_labels = ['Etosha West -- summer', 'Etosha West -- winter', 'North Sand Sea -- summer',
-              'North Sand Sea -- winter']
-years = [2015, 2016, 2017, 2017]
-months = [11, 8, 12, 6]
-days = [(10, 14), (17, 20), (5, 8), (1, 4)]
-month_calendar = {index: month for index, month in enumerate(calendar.month_name) if month}
-
-
-stations_plot = ['Adamax_Station', 'Adamax_Station', 'Deep_Sea_Station', 'Deep_Sea_Station']
[…]

### 2022-05-22 Paper_figure/Figure03.py

@@ -7,11 +7,26 @@ Figure 3
 
 import numpy as np
 import matplotlib.pyplot as plt
+import matplotlib.dates as mdates
+import calendar
+import locale
+from datetime import datetime, timedelta
 import sys
 import os
 sys.path.append('../')
 import python_codes.theme as theme
-from python_codes.plot_functions import make_nice_histogram
+from python_codes.general import smallestSignedAngleBetween, find_mode_distribution
+
+
+locale.setlocale(locale.LC_ALL, 'en_US.utf8')
+
+
+def tick_formatter(ax, fmt='%d'):
+    myFmt = mdates.DateFormatter(fmt)
+    ax.xaxis.set_major_formatter(myFmt)
+    ticklabels = ax.get_xticklabels()
+    ticklabels[0].set_ha('left')
+
 
 # Loading figure theme
 theme.load_style()
@@ -26,58 +41,85 @@ Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=Tru
 Stations = sorted(Data.keys())
 
 # Figure properties
-station = 'Deep_Sea_Station'
-#
-theta_bins = [[0, 90], [150, 230]]
-velocity_bins = [[0.05, 0.2], [0.3, 10]]
-Data_pattern = np.load(os.path.join(path_outputdata, 'Data_DEM.npy'), allow_pickle=True).item()[station]
-icon = [r'\faSun', r'\faMoon']
+variables = ['U_star', 'Orientation']
+label_var = {'U_star': r'Velocity, $u_{*}~[\textup{m}~\textup{s}^{-1}]$', 'Orientation': r'Orientation, $\theta~[^\circ]$'}
+labels = [(r'\textbf{a}', r'\textbf{b}'), (r'\textbf{c}', r'\textbf{d}'),
+          (r'\textbf{e}', r'\textbf{f}'), (r'\textbf{g}', r'\textbf{h}')]
+row_labels = ['Huab -- summer', 'Huab -- winter',
+              'Etosha West -- summer', 'Etosha West -- winter']
+years = [2018, 2018, 2015, 2016]
+months = [12, 6, 11, 8]
+days = [(3, 6), (2, 5), (9, 12), (17, 20)]
+month_calendar = {index: month for index, month in enumerate(calendar.month_name) if month}
+bbox_props = dict(boxstyle='round', facecolor='wheat', alpha=0.7)
 
-color_ax = 'purple'
 
-# ################ Figure
[…]

### 2022-05-22 Paper_figure/Figure05.py

@@ -1,88 +1,119 @@
 """
-============
+============================
 Figure 5
-============
+============================
 
 """
 
 import numpy as np
 import matplotlib.pyplot as plt
-import matplotlib.colors as mpcolors
+import matplotlib.dates as mdates
+import calendar
+import locale
+from datetime import datetime, timedelta
 import sys
 import os
 sys.path.append('../')
 import python_codes.theme as theme
 from python_codes.general import smallestSignedAngleBetween, find_mode_distribution
-from python_codes.plot_functions import plot_regime_diagram
+
+locale.setlocale(locale.LC_ALL, 'en_US.utf8')
+
+
+def tick_formatter(ax, fmt='%d'):
+    myFmt = mdates.DateFormatter(fmt)
+    ax.xaxis.set_major_formatter(myFmt)
+    ticklabels = ax.get_xticklabels()
+    ticklabels[0].set_ha('left')
+
 
 # Loading figure theme
 theme.load_style()
 
 # paths
-path_imgs = '../static/images/'
 path_savefig = '../Paper/Figures'
-path_outputdata = '../static/data/processed_data'
+path_outputdata = '../static/data/processed_data/'
 
-# ##### Loading meteo data
+# Loading wind data
 Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
-Stations = ['South_Namib_Station', 'Deep_Sea_Station']
 
-# #### Computing quantities
+# Figure properties
+variables = ['U_star', 'Orientation']
+label_var = {'U_star': r'Velocity, $u_{*}~[\textup{m}~\textup{s}^{-1}]$', 'Orientation': r'Orientation, $\theta~[^\circ]$'}
+labels = [(r'\textbf{a}', r'\textbf{b}'), (r'\textbf{c}', r'\textbf{d}'),
+          (r'\textbf{e}', r'\textbf{f}'), (r'\textbf{g}', r'\textbf{h}')]
+row_labels = ['North Sand Sea -- summer', 'North Sand Sea -- winter', 'South Sand Sea -- summer',
+              'South Sand Sea -- winter']
+years = [2017, 2017, 2017, 2017]
+months = [12, 6, 11, 5]
+days = [(5, 8), (1, 4), (3, 6), (8, 11)]
+month_calendar = {index: month for index, month in enumerate(calendar.month_name) if month}
[…]

### 2022-05-22 Paper_figure/Figure06.py

@@ -7,47 +7,11 @@ Figure 6
 
 import numpy as np
 import matplotlib.pyplot as plt
-import matplotlib.patches as ptch
-from matplotlib.colors import to_rgba
-from PIL import Image
-from datetime import datetime
 import sys
 import os
 sys.path.append('../')
 import python_codes.theme as theme
-from python_codes.meteo_analysis import quartic_transport_law, quadratic_transport_law
-from python_codes.general import Make_angular_PDF, cosd, sind, Vector_average
-from python_codes.CourrechDuPont2014 import Bed_Instability_Orientation, Elongation_direction
-from python_codes.plot_functions import plot_flux_rose, plot_arrow
-
-
-def North_arrow(fig, ax, center, length, length_small, width, radius, theta=0, color='k'):
-    y_start = radius + length - length_small
-    arrow = np.array([[0, -y_start], [width/2, -radius],
-                      [0, -(radius + length)], [-width/2, -radius], [0, -y_start]])
-    arrow = arrow + np.array(center)
-    ax.add_patch(plt.Polygon(arrow, color=color))
-    t = ax.text(center[0], center[1], 'N')
-    r = fig.canvas.get_renderer()
-    inv = ax.transData.inverted()
-    bb = t.get_window_extent(renderer=r).transformed(inv)
-    t.set_visible(False)
-    height_t = bb.height
-    t = ax.text(center[0], center[1] - height_t/2, r'\textbf{N}', color=color, ha='center')
-
-
-def plot_orientation_wedge(ax, A_F, A_BI, center, length, color_F, color_BI, alpha=0.2, **kwargs):
-    wedge_f = ptch.Wedge(center, length, -np.nanmax(A_F), -np.nanmin(A_F),
-                         edgecolor=color_F, facecolor=to_rgba(color_F, alpha), joinstyle='round', **kwargs)
-    wedge_bi = ptch.Wedge(center, length, -np.nanmax(A_BI), -np.nanmin(A_BI),
-                          edgecolor=color_BI, facecolor=to_rgba(color_BI, alpha), joinstyle='round', **kwargs)
-    wedge_bis = ptch.Wedge(center, length, -np.nanmax(A_BI) + 180, -np.nanmin(A_BI) + 180,
[…]

### 2022-05-22 Paper_figure/Figure07.py

@@ -1,0 +1,213 @@
+"""
+============
+Figure 7
+============
+
+"""
+
+import numpy as np
+import matplotlib.pyplot as plt
+from matplotlib.patches import FancyArrowPatch
+import sys
+import os
+sys.path.append('../')
+import python_codes.theme as theme
+from python_codes.general import cosd, sind
+from python_codes.linear_theory import Cisaillement_basal_rotated_wind, coeffA0, coeffB0
+
+
+def perturb(x, z, amp, lamb, shift):
+    return np.cos(2*np.pi*x/lamb)*amp*np.exp(-z/lamb) + shift
+
+
+def plot_vertical_profile(ax, height, Virtual_potential_temperature, grad_free_atm, theta_free_atm, blh, theta_ground, Hmax_fit, color='tab:blue', label=None, alpha_pt=1):
+    Hfit = np.linspace(blh, Hmax_fit, 100)
+    #
+    line = ax.vlines(theta_ground, 0, blh/1e3, color=color, label=label, zorder=-3, lw=1)
+    ax.axhline(blh/1e3, color=color, ls='--')
+    ax.plot(np.poly1d([grad_free_atm, theta_free_atm])(Hfit), Hfit/1e3, color=line.get_color(), zorder=-2, lw=1)
+    ax.plot(Virtual_potential_temperature, height/1e3, '.', color=line.get_color(), zorder=-1, alpha=alpha_pt)
+    # ax.scatter(theta_ground, blh/1e3, s=30, facecolors=line.get_color(), edgecolors='k', linewidth=2, zorder=0)
+
+
+def plot_streamlines(ax, z_pos, x, amp, lamb, **kwargs):
+    for z in z_pos:
+        ax.plot(x, perturb(x, z, amp, lamb, z), **kwargs)
+
+
+def topo(x, y, alpha, k, xi):
+    return xi*np.cos(k*(cosd(alpha)*x + sind(alpha)*y))
+
+
+# Loading figure theme
+theme.load_style()
+
+# paths
+path_imgs = '../static/images/'
+path_savefig = '../Paper/Figures'
+path_outputdata = '../static/data/processed_data'
+
+# ##### Loading meteo data
+Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
[…]

### 2022-05-22 Paper_figure/Figure09.py

@@ -1,0 +1,225 @@
+"""
+============
+Figure 8
+============
+
+"""
+
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.patches as ptch
+from matplotlib.colors import to_rgba
+from PIL import Image
+from datetime import datetime
+import sys
+import os
+sys.path.append('../')
+import python_codes.theme as theme
+from python_codes.meteo_analysis import quartic_transport_law, quadratic_transport_law
+from python_codes.general import Make_angular_PDF, cosd, sind, Vector_average
+from python_codes.CourrechDuPont2014 import Bed_Instability_Orientation, Elongation_direction
+from python_codes.plot_functions import plot_flux_rose, plot_arrow
+
+
+def North_arrow(fig, ax, center, length, length_small, width, radius, theta=0, color='k'):
+    y_start = radius + length - length_small
+    arrow = np.array([[0, -y_start], [width/2, -radius],
+                      [0, -(radius + length)], [-width/2, -radius], [0, -y_start]])
+    arrow = arrow + np.array(center)
+    ax.add_patch(plt.Polygon(arrow, color=color))
+    t = ax.text(center[0], center[1], 'N')
+    r = fig.canvas.get_renderer()
+    inv = ax.transData.inverted()
+    bb = t.get_window_extent(renderer=r).transformed(inv)
+    t.set_visible(False)
+    height_t = bb.height
+    t = ax.text(center[0], center[1] - height_t/2, r'\textbf{N}', color=color, ha='center')
+
+
+def plot_orientation_wedge(ax, A_F, A_BI, center, length, color_F, color_BI, alpha=0.2, **kwargs):
+    wedge_f = ptch.Wedge(center, length, -np.nanmax(A_F), -np.nanmin(A_F),
+                         edgecolor=color_F, facecolor=to_rgba(color_F, alpha), joinstyle='round', **kwargs)
+    wedge_bi = ptch.Wedge(center, length, -np.nanmax(A_BI), -np.nanmin(A_BI),
+                          edgecolor=color_BI, facecolor=to_rgba(color_BI, alpha), joinstyle='round', **kwargs)
+    wedge_bis = ptch.Wedge(center, length, -np.nanmax(A_BI) + 180, -np.nanmin(A_BI) + 180,
[…]

### 2022-05-22 Paper_figure/Supplementary_Figures/Figure01_supp.py

@@ -10,37 +10,67 @@ import sys
 import glob
 import numpy as np
 import matplotlib.pyplot as plt
-from PIL import Image
+import matplotlib.patches as mpatches
 sys.path.append('../../')
 import python_codes.theme as theme
-
-# Loading figure theme
+from datetime import timedelta
+#
 theme.load_style()
+
+
+def make_range_broken_barh(time, dt):
+    diff = np.diff(time)
+    t_diff = np.concatenate((time[1:][diff > dt_threshold], time[:-1][diff > dt_threshold]))
+    t_diff = np.insert(t_diff, [0, t_diff.size], [time[0], time[-1]])
+    t_diff = np.array(sorted(t_diff))
+    return [(tstart, tspan) for tstart, tspan in zip(t_diff[::2], np.diff(t_diff)[::2])]
 
 
 # paths
 path_savefig = '../../Paper/Figures'
-path_images = '../../static/images'
+path_inputdata = '../../static/data/raw_data/'
 
-# images
-list_images = sorted(glob.glob(os.path.join(path_images, '*+*')))
-order_plot = [6, 4, 5, 2, 1]
-labels = [r'\textbf{a}', r'\textbf{b}', r'\textbf{c}', r'\textbf{d}', r'\textbf{e}']
 
-axd = plt.figure(constrained_layout=True,
-                 figsize=(theme.fig_width, 0.7*theme.fig_height_max)).subplot_mosaic(
-    """
-    AABB
-    CCDD
-    .EE.
-    """,
-    gridspec_kw={'height_ratios': (1, 1.2, 1)}
-)
+Stations = ['Adamax_Station', 'Huab_Station', 'Deep_Sea_Station', 'South_Namib_Station']
+labels = ['Etosha West', 'Huab', 'North Sand Sea', 'South Sand Sea']
 
-for ax_key, i, label in zip(axd.keys(), order_plot, labels):
-    axd[ax_key].imshow(np.array(Image.open(list_images[i])))
-    axd[ax_key].set_axis_off()
-    axd[ax_key].text(0.02, 0.98, label, ha='left', va='top', transform=axd[ax_key].transAxes)
+directory_types = ['ERA5Land', 'measured_wind_data']
 
-plt.savefig(os.path.join(path_savefig, 'Figure1_supp.pdf'), dpi=400)
+colors = [theme.color_Era5Land, theme.color_insitu]
+
+dt_threshold = timedelta(minutes=60)
[…]

### 2022-05-22 Paper_figure/Supplementary_Figures/Figure02_supp.py

@@ -5,72 +5,61 @@ Figure 2 -- Online Resource
 
 """
 
-import os
-import sys
-import glob
 import numpy as np
+import os
+from datetime import datetime
+import matplotlib.dates as mdates
 import matplotlib.pyplot as plt
-import matplotlib.patches as mpatches
+import sys
 sys.path.append('../../')
 import python_codes.theme as theme
-from datetime import timedelta
-#
+
 theme.load_style()
-
-
-def make_range_broken_barh(time, dt):
-    diff = np.diff(time)
-    t_diff = np.concatenate((time[1:][diff > dt_threshold], time[:-1][diff > dt_threshold]))
-    t_diff = np.insert(t_diff, [0, t_diff.size], [time[0], time[-1]])
-    t_diff = np.array(sorted(t_diff))
-    return [(tstart, tspan) for tstart, tspan in zip(t_diff[::2], np.diff(t_diff)[::2])]
-
 
 # paths
 path_savefig = '../../Paper/Figures'
-path_inputdata = '../../static/data/raw_data/'
+path_outputdata = '../../static/data/processed_data'
+path_inputdata = '../../static/data/raw_data'
+
+# figure parameters
+station = 'South_Namib_Station'
+tmin, tmax = datetime(2017, 6, 3), datetime(2017, 6, 10)
+Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
+
+# Loading and recomputing some raw data
+path_insitu = os.path.join(path_inputdata, 'measured_wind_data/in_situ_wind_data_' + station + '.npy')
+Data_insitu = np.load(path_insitu, allow_pickle=True).item()
+#
+t_insitu = Data_insitu['time']
+U_insitu = Data_insitu['velocity']
+# putting angles in trigo. ref.
+Orientation_insitu = (270 - Data_insitu['direction']) % 360
 
 
-Stations = ['Adamax_Station', 'Huab_Station', 'Deep_Sea_Station', 'South_Namib_Station']
-labels = ['Etosha West', 'Huab', 'North Sand Sea', 'South Sand Sea']
+# ### Figure
+fig, axarr = plt.subplots(2, 1, figsize=(theme.fig_width, 0.85*theme.fig_width),
+                          constrained_layout=True, sharex=True)
 
-directory_types = ['ERA5Land', 'measured_wind_data']
[…]

### 2022-05-22 Paper_figure/Supplementary_Figures/Figure03_supp.py

@@ -7,9 +7,8 @@ Figure 3 -- Online Resource
 
 import numpy as np
 import os
-from datetime import datetime
-import matplotlib.dates as mdates
 import matplotlib.pyplot as plt
+from matplotlib.colors import Normalize
 import sys
 sys.path.append('../../')
 import python_codes.theme as theme
@@ -18,48 +17,43 @@ theme.load_style()
 
 # paths
 path_savefig = '../../Paper/Figures'
-path_outputdata = '../../static/data/processed_data'
-path_inputdata = '../../static/data/raw_data'
+path_outputdata = '../../static/data/processed_data/'
 
-# figure parameters
-station = 'South_Namib_Station'
-tmin, tmax = datetime(2017, 6, 3), datetime(2017, 6, 10)
 Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
+Data_roughness = np.load(os.path.join(path_outputdata, 'Data_calib_roughness.npy'),
+                         allow_pickle=True).item()
 
-# Loading and recomputing some raw data
-path_insitu = os.path.join(path_inputdata, 'measured_wind_data/in_situ_wind_data_' + station + '.npy')
-Data_insitu = np.load(path_insitu, allow_pickle=True).item()
+labels = [r'\textbf{a}', r'\textbf{b}', r'\textbf{c}', r'\textbf{d}']
+norm = Normalize(vmin=0.3, vmax=1, clip=True)
+
+# #### Figure
+fig, axrr = plt.subplots(2, 2, figsize=(theme.fig_width, 1.1*theme.fig_width),
+                         sharex=True, sharey=True, constrained_layout=True)
 #
-t_insitu = Data_insitu['time']
-U_insitu = Data_insitu['velocity']
-# putting angles in trigo. ref.
-Orientation_insitu = (270 - Data_insitu['direction']) % 360
+for i, (ax, p, metric, station, label) in enumerate(zip(axrr.flatten(), Data_roughness['Pvals'],
+                                                        Data_roughness['Metrics'], Data_roughness['Stations'], labels)):
[…]

### 2022-05-22 Paper_figure/Supplementary_Figures/Figure05_supp.py

@@ -9,10 +9,10 @@ import numpy as np
 import os
 import matplotlib.pyplot as plt
 import sys
-from types import SimpleNamespace
 sys.path.append('../../')
 import python_codes.theme as theme
-from python_codes.general import cosd, sind
+from python_codes.plot_functions import plot_scatter_surrounded
+
 
 theme.load_style()
 
@@ -20,59 +20,44 @@ theme.load_style()
 path_savefig = '../../Paper/Figures'
 path_outputdata = '../../static/data/processed_data/'
 
-Data_DEM = np.load(os.path.join(path_outputdata, 'Data_DEM.npy'),
-                   allow_pickle=True).item()
+Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
+labels = [r'\textbf{a}', r'\textbf{b}']
 
-labels = [(r'\textbf{a}', r'\textbf{d}'), (r'\textbf{b}', r'\textbf{e}'), (r'\textbf{c}', r'\textbf{f}')]
+# preparing data
+Stations_ref = ['Adamax_Station', 'Huab_Station']
+#
+Theta_ERA = np.concatenate([Data[station]['Orientation_era'] for station in Stations_ref])
+Theta_Station = np.concatenate([Data[station]['Orientation_insitu'] for station in Stations_ref])
+#
+U_ERA = np.concatenate([Data[station]['U_star_era'] for station in Stations_ref])
+U_Station = np.concatenate([Data[station]['U_star_insitu'] for station in Stations_ref])
 
-fig, axrr = plt.subplots(3, 2, figsize=(theme.fig_width, 0.75*theme.fig_height_max),
-                         constrained_layout=True, gridspec_kw={'width_ratios': (0.9, 1)})
-for i, station in enumerate(Data_DEM.keys()):
-    # loading into namespace from data dictionnary to shorten call
-    n = SimpleNamespace(**Data_DEM[station])
-    # ax0: Topo
-    cs = axrr[0, i].contourf(n.lon, n.lat, n.topo, levels=50)
-    for c in cs.collections:
-        c.set_edgecolor("face")
-        c.set_rasterized(True)
[…]

### 2022-05-22 Paper_figure/Supplementary_Figures/Figure08_supp.py

@@ -6,76 +6,105 @@ Figure 8 -- Online Resource
 """
 
 import numpy as np
+import os
 import matplotlib.pyplot as plt
 import sys
-import os
 sys.path.append('../../')
 import python_codes.theme as theme
 from python_codes.plot_functions import make_nice_histogram
 
-# Loading figure theme
+
+def plot_vertical_profile(ax, height, Virtual_potential_temperature, grad_free_atm, theta_free_atm, blh, theta_ground, Hmax_fit, color='tab:blue', label=None):
+    Hfit = np.linspace(blh, Hmax_fit, 100)
+    #
+    line = ax.vlines(theta_ground, 0, blh/1e3, color=color, label=label, zorder=-3)
+    ax.axhline(blh/1e3, alpha=0.5, color=color, ls='--')
+    ax.plot(np.poly1d([grad_free_atm, theta_free_atm])(Hfit), Hfit/1e3, color=line.get_color(), zorder=-2)
+    ax.plot(Virtual_potential_temperature, height/1e3, '.', color=line.get_color(), zorder=-1)
+    # ax.scatter(theta_ground, blh/1e3, s=30, facecolors=line.get_color(), edgecolors='k', linewidth=2, zorder=0)
+
+
 theme.load_style()
 
 # paths
 path_savefig = '../../Paper/Figures'
 path_outputdata = '../../static/data/processed_data/'
 
-# Loading wind data
+# Loading data
 Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
-Stations = sorted(Data.keys())
 
-# Figure properties
-station = 'South_Namib_Station'
-#
-theta_bins = [[0, 140], [150, 260]]
-velocity_bins = [[0.05, 0.2], [0.3, 10]]
-Data_pattern = np.load(os.path.join(path_outputdata, 'Data_DEM.npy'), allow_pickle=True).item()[station]
-icon = [r'\faSun', r'\faMoon']
+labels = [r'\textbf{a}', r'\textbf{b}', r'\textbf{c}', r'\textbf{d}']
 
-color_ax = 'purple'
+# ## vertical profiles parameters
+station = 'Deep_Sea_Station'
+time_steps_bad = [10856, 30266, 33463]
+time_steps_good = [2012, 30302, 30310]
+colors = ['tab:blue', 'tab:orange', 'tab:green']
+Hmax_fit = 10000  # [m]
 
-# ################ Figure
[…]

### 2022-05-22 Paper_figure/Supplementary_Figures/Figure09_supp.py

@@ -11,7 +11,9 @@ import sys
 import os
 sys.path.append('../../')
 import python_codes.theme as theme
+from python_codes.meteo_analysis import mu
 from python_codes.plot_functions import make_nice_histogram
+
 
 # Loading figure theme
 theme.load_style()
@@ -20,54 +22,48 @@ theme.load_style()
 path_savefig = '../../Paper/Figures'
 path_outputdata = '../../static/data/processed_data/'
 
-# Loading wind data
-Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'),
-               allow_pickle=True).item()
-Stations = ['Huab_Station', 'Adamax_Station']
+# ##### Loading meteo data
+Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
 
-# Figure properties
-velocity_bins = [[0.05, 0.2], [0.3, 10]]
-color_ax = 'purple'
+# ## histograms parameters
+Stations = ['South_Namib_Station', 'Deep_Sea_Station']
+g = 9.81  # [m/s2] gravitational constant
+z0_era = 1e-3  # [m] hydrodynamic roughness
+#
+labels = [r'\textbf{a}', r'\textbf{b}', r'\textbf{c}', r'\textbf{d}']
+nbins = 80
+
+# #### Figure
+fig, axarr = plt.subplots(2, 2, figsize=(theme.fig_width, 0.9*theme.fig_width),
+                          constrained_layout=True, sharey=True)
+
+for station in Stations:
+    make_nice_histogram(Data[station]['Boundary layer height'], nbins, axarr[0, 0], alpha=0.4, label=' '.join(station.split('_')[:-1]), density=False, scale_bins='log')
+    #
+    N = np.sqrt(g*Data[station]['gradient_free_atm']/Data[station]['theta_ground'])
+    make_nice_histogram(N, nbins, axarr[0, 1], alpha=0.4, label=' '.join(station.split('_')[:-1]), density=False)
+    #
+    U_H = Data[station]['U_star_era']*mu(Data[station]['Boundary layer height'], z0_era)
+    make_nice_histogram(U_H, nbins, axarr[1, 0], alpha=0.4, label=' '.join(station.split('_')[:-1]), density=False)
+    #
[…]

### 2022-05-22 Paper_figure/Supplementary_Figures/Figure12_supp.py

@@ -6,104 +6,71 @@ Figure 12 -- Online Resource
 """
 
 import numpy as np
-import os
 import matplotlib.pyplot as plt
 import sys
+import os
 sys.path.append('../../')
 import python_codes.theme as theme
 from python_codes.plot_functions import make_nice_histogram
 
 
-def plot_vertical_profile(ax, height, Virtual_potential_temperature, grad_free_atm, theta_free_atm, blh, theta_ground, Hmax_fit, color='tab:blue', label=None):
-    Hfit = np.linspace(blh, Hmax_fit, 100)
-    #
-    line = ax.vlines(theta_ground, 0, blh/1e3, color=color, label=label, zorder=-3)
-    ax.axhline(blh/1e3, alpha=0.5, color=color, ls='--')
-    ax.plot(np.poly1d([grad_free_atm, theta_free_atm])(Hfit), Hfit/1e3, color=line.get_color(), zorder=-2)
-    ax.plot(Virtual_potential_temperature, height/1e3, '.', color=line.get_color(), zorder=-1)
-    # ax.scatter(theta_ground, blh/1e3, s=30, facecolors=line.get_color(), edgecolors='k', linewidth=2, zorder=0)
-
-
+# Loading figure theme
 theme.load_style()
 
 # paths
 path_savefig = '../../Paper/Figures'
 path_outputdata = '../../static/data/processed_data/'
 
-# Loading data
+# ##### Loading meteo data
 Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
+# Stations = ['South_Namib_Station', 'Deep_Sea_Station']
+Stations = ['Deep_Sea_Station']
 
-labels = [r'\textbf{a}', r'\textbf{b}', r'\textbf{c}', r'\textbf{d}']
+# #### Computing quantities
+Orientation_era = np.concatenate([Data[station]['Orientation_era'] for station in Stations])
+U_era = np.concatenate([Data[station]['U_star_era'] for station in Stations])
+U_insitu = np.concatenate([Data[station]['U_star_insitu'] for station in Stations])
+time = np.concatenate([Data[station]['time'] for station in Stations])
+month = np.array([i.month for i in time])
+hour = np.array([i.hour for i in time])
+#
+delta_u = (U_era - U_insitu)/U_era
[…]

### 2022-05-22 Paper_figure/Supplementary_Figures/Figure13_supp.py

@@ -11,8 +11,13 @@ import sys
 import os
 sys.path.append('../../')
 import python_codes.theme as theme
-from python_codes.meteo_analysis import mu
-from python_codes.plot_functions import make_nice_histogram
+from python_codes.general import cosd, sind
+from python_codes.plot_functions import plot_regime_diagram
+from python_codes.linear_theory import Cisaillement_basal_rotated_wind
+
+
+def topo(x, y, alpha, k, xi):
+    return xi*np.cos(k*(cosd(alpha)*x + sind(alpha)*y))
 
 
 # Loading figure theme
@@ -22,48 +27,107 @@ theme.load_style()
 path_savefig = '../../Paper/Figures'
 path_outputdata = '../../static/data/processed_data/'
 
-# ##### Loading meteo data
-Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
-
-# ## histograms parameters
+# ## regime diagram properties
+# data
 Stations = ['South_Namib_Station', 'Deep_Sea_Station']
-g = 9.81  # [m/s2] gravitational constant
-z0_era = 1e-3  # [m] hydrodynamic roughness
+Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
+numbers = {key: np.concatenate([Data[station][key] for station in Stations]) for key in ('Froude', 'kH', 'kLB')}
+
+# Time series hydrodynamic coefficients
+Hydro_coeffs_time = np.load(os.path.join(path_outputdata, 'time_series_hydro_coeffs.npy'), allow_pickle=True).item()
+modulus = np.linalg.norm(np.concatenate([Hydro_coeffs_time[station] for station in Stations], axis=1), axis=0)
+
 #
-labels = [r'\textbf{a}', r'\textbf{b}', r'\textbf{c}', r'\textbf{d}']
-nbins = 80
+couples = [('Froude', 'kH'), ('kLB', 'kH')]
+labels = [r'\textbf{a}', r'\textbf{b}', r'\textbf{c}']
+#
[…]

### 2022-05-22 Paper_figure/Supplementary_Figures/Figure14_supp.py

@@ -5,14 +5,28 @@ Figure 14 -- Online Resource
 
 """
 
-import numpy as np
 import os
-import matplotlib.pyplot as plt
-from matplotlib.colors import LogNorm
 import sys
+import calendar
+import locale
+import numpy as np
+import matplotlib.dates as mdates
+import matplotlib.pyplot as plt
+from matplotlib.lines import Line2D
+from datetime import datetime, timedelta
 sys.path.append('../../')
 import python_codes.theme as theme
-from python_codes.plot_functions import make_nice_histogram
+from python_codes.meteo_analysis import mu
+
+locale.setlocale(locale.LC_ALL, 'en_US.utf8')
+
+
+def tick_formatter(ax, fmt='%d'):
+    myFmt = mdates.DateFormatter(fmt)
+    ax.xaxis.set_major_formatter(myFmt)
+    ticklabels = ax.get_xticklabels()
+    ticklabels[0].set_ha('left')
+
 
 theme.load_style()
 
@@ -20,109 +34,75 @@ theme.load_style()
 path_savefig = '../../Paper/Figures'
 path_outputdata = '../../static/data/processed_data/'
 
-# Loading data
 Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
-Stations = ['South_Namib_Station', 'Deep_Sea_Station']
 
-numbers = {key: np.concatenate([Data[station][key] for station in Stations]) for key in ('Froude', 'kH', 'kLB')}
-mask = ~np.isnan(numbers['Froude'])
-ad_hoc_quantity = np.concatenate([Data[station]['U_star_era'] for station in Stations])
+Stations = ['Deep_Sea_Station', 'Deep_Sea_Station', 'South_Namib_Station', 'South_Namib_Station']
+years = [2017, 2017, 2017, 2017]
+months = [12, 6, 11, 5]
+days = [(5, 8), (1, 4), (3, 6), (8, 11)]
+#
+z0_values = np.array([(1e-3, 1e-3), (1e-3, 1e-4), (1e-4, 1e-3)])
+#
+labels = [r'\textbf{a}', r'\textbf{b}', r'\textbf{c}', r'\textbf{d}']
+bbox2 = dict(facecolor=(1, 1, 1, 0.7), edgecolor=(1, 1, 1, 0), pad=0.5)
+month_calendar = {index: month for index, month in enumerate(calendar.month_name) if month}
 
-# Figure properties
[…]

### 2022-05-22 Paper_figure/Supplementary_Figures/Figure16_supp.py

@@ -1,77 +1,0 @@
-"""
-============================
-Figure 16 -- Online Resource
-============================
-
-"""
-
-import numpy as np
-import matplotlib.pyplot as plt
-import sys
-import os
-sys.path.append('../../')
-import python_codes.theme as theme
-from python_codes.plot_functions import make_nice_histogram
-
-
-# Loading figure theme
-theme.load_style()
-
-# paths
-path_savefig = '../../Paper/Figures'
-path_outputdata = '../../static/data/processed_data/'
-
-# ##### Loading meteo data
-Data = np.load(os.path.join(path_outputdata, 'Data_final.npy'), allow_pickle=True).item()
-# Stations = ['South_Namib_Station', 'Deep_Sea_Station']
-Stations = ['Deep_Sea_Station']
-
-# #### Computing quantities
-Orientation_era = np.concatenate([Data[station]['Orientation_era'] for station in Stations])
-U_era = np.concatenate([Data[station]['U_star_era'] for station in Stations])
-U_insitu = np.concatenate([Data[station]['U_star_insitu'] for station in Stations])
-time = np.concatenate([Data[station]['time'] for station in Stations])
-month = np.array([i.month for i in time])
-hour = np.array([i.hour for i in time])
-#
-delta_u = (U_era - U_insitu)/U_era
-limits = [0, -0.5]
-
-# #### Figure
-fig, axarr = plt.subplots(2, 2, figsize=(theme.fig_width, 0.9*theme.fig_width),
-                          constrained_layout=True)
-gs = axarr[0, 0].get_gridspec()
-# remove the underlying axes
-for ax in axarr[0, :]:
-    ax.remove()
-axbig = fig.add_subplot(gs[0, :])
-
-for limit in limits:
-    mask = delta_u < limit
-    make_nice_histogram(Orientation_era[mask], 150, axbig, vmin=0, vmax=360, alpha=0.5, density=True)
-    make_nice_histogram(month[mask], 13, axarr[1, 0], vmin=0.5, vmax=12.5, alpha=0.5, density=True)
-    make_nice_histogram(hour[mask], 13, axarr[1, 1], vmin=-0.5, vmax=23.5, alpha=0.5, density=True)
-
[…]

### 2022-05-22 python_codes/plot_functions.py

@@ -34,8 +34,10 @@ def plot_wind_rose(theta, U, bins, ax, fig, label=None,
         return the axe on which the wind rose is plotted. Can be used for further modifications.
 
     """
-    ax_rose = WindroseAxes.from_ax(fig=fig)
-    ax_rose.set_position(ax.get_position(), which='both')
+    # changing ax to have to have the windrose projection
+    subplotspec = ax.get_subplotspec()
+    ax.remove()
+    ax_rose = fig.add_subplot(subplotspec, projection='windrose')
     Angle = (90 - theta) % 360
     ax_rose.bar(Angle, U, bins=bins, normed=True, zorder=20, opening=1, edgecolor=None,
                 linewidth=0.5, nsector=60, **kwargs)
@@ -47,7 +49,6 @@ def plot_wind_rose(theta, U, bins, ax, fig, label=None,
     ax_rose.set_yticklabels([])
     if label is not None:
         fig.text(0.5, 0.05, label, ha='center', va='center', transform=ax.transAxes, bbox=props)
-    ax.remove()
     return ax_rose
 
 

### 2022-06-08 Paper/main.tex
Cyril first modifications for review

@@
 While feedbacks between these variables are well studied at the scale of a single dune, the average effect of a periodic dune pattern on atmospheric flows remains poorly constrained due to a lack of data in major sand seas.
-Here, we compare field measurements of surface wind data to the predictions of the ERA5-Land climate reanalysis at four locations in Namibia, including within the giant-dune field of the Namib sand sea.
+Here, we compare wind tower measurements of surface wind data to the predictions of the ERA5-Land climate reanalysis at four locations in Namibia, within and outside the giant-dune field of the Namib sand sea.
 In the desert plains to the north of the sand sea, observations and predictions agree well.
-This is also the case in the interdune areas of the sand sea, except for the weak winds blowing at night, which exhibit additional components aligned with the giant dune orientation, which are not predicted by the ERA5-Land reanalysis.
-We quantify these similarities and differences and provide a physical understanding of the relevant aerodynamical regimes to relate them to the daily cycle of the turbulent atmospheric boundary layer over a dune pattern of given wavelength.
-We conclude by identifying the conditions under which the ERA5-Land reanalysis data can reliably be used to study dune morphodynamics.
-We also propose that, in multidirectional wind regimes, deflections of specific winds by giant dunes could explain the occurrence of secondary dune patterns with a different orientation to the primary structures between which they develop.
-\keywords{Atmospheric boundary layer \and Sand dunes \and Fluid-structure interactions}
+This is also the case in the interdune areas of the sand sea during the day, but not during the night, where additional components aligned with the giant dune orientation are measured instead of the easterly wind predicted by the ERA5-Land reanalysis.
+%
+These discrepancies, with wind deviation and attenuation larger then $50^\circ$ and $60~\%$, are linked to the daily cycle of the turbulent atmospheric boundary layer over a complex topography, and to the associated flow regimes.
+During the night, the low wind velocities within a shallow boundary layer induces a flow confinement, leading to a strong streamline compression above the giant dunes and thus resulting in large flow deviations.
+During, the day, the flow confinement is reduced by higher wind velocities and deep boundary layers.
+%
+We finally propose that, in multidirectional wind regimes, this mechanism and the resulting wind deflections by giant dunes could explain the occurrence of secondary dune patterns elongating in different orientations compared to the primary structures between which they develop.
+
+\keywords{Atmospheric boundary layer \and Sand dunes \and Flow over hills}
 \end{abstract}
@@
 
-The description of turbulent flows over complex topography is relevant for a large variety of different environmental systems \citep{Finnigan2020}.
+The description of turbulent flows over complex topography is relevant for a large variety of different environmental systems \citep{Sherman1978, Walmsley1982, baines1995, Wood2000, Venditti2013, Finnigan2020}.
 For example, the flow over hills is of primary interest for wind power, meteorological and air pollution phenomena \citep{Taylor1987}.
-The properties of these flows are also key to the understanding of geophysical phenomena, including the formation of wind-driven waves on the ocean surface \citep{Sullivan2010}, dissolution bedforms \citep{Claudin2017}, or sedimentary ripples and dunes \citep{Charru2013, Courrech2015}.
+The properties of these flows are also key to the understanding of geophysical phenomena, including the formation of wind-driven waves on the ocean surface \citep{Sullivan2010}, dissolution bedforms \citep{Claudin2017, Guerin2020}, or sedimentary ripples and dunes \citep{Bagnold1941, Charru2013, Courrech2015}.
 Importantly, the troposphere presents a vertical structure, with a lower convective boundary layer, of typical kilometer-scale thickness, capped by a stably stratified region \citep{Stull1988}.
@@
 
-Focusing on the wind close to the surface, two related topographic feedbacks on the windflow over dunes can be commented on separately.
+Compared to mountains, aeolian sand dunes offer idealized topographies to the study of atmospheric turbulent flow over wavy bottoms, due to their smooth shape, free of canopies, resulting from their interaction with the wind \citep{Bagnold1941}.
+Then, past studies have highlighted two related topographic feedbacks on the windflow close to the dune/hill surface.
+%
 First is the effect on wind speed, with documented flow acceleration on upwind slopes \citep{Weaver2011} and deceleration on downwind slopes \citep{Baddock2007}, where the speed-up factor is essentially proportional to the obstacle aspect ratio \citep{Jackson1975}.
-Importantly, the velocity maximum is typically shifted upwind of the obstacle crest.
+Importantly, the velocity maximum is typically shifted upwind of the obstacle crest \citep{Jackson1975, Claudin2013}.
 This behaviour has been theoretically predicted by means of asymptotic analysis of a neutrally stratified boundary-layer flow over an obstacle of vanishing aspect ratio \citep{Jackson1975, Mason1979, Sykes1980, Hunt1988, Belcher1998}.
-Experiments in flumes \citep{Zilker1977, Zilker1979, Frederick1988, Poggi2007, Bristow2022}, in wind tunnels \citep{Gong1989, Finnigan1990, Gong1996} and in field conditions \citep{Taylor1987a, Claudin2013, Fernando2019, Lu2021}, have also documented this effect.
+Experiments in flumes \citep{Zilker1977, Zilker1979, Frederick1988, Poggi2007, Bristow2022}, in wind tunnels \citep{Gong1989, Finnigan1990, Gong1996} and in field conditions at all scales \citep{Taylor1987a, Claudin2013, Fernando2019, Lu2021}, have also documented this effect.
 Interestingly, a similar behaviour exists for the pressure perturbation, but with a slight downwind shift for the pressure minimum \citep{Claudin2021}.
 The second effect, much less studied, is the flow deflection that occurs when the incident wind direction is not perpendicular to the ridge crest.
-While predicted to be small (less than $10^{\circ}$) in the linear regime valid for shallow topography \citep{Gadal2019}, significant flow steering has been reported in the field on the downwind side of steep enough obstacles, such as mountain ranges \citep{Kim2000, Lewis2008, Fernando2019}, well-developed sand dunes \citep{Walker2009, Hesp2015, Walker2017, Smith2017, deWinter2020}, and valley topographies \citep{Wiggs2002, Garvey2005}.
+While predicted to be small (less than $10^{\circ}$) in the linear regime valid for shallow topography \citep{Gadal2019}, significant flow steering has been reported in the field on the downwind side of steep enough obstacles, such as well-developed sand dunes \citep{Tsoar1983, Sweet1990, Walker2002, Smith2017} and in particular coastal foredunes (e.g.~\citet{rasmussen1989}, \citet{Walker2006}, \citet{Walker2009}, \citet{Hesp2015}, \citet{Walker2017}, \citet{deWinter2020}), mountain ranges \citep{Kim2000, Lewis2008, Fernando2019}, and valley topographies \citep{Wiggs2002, Garvey2005}.
 
-For practical reasons, wind measurement over sand dunes has been performed over small bedforms, typically a few meters high (corresponding to tens of meters long) (e.g.~\citet{Lancaster1996}, \citet{MckennaNeuman1997}, \citet{Sauermann2003}, \citet{Andreotti2002}, \citet{Walker2002}, \citet{Weaver2011}).
-Giant dunes, with kilometer-scale wavelengths and heights of tens of meters, are more difficult to investigate although for several reasons they provide a choice configuration for the study of turbulent flows over a complex topography.
+Wind measurements over sand dunes has been mainly performed over small bedforms, typically a few meters high (corresponding to tens of meters long) (e.g.~\citet{Mulligan1988}, \citet{Hesp1989}, \citet{Lancaster1996}, \citet{MckennaNeuman1997}, \citet{Sauermann2003}, \citet{Andreotti2002}, \citet{Walker2002}, \citet{Weaver2011}).
+For practical reasons, fewer studies performed similar measurements on giant dunes \citep{Havholm1988}, with kilometer-scale wavelengths and heights of tens of meters.
+However, they provide a choice configuration for the study of turbulent flows over a complex topography.
 First, one expects larger wind disturbances for larger obstacles.
@@
 Third, they usually form large patterns in sand seas and thus behave as rather clean periodic perturbations, in contrast with isolated dunes.
-Finally, because the morphodynamics of aeolian bedforms are strongly dependent on the local wind regime \citep{Livingstone2019}, one can expect to see the consequences of windflow disturbance by large dunes on neighbouring small dunes.
-A similar effect is observed on the properties of impact ripple patterns due to the presence of dunes \citep{Howard1977,Hood2021}.
+Finally, because the morphodynamics of aeolian bedforms are strongly dependent on the local wind regime \citep{Livingstone2019}, one can expect to see the consequences of windflow disturbance by large dunes on neighbouring small dunes \citep{Brookfield1977, Ewing2006}.
+A similar effect is observed on the properties of impact ripple patterns due to the presence of dunes \citep{Howard1977, Hood2021}.
 
@@
 
-%---------------------------------------------
+% ---------------------------------------------
 \begin{figure}[t]
@@
 \subsection{Wind and elevation data}
-At each meteorological station (Fig.~\ref{Fig1}), wind speed and direction were sampled every 10 minutes using cup anemometers (Vector Instruments A100-LK) and wind vanes (Vector Instruments W200-P) at heights which varied between 2~m and 3~m depending on the station.
+At each meteorological station (Fig.~\ref{Fig1}), wind speed and direction were sampled every 10 minutes using cup anemometers (Vector Instruments A100-LK) and wind vanes (Vector Instruments W200-P) at a single height, which was between 2~m and 3~m depending on the station.
 The available period of measurements at each station ranged from 1 to 5 discontinuous years distributed between 2012 and 2020 (Online Resource Fig.~\ref{Fig1_supp}).
@@
 \caption{Temporal comparison between the wind data coming from the ERA5-Land climate reanalysis (orange lines) and from the local measurements (blue lines).
-Coloured swathes indicate day (between 1000 UTC and 2200 UTC) and night (before 1000 UTC or after 2200 UTC) \textbf{a--b}: Etosha West station in summer.
-\textbf{b--c}: Etosha West station in winter.
-\textbf{d--e}: Huab station in summer.
-\textbf{f--g}: Huab station in winter.
+Coloured swathes indicate day (between 1000 UTC and 2200 UTC) and night (before 1000 UTC or after 2200 UTC).
+Numbers indicate the average flow deflection $\delta_{\theta}$ and wind modulation $\delta_{u}$ over the given period, as well as the frequency of occurence of extreme events ($\delta_{\theta} > 50^\circ$, $\vert\delta_{u}\vert > 0.6$).
+\textbf{a--b}: Huab station in summer.
+\textbf{b--c}: Huab station in winter.
+\textbf{d--e}: Etosha West station in summer.
+\textbf{f--g}: Etosha West station in winter.
 Time series of the two other stations are shown in Fig.~\ref{Fig5}.}
@@
 \caption{Temporal comparison between the wind data coming from the ERA5-Land climate reanalysis (orange lines) and from the local measurements (blue lines).
-Coloured swathes indicate day (between 1000 UTC and 2200 UTC) and night (before 1000 UTC or after 2200 UTC) \textbf{a--b}: North Sand Sea station in summer.
+Coloured swathes indicate day (between 1000 UTC and 2200 UTC) and night (before 1000 UTC or after 2200 UTC).
+Numbers indicate the average flow deflection $\delta_{\theta}$ and wind modulation $\delta_{u}$ over the given period, as well as the frequency of occurence of extreme events ($\delta_{\theta} > 50^\circ$, $\vert\delta_{u}\vert > 0.6$).
+\textbf{a--b}: North Sand Sea station in summer.
 \textbf{b--c}: North Sand Sea station in winter.
@@
 \section{Influence of wind speed and circadian cycle on the atmospheric boundary layer}
-The wind deflection induced by linear dunes has previously been related to the incident angle between wind direction and crest orientation, with a maximum deflection evident for incident angles between $30^{\circ}$ and $70^{\circ}$ \citep{Walker2009, Hesp2015}.
+
+The wind deflection induced by dunes has previously been related to the incident angle between wind direction and crest orientation, with a maximum deflection evident for incident angles between $30^{\circ}$ and $70^{\circ}$ \citep{Walker2009, Hesp2015}.
 In the data analysed here, the most deflected wind at both the North and South Sand Sea stations is seen to be where the incident angle is perpendicular to the giant dunes (Figs.~\ref{Fig2} and \ref{Fig6}).
@@
 Part of this response, on which we focus here, is the flow deflection by the ridges.
-In a simplified way, it can be understood from the Bernoulli principle~\citep{Hesp2015}: as the flow approaches the ridge crest, the compression of the streamlines results in larger flow velocities, and thus lower pressures \citep{Rubin1987}.
+In a simplified way, it can be understood from the Bernoulli principle~\citep{Hesp2015}: as the flow approaches the ridge crest, the compression of the streamlines results in larger flow velocities, and thus lower pressures \citep{Jackson1975}.
 An incident flow oblique to the ridge is then deflected towards lower pressure zones, i.e towards the crest.
@@
 Flow confinement below a capping surface, which enhances streamline compression, has a strong effect on the hydrodynamic response and typically increases flow deflection.
-This is the case for bedforms forming in open channel flows such as rivers~\citep{Fourriere2010, Unsworth2018}.
+This is the case for bedforms forming in open channel flows such as rivers~\citep{Kennedy1963, Chang1970, Mizumura1995, Colombini2004, Fourriere2010, Andreotti2012, Unsworth2018}.
 This is also relevant for aeolian dunes as they evolve in the turbulent atmospheric boundary layer (ABL) capped by the stratified free atmosphere (FA) \citep{Andreotti2009}.
@@
 
+Note that, in the presented results, the only quantites depending on the choice of the hydrodynamic roughnesses (see Online Resource section~\ref{calib_z0}) are the wind shear velocities, and as such the Froude number $\mathcal{F}r$ and the relative velocity modulation $\delta_{u}$.
+Considering the possible range of realistic roughnesses values ($10^{-4}$m to $10^{-5}$), the uncertainty of velocities estimated using the law of the wall, and thus the Froude number, is at most $30$.
+It is harder to quantify the error on the computation of $\delta_{u}$ considering the method used.
+However, Figure~\ref{Fig14_supp} shows that this choice has very little influence of the tempral variation of the relative velocity modulation, but can induce a global increase of decrease of its values.
+As such, it will not change the aspect of the regime diagram presented in Figure~\ref{Fig8}b, but may only change the $\delta_{u}$--value at which the transition between regimes is observed (here taken as 0, dashed black green lines in Figure~\ref{Fig8}b).
+
 %---------------------------------------------
@@
 The theoretical computation of the wind disturbances induced by sinusoidal ridges under flow confinement has been performed in the linear limit \citep{Andreotti2009, Andreotti2012}, i.e.
-when the aspect ration of these ridges is small ($k\xi_0 \ll 1$).
+when the aspect ratio of these ridges is small ($k\xi_0 \ll 1$).
 These models are able to qualitatively reproduce the observed wind deflection (Appendix 1, Online Resource Figs.~\ref{Fig11_supp} and \ref{Fig13_supp}), and thus provide the physical support for the interpretation we propose here based on hydrodynamic regimes.
@@
 where $\mathcal{A}_{0}$ and $\mathcal{B}_{0}$ are now two coefficients independent of the dune orientation $\alpha$, corresponding to the transverse case ($\alpha=90$).
-For a fully turbulent boundary layer capped by a stratified atmosphere, these coefficients depend on $k H$, $k z_{0}$, $\mathcal{F}$ and $\mathcal{F}_{\textup{I}}$ \citep{Andreotti2009}.
-In this study, we assume a constant hydrodynamic roughness $z_{0} \simeq 1$~mm (Online Resource section 1).
+In the case of a fully turbulent boundary layer capped by a stratified atmosphere, these coefficients depend on $k H$, $k z_{0}$, $\mathcal{F}$ and $\mathcal{F}_{\textup{I}}$ \citep{Andreotti2009}.
+For their computation, we assume here a constant hydrodynamic roughness $z_{0} \simeq 1$~mm (Online Resource section 1).
 For the considered giant dunes, this leads to $k z_{0} \simeq 2 \cdot 10^{-6}$, as their wavelength is $\lambda \simeq 2.4$~km (or $k \simeq 2 \cdot 10^{-3}$~m$^{-1}$).
@@
 As the regionally predicted and locally measured velocities are available at different heights, we can not compare them directly.
-We therefore convert all velocities into shear velocities $u_{*}$, characteristic of the turbulent velocity profile \citep{Spalding1961, Stull1988}:
+We therefore convert all velocities into shear velocities $u_{*}$, characteristic the turbulent logarithmic velocity profile \citep{Spalding1961, Stull1988}:
 %
@@
 where $z$ is the vertical coordinate, $\kappa = 0.4$ the von K\'arm\'an constant and $z_{0}$ the hydrodynamic roughness.
+Note that, strickly speaking, this logarithmic profile is valid for a neutrally stratified ABL only.
+Vertical density gradients occuring in other conditions may thus induce large discrepancies~\citep{Monin1954, Garratt1994, Dyer1974}.
+However, as our wind measurements are in the flow region close enough to the surface, where these effects are negligible, this logarithmic wind profile remains a farily good approximation in all conditions~\citep{gunn2021}.
+%
 Several measurements of hydrodynamic roughnesses are available \citep{Raupach1992,Bauer1992,Brown2008,Nield2014}.

### 2022-06-08 Paper_figure/Figure05.py

@@ -96,19 +96,19 @@ for i, (subfig, yr, mth, dy, station) in enumerate(zip(subfigs, years, months, d
         if var == 'U_star':
             ax.set_ylim((0, 0.5))
             ax.text(0.5, 0.94,
-                    r"""$ f_{{u}} = {:.2f}$
-                        $\langle \delta_{{u}} \rangle = {:.2f}$""".format(
-                        mask_u.sum()/delta_angle.size,
-                        delta_u[mask_u_theta].mean()),
+                    r"""$\langle \delta_{{u}} \rangle = {:.2f}$
+                        $f_{{u}} = {:.2f}$""".format(
+                        delta_u[mask_u_theta].mean(),
+                        mask_u.sum()/delta_angle.size),
                     ha='center', va='top', transform=ax.transAxes, bbox=bbox_props)
         else:
             ax.set_ylim((0, 360))
             ax.set_yticks((0, 90, 180, 270, 360))
             ax.text(0.5, 0.94,
-                    r"""$f_{{\theta}} = {:.2f}$
-                        $\langle \delta_{{\theta}} \rangle = {:.0f}$""".format(
-                        mask_theta.sum()/delta_angle.size,
-                        delta_angle[mask_u_theta].mean()),
+                    r"""$\langle \delta_{{\theta}} \rangle = {:.0f}$
+                        $f_{{\theta}} = {:.2f}$""".format(
+                        delta_angle[mask_u_theta].mean(),
+                        mask_theta.sum()/delta_angle.size),
                     ha='center', va='top', transform=ax.transAxes, bbox=bbox_props)
 
 #

### 2022-06-08 Paper_figure/Supplementary_Figures/Figure14_supp.py

@@ -49,7 +49,7 @@ month_calendar = {index: month for index, month in enumerate(calendar.month_name
 
 
 # #### Figure
-fig, axarr = plt.subplots(2, 2, figsize=(theme.fig_width, theme.fig_width),
+fig, axarr = plt.subplots(3, 2, figsize=(theme.fig_width, 1.3*theme.fig_width),
                           sharey=True)
 
 for (ax, station, yr, mth, day, label) in zip(axarr.flatten(), Stations, years,
@@ -86,7 +86,7 @@ for (ax, station, yr, mth, day, label) in zip(axarr.flatten(), Stations, years,
                         facecolor=theme.color_night, alpha=0.1, edgecolor=None, label=theme.Icon_night)
     #
 
-for ax in axarr[:, 0]:
+for ax in axarr[:-1, 0]:
     ax.set_ylabel(r'$\delta_{\textup{u}}$')
 
 custom_lines = [Line2D([0], [0], color='tab:blue'),
@@ -102,6 +102,8 @@ fig.legend(custom_lines, ['calibrated',
            title=r'Hydrodynamic roughnesses, $\left(z_{0}^{\textup{Era5Land}}, \, z_{0}^{\textup{local}}\right)$, [m]',
            ncol=4,
            handletextpad=0.4, columnspacing=1.5, loc='upper center')
+
+# ######### Froude number
 
 plt.subplots_adjust(top=0.86, right=0.98, bottom=0.08, hspace=0.3, wspace=0.1)
 plt.savefig(os.path.join(path_savefig, 'Figure14_supp.pdf'), dpi=400)

### 2022-07-04 Paper/main.tex
Philippe round of review

@@
 
-\title{Local wind regime induced by giant linear dunes: comparison of ERA5-Land re-analysis with surface measurements}
+\title{Local wind regime induced by giant linear dunes: comparison of ERA5-Land reanalysis with surface measurements}
 
@@
 Emergence and growth of sand dunes results from the dynamic interaction between topography, wind flow and sediment transport.
-While feedbacks between these variables are well studied at the scale of a single dune, the average effect of a periodic dune pattern on atmospheric flows remains poorly constrained due to a lack of data in major sand seas.
-Here, we compare wind tower measurements of surface wind data to the predictions of the ERA5-Land climate reanalysis at four locations in Namibia, within and outside the giant-dune field of the Namib sand sea.
+While feedbacks between these variables are well studied at the scale of a single and relatively small dune, the average effect of a periodic large-scale dune pattern on atmospheric flows remains poorly constrained, due to a lack of data in major sand seas.
+Here, we compare local measurements of surface winds to the predictions of the ERA5-Land climate reanalysis at four locations in Namibia, within and outside the giant linear dune field of the Namib sand sea.
 In the desert plains to the north of the sand sea, observations and predictions agree well.
-This is also the case in the interdune areas of the sand sea during the day, but not during the night, where additional components aligned with the giant dune orientation are measured instead of the easterly wind predicted by the ERA5-Land reanalysis.
+This is also the case in the interdune areas of the sand sea during the day.
+During the night, however, an additional wind component aligned with the giant dune orientation is measured, in contrast to the easterly wind predicted by the ERA5-Land reanalysis.
 %
-These discrepancies, with wind deviation and attenuation larger then $50^\circ$ and $60~\%$, are linked to the daily cycle of the turbulent atmospheric boundary layer over a complex topography, and to the associated flow regimes.
-During the night, the low wind velocities within a shallow boundary layer induces a flow confinement, leading to a strong streamline compression above the giant dunes and thus resulting in large flow deviations.
-During, the day, the flow confinement is reduced by higher wind velocities and deep boundary layers.
+We link these discrepancies, with wind deviation and velocity attenuation larger then $50^\circ$ and $60~\%$, to the daily cycle of the turbulent atmospheric boundary layer over a complex topography, and to the associated flow regimes.
+During the night, a shallow boundary layer induces a flow confinement associated with a strong streamline compression above the giant dunes, resulting in large flow deviations, especially for the the lower winds.
+During the day, the flow confinement is reduced by a thicker boundary layer and higher wind velocities, and the feedback of the giant dunes on the atmospheric flow is negligible.
 %
-We finally propose that, in multidirectional wind regimes, this mechanism and the resulting wind deflections by giant dunes could explain the occurrence of secondary dune patterns elongating in different orientations compared to the primary structures between which they develop.
+We finally propose that this mechanism and the resulting wind deflections by the giant dunes could explain the occurrence of smaller-scale secondary dune patterns, elongating along a different orientation compared to the primary dunes between which they develop.
 
@@
 \newpage
+
+CITE \citet{Chand22}
 
@@
 
-Compared to mountains, aeolian sand dunes offer idealized topographies to the study of atmospheric turbulent flow over wavy bottoms, due to their smooth shape, free of canopies, resulting from their interaction with the wind \citep{Bagnold1941}.
-Then, past studies have highlighted two related topographic feedbacks on the windflow close to the dune/hill surface.
-%
+Compared to hills and mountains, aeolian sand dunes offer idealized elevation profiles for the study of atmospheric turbulent flow over topographies, due to their smooth shape, free of canopies.
+Besides, dunes provide a rather wide range of scales, from decameters to kilometers, and very often come in a fairly regular pattern, which further simplifies the flow structure analysis.
+Past studies have highlighted two important topographic feedbacks on the wind flow close to the dune/hill surface.
 First is the effect on wind speed, with documented flow acceleration on upwind slopes \citep{Weaver2011} and deceleration on downwind slopes \citep{Baddock2007}, where the speed-up factor is essentially proportional to the obstacle aspect ratio \citep{Jackson1975}.
 Importantly, the velocity maximum is typically shifted upwind of the obstacle crest \citep{Jackson1975, Claudin2013}.
-This behaviour has been theoretically predicted by means of asymptotic analysis of a neutrally stratified boundary-layer flow over an obstacle of vanishing aspect ratio \citep{Jackson1975, Mason1979, Sykes1980, Hunt1988, Belcher1998}.
+This behaviour has been theoretically predicted by means of asymptotic analysis of a neutrally stratified boundary-layer flow over an obstacle of vanishing aspect ratio \citep{Jackson1975, Mason1979, Sykes1980, Hunt1988, Belcher1998, Kroy2002}.
 Experiments in flumes \citep{Zilker1977, Zilker1979, Frederick1988, Poggi2007, Bristow2022}, in wind tunnels \citep{Gong1989, Finnigan1990, Gong1996} and in field conditions at all scales \citep{Taylor1987a, Claudin2013, Fernando2019, Lu2021}, have also documented this effect.
@@
 The second effect, much less studied, is the flow deflection that occurs when the incident wind direction is not perpendicular to the ridge crest.
-While predicted to be small (less than $10^{\circ}$) in the linear regime valid for shallow topography \citep{Gadal2019}, significant flow steering has been reported in the field on the downwind side of steep enough obstacles, such as well-developed sand dunes \citep{Tsoar1983, Sweet1990, Walker2002, Smith2017} and in particular coastal foredunes (e.g.~\citet{rasmussen1989}, \citet{Walker2006}, \citet{Walker2009}, \citet{Hesp2015}, \citet{Walker2017}, \citet{deWinter2020}), mountain ranges \citep{Kim2000, Lewis2008, Fernando2019}, and valley topographies \citep{Wiggs2002, Garvey2005}.
+While predicted to be small (less than $10^{\circ}$) in the linear regime valid for shallow topography \citep{Gadal2019}, significant flow steering has been reported in the field on the downwind side of steep enough obstacles, such as well-developed sand dunes \citep{Tsoar1983, Sweet1990, Walker2002, Smith2017} and in particular coastal foredunes (e.g.~ \citet{Hunter1983}, \citet{rasmussen1989}, \citet{Walker2006}, \citet{Walker2009}, \citet{Hesp2015}, \citet{Walker2017}, \citet{deWinter2020}), mountain ranges \citep{Kim2000, Lewis2008, Fernando2019}, and valley topographies \citep{Wiggs2002, Garvey2005}.
 
-Wind measurements over sand dunes has been mainly performed over small bedforms, typically a few meters high (corresponding to tens of meters long) (e.g.~\citet{Mulligan1988}, \citet{Hesp1989}, \citet{Lancaster1996}, \citet{MckennaNeuman1997}, \citet{Sauermann2003}, \citet{Andreotti2002}, \citet{Walker2002}, \citet{Weaver2011}).
+Wind measurements over sand dunes has been mainly performed over small bedforms, typically a few meters high (corresponding to several tens of meters long) (e.g.~\citet{Mulligan1988}, \citet{Hesp1989}, \citet{Lancaster1996}, \citet{MckennaNeuman1997}, \citet{Sauermann2003}, \citet{Andreotti2002}, \citet{Walker2002}, \citet{Weaver2011}).
 For practical reasons, fewer studies performed similar measurements on giant dunes \citep{Havholm1988}, with kilometer-scale wavelengths and heights of tens of meters.
-However, they provide a choice configuration for the study of turbulent flows over a complex topography.
+However, such large dunes provide a choice configuration for the study of turbulent flows over a complex topography.
 First, one expects larger wind disturbances for larger obstacles.
-Secondly, their large size makes them interact with the vertical structure of the atmosphere \citep{Andreotti2009}.
+Secondly, their large size can make them interact with the vertical structure of the atmosphere \citep{Andreotti2009}.
 Third, they usually form large patterns in sand seas and thus behave as rather clean periodic perturbations, in contrast with isolated dunes.
-Finally, because the morphodynamics of aeolian bedforms are strongly dependent on the local wind regime \citep{Livingstone2019}, one can expect to see the consequences of windflow disturbance by large dunes on neighbouring small dunes \citep{Brookfield1977, Ewing2006}.
+Finally, because the morphodynamics of aeolian bedforms is strongly dependent on the local wind regime \citep{Livingstone2019}, one can expect to see the consequences of windflow disturbance by large dunes on neighbouring small dunes \citep{Brookfield1977, Ewing2006}.
 A similar effect is observed on the properties of impact ripple patterns due to the presence of dunes \citep{Howard1977, Hood2021}.
 
-Atmospheric flows have been much studied at the desert-scale with climate reanalyses based on global atmospheric models \citep{Blumberg1996, Livingstone2010, Ashkenazy2012, Jolivet2021, Hu2021}, such as ERA-40, ERA-Interim or ERA-5 \citep{Uppala2005, Dee2011, Hersbach2020}.
-However, the spatial resolution (tens of kilometers) of these reanalyses implies average quantities that do not resolve the smaller scales of interest, which range from individual dunes to small mountains \citep{Livingstone2010}.
-Recently, the release of ERA5-Land has resolved this limitation by providing up to 70 years of hourly wind predictions at a $9$~km spatial resolution \citep{munoz2021}.
+Atmospheric flows have been much studied at the desert-scale with climate reanalyses based on global atmospheric models \citep{Blumberg1996, Livingstone2010, Ashkenazy2012, Jolivet2021, Hu2021, gunn2021}, such as ERA-40, ERA-Interim or ERA5 \citep{Uppala2005, Dee2011, Hersbach2020}.
+However, the spatial resolution of these reanalyses (tens of kilometers) implies average quantities that do not resolve the smaller scales of interest, which range from individual dunes to small mountains \citep{Livingstone2010}.
+Recently, the release of ERA5-Land has partly resolved this limitation by providing up to 70 years of hourly wind predictions at a $9$~km spatial resolution \citep{munoz2021}.
 However, its validity remains to be studied, especially in remote desert areas where assimilation of measured data is very low.
 
-In this work, we compare local wind speeds and directions measured by meteorological stations at four different locations inside and north of the giant-dune field of the Namib sand sea to the regional predictions of the ERA5-Land climate reanalysis.
+In this work, we compare local wind speeds and directions measured by meteorological stations at four different locations inside and north of the giant linear dune field of the Namib sand sea to the regional predictions of the ERA5-Land climate reanalysis.
 Where the meteorological stations are surrounded by a relatively flat environment, we show that local measurements and regional predictions agree well.
@@
 \caption{Wind data used in this study.
-Satellite images of these different environements (Google-Earth, Maxar Technologies, CNES/Airbus) are shown on the left.
+Satellite images of the different environments (Google-Earth, Maxar Technologies, CNES/Airbus) are shown on the left.
 The black dots show the location of the wind measurements stations.
-On the center and on the right, wind roses representing the data from the ERA5-Land climate reanalysis and the local wind stations are respectively shown.
-Note that the bars show the direction towards which the wind blows.}
+On the right of the photos, the corresponding wind roses representing the data from the ERA5-Land climate reanalysis and the local wind stations are displayed.
+Note: the graphical convention for the wind roses is that the bars show the direction towards which the wind blows (see color bar for velocity scale).}
 \label{Fig2}
@@
 \caption{Temporal comparison between the wind data coming from the ERA5-Land climate reanalysis (orange lines) and from the local measurements (blue lines).
-Coloured swathes indicate day (between 1000 UTC and 2200 UTC) and night (before 1000 UTC or after 2200 UTC).
-Numbers indicate the average flow deflection $\delta_{\theta}$ and wind modulation $\delta_{u}$ over the given period, as well as the frequency of occurence of extreme events ($\delta_{\theta} > 50^\circ$, $\vert\delta_{u}\vert > 0.6$).
+Coloured swathes indicate day (between 10.00 UTC and 22.00 UTC) and night (before 10.00 UTC or after 22.00 UTC).
+Numbers in legends indicate the average flow deflection $\delta_{\theta}$ and relative wind modulation $\delta_{u}$ over the displayed period (see section~\ref{DataDistribution} for their definitions), as well as the percentage $f_\theta$ and $f_u$ of occurrence of extreme events ($\delta_{\theta} > 50^\circ$, $\vert\delta_{u}\vert > 0.6$).
 \textbf{a--b}: Huab station in summer.
@@
 \includegraphics[scale=1]{Figures/Figure4.pdf}
-\caption{Distributions of wind direction at the Huab and Etosha West stations for the
-ERA5-Land climate reanalysis (orange) and the local measurements (blue). In each subplot,
-both distributions are plotted from the same time steps, selected with constraints on the
-wind wind velocity (rows) in the ERA5-Land dataset. The numbers at the top centre give
-the percentage of time steps selected in each sub-range, as well as the percentage of them
-corresponding to the day (between 1000 UTC and 2200 UTC).}
+\caption{Distributions of wind direction at Huab and Etosha West stations for the ERA5-Land climate reanalysis (orange) and the local measurements (blue).
+In each subplot, both distributions are plotted from the same time steps, selected for different ranges of the wind wind velocity (rows) in the ERA5-Land dataset.
+The numbers at the upper right corners give the percentage of time steps selected in each sub-range (top), as well as the percentage of them corresponding to the day -- defined between 10.00 UTC and 22.00 UTC (bottom).}
 \label{Fig4}
@@
 In the case of the Etosha West and Huab stations, the time series of wind speed and direction from the regional predictions quantitatively match those corresponding to the local measurements (Figs.~\ref{Fig3}, ~\ref{Fig4} and Online Resource Fig.~\ref{Fig5_supp}).
-For the North Sand Sea and South Sand Sea stations within the giant dune field, we observe that this agreement is also good, but limited to the October-March time period (Fig.~\ref{Fig4}a, b, e, f).
-However, the field-measured wind roses exhibit additional wind components aligned with the giant dune orientation, as evidenced on the satellite images (Fig.~\ref{Fig2}c,d).
+For the North Sand Sea and South Sand Sea stations within the giant linear dune field, we observe that this agreement is also good, but limited to the October-March time period (Fig.~\ref{Fig4}a, b, e, f).
+However, the field-measured wind roses exhibit additional wind components aligned with the dune orientation, as evidenced on the satellite images (Fig.~\ref{Fig2}c,d).
 
@@
 In the late afternoon and during the night, when the easterly `berg' and katabatic winds blow, measurements and predictions differ.
-In this case, the angular wind distribution of the local measurements exhibits two additional modes corresponding to reversing winds aligned with the giant dune orientation (purple frame in Fig.~\ref{Fig6}, Online Resource Fig.~\ref{Fig6_supp}).
+In this case, the angular wind distribution of the local measurements exhibits two additional modes corresponding to reversing winds aligned with the dune orientation (purple frame in Fig.~\ref{Fig6}, Online Resource Fig.~\ref{Fig6_supp}).
 This deviation is also associated with a general attenuation of the wind strength (Online Resource Fig.~\ref{Fig7_supp}).
-Remarkably, all these figures show that these wind reorientation and attenuation processes occur only at low velocities of the regional wind, typically for $u_{*, \, \textup{ERA}} \lesssim 0.2~\textrm{m}~\textrm{s}^{-1}$.
-For shear velocities larger than $u_{*, \, \textup{ERA}} \simeq 0.3~\textrm{m}~\textrm{s}^{-1}$, the wind reorientation is not apparent.
+Remarkably, all these figures show that these wind reorientation and attenuation processes occur only at low velocities of the regional wind, typically for $u_{*}^{\textup{ERA5-Land}} \lesssim 0.2~\textrm{m}~\textrm{s}^{-1}$.
+For shear velocities larger than $u_{*}^{\textup{ERA5-Land}} \simeq 0.3~\textrm{m}~\textrm{s}^{-1}$, the wind reorientation is not apparent.
 Finally, for intermediate shear velocities, both situations of wind flow reoriented along the dune crest and not reoriented can be successively observed (Online Resource Fig.~\ref{Fig6_supp}).
-Importantly, these values are not precise thresholds, but indicative of a crossover between regimes, whose physical interpretation is discussed in the next section.
+Importantly, these values are not precise thresholds (and certainly not related to the threshold for sediment transport), but indicative of a crossover between regimes, whose physical interpretation is discussed in the next section.
 
@@
 \includegraphics[scale=1]{Figures/Figure5.pdf}
-\caption{Temporal comparison between the wind data coming from the ERA5-Land climate reanalysis (orange lines) and from the local measurements (blue lines).
-Coloured swathes indicate day (between 1000 UTC and 2200 UTC) and night (before 1000 UTC or after 2200 UTC).
-Numbers indicate the average flow deflection $\delta_{\theta}$ and wind modulation $\delta_{u}$ over the given period, as well as the frequency of occurence of extreme events ($\delta_{\theta} > 50^\circ$, $\vert\delta_{u}\vert > 0.6$).
-\textbf{a--b}: North Sand Sea station in summer.
-\textbf{b--c}: North Sand Sea station in winter.
-\textbf{d--e}: South Sand Sea in summer.
-\textbf{f--g}: South Sand Sea station in winter.
-Time series of the two other stations are shown in Fig.~\ref{Fig3}.}
+\caption{Same as Fig.~\ref{Fig3} for North Sand Sea station in summer (\textbf{a--b}), North Sand Sea station in winter (\textbf{b--c}), South Sand Sea station in summer (\textbf{d--e}) and South Sand Sea station in winter (\textbf{f--g}).}
 \label{Fig5}
@@
 \includegraphics[scale=1]{Figures/Figure6.pdf}
-\caption{Distributions of wind direction at the North Sand Sea (\textbf{a}) and South Sand Sea (\textbf{b}) stations for the ERA5-Land climate reanalysis (orange) and the local measurements (blue).
-In each subplot, both distributions are plotted from the same time steps, selected with constraints on the wind direction (columns) and/or wind velocity (rows) of the ERA5-Land dataset.
-The grey vertical dashed lines indicate the dune orientation.
-The numbers at the top right give the percentage of time steps selected in each sub-range, as well as the percentage corresponding to the daytime (between 1000 UTC and 2200 UTC).
-Contrary to the Huab and Etosha West stations (Fig.~\ref{Fig4}), histograms does not match low wind velocities.
-More specifically, the purple frame highlights the regime (low wind velocities, nocturnal easterly wind) in which the data from both datasets differ.}
+\caption{Same as Fig.~\ref{Fig4} but for North Sand Sea (\textbf{a}) and South Sand Sea (\textbf{b}) stations.
+Here, subplots correspond to different ranges for the wind direction (columns) and wind velocity (rows) of the ERA5-Land dataset.
+The grey vertical dashed lines indicate the main dune orientation.
+In contrast with observations at the Huab and Etosha West stations (Fig.~\ref{Fig4}), histograms do not match well at low wind velocities, and the purple frame highlights the regime (low wind velocities, nocturnal easterly wind) in which the data from both datasets differ most.}
 \label{Fig6}
@@
 \includegraphics[scale=1]{Figures/Figure7.pdf}
-\caption{\textbf{a--b}: Vertical profiles of the virtual potential temperature at 2 different time steps (day - 03/11/2015 - 1200 UTC, night - 01/13/2013 - 0900 UTC) at the North Sand Sea station.
+\caption{\textbf{a--b}: Vertical profiles of the virtual potential temperature $T_{\rm vp}$ at two different time steps (day - 03/11/2015 - 12.00 UTC, night - 01/13/2013 - 09.00 UTC) at the North Sand Sea station.
 Dots: data from the ERA5 reanalysis.
@@
 As flow confinement is characterised by a thickness $H$, the interaction between the dunes and the wind in the ABL is well captured by the parameter $k H$.
-The height $H$ is directly related to the radiative fluxes at the Earth surface.
+The height $H$ is directly related to the sensitive heat flux from the Earth surface.
 It is typically on the order of a kilometre, but significantly varies with the circadian and seasonal cycles.
-Emerging and small dunes, with wavelengths in the range $20$ to $100$~m, are not affected by the confinement, corresponding to $k H \gg 1$.
-For giant dunes with kilometric wavelengths, however, their interaction with the FA is significant \citep{Andreotti2009}.
+Emerging and small dunes, with wavelengths in the range $20$ to $100$~m, are not affected by the flow confinement, corresponding to $k H \gg 1$.
+For giant dunes with kilometer-scale wavelengths, however, their interaction with the FA can be significant \citep{Andreotti2009}.
 This translates into a parameter $kH$ in the range $0.02$--$5$, depending on the moment of the day and the season.
@@
 The intensity of the stratification, i.e.
-the amplitude of the gradient $\left | \partial_z \rho \right|$, also impacts its ability to deform the capping layer under the presence of an underlying obstacle, and thus affects the influence of flow confinement.
+the amplitude of the gradient $\left | \partial_z \rho \right|$ in the FA, also impacts the ability to deform the capping layer under the presence of an underlying obstacle, and thus affects the influence of flow confinement.
 This can be quantified using the internal Froude number~\citep{Vosper2004, Stull2006, Sheridan2006, Hunt2006, Jiang2014} $\mathcal{F}_{\textup{I}} = k U/N$, where $N = \sqrt{-g \partial_z \rho / \rho_{0}}$ is the the Brunt-V\"ais\"al\"a frequency \citep{Stull1988}.
-Both Froude numbers have in practice the same qualitative effect on flow confinement, and we shall restrict the main discussion to $\mathcal{F}$ only.
+Both Froude numbers have in practice the same qualitative effect on flow confinement (a smaller Froude corresponding to a stiffer interface), and we shall restrict the main discussion to $\mathcal{F}$ only.
 
@@
 \caption{Regime diagrams of the wind deviation $\delta_{\theta}$ (\textbf{a}) and relative attenuation/amplification $\delta_{u}$ (\textbf{b}) in the space $(\mathcal{F}, \, kH)$, containing the data from both the North Sand Sea and South Sand Sea stations.
-Green dashed lines empirically delimit the different regimes.
+The green dashed lines empirically delimit the different regimes.
 The point density in each bin of the diagrams is shown in Online Resource Fig.~\ref{Fig10_supp} -- 95\% of the data occur in the range $-1 < \delta u < 1$.
@@
 \subsection{Data distribution in the flow regimes}
+\label{DataDistribution}
 We can go one step further and analyse how our data quantitatively spread over the different regimes discussed above.
@@
 \begin{equation}
-\delta_{\textup{u}} = \frac{u_{*}^{\textup{ERA}} -  u_{*}^{\textup{station}}}{u_{*}^{\textup{ERA}}}.
+\delta_{\textup{u}} = \frac{u_{*}^{\textup{ERA5-Land}} -  u_{*}^{\textup{Local mes.}}}{u_{*}^{\textup{ERA5-Land}}}.
 \end{equation}
@@
 Below a crossover value $k H \simeq 0.3$, wind disturbance is less sensitive to the $\mathcal{F}$-value.
-This is probably due to enhanced non-linear effects linked to flow modulation by the obstacle when confinement is strong.
+This is probably due to enhanced non-linear effects linked to flow modulation by the obstacle when confinement is strong (e.g.
+wakes and flow recirculations).
 The Froude number also controls a transition from damped to amplified wind velocities in the interdune, with a crossover around $\mathcal{F} \simeq 0.4$ (Fig.~\ref{Fig8}b).
@@
 
-Note that, in the presented results, the only quantites depending on the choice of the hydrodynamic roughnesses (see Online Resource section~\ref{calib_z0}) are the wind shear velocities, and as such the Froude number $\mathcal{F}r$ and the relative velocity modulation $\delta_{u}$.
-Considering the possible range of realistic roughnesses values ($10^{-4}$m to $10^{-5}$), the uncertainty of velocities estimated using the law of the wall, and thus the Froude number, is at most $30$.
-It is harder to quantify the error on the computation of $\delta_{u}$ considering the method used.
-However, Figure~\ref{Fig14_supp} shows that this choice has very little influence of the tempral variation of the relative velocity modulation, but can induce a global increase of decrease of its values.
-As such, it will not change the aspect of the regime diagram presented in Figure~\ref{Fig8}b, but may only change the $\delta_{u}$--value at which the transition between regimes is observed (here taken as 0, dashed black green lines in Figure~\ref{Fig8}b).
+It is important to discuss the sensitivity of the results with respect to the choice of the hydrodynamic roughnesses (see Online Resource section~\ref{calib_z0}).
+In fact, the only quantities dependent on this choice are those which involve the amplitude of the velocities: wind shear velocities, Froude number $\mathcal{F}$ and relative velocity modulation $\delta_{u}$.
+Those associated with wind direction are independent of this choice.
+Considering the possible range of realistic roughnesses values, the uncertainty on velocities estimated using the law of the wall is at most $30 \%$.
+A similar maximum uncertainty applies to the Froude number.
+This uncertainty also propagates to $\delta_{u}$, for which Figure~\ref{Fig14_supp} shows that the choice of roughness has little influence of its temporal variations, even if it can induce a global increase or decrease of its values.
+As such, the choice of $z_0$ will not qualitatively affect the overall aspect of the regime diagram presented in Figure~\ref{Fig8}b.
+It may only change the crossover value of $\delta_{u}$ at which the transition between regimes is observed (dashed green lines in that figure).
+Our conclusions are thus robust with respect to the somewhat arbitrary choice of the hydrodynamic roughnesses in the use of the data from ERA5-Land reanalysis.
 
@@
 The ellipses indicate the different types of elongating dunes, at large (plain line) and small (dashed line) scales.
-Dune orientation are predicted using the model of \citet{Courrech2014} from the sand flux angular distributions, shown here (roses and resultant transport direction) for typical values (grain size $180~\mu$m, flux-up ratio of $1.6$).
+Dune orientation are predicted using the model of \citet{Courrech2014} from the sand flux angular distributions, shown here (roses) along with the resultant transport direction (brown arrow) for typical values (grain size $180~\mu$m, flux-up ratio of $1.6$).
 Code for corresponding arrows: use of ERA5-Land data (plain line), use of local measurements (dashed line), prediction for the bed instability growth mechanism (blue), prediction for the elongation growth mechanism (red).
 Wedges show the uncertainty on the orientation calculation when parameters are varied.
-The black dots indicate the location of the meteorological stations with respect to the dunes.
+The black dots indicate the location of the meteorological stations in the interdune.
 See Appendix 2 for additional details.}
@@
 As illustrated in Fig.~\ref{Fig9}, small linear dunes ($\sim50$~m wide) are often present in the 1--2~km interdune between giant linear dunes in the Namib Sand Sea \citep{Livingstone2010}.
-These smaller dunes do not exhibit the same orientation as the large ones, and are sometimes named `crossing dunes'.
-Whilst differences between large and small scale dune patterns are observed ubiquitously, they are largely attributed to the presence of two different dune growth mechanisms, leading to two different dune patterns (orientations and/or morphologies) for the same wind regime~\citep{Courrech2014, Runyon2017, lu2017, Song2019, Gadal2020, Hu2021}.
-Here, however, our arguments enable the development of differing orientations for the small and giant linear dunes whilst also imposing the same dune growth mechanism (elongating mode).
+These smaller dunes do not exhibit the same orientation as the large ones, and are sometimes named `crossing dunes'~\citep{Chand22}.
+Whilst differences between large and small scale dune patterns are observed ubiquitously, they are usually attributed to the presence of two different dune growth mechanisms, leading to two different dune patterns (orientations and/or morphologies) for the same wind regime~\citep{Courrech2014, Runyon2017, lu2017, Song2019, Gadal2020, Hu2021}.
+Here, however, our arguments enable the development of differing orientations for the small and giant linear dunes governed by the same dune growth mechanism (elongating mode).
 Figure~\ref{Fig9} shows how the orientations for the small and giant dunes can be derived from the locally measured and regionally predicted winds respectively (red arrows in Fig.~\ref{Fig9}).
-These predictions require a specification for the threshold of eolian sand transport.
-Importantly, its value expressed as a shear velocity $u_{\textup{th}} \simeq 0.15~\textup{m}\textup{s}^{-1}$ is reached in the deflected wind regime already.
+These predictions require to specify the threshold of aeolian sand transport.
+Importantly, its value (a shear velocity estimated $u_{\textup{th}} \simeq 0.15~\textup{m}~\textup{s}^{-1}$) is reached in the winds which are deflected -- recall that the larger winds are not deflected, see Fig.~\ref{Fig6}.
 The feedback of the giant dunes on the wind described in this study thus provides a potential explanation for the existence of these small linear dunes elongating across the interdune, a dynamic which has remained unresolved to date.
@@
 This study presents the evidence that wind flow patterns around giant dunes are influenced by the atmospheric boundary layer, particularly during nocturnal conditions.
-It leaves open the debate as to whether the size of giant dunes is limited by the depth of this layer \citep{Andreotti2009}, in contrast to an unconstrained dune growth, ever-slower with size \citep{Gunn2021b}.
-More field evidence is needed from additional dune fields, but this mechanism would then allow for the inference of the ABL depth from giant bedform wavelengths where measurements are not feasible or available, such as Titan \citep{Lorenz2010}.
+However, we do not address here the question of the limitation of their pattern coarsening, and leave open the debate as to whether the size of giant dunes is controlled by the depth of this layer \citep{Andreotti2009}, in contrast to sediment supply limited and ever-slower growth with size \citep{Werner1999, Gunn2021b}.
+More field evidence is definitively needed from additional dune fields, but this mechanism would allow for the inference of the ABL depth from giant bedform wavelengths where measurements are not feasible or available, such as Titan \citep{Lorenz2010}.
 
-To conclude on conditions under which the ERA5-Land reanalysis data can reliably be used to study dune morphodynamics, we summarise the comparison of local (direct measurements) and regional (climate reanalysis) wind data.
+To conclude on conditions under which the ERA5-Land reanalysis data can reliably be used to study dune morphodynamics, we summarise the comparison of local (direct measurements) and regional (climate reanalysis) wind data as follows.
 In flat areas, the agreement between the two confirms the ability of the ERA5-Land climate reanalysis to predict the wind regime down to scales $\sim10$~km, i.e the model grid.
@@
 These models are able to qualitatively reproduce the observed wind deflection (Appendix 1, Online Resource Figs.~\ref{Fig11_supp} and \ref{Fig13_supp}), and thus provide the physical support for the interpretation we propose here based on hydrodynamic regimes.
-However, these models cannot quantitatively predict the magnitude of these observations, probably due to the presence of expected non-linearities in high confinement situations linked to strong flow modulations.
+However, these models cannot quantitatively predict the magnitude of our observations, probably due to the presence of expected non-linearities in high confinement situations linked to strong flow modulations.
 Besides, these linear calculations only predict wind attenuation in the interdune, in contrast with the observed enhanced velocities associated with particular evening winds from the South during the period October-March (Online Resource Fig.~\ref{Fig12_supp}).
 Some other models predict different spatial flow structures in response to a modulated topography, such as lee waves and rotors \citep{baines1995, Vosper2004}.
-However, our measurements are located at a single point in the interdune, so we are unable to explore these types of responses.
+However, our measurements are located at a single point in the interdune, and we are thus unable to explore these types of responses.
 Data at different places along and across the ridges are needed to investigate and possibly map such flow structures, and for further comparisons with the models.
@@
 Neither the European Commission nor ECMWF is responsible for any use that may be made of the Copernicus Information or Data it contains.
-Fully documented codes used to analyse this study are available at https://github.com/Cgadal/GiantDunes (will be made public upon acceptance of this manuscript for publication).
-
-% ERA5 and ERA5-Land datasets are publicly available at the Copernicus Climate Change Service (C3S) Climate Data Store. As such, It contains modified Copernicus Climate Change
-% Service Information (2021). Neither the European Commission nor ECMWF is
-% responsible for any use that may be made of the Copernicus Information or Data it
-% contains.
-The locally measured wind data can be found at \cg{upload on public data repository}.
-The digital elevation models from the Shuttle Radar Topography Mission are publicly available from Nasa servers, and can be downloaded at https://dwtkns.com/srtm30m/.
-Fully documented codes used to analyse this study are available at https://github.com/Cgadal/GiantDunes (will be made public upon acceptance of this manuscript for publication).
+Documented codes used in this study to analyse this data are available at https://github.com/Cgadal/GiantDunes (will be made public upon acceptance of this manuscript for publication).
 
 Multiple grants have supported the collection of wind data through visits to the four sites between 2013 and 2020 (John Fell Oxford University Press (OUP) Research Fund (121/474); National Geographic (CP-029R-17); Natural Environment Research Council UK (NE/R010196/1 and NE/H021841/1 NSFGEO-NERC); Southampton Marine and Maritime Institute SMMI EPSRC-GCRF UK), along with research permits (1978/2014, 2140/2016, 2304/2017, 2308/2017, RPIV00022018, RPIV0052018, RPIV00230218).
-The authors are very grateful for support from Etosha National Park (especially Shyane Kötting, Boas Erckie, Pierre du Preez, Claudine Cloete, Immanuel Kapofi, Wilferd Versfeld, and Werner Kilian), Gobabeb Namib Research Institute (Gillian Maggs-Kölling and Eugene Marais), The Skeleton Coast National Park (Joshua Kazeurua).
+The authors are very grateful for support from Etosha National Park (especially Shyane K\"otting, Boas Erckie, Pierre du Preez, Claudine Cloete, Immanuel Kapofi, Wilferd Versfeld, and Werner Kilian), Gobabeb Namib Research Institute (Gillian Maggs-K\"olling and Eugene Marais), The Skeleton Coast National Park (Joshua Kazeurua).
 Various researchers and desert enthusiasts have assisted with instruments and the logistics of expeditions, especially Mary Seely for expert guidance at the North Sand Sea site.
 
-Finally, we acknowledge financial support from the Laboratoire d’Excellence UnivEarthS Grant ANR-10-LABX-0023, the Initiative d’Excellence Universite de Paris Grant ANR-18-IDEX-0001, the French National Research Agency Grants ANR-17-CE01-0014/SONO and the National Science Center of Poland Grant 2016/23/B/ST10/01700.
+We thank B.
+Andreotti for useful discussions.
+Finally, we acknowledge financial support from the Laboratoire d'Excellence UnivEarthS Grant ANR-10-LABX-0023, the Initiative d'Excellence Universit\'e de Paris Grant ANR-18-IDEX-0001, the French National Research Agency Grants ANR-17-CE01-0014/SONO and the National Science Center of Poland Grant 2016/23/B/ST10/01700.
 
@@
 \includegraphics[scale=1]{Figures/Figure2_supp.pdf}
-\caption{Comparison between raw local wind measurements, and hourly-averaged data for South Sand Sea station. \textbf{a}: wind direction. \textbf{b}: wind velocity at height $2.6$~m.}
+\caption{Local wind measurements: comparison between raw (blue) and hourly-averaged (light blue) data from South Sand Sea station.
+\textbf{a}: wind direction.
+\textbf{b}: wind velocity at height $2.6$~m.}
 \label{Fig2_supp}
@@
 \caption{Calibration of hydrodynamic roughness.
-The parameter $\delta$ \eqref{metric_roughness} quantifying the difference between local and predicted winds is shown in colorscale as a function of the hydrodynamic roughnesses chosen for the ERA5-Land and for local winds, for the (\textbf{a}) Etosha West, (\textbf{b}) North Sand Sea, (\textbf{c}) Huab and (\textbf{d}) South Sand Sea stations.
+The parameter $\delta$ \eqref{metric_roughness} quantifying the difference between local and predicted winds is shown in color scale as a function of the hydrodynamic roughnesses chosen for the ERA5-Land and for local winds, for the (\textbf{a}) Etosha West, (\textbf{b}) North Sand Sea, (\textbf{c}) Huab and (\textbf{d}) South Sand Sea stations.
 The red dashed and plain lines shows the minima of $\delta$ and the identity line, respectively.
@@
 \textbf{a--d}: Bed elevation detrended by a fitted second order polynomial base-line.
-\textbf{b--e}: Autocorrelation matrix shown in colorscale.
+\textbf{b--e}: Autocorrelation matrix shown in color scale.
 The black line shows the detected dune orientation, and the red line represents the autocorrelation profile along which the dune wavelength is calculated, displayed in \textbf{c--f}.
@@
 \caption{Statistical comparison of the wind orientation between the ERA5-Land dataset and the local measurements for the South Sand Sea and North Sand Sea stations, for different velocity ranges.
-\textbf{a}: $u_{*, \textup{\textup{ERA}}} < 0.1~\textup{m}~\textup{s}^{-1}$.
-\textbf{b}: $0.1 < u_{*, \textup{\textup{ERA}}} \leq 0.25~\textup{m}~\textup{s}^{-1}$.
-\textbf{c}: $u_{*, \textup{\textup{ERA}}} \geq 0.25~\textup{m}~\textup{s}^{-1}$.
+\textbf{a}: $u_{*}^{\textup{ERA5-Land}} < 0.1~\textup{m}~\textup{s}^{-1}$.
+\textbf{b}: $0.1 < u_{*}^{\textup{ERA5-Land}} \leq 0.25~\textup{m}~\textup{s}^{-1}$.
+\textbf{c}: $u_{*}^{\textup{ERA5-Land}} \geq 0.25~\textup{m}~\textup{s}^{-1}$.
 The measured dune orientations are subtracted to the wind orientation, which allows us to plot both stations on the same graph.
@@
 \includegraphics[scale=1]{Figures/Figure8_supp.pdf}
-\caption{\textbf{a}: Vertical profiles of the virtual potential temperature at three different times (blue: 29/11/2012 - 1100 UTC, orange: 21/03/2017 - 1200 UTC, green: 21/03/2017 - 2000 UTC) at the South Sand Sea station.
+\caption{\textbf{a}: Vertical profiles of the virtual potential temperature at three different times (blue: 29/11/2012 - 11.00 UTC, orange: 21/03/2017 - 12.00 UTC, green: 21/03/2017 - 20.00 UTC) at the South Sand Sea station.
 Dots: data from the ERA5 reanalysis.
@@
 Plain lines: vertical (ABL) and linear (FA) fits to estimate the quantities displayed in Online Resource Fig.~\ref{Fig9_supp}.
-\textbf{b}: Examples of ill-processed vertical profiles at three different times (blue: 2/12/2013 - 2300 UTC, orange: 20/03/2017 - 0000 UTC, green: 14/07/2017 - 1400 UTC) at the South Sand Sea station.
-\textbf{c}: Hourly distribution of ill-processed vertical profiles.
-\textbf{d}: Monthly distribution of ill-processed vertical profiles.}
+\textbf{b}: Examples of ill-processed vertical profiles at three different times (blue: 2/12/2013 - 23.00 UTC, orange: 20/03/2017 - 00.00 UTC, green: 14/07/2017 - 14.00 UTC) at the South Sand Sea station.
+Distribution of ill-processed vertical profiles at South (orange) and North (light orange) Sand Sea station: hourly (\textbf{c}) and monthly (\textbf{d}) counts.}
 \label{Fig8_supp}
@@
 \includegraphics[scale=1]{Figures/Figure9_supp.pdf}
-\caption{Distributions of the meteorological parameters resulting from the processing of the ERA5-Land data for the South Sand Sea (blue) and the North Sand Sea (orange) stations.}
+\caption{Distributions of the meteorological parameters resulting from the processing of the ERA5-Land data for the South Sand Sea (orange) and the North Sand Sea (light orange) stations.}
 \label{Fig9_supp}
@@
 \includegraphics[scale=1]{Figures/Figure10_supp.pdf}
-\caption{Non-dimensional parameters distributions. For the marginal distributions, the orange correspond to the South Sand Sea station, and the blue to the North Sand Sea station.}
+\caption{Non-dimensional parameters distributions. For the marginal distributions, the light orange corresponds to the South Sand Sea station, and the orange to the North Sand Sea station.}
 \label{Fig10_supp}
@@
 \includegraphics[scale=1]{Figures/Figure12_supp.pdf}
-\caption{Normalized distributions of amplified velocities for the North sand Sea (blue: $\delta_u < 0$, orange: $\delta_u < -0.5$).
+\caption{Normalized distributions of amplified velocities for the North sand Sea (light orange: $\delta_u < 0$, orange: $\delta_u < -0.5$).
 \textbf{a}: Angular distributions.
@@
 \includegraphics[scale=1]{Figures/Figure14_supp.pdf}
-\caption{Time series of the relative velocity disturbance $\delta_{u}$ corresponding to Fig.~\ref{Fig5}, for different values of the hydroydnamic roughnesses.
-\textbf{a}: North Sand Sea – summer, \textbf{b}: North Sand Sea – winter, \textbf{d}: South Sand Sea – summer, \textbf{e}: South Sand Sea – winter.
-Note that $\delta_{\theta}$ is independant of the choice of $z_{0}^{\textup{Era5Land}}$ and $z_{0}^{\textup{local}}$.}
+\caption{Time series of the relative velocity disturbance $\delta_{u}$ corresponding to Fig.~\ref{Fig5}, for different values of the hydrodynamic roughnesses.
+\textbf{a}: North Sand Sea -- summer, \textbf{b}: North Sand Sea -- winter, \textbf{d}: South Sand Sea -- summer, \textbf{e}: South Sand Sea -- winter.
+Note that $\delta_{\theta}$ is independent of the choice of $z_{0}^{\textup{ERA5-Land}}$ and $z_{0}^{\textup{Local mes.}}$.}
 \label{Fig14_supp}

### 2022-07-04 Paper_figure/Figure05.py

@@ -16,7 +16,7 @@ import matplotlib.transforms as mtransforms
 from datetime import datetime, timedelta
 sys.path.append('../')
 import python_codes.theme as theme
-from python_codes.general import smallestSignedAngleBetween, find_mode_distribution
+from python_codes.general import smallestSignedAngleBetween
 
 
 locale.setlocale(locale.LC_ALL, 'en_US.utf8')
@@ -77,7 +77,6 @@ for i, (subfig, yr, mth, dy, station) in enumerate(zip(subfigs[1:], years, month
     mask = (Data[station]['time'] >= tmin) & (Data[station]['time'] < tmax)
     delta_u = np.abs((Data[station]['U_star_era'][mask] - Data[station]['U_star_insitu'][mask])/Data[station]['U_star_era'][mask])
     Delta = smallestSignedAngleBetween(Data[station]['Orientation_era'][mask], Data[station]['Orientation_insitu'][mask])
-    mode_delta = np.array([find_mode_distribution(Delta, i) for i in np.arange(150, 350)]).mean()
     delta_angle = np.abs(Delta)
     #
     mask_u_theta = (np.abs(delta_u) < 1) & (delta_angle < 85)

### 2022-07-04 Paper_figure/Supplementary_Figures/Figure06_supp.py

@@ -63,7 +63,7 @@ fig, axarr = plt.subplots(3, 1, figsize=(theme.fig_width, 1.3*theme.fig_width),
 for i, (ax, label, x, y) in enumerate(zip(axarr.flatten(), labels, X, Y)):
     plt.sca(ax)
     plot_scatter_surrounded(x % 360, y % 360, color='tab:blue', alpha=0.2)
-    ax.set_ylabel(r'$\theta_{\textup{Local mes.}} - \alpha_{\textup{dune}}$')
+    ax.set_ylabel(r'$\theta^{\textup{Local mes.}} - \alpha^{\textup{dune}}$')
     ax.text(-0.1, 0.98, label, ha='center', va='center', transform=ax.transAxes)
     if i in [0, 1]:
         ax.axhline(180, color='k', linestyle='--')
@@ -74,7 +74,7 @@ for i, (ax, label, x, y) in enumerate(zip(axarr.flatten(), labels, X, Y)):
 
 ax.set_xlim(0, 360)
 ax.set_ylim(0, 360)
-ax.set_xlabel(r'$\theta_{\textup{ERA5-Land}} - \alpha_{\textup{dune}}$')
+ax.set_xlabel(r'$\theta^{\textup{ERA5-Land}} - \alpha^{\textup{dune}}$')
 
 plt.savefig(os.path.join(path_savefig, 'Figure6_supp.pdf'), dpi=400)
 plt.show()

### 2022-07-04 Paper/main.tex
A few modifications on Philippe's round

@@
 \newpage
-
-CITE \citet{Chand22}
 
@@
 These predictions require to specify the threshold of aeolian sand transport.
-Importantly, its value (a shear velocity estimated $u_{\textup{th}} \simeq 0.15~\textup{m}~\textup{s}^{-1}$) is reached in the winds which are deflected -- recall that the larger winds are not deflected, see Fig.~\ref{Fig6}.
-The feedback of the giant dunes on the wind described in this study thus provides a potential explanation for the existence of these small linear dunes elongating across the interdune, a dynamic which has remained unresolved to date.
+Importantly, its value (a shear velocity estimated at $u_{\textup{th}} \simeq 0.15~\textup{m}~\textup{s}^{-1}$) can be reached in periods during which winds are deflected (recall that the stronger winds, responsible for most of the sediment transport and associated dune morphodynamics, are not deflected -- see Fig.~\ref{Fig6}).
+The feedback of the giant dunes on the wind described in this study, through wind delection and attenuation, thus provides a potential explanation for the existence of these small linear dunes elongating across the interdune, a dynamic which has remained unresolved to date.
 These crossing dunes could provide additional constraints for the inference of local winds from bedforms, similarly to that currently performed on Mars using ripple orientations \citep{Liu2015, Hood2021}.
@@
 We would like to acknowledge the contributors of the following open-source python librairies, Matplotlib \citep{Hunter2007}, Numpy \citep{Harris2020} and Scipy \citep{Virtanen2020}, which provide an incredibly efficient ecosystem allowing scientific research in Python.
+We also thank B.
+Andreotti and A.
+Gunn for useful discussions.
 
@@
 
-We thank B.
-Andreotti for useful discussions.
 Finally, we acknowledge financial support from the Laboratoire d'Excellence UnivEarthS Grant ANR-10-LABX-0023, the Initiative d'Excellence Universit\'e de Paris Grant ANR-18-IDEX-0001, the French National Research Agency Grants ANR-17-CE01-0014/SONO and the National Science Center of Poland Grant 2016/23/B/ST10/01700.
@@
 The red dashed and plain lines shows the minima of $\delta$ and the identity line, respectively.
-The orange lines and dots highlight the chosen hydrodynamic roughnesses for the local winds deduced from setting $z_{0}^{\textup{ERA5Land}} = 1$~mm.}
+The orange lines and dots highlight the chosen hydrodynamic roughnesses for the local winds deduced from setting $z_{0}^{\textup{ERA5-Land}} = 1$~mm.}
 \label{Fig3_supp}

### 2022-07-04 Paper/main.tex
fix typos

@@
 Importantly, its value (a shear velocity estimated at $u_{\textup{th}} \simeq 0.15~\textup{m}~\textup{s}^{-1}$) can be reached in periods during which winds are deflected (recall that the stronger winds, responsible for most of the sediment transport and associated dune morphodynamics, are not deflected -- see Fig.~\ref{Fig6}).
-The feedback of the giant dunes on the wind described in this study, through wind delection and attenuation, thus provides a potential explanation for the existence of these small linear dunes elongating across the interdune, a dynamic which has remained unresolved to date.
+The feedback of the giant dunes on the wind described in this study, through wind deflection and attenuation, thus provides a potential explanation for the existence of these small linear dunes elongating across the interdune, a dynamic which has remained unresolved to date.
 These crossing dunes could provide additional constraints for the inference of local winds from bedforms, similarly to that currently performed on Mars using ripple orientations \citep{Liu2015, Hood2021}.

### 2022-07-07 Paper/main.tex
Adding correction from Clément

@@
 %
-We link these discrepancies, with wind deviation and velocity attenuation larger then $50^\circ$ and $60~\%$, to the daily cycle of the turbulent atmospheric boundary layer over a complex topography, and to the associated flow regimes.
-During the night, a shallow boundary layer induces a flow confinement associated with a strong streamline compression above the giant dunes, resulting in large flow deviations, especially for the the lower winds.
-During the day, the flow confinement is reduced by a thicker boundary layer and higher wind velocities, and the feedback of the giant dunes on the atmospheric flow is negligible.
-%
-We finally propose that this mechanism and the resulting wind deflections by the giant dunes could explain the occurrence of smaller-scale secondary dune patterns, elongating along a different orientation compared to the primary dunes between which they develop.
+For the given dune orientation and measured wind regime, we link
+the observed wind deviation (up to 50\textdegree) to the daily cycle of the turbulent
+atmospheric boundary layer. During the night, a shallow boundary layer
+induces a flow confinement above the giant dunes, resulting in large
+flow deviations, especially for the lower easterly winds. During the
+day, the feedback of the giant dunes on the atmospheric flow is much weaker
+due to the thicker boundary layer and higher wind speeds. Finally, we
+propose that the confinement mechanism and the associated wind
+deflections induced by giant dunes could explain the development of smaller-scale secondary
+dunes, elongating obliquely in the interdune areas of the primary dune
+patterns.
 
@@
 Past studies have highlighted two important topographic feedbacks on the wind flow close to the dune/hill surface.
+%
 First is the effect on wind speed, with documented flow acceleration on upwind slopes \citep{Weaver2011} and deceleration on downwind slopes \citep{Baddock2007}, where the speed-up factor is essentially proportional to the obstacle aspect ratio \citep{Jackson1975}.
-Importantly, the velocity maximum is typically shifted upwind of the obstacle crest \citep{Jackson1975, Claudin2013}.
+Under multidirectional wind regimes with frequent wind reversals, this speed-up effect induces large differences in the amplitude and orientation of the resultant sediment transport between flat sand beds and the dune crests \citep{Zhang2014, Rozir2019, Geo2021}.
+In addition, the maximum in wind velocity is typically shifted upwind of the obstacle crest \citep{Jackson1975, Claudin2013}.
 This behaviour has been theoretically predicted by means of asymptotic analysis of a neutrally stratified boundary-layer flow over an obstacle of vanishing aspect ratio \citep{Jackson1975, Mason1979, Sykes1980, Hunt1988, Belcher1998, Kroy2002}.
@@
 Interestingly, a similar behaviour exists for the pressure perturbation, but with a slight downwind shift for the pressure minimum \citep{Claudin2021}.
+
 The second effect, much less studied, is the flow deflection that occurs when the incident wind direction is not perpendicular to the ridge crest.
-While predicted to be small (less than $10^{\circ}$) in the linear regime valid for shallow topography \citep{Gadal2019}, significant flow steering has been reported in the field on the downwind side of steep enough obstacles, such as well-developed sand dunes \citep{Tsoar1983, Sweet1990, Walker2002, Smith2017} and in particular coastal foredunes (e.g.~ \citet{Hunter1983}, \citet{rasmussen1989}, \citet{Walker2006}, \citet{Walker2009}, \citet{Hesp2015}, \citet{Walker2017}, \citet{deWinter2020}), mountain ranges \citep{Kim2000, Lewis2008, Fernando2019}, and valley topographies \citep{Wiggs2002, Garvey2005}.
+While predicted to be small (less than $10^{\circ}$) in the linear regime valid for shallow topography \citep{Gadal2019}, significant flow steering has been reported in the field on the downwind side of steep enough obstacles, such as well-developed sand dunes \citep{Tsoar1983, Sweet1990, Walker2002, Smith2017} and in particular coastal foredunes \citep[e.g.][]{Hunter1983, rasmussen1989, Walker2006, Walker2009, Hesp2015, Walker2017, deWinter2020}, mountain ranges \citep{Kim2000, Lewis2008, Fernando2019}, and valley topographies \citep{Wiggs2002, Garvey2005}.
 
-Wind measurements over sand dunes has been mainly performed over small bedforms, typically a few meters high (corresponding to several tens of meters long) (e.g.~\citet{Mulligan1988}, \citet{Hesp1989}, \citet{Lancaster1996}, \citet{MckennaNeuman1997}, \citet{Sauermann2003}, \citet{Andreotti2002}, \citet{Walker2002}, \citet{Weaver2011}).
+Wind measurements over sand dunes has been mainly performed over small bedforms, typically a few meters high (corresponding to several tens of meters long) \citep[e.g.][]{Mulligan1988, Hesp1989, Lancaster1996, MckennaNeuman1997, Sauermann2003, Andreotti2002, Walker2002, Weaver2011}.
 For practical reasons, fewer studies performed similar measurements on giant dunes \citep{Havholm1988}, with kilometer-scale wavelengths and heights of tens of meters.
@@
 
-It is important to discuss the sensitivity of the results with respect to the choice of the hydrodynamic roughnesses (see Online Resource section~\ref{calib_z0}).
-In fact, the only quantities dependent on this choice are those which involve the amplitude of the velocities: wind shear velocities, Froude number $\mathcal{F}$ and relative velocity modulation $\delta_{u}$.
-Those associated with wind direction are independent of this choice.
-Considering the possible range of realistic roughnesses values, the uncertainty on velocities estimated using the law of the wall is at most $30 \%$.
-A similar maximum uncertainty applies to the Froude number.
-This uncertainty also propagates to $\delta_{u}$, for which Figure~\ref{Fig14_supp} shows that the choice of roughness has little influence of its temporal variations, even if it can induce a global increase or decrease of its values.
-As such, the choice of $z_0$ will not qualitatively affect the overall aspect of the regime diagram presented in Figure~\ref{Fig8}b.
-It may only change the crossover value of $\delta_{u}$ at which the transition between regimes is observed (dashed green lines in that figure).
-Our conclusions are thus robust with respect to the somewhat arbitrary choice of the hydrodynamic roughnesses in the use of the data from ERA5-Land reanalysis.
+As the hydrodynamic roughness $z_0$ determine the magnitude of wind shear velocities, Froude
+number $\mathcal{F}$ and relative velocity modulation $\delta_{u}$, it is important to discuss
+the sensitivity of the results to the $z_0$-values chosen for both the ERA5-Land and the field
+data (see Online Resource section~\ref{calib_z0}). Other quantities associated with wind direction are
+independent of this choice. Considering the possible range of realistic roughness values, the
+uncertainty on velocities estimated using the law of the wall is at most 30~\%. A similar
+maximum uncertainty applies to the Froude number. This uncertainty also propagates to
+$\delta_{u}$, for which Figure~\ref{Fig14_supp} shows that the choice of roughness has little influence on
+its temporal variations,even if it can induce a global increase or decrease of its values. Hence, the choice
+of the $z_0$-values will not qualitatively affect the overall aspect of the regime diagram
+presented in Figure 8b. It may only change the value of $\delta_u$ for which the transition
+between regimes is observed (dashed green lines in Figure~\ref{Fig8}b). Our conclusions are thus robust
+with respect to the somewhat arbitrary choice of the hydrodynamic roughness values.
 

### 2022-07-07 Paper/main.tex
adding comments from Matt and Jo

@@
 \and
-M. Baddock \at
+M.C. Baddock \at
 Geography and Environment, Loughborough University, Loughborough, UK.
@@
 Emergence and growth of sand dunes results from the dynamic interaction between topography, wind flow and sediment transport.
-While feedbacks between these variables are well studied at the scale of a single and relatively small dune, the average effect of a periodic large-scale dune pattern on atmospheric flows remains poorly constrained, due to a lack of data in major sand seas.
-Here, we compare local measurements of surface winds to the predictions of the ERA5-Land climate reanalysis at four locations in Namibia, within and outside the giant linear dune field of the Namib sand sea.
+While feedbacks between these variables are well studied at the scale of a single and relatively small dune, the average effect of a periodic large-scale dune pattern on atmospheric flows remains poorly constrained, due to a pressing lack of data in major sand seas.
+Here, we compare local measurements of surface winds to the predictions of the ERA5-Land climate reanalysis at four locations in Namibia, both within and outside the giant linear dune field of the Namib Sand Sea.
 In the desert plains to the north of the sand sea, observations and predictions agree well.
@@
 For the given dune orientation and measured wind regime, we link
-the observed wind deviation (up to 50\textdegree) to the daily cycle of the turbulent
+the observed wind deviation (over 50\textdegree) to the daily cycle of the turbulent
 atmospheric boundary layer. During the night, a shallow boundary layer
 induces a flow confinement above the giant dunes, resulting in large
-flow deviations, especially for the lower easterly winds. During the
+flow deviations, especially for the slower easterly winds. During the
 day, the feedback of the giant dunes on the atmospheric flow is much weaker
@@
 deflections induced by giant dunes could explain the development of smaller-scale secondary
-dunes, elongating obliquely in the interdune areas of the primary dune
-patterns.
+dunes, which elongate obliquely in the interdune areas of the primary dune
+pattern.
 
@@
 First is the effect on wind speed, with documented flow acceleration on upwind slopes \citep{Weaver2011} and deceleration on downwind slopes \citep{Baddock2007}, where the speed-up factor is essentially proportional to the obstacle aspect ratio \citep{Jackson1975}.
-Under multidirectional wind regimes with frequent wind reversals, this speed-up effect induces large differences in the amplitude and orientation of the resultant sediment transport between flat sand beds and the dune crests \citep{Zhang2014, Rozir2019, Geo2021}.
-In addition, the maximum in wind velocity is typically shifted upwind of the obstacle crest \citep{Jackson1975, Claudin2013}.
+Under multidirectional wind regimes with frequent wind reversals, this speed-up effect induces large differences in the amplitude and orientation of the resultant sediment transport between flat sand beds and the dune crests \citep{Zhang2014, Rozier2019, Gao2021}.
+In addition, the position of maximum velocity is typically shifted upwind of the obstacle crest \citep{Jackson1975, Claudin2013}.
 This behaviour has been theoretically predicted by means of asymptotic analysis of a neutrally stratified boundary-layer flow over an obstacle of vanishing aspect ratio \citep{Jackson1975, Mason1979, Sykes1980, Hunt1988, Belcher1998, Kroy2002}.
@@
 
-Wind measurements over sand dunes has been mainly performed over small bedforms, typically a few meters high (corresponding to several tens of meters long) \citep[e.g.][]{Mulligan1988, Hesp1989, Lancaster1996, MckennaNeuman1997, Sauermann2003, Andreotti2002, Walker2002, Weaver2011}.
-For practical reasons, fewer studies performed similar measurements on giant dunes \citep{Havholm1988}, with kilometer-scale wavelengths and heights of tens of meters.
+Wind measurements over sand dunes have been mainly performed over small bedforms, typically a few meters high (corresponding to several tens of meters long) \citep[e.g.][]{Mulligan1988, Hesp1989, Lancaster1996, MckennaNeuman1997, Sauermann2003, Andreotti2002, Walker2002, Weaver2011}.
+For practical reasons, fewer studies have performed similar measurements on giant dunes \citep{Havholm1988}, with kilometer-scale wavelengths and heights of tens of meters.
 However, such large dunes provide a choice configuration for the study of turbulent flows over a complex topography.
@@
 The agreement is also good in the interdune areas of the sand sea, except for some weak winds blowing at night, which exhibit an additional component aligned with the giant dune orientation.
-These winds are not predicted by the ERA5-Land reanalysis (section 2).
-Further, we are able to link the magnitude of these differences to the circadian cycle of the atmospheric boundary layer (section 3).
-Finally, we draw implications for the wind disturbances on smaller-scale dunes (section 4), suggesting a possible origin for crossing dunes.
+These winds are not predicted by the ERA5-Land reanalysis (section~\ref{sec:sec2}).
+Further, we are able to link the magnitude of these differences to the circadian cycle of the atmospheric boundary layer (section~\ref{sec:sec3}).
+Finally, we draw implications for the wind disturbances on smaller-scale dunes (section~\ref{sec:sec4}), suggesting a possible origin for crossing dunes, a distinctive secondary dune form observed in the Namib and other sand seas.
 
@@
 \section{Wind regimes across the Namib Sand Sea}
+\label{sec:sec2}
 We measured the wind regime at four different locations in Namibia, representative of various arid environments across the Namib desert (Fig.~\ref{Fig1}, Fig.~\ref{Fig2}).
 The Etosha West station was located at the Adamax waterhole to the west of Etosha Pan in northern Namibia, in a sparsely vegetated area.
-The Huab station was near the coast on a hyper-arid flat gravel plain lying north the ephemeral Huab river.
+The Huab station was near the coast on a hyper-arid flat gravel plain lying north of the ephemeral Huab river.
 Here, barchan dunes up to a few meters in height develop from the sediment blowing out of the river valley \citep{Nield2017, Hesp1998}.
@@
 We checked that at least one complete seasonal cycle was available for each station.
-Regional winds were extracted at the same locations and periods from the ERA5-Land dataset, which is a replay at a smaller spatial resolution of ERA5, the latest climate reanalysis from the ECMWWF \citep{Hersbach2020, munoz2021}.
+Regional winds were extracted at the same locations and periods from the ERA5-Land dataset, which is a replay at a smaller spatial resolution of ERA5, the latest climate reanalysis from the ECMWF \citep{Hersbach2020, munoz2021}.
 This dataset provided hourly predictions of the 10-m wind velocity and direction at a spatial resolution of $0.1^\circ\times0.1^\circ$ ($\simeq 9~\textup{km}$ in Namibia).
@@
 \section{Influence of wind speed and circadian cycle on the atmospheric boundary layer}
-
+\label{sec:sec3}
 The wind deflection induced by dunes has previously been related to the incident angle between wind direction and crest orientation, with a maximum deflection evident for incident angles between $30^{\circ}$ and $70^{\circ}$ \citep{Walker2009, Hesp2015}.
@@
 \textbf{c--f}: Sketches representing the interaction between the giant dunes and the atmospheric flow for different meteorological conditions.
-\textbf{g}: Streamlines over a sinusoidal topography $\xi(x,y)$ qualitatively representing the effect of low, medium and strong flow confinement, in relation to the above panels (see Appendix~1 for more details).
+\textbf{g}: Streamlines over a sinusoidal topography $\xi(x,y)$ qualitatively representing the effect of low, medium and strong flow confinement, in relation to the above panels (see Appendix~\ref{turbulent_wind_model} for more details).
 The blue arrow indicates the undisturbed wind direction.}
@@
 %
-These two quantities are represented as maps in the plane $(\mathcal{F}, \, kH)$ (Fig.~\ref{Fig8}a,b), and one can clearly identify different regions in these graphs.
+These two quantities are represented as maps in the plane $(\mathcal{F}, \, kH)$ (Fig.~\ref{Fig8}a,b), and one can clearly identify different regions in these plots.
 Small wind disturbances (small $\delta_{\theta}$ and $\delta_{\textup{u}}$) are located in the top-right part of the diagrams, corresponding to a regime with low-interaction as well as low-confinement ($k H$ and $\mathcal{F}$ large enough, Fig.~\ref{Fig7}d).
@@
 The black dots indicate the location of the meteorological stations in the interdune.
-See Appendix 2 for additional details.}
+See Appendix~\ref{app:app2} for additional details.}
 \label{Fig9}
@@
 \section{Discussion and conclusion}
-The feedback of the giant dunes on the wind flow has important implications for smaller scales bedforms.
-As illustrated in Fig.~\ref{Fig9}, small linear dunes ($\sim50$~m wide) are often present in the 1--2~km interdune between giant linear dunes in the Namib Sand Sea \citep{Livingstone2010}.
+\label{sec:sec4}
+
+The feedback of the giant dunes on the wind flow has important implications for smaller scale bedforms.
+As illustrated in Fig.~\ref{Fig9}, small linear dunes ($\sim50$~m wide) are often present within the 1--2~km interdune spaces between giant linear dunes in the Namib Sand Sea \citep{Livingstone2010}.
 These smaller dunes do not exhibit the same orientation as the large ones, and are sometimes named `crossing dunes'~\citep{Chand22}.
@@
 Figure~\ref{Fig9} shows how the orientations for the small and giant dunes can be derived from the locally measured and regionally predicted winds respectively (red arrows in Fig.~\ref{Fig9}).
-These predictions require to specify the threshold of aeolian sand transport.
-Importantly, its value (a shear velocity estimated at $u_{\textup{th}} \simeq 0.15~\textup{m}~\textup{s}^{-1}$) can be reached in periods during which winds are deflected (recall that the stronger winds, responsible for most of the sediment transport and associated dune morphodynamics, are not deflected -- see Fig.~\ref{Fig6}).
+These predictions require the threshold of aeolian sand transport to be specified.
+Importantly, its value (a shear velocity estimated at $u_{\textup{th}} \simeq 0.15~\textup{m}~\textup{s}^{-1}$ -- see Appendix~\ref{app:app2}) can be reached in periods during which deflected winds are observed (recall that the stronger winds, responsible for most of the sediment transport and associated dune morphodynamics, are not deflected -- see Fig.~\ref{Fig6}).
 The feedback of the giant dunes on the wind described in this study, through wind deflection and attenuation, thus provides a potential explanation for the existence of these small linear dunes elongating across the interdune, a dynamic which has remained unresolved to date.
@@
 
-This study presents the evidence that wind flow patterns around giant dunes are influenced by the atmospheric boundary layer, particularly during nocturnal conditions.
-However, we do not address here the question of the limitation of their pattern coarsening, and leave open the debate as to whether the size of giant dunes is controlled by the depth of this layer \citep{Andreotti2009}, in contrast to sediment supply limited and ever-slower growth with size \citep{Werner1999, Gunn2021b}.
+This study presents field and reanalysis-based evidence that wind flow patterns around giant dunes are influenced by the atmospheric boundary layer, particularly during nocturnal conditions.
+However, we do not address here the question of the limitation of the giant dune pattern coarsening, and leave open the debate as to whether their size is controlled by the depth of this layer \citep{Andreotti2009}, in contrast to sediment supply limited and ever-slower growth with size \citep{Werner1999, Gunn2022}.
 More field evidence is definitively needed from additional dune fields, but this mechanism would allow for the inference of the ABL depth from giant bedform wavelengths where measurements are not feasible or available, such as Titan \citep{Lorenz2010}.
@@
 when the aspect ratio of these ridges is small ($k\xi_0 \ll 1$).
-These models are able to qualitatively reproduce the observed wind deflection (Appendix 1, Online Resource Figs.~\ref{Fig11_supp} and \ref{Fig13_supp}), and thus provide the physical support for the interpretation we propose here based on hydrodynamic regimes.
+These models are able to qualitatively reproduce the observed wind deflection (Appendix~\ref{turbulent_wind_model}, Online Resource Figs.~\ref{Fig11_supp} and \ref{Fig13_supp}), and thus provide the physical support for the interpretation we propose here based on hydrodynamic regimes.
 However, these models cannot quantitatively predict the magnitude of our observations, probably due to the presence of expected non-linearities in high confinement situations linked to strong flow modulations.
-Besides, these linear calculations only predict wind attenuation in the interdune, in contrast with the observed enhanced velocities associated with particular evening winds from the South during the period October-March (Online Resource Fig.~\ref{Fig12_supp}).
+Besides, these linear calculations only predict wind attenuation in the interdune, in contrast with the observed enhanced velocities associated with particular evening winds from the south during the period October-March (Online Resource Fig.~\ref{Fig12_supp}).
 Some other models predict different spatial flow structures in response to a modulated topography, such as lee waves and rotors \citep{baines1995, Vosper2004}.
@@
 \section*{Appendix 2: Sediment transport and dune morphodynamics}
+\label{app:app2}
 
@@
 Two orientations are possible depending on the mechanism dominating the dune growth: elongation or bed instability.
-The orientation $\alpha$ corresponding the bed instability is then the one that maximises the following growth rate \citep{Rubin1987}:
+The orientation $\alpha$ corresponding to the bed instability is then the one that maximises the following growth rate \citep{Rubin1987}:
 %
@@
 
-Examples of vertical profiles of the virtual potential temperature deduced from ERA5 are shown in Online Resource Fig.~\ref{Fig8_supp}a.
+Examples of vertical virtual potential temperature profiles deduced from ERA5 are shown in Online Resource Fig.~\ref{Fig8_supp}a.
 For each of them, an average temperature is computed below the ABL depth ($z<H$), and a linear function is fitted above, allowing us to extract the temperature jump $\Delta T_{\textup{vp}}$.
@@
 \includegraphics[scale=1]{Figures/Figure1_supp.pdf}
-\caption{Gant chart representing the valid time steps for the two data sets, for all stations.}
+\caption{Gantt chart representing the valid time steps for the two data sets, for all stations.}
 \label{Fig1_supp}

### 2022-07-13 Paper/main.tex
Splitting acknowledgements section

@@
 
-All data used in this study can be found in \citet{DataPaper}.
+\end{acknowledgements}
+
+{\footnotesize\noindent\textbf{Data Availability}~All data used in this study can be found in \citet{DataPaper}.
 Note that it contains modified Copernicus Climate Change Service Information (2021).
 Neither the European Commission nor ECMWF is responsible for any use that may be made of the Copernicus Information or Data it contains.
-Documented codes used in this study to analyse this data are available at https://github.com/Cgadal/GiantDunes (will be made public upon acceptance of this manuscript for publication).
+Documented codes used in this study to analyse this data are available at https://github.com/Cgadal/GiantDunes (will be made public upon acceptance of this manuscript for publication).}
 
-Multiple grants have supported the collection of wind data through visits to the four sites between 2013 and 2020 (John Fell Oxford University Press (OUP) Research Fund (121/474); National Geographic (CP-029R-17); Natural Environment Research Council UK (NE/R010196/1 and NE/H021841/1 NSFGEO-NERC); Southampton Marine and Maritime Institute SMMI EPSRC-GCRF UK), along with research permits (1978/2014, 2140/2016, 2304/2017, 2308/2017, RPIV00022018, RPIV0052018, RPIV00230218).
+\vspace{.2cm }
+
+{\footnotesize\noindent\textbf{Fundings}~Multiple grants have supported the collection of wind data through visits to the four sites between 2013 and 2020 (John Fell Oxford University Press (OUP) Research Fund (121/474); National Geographic (CP-029R-17); Natural Environment Research Council UK (NE/R010196/1 and NE/H021841/1 NSFGEO-NERC); Southampton Marine and Maritime Institute SMMI EPSRC-GCRF UK), along with research permits (1978/2014, 2140/2016, 2304/2017, 2308/2017, RPIV00022018, RPIV0052018, RPIV00230218).
 The authors are very grateful for support from Etosha National Park (especially Shyane K\"otting, Boas Erckie, Pierre du Preez, Claudine Cloete, Immanuel Kapofi, Wilferd Versfeld, and Werner Kilian), Gobabeb Namib Research Institute (Gillian Maggs-K\"olling and Eugene Marais), The Skeleton Coast National Park (Joshua Kazeurua).
 Various researchers and desert enthusiasts have assisted with instruments and the logistics of expeditions, especially Mary Seely for expert guidance at the North Sand Sea site.
+%
+Finally, we acknowledge financial support from the Laboratoire d'Excellence UnivEarthS Grant ANR-10-LABX-0023, the Initiative d'Excellence Universit\'e de Paris Grant ANR-18-IDEX-0001, the French National Research Agency Grants ANR-17-CE01-0014/SONO and the National Science Center of Poland Grant 2016/23/B/ST10/01700.}
 
-Finally, we acknowledge financial support from the Laboratoire d'Excellence UnivEarthS Grant ANR-10-LABX-0023, the Initiative d'Excellence Universit\'e de Paris Grant ANR-18-IDEX-0001, the French National Research Agency Grants ANR-17-CE01-0014/SONO and the National Science Center of Poland Grant 2016/23/B/ST10/01700.
-
-\end{acknowledgements}
+\vspace{.2cm }
+{\footnotesize\noindent\textbf{Conflict of interest}~All authors declare that they have no conflicts of interest.}
 

### 2022-07-28 Paper/main.tex
Updating paper with proofs

@@
 
-\title{Local wind regime induced by giant linear dunes: comparison of ERA5-Land reanalysis with surface measurements}
+\title{Local Wind Regime Induced by Giant Linear Dunes: Comparison of ERA5-Land Reanalysis with Surface Measurements}
 
@@
 Energy and Environment Institute, University of Hull, Hull, UK.
+\at
+School of Geography and Environmental Science, University of Southampton, Southampton, UK.
 \and
 C. Narteau \at
-Institut de Physique du Globe de Paris, Universit\'e de Paris, CNRS, Paris, France.
+Université Paris Cité, Institut de physique du globe de Paris, CNRS, Paris, France
 \and
-G. Wiggs \at
+Giles F.S. Wiggs \at
 School of Geography and the Environment, University of Oxford, Oxford, UK.
@@
 The agreement is also good in the interdune areas of the sand sea, except for some weak winds blowing at night, which exhibit an additional component aligned with the giant dune orientation.
-These winds are not predicted by the ERA5-Land reanalysis (section~\ref{sec:sec2}).
-Further, we are able to link the magnitude of these differences to the circadian cycle of the atmospheric boundary layer (section~\ref{sec:sec3}).
-Finally, we draw implications for the wind disturbances on smaller-scale dunes (section~\ref{sec:sec4}), suggesting a possible origin for crossing dunes, a distinctive secondary dune form observed in the Namib and other sand seas.
+These winds are not predicted by the ERA5-Land reanalysis (Sect.~\ref{sec:sec2}).
+Further, we are able to link the magnitude of these differences to the circadian cycle of the atmospheric boundary layer (Sect.~\ref{sec:sec3}).
+Finally, we draw implications for the wind disturbances on smaller-scale dunes (Sect.~\ref{sec:sec4}), suggesting a possible origin for crossing dunes, a distinctive secondary dune form observed in the Namib and other sand seas.
 
@@
   \includegraphics[scale=1]{Figures/Figure1.pdf}
-  \caption{Studied field sites. \textbf{a}: Location of the different sites in Namibia. \textbf{b--e}: Photographs of the meteorological stations.}
+  \caption{Studied field sites. \textbf{a} Location of the different sites in Namibia. \textbf{b-e}: Photographs of the meteorological stations.}
   \label{Fig1}
@@
 %---------------------------------------------------------------------------------------
-\section{Wind regimes across the Namib Sand Sea}
+\section{Wind Regimes Across The Namib Sand Sea}
 \label{sec:sec2}
-We measured the wind regime at four different locations in Namibia, representative of various arid environments across the Namib desert (Fig.~\ref{Fig1}, Fig.~\ref{Fig2}).
+We measured the wind regime at four different locations in Namibia, representative of various arid environments across the Namib desert (Figs.~\ref{Fig1}~and~\ref{Fig2}).
 The Etosha West station was located at the Adamax waterhole to the west of Etosha Pan in northern Namibia, in a sparsely vegetated area.
@@
 %------------------------
-\subsection{Wind and elevation data}
+\subsection{Wind and Elevation Data}
 At each meteorological station (Fig.~\ref{Fig1}), wind speed and direction were sampled every 10 minutes using cup anemometers (Vector Instruments A100-LK) and wind vanes (Vector Instruments W200-P) at a single height, which was between 2~m and 3~m depending on the station.
@@
 Coloured swathes indicate day (between 10.00 UTC and 22.00 UTC) and night (before 10.00 UTC or after 22.00 UTC).
-Numbers in legends indicate the average flow deflection $\delta_{\theta}$ and relative wind modulation $\delta_{u}$ over the displayed period (see section~\ref{DataDistribution} for their definitions), as well as the percentage $f_\theta$ and $f_u$ of occurrence of extreme events ($\delta_{\theta} > 50^\circ$, $\vert\delta_{u}\vert > 0.6$).
-\textbf{a--b}: Huab station in summer.
-\textbf{b--c}: Huab station in winter.
-\textbf{d--e}: Etosha West station in summer.
-\textbf{f--g}: Etosha West station in winter.
+Numbers in legends indicate the average flow deflection $\delta_{\theta}$ and relative wind modulation $\delta_{u}$ over the displayed period (see Sect.~\ref{DataDistribution} for their definitions), as well as the percentage $f_\theta$ and $f_u$ of occurrence of extreme events ($\delta_{\theta} > 50^\circ$, $\vert\delta_{u}\vert > 0.6$).
+\textbf{a,b}: Huab station in summer.
+\textbf{b,c}: Huab station in winter.
+\textbf{d,e}: Etosha West station in summer.
+\textbf{f,g}: Etosha West station in winter.
 Time series of the two other stations are shown in Fig.~\ref{Fig5}.}
@@
 %------------------------
-\subsection{Comparison of local and regional winds}
+\subsection{Comparison of Local and Regional Winds}
 \label{section_data_feedback}
@@
 In the Namib, the regional wind patterns are essentially controlled by the sea breeze, resulting in strong northward components (sometimes slightly deviated by the large scale topography) present in all regional wind roses \citep{lancaster1985}.
-These daytime winds are dominant during the period October-March (Fig.~\ref{Fig3}f and Online Resource Fig.~\ref{Fig4}f).
-During April-September, an additional (and often nocturnal) easterly component can also be recorded, induced by the combination of katabatic winds forming in the mountains, and infrequent `berg' winds, which are responsible for the high wind velocities observed \citep{lancaster1984}.
+These daytime winds are dominant during the period October--March (Fig.~\ref{Fig3}f and Online Resource Fig.~\ref{Fig4}f).
+During April--September, an additional (and often nocturnal) easterly component can also be recorded, induced by the combination of katabatic winds forming in the mountains, and infrequent `berg' winds, which are responsible for the high wind velocities observed \citep{lancaster1984}.
 The frequency of these easterly components decreases from inland to the coast.
-As a result, bidirectional wind regimes within the Namib Sand Sea and at the west Etosha site (Fig.~\ref{Fig2}a,c,d) and a unidirectional wind regime on the coast at the outlet of the Huab River (Fig.~\ref{Fig2}b) are observed.
+As a result, bidirectional wind regimes within the Namib Sand Sea and at the west Etosha site (Fig.~\ref{Fig2}a,c and d) and a unidirectional wind regime on the coast at the outlet of the Huab River (Fig.~\ref{Fig2}b) are observed.
 
 In the case of the Etosha West and Huab stations, the time series of wind speed and direction from the regional predictions quantitatively match those corresponding to the local measurements (Figs.~\ref{Fig3}, ~\ref{Fig4} and Online Resource Fig.~\ref{Fig5_supp}).
-For the North Sand Sea and South Sand Sea stations within the giant linear dune field, we observe that this agreement is also good, but limited to the October-March time period (Fig.~\ref{Fig4}a, b, e, f).
-However, the field-measured wind roses exhibit additional wind components aligned with the dune orientation, as evidenced on the satellite images (Fig.~\ref{Fig2}c,d).
+For the North Sand Sea and South Sand Sea stations within the giant linear dune field, we observe that this agreement is also good, but limited to the October-March time period (Fig.~\ref{Fig4}a, b and e, f).
+However, the field-measured wind roses exhibit additional wind components aligned with the dune orientation, as evidenced on the satellite images (Fig.~\ref{Fig2}c and d).
 
-More precisely, during the April-September period, the local and regional winds in the interdune match during daytime only, i.e when the southerly/southwesterly sea breeze dominates (Figs.~\ref{Fig5}c,d,g,h and \ref{Fig6}).
+More precisely, during the April--September period, the local and regional winds in the interdune match during daytime only, i.e when the southerly/southwesterly sea breeze dominates (Figs.~\ref{Fig5}c,d,g,h and \ref{Fig6}).
 In the late afternoon and during the night, when the easterly `berg' and katabatic winds blow, measurements and predictions differ.
@@
 \includegraphics[scale=1]{Figures/Figure5.pdf}
-\caption{Same as Fig.~\ref{Fig3} for North Sand Sea station in summer (\textbf{a--b}), North Sand Sea station in winter (\textbf{b--c}), South Sand Sea station in summer (\textbf{d--e}) and South Sand Sea station in winter (\textbf{f--g}).}
+\caption{Same as Fig.~\ref{Fig3} for North Sand Sea station in summer (\textbf{a,b}), North Sand Sea station in winter (\textbf{b,c}), South Sand Sea station in summer (\textbf{d,e}) and South Sand Sea station in winter (\textbf{f,g}).}
 \label{Fig5}
@@
 %---------------------------------------------------------------------------------------
-\section{Influence of wind speed and circadian cycle on the atmospheric boundary layer}
+\section{Influence of Wind Speed and Circadian Cycle on the Atmospheric Boundary Layer}
 \label{sec:sec3}
@@
 \includegraphics[scale=1]{Figures/Figure7.pdf}
-\caption{\textbf{a--b}: Vertical profiles of the virtual potential temperature $T_{\rm vp}$ at two different time steps (day - 03/11/2015 - 12.00 UTC, night - 01/13/2013 - 09.00 UTC) at the North Sand Sea station.
+\caption{\textbf{a,b}: Vertical profiles of the virtual potential temperature $T_{\rm vp}$ at two different time steps (day - 03/11/2015 - 12.00 UTC, night - 01/13/2013 - 09.00 UTC) at the North Sand Sea station.
 Dots: data from the ERA5 reanalysis.
@@
 %------------------------
-\subsection{Flow over a modulated bed}
+\subsection{Flow Over a Modulated Bed}
 \label{theoretical_framework}
@@
 
-With this theoretical framework in mind, and in the context of the measured wind data in the North and South Sand Sea stations, the smallest wind disturbances are expected to occur during the day, when the ABL depth is the largest and comparable to the dune wavelength ($k H \gtrsim 1$), which corresponds to a weak confinement situation (Fig.~\ref{Fig7}c,d).
+With this theoretical framework in mind, and in the context of the measured wind data in the North and South Sand Sea stations, the smallest wind disturbances are expected to occur during the day, when the ABL depth is the largest and comparable to the dune wavelength ($k H \gtrsim 1$), which corresponds to a weak confinement situation (Fig.~\ref{Fig7}c and d).
 In contrast, large wind disturbances are expected to occur during the night, when the confinement is mainly induced by a shallow ABL (Fig.~\ref{Fig7}e).
@@
 %------------------------
-\subsection{Data distribution in the flow regimes}
+\subsection{Data Distribution in the Flow Regimes}
 \label{DataDistribution}
@@
 %
-These two quantities are represented as maps in the plane $(\mathcal{F}, \, kH)$ (Fig.~\ref{Fig8}a,b), and one can clearly identify different regions in these plots.
+These two quantities are represented as maps in the plane $(\mathcal{F}, \, kH)$ (Fig.~\ref{Fig8}a and b), and one can clearly identify different regions in these plots.
 Small wind disturbances (small $\delta_{\theta}$ and $\delta_{\textup{u}}$) are located in the top-right part of the diagrams, corresponding to a regime with low-interaction as well as low-confinement ($k H$ and $\mathcal{F}$ large enough, Fig.~\ref{Fig7}d).
@@
 the sensitivity of the results to the $z_0$-values chosen for both the ERA5-Land and the field
-data (see Online Resource section~\ref{calib_z0}). Other quantities associated with wind direction are
+data (see Online Resource Sect.~\ref{calib_z0}). Other quantities associated with wind direction are
 independent of this choice. Considering the possible range of realistic roughness values, the
@@
 %---------------------------------------------------------------------------------------
-\section{Discussion and conclusion}
+\section{Discussion and Conclusion}
 \label{sec:sec4}
@@
 \vspace{.2cm }
-{\footnotesize\noindent\textbf{Conflict of interest}~All authors declare that they have no conflicts of interest.}
+{\footnotesize\noindent\textbf{Conflict of interest}~The authors declare that they have no conflict of interest.}
 
 %---------------------------------------------------------------------------------------
-\section*{Appendix 1: Linear theory of wind response to topographic perturbation}
+\section*{Appendix 1: Linear Theory of Wind Response to Topographic Perturbation}
 \label{turbulent_wind_model}
@@
 %---------------------------------------------------------------------------------------
-\section*{Appendix 2: Sediment transport and dune morphodynamics}
+\section*{Appendix 2: Sediment Transport and Dune Morphodynamics}
 \label{app:app2}
@@
 %------------------------
-\paragraph{Sediment transport} ---
+\paragraph{Sediment Transport} ---
 The prediction of sand fluxes from wind data has been a long standing issue in aeolian geomorphological studies~\citep{Fryberger79, Pearce2005, Sherman2012, Shen2019}.
@@
 %------------------------
-\paragraph{Dune orientations} ---
+\paragraph{Dune Orientations} ---
 Dune orientations are predicted with the dimensional model of \citet{Courrech2014}, from the sand flux time series computed with the above transport law.
